### 2020-11-28 article.tex
Initial Overleaf Import

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Anonymous Giraffe\\Universit\"at Innsbruck \And 
        Second Author\\Plus Affiliation}
\title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Anonymous Giraffe, Second Author} %% comma-separated
\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using \texttt{scoringutils} in R} %% without formatting
\Shorttitle{\pkg{foo}: A Capitalized Title} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  Forecasts play a role in many scientific fields from finance to meteorology and epidemiology. With the emergence of Covid-19 the role of forecasting to inform public policy has again attracted increased attention. In this paper we introduce a complete framework for evaluating different types of forecasts using the R package scoringutils. We discuss different appropriate evaluation metrics and apply the framework to XXXXX. 
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Nikos Bosse\\
  Centre for Mathematical Modelling of Infectious Diseases\\
%%  Faculty of Economics and Statistics\\
  London School of Hygiene and Tropical Medicine\\
%%  6020 Innsbruck, Austria\\
  E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
  %% URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section[introduction]{Introduction}
Good forecasts are of great interest to decision makers in almost every field. An integral part of assessing and improving their usefulness is forecast evaluation. For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts [some citations]. Yet even with this rich body of research available, implementing a consistent framework in R to evaluate forecasts is not trivial. We therefore present the scoringutils pacakge. The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users access to a consistent evaluation framework. The remainder of the paper is structured as follows. First, different types of forecasts will be discussed. Then the general forecasting paradigm will be reviewed and discussed. Next, the scoringutils package will be introduced and the different metrics included will be discussed. Lastly, the evaluation framework will be applied to forecasts of death numbers from Covid-19 in the United States.

\end{document}

### 2020-12-07 article.tex
Updates from Overleaf

@@
 \documentclass[article]{jss}
-
+\usepackage{amsmath}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Yet even with this rich body of research available, implementing a consistent framework in R to evaluate forecasts is not trivial.
-We therefore present the scoringutils pacakge.
-The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users access to a consistent evaluation framework.
-The remainder of the paper is structured as follows.
-First, different types of forecasts will be discussed.
-Then the general forecasting paradigm will be reviewed and discussed.
-Next, the scoringutils package will be introduced and the different metrics included will be discussed.
-Lastly, the evaluation framework will be applied to forecasts of death numbers from Covid-19 in the United States.
+We therefore present the \pkg{scoringutils} pacakge.
+The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts.
+In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK.
+
+\subsection{Forecast types}
+
+In its most general sense, a forecast is the forecaster’s stated belief about the future.
+For conceptual clarity, it is however useful to distinguish between different times of forecasts.
+Forecasts can be either quantitative or qualitative, but \pkg{scoringutils} only deals with the former.
+In terms of quantitative forecasts we can distinguish point forecasts from probabilistic forecasts.
+A point forecast states a single number and is the simplest form of a forecast.
+A point forecast is limited in its usefulness, as it does not give information about the uncertainty of the foreacst.
+A very certain forecast may, for example, warrant a very different course of actions than does a very uncertain one.
+A probabilistic forecast, in contrast, is a forecast made in terms of the entire predictive distribution.
+Providing the entire predictive distribution allows the forecaster to express their belief about all aspects of the underlying data-generating distribution (including e.g.
+skewness or the width of its tails).
+Probabilistic forecasts are therefore the focus of this paper as well as the \pkg{scoringutils} package.
+
+\subsection{Forecast formats}
+
+Not only can we distinguish different forecast types, but also different formats in which predictions may be reported.
+This distinction is important, as different formats and prediction types require slightly different evaluation approaches.
+Forecasts are usually expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
+Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
+They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
+For that reason, usually quantiles are reported instead [citation FORECAST HUBS].
+As an alternative representation of a quantile forecast, one can also express predictions using central prediction intervals.
+In this case, the forecaster reports a range to denote the interval as well as a value for the lower and upper bound.
+A forecaster could also in principle state their forecasts in a binary way by defining an outcome and assigning a probability that the outcome will come true.
+This type of forecasting is common in many classification problems.
+If we think of the outcome of some number being larger or smaller than a certain value, we can immediately see the connection between binary predictions and the concept of a cumulative distribution functions (CDF).
+All of these forecast types can be evaluated using \pkg{scoringutils}.
+The general forecasting paradigm [GNEITING] that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+
+\subsection{The forecasting paradigm}
+
+Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007].
+For an ideal forecast, we therefore have
+$$ P_t = F_t, $$
+where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution.
+As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
+[CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
+Calibration refers to the statistical consistency between the predictive distribution and the observations.
+A well calibrated forecast does not systematically differ deviate from the observed values.
+For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007].
+Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
+how precise the forecasts are.
+The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
+Take for example the task of predicting rainfall in a city like London.
+A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time.
+On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
+Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
+DO I WANT THAT FIGURE?
+
+\section[data]{Data: UK short-term forecasts}
+To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
+Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
+The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
+Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
+The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL]
+
+- timing (weekly?) and number of forecast dates
+- the four targets
+- the models. 
+- is the set complete?
+
+PLOT WITH OVERVIEW OF MISSING FORECASTS LIKE THE ONE THEY MADE FOR THE FORECAST HUB.
+
+\section[run-evaluation]{Running an automated evaluation}
+A full evaluation of all forecasts based on observed values can be performed using the function \code{eval_forecasts}.
+The function automatically recognises the prediction type and format and applies the appropriate scoring metrics.
+Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
+
+All it requires is a \code{data.frame} or similar that has a column called "prediction" and one called "true\_value".
+Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
+To get a sense of the required input formats, example data for each format is provided with the package and shown in the package vignette.
+We also included functionality to transform between various formats.
+Additional columns may be present to indicate a grouping of forecasts.
+One common use case is a set of forecasts made by different models for various locations at different time points, each for several weeks into the future.
+In this case, we would expect additional columns that could be called "model", "date", "forecast\_date", "forecast\_horizon" and "location".
+Through the by argument, the user needs to to specify the unit of a single forecast.
+In the above example we would set by = c("model", "date", "forecast\_date", "forecast\_horizon", "location").
+Columns such as quantile or sample should not be included, because several quantiles or samples make up one forecast.
+By default, by = NULL and scoringutils will automatically use all appropriate present columns.
+
+The \code{summarise_by} argument allows the user to choose categories to aggregate over.
+If we were only interested in scores for the different models, we would for example specify \code{summarise_by = c("model")}.
+If we wanted to have scores for every model in every location, we would specify \code{summarise_by = c("model", "location")}.
+If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
+
+When aggregating, \code{eval_forecasts} takes the mean according to the group defined in \code{summarise_by}.
+In the above example, if \code{summarise_by = c("model", "location")}, scores would be averaged over all forecast dates, forecast horizons and quantiles to yield one score per model and location.
+In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying sd = TRUE and for example quantiles = c(0.5) for the median.
+
+To evaluate the UK short term forecast data with scoringutils, we first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands:
+
+\begin{CodeInput}
+# install and load data pacakge from external repository
+remotes::install_github("sbfnk/covid19.forecasts.uk")
+library(covid19.forecasts.uk)
+
+# load truth data
+data(covid_uk_data)
+
+# load forecasts
+data(uk_forecasts)
+\end{CodeInput}
+
+ADD OUTPUT HERE TO SHOW THE FIRST COUPLE OF ROWS (MUST BE DONE WITH SWEAVE PROBABLY)
+
+We also need to make some some minor changes to the data.
+Both data sets need to be joined and variable names for predictions and forecasts need to be changed.
+This can be achieved using the following commands:
+
+\begin{CodeInput}
+combined <- dplyr::inner_join(uk_forecasts %>%
+                                dplyr::rename(prediction = value), 
+                              covid_uk_data %>%
+                                dplyr::rename(true_value = value))
+\end{CodeInput}
+MAYBE USE DATA.TABLE INSTEAD OF DPLYR IN ORDER TO AVOID MIXING? ALSO MAYBE JUST IMPLEMENT THE JOINING WITHIN SCORINGUTILS DIRECTLY
+
+The scoringutils package has some built-in functionality for visualisation (also see section [visualisation]). In order to plot predictions we are interested in, we could call 
+\begin{CodeInput}
+additional_observations <- covid_uk_data %>%
+  dplyr::rename(true_value = value) %>%
+  dplyr::filter(value_date <= "2020-06-22", 
+                value_date > "2020-06-01", 
+                value_type %in% unique(example_subset$value_type))
+
+scoringutils::plot_predictions(example_subset, 
+                               additional_observations,
+                               x = "value_date",
+                               facet_formula = geography ~ value_type)
+\end{CodeInput}
+MAYBE CHANGE THE SCORINGUTILS CODE TO MAKE THAT A BIT SLICKER. FOR EXAMPLE: HAVE THE ENTIRE DATA SET AND RESTRICT FORECASTS TO EVERYTHING AFTER A GIVEN DATE.
+
+An evaluation that returns scores for every model and every prediction target can now be performed by simply calling 
+\begin{CodeInput}
+library(scoringutils)
+scores <- eval_forecasts(combined, 
+                         summarise_by = c("model", "value_type"))
+
+scores
+\end{CodeInput}
+
+Pairwise comparisons between models [CITATION] can be made using 
+\begin{CodeInput}
+pairwise <- pairwise_comparison(scores)
+\end{CodeInput}
+This will be explained in more detail in section [visualisation and interpretation of the evaluation].
+
+The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
+
+\section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}
+
+The metrics included in the \pkg{scoringuitls} package can be divided into two categories.
+The first consists of metrics that aim to capture different aspects of sharpness and calibration, the second comprises various proper scoring rules.
+We begin with the latter.
+
+\subsection{Proper scoring rules}
+
+Proper scoring rules [CITATION] jointly assess sharpness and calibration and assign a single numeric value to a forecast.
+A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
+This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state state their true best belief.
+The following scoring rules are implemented in \pkg{scoringutils}: The (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], the Dawid-Sebastiani-score (dss) [CITATION], the (weighted) interval score (wis), and the Brier score (bs).
+The first three proper scoring rules are implemented as wrappers around functions from the \pkg{scoringRules} package.
+They are suitable for any sample-based prediction format.
+The Log Score, however, is not applied if integer-valued forecasts are supplied, as the implementation in \pkg{scoringRules} requires a kernel density estimation that may be inappropriate for integer values.
+The interval score is appropriate for forecasts in a quantile formats.
+For every central prediction interval, the interval score is computed as the sum of separate penalties for overprediction, underprediction and sharpness (i.e.
+width) of the forecast.
+By default, the weighted interval score is returned.
+The overall score for a forecast is obtained as the mean of all scores for the individual central prediction intervals, weighted according to the range of the prediction interval.
+This ensures that the weighted interval score converges to the continuous ranked probability score for an increasing number of available prediction intervals.
+DO I WANT FORMULAS FOR ALL THE SCORES? The Brier score is used for binary predictions.
+
+TALK ABOUT CAVEATS (LATER?) E.G. THAT THE MEAN WIS IS INFLUENCED BY THE ABSOLUTE VALUE OF THE TARGET?
+
+\subsection{Evaluating calibration and sharpness independently}
+
+In addition to the proper scoring rules outlined above, \pkg{scoringutils} makes numerous metrics available to evaluate calibration and sharpness independently.
+This is especially helpful for model diagnostics.
+
+\subsubsection{Assessing calibration}
+
+Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g.
+@funkAssessingPerformanceRealtime2019; @gneitingProbabilisticForecastsCalibration2007; @gneitingStrictlyProperScoring2007).
+Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
+
+Bias, i.e.
+systematic over- or underprediction, is a very common form of miscalibration.
+For continuous forecasts, assessing whether a predictive distribution has a tendency to over- or underpredict can be very easily achieved by simply evaluating the predictive distribution at the true observed value.
+This metric is a generalisation of the integer-valued one @funkAssessingPerformanceRealtime2019 have proposed.
+It is also closely related to the probability integral transform (PIT) discussed later in this chapter.
+To improve the interpretability of the score we can transform it to a value between -1 (under-prediction) and 1 (over-prediction).
+Consequently, we measure bias as
+$$B_t (P_t, x_t) = 1 - 2 \cdot (P_t (x_t)),$$
+where $P_t$ is the cumulative distribution function of the predictive distribution for the true value $x_t$.
+When using predictive samples, $P_t (x_t)$ is simply the fraction of predictive samples for $x_t$ that are smaller than the true observed $x_t$.
+
+For integer valued forecasts, we use the metric proposed by @funkAssessingPerformanceRealtime2019: 
+$$B_t (P_t, x_t) = 1 - (P_t (x_t) + P_t (x_t + 1)).$$
+Bias can again assume values between -1 (under-prediction) and 1 (over-prediction) and is 0 ideally.
+
+For quantile forecasts, we propose the following metric to assess bias: 
+\begin{align*}
+  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t\}) \mathbb{1}( x_t \leq q_{t, 0.5}) \\
+  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbb{1}( x_t \geq q_{t, 0.5}),
+\end{align*}
+where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$.
+They represent our belief about what the true value $x_t$ will be.
+For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$.
+$\mathbb{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
+In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution.
+If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
+If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
+For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
+For quantile forecasts, an alternative approach is to look at the over- and underprediction components of the weighted interval score.
+These however, capture the bias in absolute terms, while the above proposed metric captures a tendency to over- and underpredict that is less sensitive to outliers.
+
+I ASSUME ALL OF THIS IS A BIT TOO LONG?
+
+Another way to look at calibration[precisely: probabilistic calibration in @gneitingProbabilisticForecastsCalibration2007] is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
+This is most easily understood in the context of quantile forecasts, but can in principle be transferred to continuous and integer forecasts as well.
+
+To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution.
+If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles.
+We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
+
+To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
+
+A third way to look at calibration is to look a the probability integral tranform.
+As explained previously, the CDF of predictice distribution $P_t$ should ideally be equal to the CDF of the true unknown distribution $F_t$ that generated the observed value $x_t$.
+In order to assess whether there are substantial deviations between the two, @dawidPresentPositionPotential1984 suggested to transform the observed values using the probability integral transform (PIT).
+Agreement between the forecasts and the observed values can then be examined by observing whether or not the transformed values follow a uniform distribution.
+The PIT is given by
+$$u_t = P_t (x_t),$$
+where $u_t$ is the transformed variable and $P_t(x_t)$ is the predictive distribution evaluated at the true observed value $x_t$.
+If $P_t = F_t$ at all times $t$, then $u_t, t = 1 \dots T$ follows a uniform distribution (for a proof see e.g.
+@angusProbabilityIntegralTransform1994).
+
+In the case of discrete outcomes, the PIT is no longer uniform even when forecasts are ideal. As @funkAssessingPerformanceRealtime2019 suggest, we use a randomised PIT instead by redefining 
+$$u_t = P_t(x_t) + vt \cdot (P_t(x_t) - P_t(x_t - 1) ),$$
+where $x_t$ is again the observed value at time $t$, $P_t()$ is the CDF of the predictive distribution function, $P_t (-1) = 0$ by definition, and $v_t$ is a standard uniform variable independent of $x_t$.
+If $P_t$ is equal to the true data-generating distribution function, then $u_t$ is standard uniform.
+@czadoPredictiveModelAssessment2009 also propose a non-randomised version of the PIT for count data that could be used alternatively.
+
+One can then plot a histogram of $u_t$ values to look for deviations from uniformity.
+U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
+Biased predictions will usually result in a triangle-shaped histogram.
+
+\subsubsection{Assessing sharpness}
+
+Sharpness is the second property central to model evaluation.
+The ability to produce narrow forecasts is a quality of the forecasts only and does not depend on the observations.
+Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
+Again we need to take slightly different approaches for continuous, integer and quantile forecasts.
+For continuous and integer forecasts we follow the suggestion from @funkAssessingPerformanceRealtime2019.
+For quantile forecasts we propose a novel metric.
+
+For continuous and integer forecasts, @funkAssessingPerformanceRealtime2019 suggest to measure sharpness as the normalised median absolute deviation about the median (MADN), i.e. 
+$$ S_t(P_t) = \frac{1}{0.675} \cdot \text{median}(|y - \text{median(y)}|), $$ 
+where $y$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant that ensures that sharpness will equal the standard deviation of the predictive distribution if $P_t$ is the CDF of a normal distribution.
+
+For quantile forecasts, we propose to measure sharpness as a weighted mean of the width of the interval ranges.
+Let $Q_t$ be a set of predicted quantiles for a true $x_t$ at time $t$.
+This set of quantiles is assumed to be symmetric, such that there exist $K$ corresponding pairs of elements $q_{t, \frac{\alpha}{2}}$ and $q_{t, 1-\frac{\alpha}{2}}$.
+These $K$ corresponding pairs of quantiles cover a $(1 - \alpha) \cdot 100$ prediction interval.
+We can, accordingly, also denote $q_{t, \frac{\alpha}{2}}$ as $l_t$ the lower bound of the prediction interval at time $t$ and $q_{t, 1-\frac{\alpha}{2}}$ as $u_t$, the upper bound.
+We measure the sharpness of a quantile forecast at time $t$ as
+\begin{align*}
+\text{sharpness}_t &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (q_{t, 1 - \frac{\alpha}{2}} - q_{t, \frac{\alpha}{2}}) \\
+                   &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (u_t - l_t).
+\end{align*}
+Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
+We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score  described in the following section.
+
+\section[visualisation]{Visualisation and interpretation of evaluation results}
+
+\begin{CodeInput}
+.
+\end{CodeInput}
 

### 2020-12-08 article.Rnw
update draft and create Rnw structure

--- 2020-12-07 article.tex
@@
 \documentclass[article]{jss}
-\usepackage{amsmath}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% almost as usual
-\author{Anonymous Giraffe\\Universit\"at Innsbruck \And 
-        Second Author\\Plus Affiliation}
+%% -- LaTeX packages and custom commands ---------------------------------------
+
+%% recommended packages
+\usepackage{thumbpdf,lmodern}
+\usepackage{amsmath, amsfonts}
+\usepackage{bbm}% to get \mathbbmm{1}
+
+%% another package (only for this demo article)
+\usepackage{framed}
+
+%% new custom commands
+\newcommand{\class}[1]{`\code{#1}'}
+\newcommand{\fct}[1]{\code{#1()}}
+
+%% For Sweave-based articles about R packages:
+%% need no \usepackage{Sweave}
+\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE, prefix.string=plots/plot}
+<<preliminaries, echo=FALSE, results=hide>>=
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+library("MASS")
+@
+
+%% -- Article metainformation (author, title, ...) -----------------------------
+
+%% - \author{} with primary affiliation
+%% - \Plainauthor{} without affiliations
+%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
+%% - \AND starts a new line, \And does not.
+\author{Nikos I. Bosse\\London School of Hygiene and Tropical Medicine
+   \And Second Author\\Plus Affiliation}
+\Plainauthor{Nikos I. Bosse, Second Author}
+
+%% - \title{} in title case
+%% - \Plaintitle{} without LaTeX markup (if any)
+%% - \Shorttitle{} with LaTeX markup (if any), used as running title
 \title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}
+\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using scoringutils in R}
+\Shorttitle{Evaluating Forecasts with \pkg{scoringutils} in \proglang{R}}
 
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Anonymous Giraffe, Second Author} %% comma-separated
-\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using \texttt{scoringutils} in R} %% without formatting
-\Shorttitle{\pkg{foo}: A Capitalized Title} %% a short title (if necessary)
-
-%% an abstract and keywords
+%% - \Abstract{} almost as usual
 \Abstract{
@@
 }
-\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
-\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
-%% at least one keyword must be supplied
 
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
+%% - \Keywords{} with LaTeX markup, at least one required
+%% - \Plainkeywords{} without LaTeX markup (if necessary)
+%% - Should be comma-separated and in sentence case.
+\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
+\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
 
-%% The address of (at least) one author should be given
-%% in the following format:
+%% - \Address{} of at least one author
+%% - May contain multiple affiliations for each author
+%%   (in extra lines, separated by \emph{and}\\).
+%% - May contain multiple authors for the same affiliation
+%%   (in the same first line, separated by comma).
 \Address{
@@
   Centre for Mathematical Modelling of Infectious Diseases\\
-%%  Faculty of Economics and Statistics\\
   London School of Hygiene and Tropical Medicine\\
-%%  6020 Innsbruck, Austria\\
+  Keppel Street \\ 
+  London WC1E 7HT \\
   E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
-  %% URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
+  % URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
 }
-%% It is also possible to add a telephone and fax number
-%% before the e-mail in the following format:
-%% Telephone: +43/512/507-7103
-%% Fax: +43/512/507-2851
-
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
+\SweaveOpts{concordance=FALSE}
 
-%% include your article here, just as usual
-%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.
+%% -- Introduction -------------------------------------------------------------
 
-\section[introduction]{Introduction}
+%% - In principle "as usual".
+%% - But should typically have some discussion of both _software_ and _methods_.
+%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
+%% - If such markup is in (sub)section titles, a plain text version has to be
+%%   added as well.
+%% - All software mentioned should be properly \cite-d.
+%% - All abbreviations should be introduced.
+%% - Unless the expansions of abbreviations are proper names (like "Journal
+%%   of Statistical Software" above) they should be in sentence case (like
+%%   "generalized linear models" below).
+
+\section[Introduction]{Introduction}
 Good forecasts are of great interest to decision makers in almost every field.
@@
 For an ideal forecast, we therefore have
+%
 $$ P_t = F_t, $$
+%
 where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution.
@@
 
+% ==============================================================================
 \section[run-evaluation]{Running an automated evaluation}
-A full evaluation of all forecasts based on observed values can be performed using the function \code{eval_forecasts}.
+
+A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
 The function automatically recognises the prediction type and format and applies the appropriate scoring metrics.
@@
 
-\begin{CodeInput}
+<<load-data>>=
 # install and load data pacakge from external repository
-remotes::install_github("sbfnk/covid19.forecasts.uk")
+# remotes::install_github("sbfnk/covid19.forecasts.uk")
 library(covid19.forecasts.uk)
@@
 data(covid_uk_data)
+head(covid_uk_data)
 
@@
 data(uk_forecasts)
-\end{CodeInput}
-
-ADD OUTPUT HERE TO SHOW THE FIRST COUPLE OF ROWS (MUST BE DONE WITH SWEAVE PROBABLY)
+head(uk_forecasts)
+@
 
@@
 
-\begin{CodeInput}
+<<modify-data>>=
 combined <- dplyr::inner_join(uk_forecasts %>%
@@
                                 dplyr::rename(true_value = value))
-\end{CodeInput}
+@
+
 MAYBE USE DATA.TABLE INSTEAD OF DPLYR IN ORDER TO AVOID MIXING? ALSO MAYBE JUST IMPLEMENT THE JOINING WITHIN SCORINGUTILS DIRECTLY
@@
 The scoringutils package has some built-in functionality for visualisation (also see section [visualisation]). In order to plot predictions we are interested in, we could call
-\begin{CodeInput}
+
+<<show-forecasts, eval=FALSE>>=
+example_subset <- combined %>%
+  dplyr::filter(model == "SIRCOVID", 
+                creation_date == "2020-06-22")
+
 additional_observations <- covid_uk_data %>%
@@
                                facet_formula = geography ~ value_type)
-\end{CodeInput}
+@
+
+The output is shown in Figure \ref{fig:forecast-visualisation}.
+
 MAYBE CHANGE THE SCORINGUTILS CODE TO MAKE THAT A BIT SLICKER. FOR EXAMPLE: HAVE THE ENTIRE DATA SET AND RESTRICT FORECASTS TO EVERYTHING AFTER A GIVEN DATE.
 
+\begin{figure}[h]
+\centering
+<<show-forecasts, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
+example_subset <- combined %>%
+  dplyr::filter(model == "SIRCOVID", 
+                creation_date == "2020-06-22")
+
+additional_observations <- covid_uk_data %>%
+  dplyr::rename(true_value = value) %>%
+  dplyr::filter(value_date <= "2020-06-22", 
+                value_date > "2020-06-01", 
+                value_type %in% unique(example_subset$value_type))
+
+scoringutils::plot_predictions(example_subset, 
+                               additional_observations,
+                               x = "value_date",
+                               facet_formula = geography ~ value_type)
+@
+\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
+\end{figure}
+
 An evaluation that returns scores for every model and every prediction target can now be performed by simply calling 
-\begin{CodeInput}
+<<scoring>>=
 library(scoringutils)
@@
 
-scores
-\end{CodeInput}
+dplyr::glimpse(scores)
+@
 
 Pairwise comparisons between models [CITATION] can be made using 
-\begin{CodeInput}
+<<pairwise-comparison>>=
 pairwise <- pairwise_comparison(scores)
-\end{CodeInput}
+@
+
 This will be explained in more detail in section [visualisation and interpretation of the evaluation].
@@
 
+% ==============================================================================
 \section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}
@@
 \begin{align*}
-  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t\}) \mathbb{1}( x_t \leq q_{t, 0.5}) \\
-  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbb{1}( x_t \geq q_{t, 0.5}),
+  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t \}) \mathbbm{1}( x_t \leq q_{t, 0.5}) \\
+  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbbm{1}( x_t \geq q_{t, 0.5}),
 \end{align*}
@@
 For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$.
-$\mathbb{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
+$\mathbbm{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
 In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution.
@@
 
-\section[visualisation]{Visualisation and interpretation of evaluation results}
+\section{Visualisation and interpretation of evaluation results}
+
+A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
+
+\begin{figure}[h]
+\centering
+<<score-table, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+scoringutils::score_table(scores, y = "model", facet_formula = ~ value_type)
+@
+\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
+\end{figure}
+
+If all we care about is which model performed best, the pairwise comparison is most suitable. This approach compares all possible pairs of models based on an overlapping set of forecasts.
+
+\begin{figure}[h]
+\centering
+<<pairwise-plot, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+scoringutils::plot_pairwise_comparison(pairwise)
+@
+\caption{\label{fig:pairwise-comparison} XXXX}
+\end{figure}
 
@@
 
+%% -- Summary/conclusions/discussion -------------------------------------------
+
+\section{Summary and discussion} \label{sec:summary}
+
+% The results in this paper were obtained using
+% \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
+% \pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
+% and all packages used are available from the Comprehensive
+% \proglang{R} Archive Network (CRAN) at
+% \url{https://CRAN.R-project.org/}.
+
+\section*{Acknowledgments}
+
+% All acknowledgments (note the AE spelling) should be collected in this
+% unnumbered section before the references. It may contain the usual information
+% about funding and feedback from colleagues/reviewers/etc. Furthermore,
+% information such as relative contributions of the authors may be added here
+% (if any).
+
+%% -- Bibliography -------------------------------------------------------------
+%% - References need to be provided in a .bib BibTeX database.
+%% - All references should be made with \cite, \citet, \citep, \citealp etc.
+%%   (and never hard-coded). See the FAQ for details.
+%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
+%% - Titles in the .bib should be in title case.
+%% - DOIs should be included where available.
+
+\bibliography{refs}
+
+%% -- Appendix (if any) --------------------------------------------------------
+%% - After the bibliography with page break.
+%% - With proper section titles and _not_ just "Appendix".
+
+\newpage
+
+\begin{appendix}
+
+\section{Appendix section} \label{app:technical}
+
+\end{appendix}
+
+%% -----------------------------------------------------------------------------
+
 \end{document}

### 2020-12-08 R/analysis_script.R

@@ -22,6 +22,25 @@ combined <- dplyr::inner_join(uk_forecasts %>%
                               covid_uk_data %>%
                                 dplyr::rename(true_value = value))
 
+example_subset <- combined %>%
+  dplyr::filter(model == "SIRCOVID", 
+                creation_date == "2020-06-22")
+
+# plot predictions -------------------------------------------------------------
+
+additional_observations <- covid_uk_data %>%
+  dplyr::rename(true_value = value) %>%
+  dplyr::filter(value_date <= "2020-06-22", 
+                value_date > "2020-06-01", 
+                value_type %in% unique(example_subset$value_type))
+
+scoringutils::plot_predictions(example_subset, 
+                               additional_observations,
+                               x = "value_date",
+                               facet_formula = geography ~ value_type)
+ggplot2::ggsave("plots/forecast_visualistion.png")
+
+
 
 # create scoring table ---------------------------------------------------------
 scores <- combined %>%

### 2021-02-06 article.Rnw
update paper draft

@@
 library("MASS")
+
+system("bibtex article")
+
+knitr::opts_chunk$set(
+  out.width = "0.9\\textwidth"
+)
+
 @
@@
 \Abstract{
-  Forecasts play a role in many scientific fields from finance to meteorology and epidemiology.
-With the emergence of Covid-19 the role of forecasting to inform public policy has again attracted increased attention.
-In this paper we introduce a complete framework for evaluating different types of forecasts using the R package scoringutils.
-We discuss different appropriate evaluation metrics and apply the framework to XXXXX.
+  Forecasts play an important role in a variety of fields.
+Its role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
+Even with a plethora of tools and models available to make forecasts, evaluating the predictions is not trivial.
+Even though scoring methods such as proper scoring rules have been widely studied in the past, there is a lack of software implementation that allows a forecaster to conveniently evaluate their forecasts.
+In this paper we introduce \pkg{scoringutils}, an \proglang{R} package that facilitates automated forecast evaluation.
+It gives the user access to a wide range of scoring metrics for various types of forecasts as well as a variety of ways to visualise the evaluation.
+We give an overview of the rationale behind forecast evaluation and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
 }
@@
 \section[Introduction]{Introduction}
-Good forecasts are of great interest to decision makers in almost every field.
+Good forecasts are of great interest to decision makers in various fields like finance (), weather predictions or infectious disease modeling \citep{funkShorttermForecastsInform2020}.
 An integral part of assessing and improving their usefulness is forecast evaluation.
-For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts [some citations].
-Yet even with this rich body of research available, implementing a consistent framework in R to evaluate forecasts is not trivial.
-We therefore present the \pkg{scoringutils} pacakge.
+For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
+\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{FUNK}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
+Yet even with this rich body of research available, implementing a complete forecast evaluation in \proglang{R} is not trivial.
+We therefore present the \pkg{scoringutils} package.
 The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts.
-In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK.
+In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK \citep{funkShorttermForecastsInform2020}.
 
@@
 
-In its most general sense, a forecast is the forecaster’s stated belief about the future.
-For conceptual clarity, it is however useful to distinguish between different times of forecasts.
-Forecasts can be either quantitative or qualitative, but \pkg{scoringutils} only deals with the former.
-In terms of quantitative forecasts we can distinguish point forecasts from probabilistic forecasts.
+In its most general sense, a forecast is the forecaster’s stated belief about the future \citep{gneitingStrictlyProperScoring2007}.
+For conceptual clarity, it is however useful to distinguish between different types of forecasts.
+In terms of the quantitative forecasts that can be evaluated with \pkg{soringutils}, we can distinguish point forecasts from probabilistic forecasts.
 A point forecast states a single number and is the simplest form of a forecast.
-A point forecast is limited in its usefulness, as it does not give information about the uncertainty of the foreacst.
+It is limited in its usefulness, as it does not give information about the uncertainty of the forecast.
 A very certain forecast may, for example, warrant a very different course of actions than does a very uncertain one.
@@
 This distinction is important, as different formats and prediction types require slightly different evaluation approaches.
-Forecasts are usually expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
+Forecasts can be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
 Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
 They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
-For that reason, usually quantiles are reported instead [citation FORECAST HUBS].
+For that reason, often quantiles are reported instead [citation FORECAST HUBS].
 As an alternative representation of a quantile forecast, one can also express predictions using central prediction intervals.
@@
 On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
-Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
+%Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
 DO I WANT THAT FIGURE?
-
-\section[data]{Data: UK short-term forecasts}
-To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
-Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
-The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
-Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
-The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL]
-
-- timing (weekly?) and number of forecast dates
-- the four targets
-- the models. 
-- is the set complete?
-
-PLOT WITH OVERVIEW OF MISSING FORECASTS LIKE THE ONE THEY MADE FOR THE FORECAST HUB.
-
-% ==============================================================================
-\section[run-evaluation]{Running an automated evaluation}
-
-A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
-The function automatically recognises the prediction type and format and applies the appropriate scoring metrics.
-Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
-
-All it requires is a \code{data.frame} or similar that has a column called "prediction" and one called "true\_value".
-Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
-To get a sense of the required input formats, example data for each format is provided with the package and shown in the package vignette.
-We also included functionality to transform between various formats.
-Additional columns may be present to indicate a grouping of forecasts.
-One common use case is a set of forecasts made by different models for various locations at different time points, each for several weeks into the future.
-In this case, we would expect additional columns that could be called "model", "date", "forecast\_date", "forecast\_horizon" and "location".
-Through the by argument, the user needs to to specify the unit of a single forecast.
-In the above example we would set by = c("model", "date", "forecast\_date", "forecast\_horizon", "location").
-Columns such as quantile or sample should not be included, because several quantiles or samples make up one forecast.
-By default, by = NULL and scoringutils will automatically use all appropriate present columns.
-
-The \code{summarise_by} argument allows the user to choose categories to aggregate over.
-If we were only interested in scores for the different models, we would for example specify \code{summarise_by = c("model")}.
-If we wanted to have scores for every model in every location, we would specify \code{summarise_by = c("model", "location")}.
-If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
-
-When aggregating, \code{eval_forecasts} takes the mean according to the group defined in \code{summarise_by}.
-In the above example, if \code{summarise_by = c("model", "location")}, scores would be averaged over all forecast dates, forecast horizons and quantiles to yield one score per model and location.
-In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying sd = TRUE and for example quantiles = c(0.5) for the median.
-
-To evaluate the UK short term forecast data with scoringutils, we first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands:
-
-<<load-data>>=
-# install and load data pacakge from external repository
-# remotes::install_github("sbfnk/covid19.forecasts.uk")
-library(covid19.forecasts.uk)
-
-# load truth data
-data(covid_uk_data)
-head(covid_uk_data)
-
-# load forecasts
-data(uk_forecasts)
-head(uk_forecasts)
-@
-
-We also need to make some some minor changes to the data.
-Both data sets need to be joined and variable names for predictions and forecasts need to be changed.
-This can be achieved using the following commands:
-
-<<modify-data>>=
-combined <- dplyr::inner_join(uk_forecasts %>%
-                                dplyr::rename(prediction = value), 
-                              covid_uk_data %>%
-                                dplyr::rename(true_value = value))
-@
-
-MAYBE USE DATA.TABLE INSTEAD OF DPLYR IN ORDER TO AVOID MIXING? ALSO MAYBE JUST IMPLEMENT THE JOINING WITHIN SCORINGUTILS DIRECTLY
-
-The scoringutils package has some built-in functionality for visualisation (also see section [visualisation]). In order to plot predictions we are interested in, we could call
-
-<<show-forecasts, eval=FALSE>>=
-example_subset <- combined %>%
-  dplyr::filter(model == "SIRCOVID", 
-                creation_date == "2020-06-22")
-
-additional_observations <- covid_uk_data %>%
-  dplyr::rename(true_value = value) %>%
-  dplyr::filter(value_date <= "2020-06-22", 
-                value_date > "2020-06-01", 
-                value_type %in% unique(example_subset$value_type))
-
-scoringutils::plot_predictions(example_subset, 
-                               additional_observations,
-                               x = "value_date",
-                               facet_formula = geography ~ value_type)
-@
-
-The output is shown in Figure \ref{fig:forecast-visualisation}.
-
-MAYBE CHANGE THE SCORINGUTILS CODE TO MAKE THAT A BIT SLICKER. FOR EXAMPLE: HAVE THE ENTIRE DATA SET AND RESTRICT FORECASTS TO EVERYTHING AFTER A GIVEN DATE.
-
-\begin{figure}[h]
-\centering
-<<show-forecasts, echo=FALSE, fig=TRUE, height=5.2, width=7>>=
-example_subset <- combined %>%
-  dplyr::filter(model == "SIRCOVID", 
-                creation_date == "2020-06-22")
-
-additional_observations <- covid_uk_data %>%
-  dplyr::rename(true_value = value) %>%
-  dplyr::filter(value_date <= "2020-06-22", 
-                value_date > "2020-06-01", 
-                value_type %in% unique(example_subset$value_type))
-
-scoringutils::plot_predictions(example_subset, 
-                               additional_observations,
-                               x = "value_date",
-                               facet_formula = geography ~ value_type)
-@
-\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
-\end{figure}
-
-An evaluation that returns scores for every model and every prediction target can now be performed by simply calling 
-<<scoring>>=
-library(scoringutils)
-scores <- eval_forecasts(combined, 
-                         summarise_by = c("model", "value_type"))
-
-dplyr::glimpse(scores)
-@
-
-Pairwise comparisons between models [CITATION] can be made using 
-<<pairwise-comparison>>=
-pairwise <- pairwise_comparison(scores)
-@
-
-This will be explained in more detail in section [visualisation and interpretation of the evaluation].
-
-The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
 
@@
 
+MAKE THE FOLLOWING PARAGRAPH SHORTER AND JUST SAY WE USE THE SHARPNESS COMPONENT OF THE WIS.
+
 For quantile forecasts, we propose to measure sharpness as a weighted mean of the width of the interval ranges.
@@
 Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
-We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score  described in the following section.
+We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score.
+
+\subsection{Pairwise comparisons}
+
+If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION].
+In turn, each pair of models is evaluated based on the targets that that both models have predicted.
+The mean score by one model is divided by the mean score of the other model to obtain a measure of relative performance.
+Relative skill scores [citation (not sure there is something published, preprint fron Estee Cramer and Nick Reich)] for every model can then be obtained by taking the geometric mean of all pairwise ratios calculated for each model (omitting comparisons where there is no overlapping set of forecasts).
+ADD FORMULAS.
+A skill score smaller than 1 indicates that a model is performing better than the average model.
+One can also scale the skill by a reference baseline.
+A scaled relative skill smaller than one then means that the model in question performed better than the baseline.
+
+By applying a permutation test or non-parametric test like the Wilcoxon rank sum test we can also obtain p-values that help determine whether or not models perform significantly differently.
+THERE IS AN ISSUE WITH CORRELATION: THE TEST WOULD TREAT ALL OBSERVATIONS AS INDEPENDENT, WHICH THEY ARENT'T, SO I WOULD ASSUME THAT IT IS TOO LIBERAL.
+CURRENTLY, NO ADJUSTMENTS ARE IMPLEMENTED, I'M NOT REALLY SURE WHETHER TO DISCUSS THIS HERE OR NOT.
+
+% ==============================================================================
+\section{Evaluating UK short-term forecasts}
+\subsection{The data}
+
+To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
+Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
+The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
+Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
+The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL].
+
+- timing (weekly?) and number of forecast dates
+- the four targets
+- the models. 
+- is the set complete?
+
+We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
+% 
+<<load-data>>=
+# install and load data pacakge from external repository
+# remotes::install_github("sbfnk/covid19.forecasts.uk")
+
+# load packages
+library(covid19.forecasts.uk)
+library(dplyr)
+library(scoringutils)
+
+# load truth data
+data(covid_uk_data)
+# head(covid_uk_data)
+
+# load forecasts
+data(uk_forecasts)
+# head(uk_forecasts)
+@
+% 
+Let us take a first look a the data:
+% 
+<<glimpse-data>>=
+glimpse(covid_uk_data)
+glimpse(uk_forecasts)
+@
+% 
+To bring the forecasts into the format needed for the evaluation, some minor changes need to be made to the data.
+The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
+While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public.
+The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way.
+% 
+<<modify-data>>=
+uk_forecasts <- rename(uk_forecasts, prediction = value)
+covid_uk_data <-rename(covid_uk_data, true_value = value)
+
+combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
+@
+% 
+Before we start with scoring the forecasts, it makes sense to start the evaluation process by visualising the data.
+To get a feeling for how complete the data set is, we can run the following code to obtain a heatmap with the number of available forecasts:
+% 
+\begin{figure}[h]
+\centering
+<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12>>=
+show_avail_forecasts(combined, 
+                     x = "creation_date", 
+                     show_numbers = FALSE,
+                     legend_position = "bottom",
+                     facet_formula = ~ value_desc)
+@
+\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
+\end{figure}
+%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
+% 
+Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
+By default, the function treats a set of different quantiles or samples as one forecast.
+However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.
+
+Forecasts can be visualised using the \fct{plot\_predictions} function.
+The function accepts either a single combined data set or two separate truth data sets that can be merged together.
+Forecasts and observed values can be filtered independently, to show for example only predictions from a certain model or a specific number of weeks of true data before the forecast.
+Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
+To obtain, for example, a specific forecast from a model we are interest in, we can call:
+% 
+\begin{figure}[h!]
+\centering
+<<show-forecasts, echo=TRUE, fig=TRUE, height=5.2, width=7>>=
+locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
+plot_predictions(truth = covid_uk_data,
+                 forecasts = uk_forecasts,
+                 filter_both = list(paste("geography %in%", locations, collapse = "")),
+                 filter_truth = list('value_date <= "2020-06-22"',
+                                     'value_date > "2020-06-01"'), 
+                 filter_forecasts = list('model == "SIRCOVID"', 
+                                         'creation_date == "2020-06-22"'),
+                 x = "value_date",
+                 facet_formula = geography ~ value_type) + 
+  ggplot2::theme(legend.position = "bottom")
+@
+\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
+\end{figure}
+% 
+The output is shown in Figure \ref{fig:forecast-visualisation}.
+
+\subsection[Scoring forecasts with eval_forecasts]{Scoring forecasts with \fct{eval\_forecasts}}
+
+A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
+This requires a \code{data.frame} or similar which has at least a column called "prediction" and one called "true\_value".
+Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
+Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
+We will not discuss all possible input formats here, but instead refer to the example data for each format that is provided with the package.
+Where possible, \pkg{scoringutils} also provides functionality to transform between various formats, e.g.
+from a sample based format to a quantile format.
+The \fct{eval\_forecasts} function automatically recognises the prediction type and input format, applies the appropriate scoring metrics and aggregates results as desired by the user.
+Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
+
+As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
+% 
+<<scoring>>=
+scores <- eval_forecasts(combined, 
+                         summarise_by = c("model", "value_desc"))
+
+glimpse(scores)
+@
+% 
+If a more detailed analysis is desired, the level of aggregation can of course be changed to show for example separate scores for the different locations as well.
+This can be achieved using the \code{summarise_by} argument.
+To additionally stratify by location, we could specify \code{summarise_by = c("mode", "value_type", "geography")}.
+If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
+This can be useful if we, for example, want to analyse what proportion of true values are covered by certain interval ranges, or if we want to analyse the accuracy of the tails of the forecasts.
+When aggregating, \fct{eval\_forecasts} takes the mean according to the group defined in \code{summarise_by}.
+In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
+In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.
+
+The user must, however, still exercise some caution when aggregating scores, as many of the metrics are absolute and scale with the magnitude of the quantity to forecast.
+Looking at one score per model (i.e.
+specifying \code{summarise_by = c("model")}) may not be so useful in this instance, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
+
+In the above example, we did not have to explicitly specify the \code{by} argument, but this may be necessary if additional columns are present in the data that do not indicate a grouping of forecasts.
+The \code{by} argument must then be used to denote the unit of a single forecast.
+In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")}.
+Quantiles should not be included, as several quantiles make up one forecast (and similarly for samples).
+If we had additional columns that do not serve to group forecasts (like for example the number of inhabitants in a certain location over time), these should also not be included.
+By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
+
+\subsection{Pairwise comparisons}
+
+Pairwise comparisons between models [CITATION] can be obtained in two different ways.
+First, relative skill scores based on pairwise comparisons are by default returned from \fct{eval\_forecasts}.
+These will be computed separately for the categories defined in the \code{summarise_by} argument (excluding the category 'model').
+Alternatively, a set of scores can be post-processed using the separate function \fct{pairwise\_comparison}.
+This approach is to be used for visualisation and if p-values for the pairwise comparisons are needed, as those are not returned from \fct{eval\_forecasts}.
+Usually, one would compute scores without specifying a \code{summarise_by} argument, but sometimes it may be sensible to average over certain scores, for example for predictions generated at a certain date.
+This allows to reduce the correlation between observations that enter the computation of p-values, which in turn makes the test less liberal.
+Using the function \fct{plot\_pairwise\_comparison} we can visualise the mean score ratios between all models as well as the
+
+<<pairwise-comparison>>=
+# unsummarised scores
+unsummarised_scores <- eval_forecasts(combined)
+pairwise <- pairwise_comparison(unsummarised_scores, 
+                                summarise_by = "value_desc")
+@
+
+The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
 
@@
 
+\subsection{Visualising aggregate scores and rankings}
 A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
@@
 \centering
-<<score-table, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
-scoringutils::score_table(scores, y = "model", facet_formula = ~ value_type)
+<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=14>>=
+score_table(scores, y = "model", facet_formula = ~ value_desc)
 @
@@
 
-If all we care about is which model performed best, the pairwise comparison is most suitable. This approach compares all possible pairs of models based on an overlapping set of forecasts.
+The most informative metric in terms of model ranking is the relative\_skill.
+However, interpretation is not always straightforward and has to be done carefully.
+We can see that performance varied quite a bit across different metrics, where some models did well on one target, but poorly on another.
+Especially the Exponential growth/decline model stands out as it received the lowest relative skill score for hospital admissions, but the highest for the total number of beds occupied.
+Looking back at Figure \ref{fig:avail-forecasts}, we see that the model has only submitted very few forecasts over all.
+It may therefore be sensible to require all models to have submitted forecasts for at least 50\% of all forecast targets in order to enter the pairwise comparisons.
+For similar reasons, the interval score may be deceiving if looked at in isolation.
+As can be seen, the DetSEIRwithNB MLE model received a lower relative skill score, but a higher interval score than the DetSEIRwithNB MCMC model.
+This, again, can be explained by the fact that they forecasted different targets.
+The interval score, as an absolute metric, is highly influenced by the absolute value of the quantity that is forecasted.
+For the same reason, one should be careful when summarising interval scores from different locations or forecast targets, as the average score will be dominated by outliers as well as differences in the absolute level.
+Assuming a large enough set of available overlapping forecasts, the relative skill score is more robust.
+It therefore is reasonable to assume that the DetSEIRwithNB MLE forecasted quantities with a higher absolute value, but tended to perform worse than the DetSEIRwithNB MCMC model as far as we can tell based on the set of all pariwise comparisons.
+This can be confirmed for the direct comparison between the two by looking at the mean score ratios from the pairwise comparisons.
+These can be obtained by calling
+
+\begin{figure}[h!]
+\centering
+<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14>>=
+plot_pairwise_comparison(pairwise) +
+  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
+@
+\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
+If a tile is blue, then the model on the y-axis performed better.
+If it is red, the model on the x-axis performed better in direct comparison.
+}
+\end{figure}
+
+we can also look at p-values in Figure \ref{fig:pairwise-comparison-pval} PROBABLY REMOVE THAT FROM THE PAPER
+\begin{figure}[h!]
+\centering
+<<pairwise-plot-ratio, echo=TRUE, fig=TRUE, height=7.2, width=14>>=
+plot_pairwise_comparison(pairwise, type = "pval") +
+  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
+@
+\caption{\label{fig:pairwise-comparison-pval} XXXX}
+\end{figure}
+
+Both plots can be combined in one: PLOT NEEDS FIXING AND ALSO PROBABLY REMOVE THAT FROM THE PAPER? Figure \ref{fig:pairwise-comparison-both}
+\begin{figure}[h!]
+\centering
+<<pairwise-plot-pval, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+plot_pairwise_comparison(pairwise, type = "together") +
+  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
+@
+\caption{\label{fig:pairwise-comparison-both} p-values and ratios together.}
+\end{figure}
+
+In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
+
+\subsection{Visual model diagnostics}
+
+For forecasts in an interval format, looking at the components of the weighted interval score separately is a natural next step.
+We can see in Figure \ref{fig:wis-components} that the majority of penalties come from over-and underprediction, instead of the sharpness component.
+We also see that most models tended to either over- or underpredict actual numbers.
+
+\begin{figure}[h!]
+\centering
+<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+wis_components(scores, 
+               facet_formula = ~ value_desc, 
+               scales = "free_x") + 
+  ggplot2::coord_flip() + 
+  ggplot2::theme(legend.position = "bottom")
+@
+\caption{\label{fig:wis-components} p-values and ratios together.}
+\end{figure}
+
+We can have a closer look at calibration using the functions \fct{interval\_coverage} and \fct{quantile\_coverage}.
+The interval coverage plot shows the proportion of all true values that fall within all the different prediction intervals.
+This gives a visual impression of probabilistic calibration [GNEITING].
+Ideally, $x$ percent of true values should be covered by the $x$\%-prediction intervals, resulting in a 45° line.
+Areas shaded in green indicate that the model is covering more true values than it actually should, while areas in white indicate that the model fails to cover the desired proportion of true values with its prediction intervals.
+The majority of the models were too confident in their predictions, while some showed showed good calibration.
+The quantile coverage plot shows the proportion of all true values below certain predictive quantiles.
+While this plot is slightly harder to interpret, it also includes information about bias as and allows to separate the lower and upper boundaries of the prediction intervals.
+We can see, for example, that the Exponential growth/decline model was consistently biased downwards.
+Figure \ref{fig:coverage}
+
+\begin{figure}[h!]
+\centering
+<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold'>>=
+cov_scores <- eval_forecasts(combined, 
+                             summarise_by = c("model", 
+                                              "range", "quantile"))
+scoringutils::interval_coverage(cov_scores)
+scoringutils::quantile_coverage(cov_scores)
+@
+
+<<quantile-coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold'>>=
+cov_scores <- eval_forecasts(combined, 
+                             summarise_by = c("model", 
+                                              "range", "quantile"))
+scoringutils::quantile_coverage(cov_scores)
+@
+
+\caption{\label{fig:coverage} quantile and interval coverage}
+\end{figure}
+
+DO I WANT TO INCLUDE PIT PLOTS AS WELL? I GUESS? NEED TO LOOK AT THE IMPLEMENTATION FOR QUANTILE FORECASTS
+
+% \begin{figure}[h]
+% \centering
+% <<
+% pit-plots, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+% # out <- eval_forecasts(combined, 
+% #                       summarise_by = c("model"), 
+% #                       pit_plots = TRUE)
+% 
+% @
+% \caption{\label{fig:calibration-pit} pit plots}
+% \end{figure}
+
+Look at e.g. bias by location? Figure \ref{fig:bias-heatmap}
 
@@
 \centering
-<<pairwise-plot, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
-scoringutils::plot_pairwise_comparison(pairwise)
+<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15>>=
+scores <- eval_forecasts(combined, 
+                         summarise_by = c("model",
+                                          "value_desc", 
+                                          "geography"), 
+                         compute_relative_skill = FALSE)
+scoringutils::score_heatmap(scores, metric = "bias", 
+                            x = "geography", facet_formula = ~ value_desc, 
+                            scale = "free_x")
 @
-\caption{\label{fig:pairwise-comparison} XXXX}
+\caption{\label{fig:bias-heatmap} bias by location}
 \end{figure}
+
+WHAT IS NEEDED HERE IS A BIT OF THINKING WITH REGARDS TO WHAT VISUALISATION I WANT TO SHOW AND IN HOW MUCH DETAIL I WANT TO ANALYSE THE MODELS.
 
@@
 \section{Summary and discussion} \label{sec:summary}
+
+COMING SOON.
 

### 2021-02-06 article.tex
Updates from Overleaf
update paper draft

@@
 \documentclass[article]{jss}
-\usepackage{amsmath}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% almost as usual
-\author{Anonymous Giraffe\\Universit\"at Innsbruck \And 
-        Second Author\\Plus Affiliation}
+%% -- LaTeX packages and custom commands ---------------------------------------
+
+%% recommended packages
+\usepackage{thumbpdf,lmodern}
+\usepackage{amsmath, amsfonts}
+\usepackage{bbm}% to get \mathbbmm{1}
+
+%% another package (only for this demo article)
+\usepackage{framed}
+
+%% new custom commands
+\newcommand{\class}[1]{`\code{#1}'}
+\newcommand{\fct}[1]{\code{#1()}}
+
+%% For Sweave-based articles about R packages:
+%% need no \usepackage{Sweave}
+
+%% -- Article metainformation (author, title, ...) -----------------------------
+
+%% - \author{} with primary affiliation
+%% - \Plainauthor{} without affiliations
+%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
+%% - \AND starts a new line, \And does not.
+\author{Nikos I. Bosse\\London School of Hygiene and Tropical Medicine
+   \And Second Author\\Plus Affiliation}
+\Plainauthor{Nikos I. Bosse, Second Author}
+
+%% - \title{} in title case
+%% - \Plaintitle{} without LaTeX markup (if any)
+%% - \Shorttitle{} with LaTeX markup (if any), used as running title
 \title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}
+\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using scoringutils in R}
+\Shorttitle{Evaluating Forecasts with \pkg{scoringutils} in \proglang{R}}
 
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Anonymous Giraffe, Second Author} %% comma-separated
-\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using \texttt{scoringutils} in R} %% without formatting
-\Shorttitle{\pkg{foo}: A Capitalized Title} %% a short title (if necessary)
+%% - \Abstract{} almost as usual
+\Abstract{
+  Forecasts play an important role in a variety of fields.
+Its role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
+Even with a plethora of tools and models available to make forecasts, evaluating the predictions is not trivial.
+Even though scoring methods such as proper scoring rules have been widely studied in the past, there is a lack of software implementation that allows a forecaster to conveniently evaluate their forecasts.
+In this paper we introduce \pkg{scoringutils}, an \proglang{R} package that facilitates automated forecast evaluation.
+It gives the user access to a wide range of scoring metrics for various types of forecasts as well as a variety of ways to visualise the evaluation.
+We give an overview of the rationale behind forecast evaluation and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
+}
 
-%% an abstract and keywords
-\Abstract{
-  Forecasts play a role in many scientific fields from finance to meteorology and epidemiology.
-With the emergence of Covid-19 the role of forecasting to inform public policy has again attracted increased attention.
-In this paper we introduce a complete framework for evaluating different types of forecasts using the R package scoringutils.
-We discuss different appropriate evaluation metrics and apply the framework to XXXXX.
-}
-\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
-\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
-%% at least one keyword must be supplied
+%% - \Keywords{} with LaTeX markup, at least one required
+%% - \Plainkeywords{} without LaTeX markup (if necessary)
+%% - Should be comma-separated and in sentence case.
+\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
+\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
 
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
-
-%% The address of (at least) one author should be given
-%% in the following format:
+%% - \Address{} of at least one author
+%% - May contain multiple affiliations for each author
+%%   (in extra lines, separated by \emph{and}\\).
+%% - May contain multiple authors for the same affiliation
+%%   (in the same first line, separated by comma).
 \Address{
@@
   Centre for Mathematical Modelling of Infectious Diseases\\
-%%  Faculty of Economics and Statistics\\
   London School of Hygiene and Tropical Medicine\\
-%%  6020 Innsbruck, Austria\\
+  Keppel Street \\ 
+  London WC1E 7HT \\
   E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
-  %% URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
+  % URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
 }
-%% It is also possible to add a telephone and fax number
-%% before the e-mail in the following format:
-%% Telephone: +43/512/507-7103
-%% Fax: +43/512/507-2851
-
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%% include your article here, just as usual
-%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.
+%% -- Introduction -------------------------------------------------------------
 
-\section[introduction]{Introduction}
-Good forecasts are of great interest to decision makers in almost every field.
+%% - In principle "as usual".
+%% - But should typically have some discussion of both _software_ and _methods_.
+%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
+%% - If such markup is in (sub)section titles, a plain text version has to be
+%%   added as well.
+%% - All software mentioned should be properly \cite-d.
+%% - All abbreviations should be introduced.
+%% - Unless the expansions of abbreviations are proper names (like "Journal
+%%   of Statistical Software" above) they should be in sentence case (like
+%%   "generalized linear models" below).
+
+\section[Introduction]{Introduction}
+Good forecasts are of great interest to decision makers in various fields like finance (), weather predictions or infectious disease modeling \citep{funkShorttermForecastsInform2020}.
 An integral part of assessing and improving their usefulness is forecast evaluation.
-For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts [some citations].
-Yet even with this rich body of research available, implementing a consistent framework in R to evaluate forecasts is not trivial.
-We therefore present the \pkg{scoringutils} pacakge.
+For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
+\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{FUNK}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
+Yet even with this rich body of research available, implementing a complete forecast evaluation in \proglang{R} is not trivial.
+We therefore present the \pkg{scoringutils} package.
 The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts.
-In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK.
+In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK \citep{funkShorttermForecastsInform2020}.
 
@@
 On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
-Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
+%Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
 DO I WANT THAT FIGURE?
 
-\section[data]{Data: UK short-term forecasts}
-To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
-Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
-The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
-Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
-The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL]
-
-- timing (weekly?) and number of forecast dates
-- the four targets
-- the models. 
-- is the set complete?
-
-PLOT WITH OVERVIEW OF MISSING FORECASTS LIKE THE ONE THEY MADE FOR THE FORECAST HUB.
-
-\section[run-evaluation]{Running an automated evaluation}
-A full evaluation of all forecasts based on observed values can be performed using the function \code{eval_forecasts}.
-The function automatically recognises the prediction type and format and applies the appropriate scoring metrics.
-Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
-
-All it requires is a \code{data.frame} or similar that has a column called "prediction" and one called "true\_value".
-Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
-To get a sense of the required input formats, example data for each format is provided with the package and shown in the package vignette.
-We also included functionality to transform between various formats.
-Additional columns may be present to indicate a grouping of forecasts.
-One common use case is a set of forecasts made by different models for various locations at different time points, each for several weeks into the future.
-In this case, we would expect additional columns that could be called "model", "date", "forecast\_date", "forecast\_horizon" and "location".
-Through the by argument, the user needs to to specify the unit of a single forecast.
-In the above example we would set by = c("model", "date", "forecast\_date", "forecast\_horizon", "location").
-Columns such as quantile or sample should not be included, because several quantiles or samples make up one forecast.
-By default, by = NULL and scoringutils will automatically use all appropriate present columns.
-
-The \code{summarise_by} argument allows the user to choose categories to aggregate over.
-If we were only interested in scores for the different models, we would for example specify \code{summarise_by = c("model")}.
-If we wanted to have scores for every model in every location, we would specify \code{summarise_by = c("model", "location")}.
-If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
-
-When aggregating, \code{eval_forecasts} takes the mean according to the group defined in \code{summarise_by}.
-In the above example, if \code{summarise_by = c("model", "location")}, scores would be averaged over all forecast dates, forecast horizons and quantiles to yield one score per model and location.
-In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying sd = TRUE and for example quantiles = c(0.5) for the median.
-
-To evaluate the UK short term forecast data with scoringutils, we first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands:
-
-\begin{CodeInput}
-# install and load data pacakge from external repository
-remotes::install_github("sbfnk/covid19.forecasts.uk")
-library(covid19.forecasts.uk)
-
-# load truth data
-data(covid_uk_data)
-
-# load forecasts
-data(uk_forecasts)
-\end{CodeInput}
-
-ADD OUTPUT HERE TO SHOW THE FIRST COUPLE OF ROWS (MUST BE DONE WITH SWEAVE PROBABLY)
-
-We also need to make some some minor changes to the data.
-Both data sets need to be joined and variable names for predictions and forecasts need to be changed.
-This can be achieved using the following commands:
-
-\begin{CodeInput}
-combined <- dplyr::inner_join(uk_forecasts %>%
-                                dplyr::rename(prediction = value), 
-                              covid_uk_data %>%
-                                dplyr::rename(true_value = value))
-\end{CodeInput}
-MAYBE USE DATA.TABLE INSTEAD OF DPLYR IN ORDER TO AVOID MIXING? ALSO MAYBE JUST IMPLEMENT THE JOINING WITHIN SCORINGUTILS DIRECTLY
-
-The scoringutils package has some built-in functionality for visualisation (also see section [visualisation]). In order to plot predictions we are interested in, we could call 
-\begin{CodeInput}
-additional_observations <- covid_uk_data %>%
-  dplyr::rename(true_value = value) %>%
-  dplyr::filter(value_date <= "2020-06-22", 
-                value_date > "2020-06-01", 
-                value_type %in% unique(example_subset$value_type))
-
-scoringutils::plot_predictions(example_subset, 
-                               additional_observations,
-                               x = "value_date",
-                               facet_formula = geography ~ value_type)
-\end{CodeInput}
-MAYBE CHANGE THE SCORINGUTILS CODE TO MAKE THAT A BIT SLICKER. FOR EXAMPLE: HAVE THE ENTIRE DATA SET AND RESTRICT FORECASTS TO EVERYTHING AFTER A GIVEN DATE.
-
-An evaluation that returns scores for every model and every prediction target can now be performed by simply calling 
-\begin{CodeInput}
-library(scoringutils)
-scores <- eval_forecasts(combined, 
-                         summarise_by = c("model", "value_type"))
-
-scores
-\end{CodeInput}
-
-Pairwise comparisons between models [CITATION] can be made using 
-\begin{CodeInput}
-pairwise <- pairwise_comparison(scores)
-\end{CodeInput}
-This will be explained in more detail in section [visualisation and interpretation of the evaluation].
-
-The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
-
+% ==============================================================================
 \section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}
@@
 Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
-We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score  described in the following section.
+We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score.
 
-\section[visualisation]{Visualisation and interpretation of evaluation results}
+\subsection{Pairwise comparisons}
+
+If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION].
+In turn, each pair of models is evaluated based on the targets that that both models have predicted.
+The mean score by one model is divided by the mean score of the other model to obtain a measure of relative performance.
+Relative skill scores [citation (not sure there is something published, preprint fron Estee Cramer and Nick Reich)] for every model can then be obtained by taking the geometric mean of all pairwise ratios calculated for each model (omitting comparisons where there is no overlapping set of forecasts).
+ADD FORMULAS.
+A skill score smaller than 1 indicates that a model is performing better than the average model.
+One can also scale the skill by a reference baseline.
+A scaled relative skill smaller than one then means that the model in question performed better than the baseline.
+
+By applying a permutation test or non-parametric test like the Wilcoxon rank sum test we can also obtain p-values that help determine whether or not models perform significantly differently.
+THERE IS AN ISSUE WITH CORRELATION: THE TEST WOULD TREAT ALL OBSERVATIONS AS INDEPENDENT, WHICH THEY ARENT'T, SO I WOULD ASSUME THAT IT IS TOO LIBERAL.
+CURRENTLY, NO ADJUSTMENTS ARE IMPLEMENTED, I'M NOT REALLY SURE WHETHER TO DISCUSS THIS HERE OR NOT.
+
+% ==============================================================================
+\section{Evaluating UK short-term forecasts}
+\subsection{The data}
+
+To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
+Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
+The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
+Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
+The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL].
+
+- timing (weekly?) and number of forecast dates
+- the four targets
+- the models. 
+- is the set complete?
+
+We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
+% 
+\begin{Schunk}
+\begin{Sinput}
+R> # install and load data pacakge from external repository
+R> # remotes::install_github("sbfnk/covid19.forecasts.uk")
+R> 
+R> # load packages
+R> library(covid19.forecasts.uk)
+R> library(dplyr)
+R> library(scoringutils)
+R> # load truth data
+R> data(covid_uk_data)
+R> # head(covid_uk_data)
+R> 
+R> # load forecasts
+R> data(uk_forecasts)
+R> # head(uk_forecasts)
+\end{Sinput}
+\end{Schunk}
+% 
+Let us take a first look a the data:
+% 
+\begin{Schunk}
+\begin{Sinput}
+R> glimpse(covid_uk_data)
+\end{Sinput}
+\begin{Soutput}
+Rows: 4,174
+Columns: 6
+$ geography  <fct> London, London, London, London, London, London, …
+$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospit…
+$ value_desc <fct> Hospital admissions, Hospital admissions, Hospit…
+$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …
+$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23,…
+$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546,…
+\end{Soutput}
+\begin{Sinput}
+R> glimpse(uk_forecasts)
+\end{Sinput}
+\begin{Soutput}
+Rows: 1,254,513
+Columns: 8
+$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, …
+$ geography     <chr> "East of England", "East of England", "East o…
+$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hos…
+$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-…
+$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-…
+$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.0…
+$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, …
+$ value_desc    <fct> Hospital admissions, Hospital admissions, Hos…
+\end{Soutput}
+\end{Schunk}
+% 
+To bring the forecasts into the format needed for the evaluation, some minor changes need to be made to the data.
+The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
+While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public.
+The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way.
+% 
+\begin{Schunk}
+\begin{Sinput}
+R> uk_forecasts <- rename(uk_forecasts, prediction = value)
+R> covid_uk_data <-rename(covid_uk_data, true_value = value)
+R> combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
+\end{Sinput}
+\end{Schunk}
+% 
+Before we start with scoring the forecasts, it makes sense to start the evaluation process by visualising the data.
+To get a feeling for how complete the data set is, we can run the following code to obtain a heatmap with the number of available forecasts:
+% 
+\begin{figure}[h]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> show_avail_forecasts(combined, 
++                       x = "creation_date", 
++                       show_numbers = FALSE,
++                       legend_position = "bottom",
++                       facet_formula = ~ value_desc)
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-show-availability}
+\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
+\end{figure}
+%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
+% 
+Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
+By default, the function treats a set of different quantiles or samples as one forecast.
+However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.
+
+Forecasts can be visualised using the \fct{plot\_predictions} function.
+The function accepts either a single combined data set or two separate truth data sets that can be merged together.
+Forecasts and observed values can be filtered independently, to show for example only predictions from a certain model or a specific number of weeks of true data before the forecast.
+Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
+To obtain, for example, a specific forecast from a model we are interest in, we can call:
+% 
+\begin{figure}[h!]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
+R> plot_predictions(truth = covid_uk_data,
++                   forecasts = uk_forecasts,
++                   filter_both = list(paste("geography %in%", locations, collapse = "")),
++                   filter_truth = list('value_date <= "2020-06-22"',
++                                       'value_date > "2020-06-01"'), 
++                   filter_forecasts = list('model == "SIRCOVID"', 
++                                           'creation_date == "2020-06-22"'),
++                   x = "value_date",
++                   facet_formula = geography ~ value_type) + 
++    ggplot2::theme(legend.position = "bottom")
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-show-forecasts}
+\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
+\end{figure}
+% 
+The output is shown in Figure \ref{fig:forecast-visualisation}.
+
+\subsection[Scoring forecasts with eval_forecasts]{Scoring forecasts with \fct{eval\_forecasts}}
+
+A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
+This requires a \code{data.frame} or similar which has at least a column called "prediction" and one called "true\_value".
+Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
+Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
+We will not discuss all possible input formats here, but instead refer to the example data for each format that is provided with the package.
+Where possible, \pkg{scoringutils} also provides functionality to transform between various formats, e.g.
+from a sample based format to a quantile format.
+The \fct{eval\_forecasts} function automatically recognises the prediction type and input format, applies the appropriate scoring metrics and aggregates results as desired by the user.
+Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
+
+As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
+% 
+\begin{Schunk}
+\begin{Sinput}
+R> scores <- eval_forecasts(combined, 
++                           summarise_by = c("model", "value_desc"))
+R> glimpse(scores)
+\end{Sinput}
+\begin{Soutput}
+Rows: 29
+Columns: 11
+$ model              <chr> "EpiSoon", "StructuredODE", "Microsimula…
+$ value_desc         <fct> Hospital admissions, Hospital admissions…
+$ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876,…
+$ sharpness          <dbl> 22.886153, 6.133974, 33.985553, 10.36508…
+$ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.…
+$ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.5…
+$ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174,…
+$ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0…
+$ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563,…
+$ relative_skill     <dbl> 0.9085917, 1.2479010, 0.7405995, 0.82232…
+$ scaled_rel_skill   <dbl> NaN, NaN, NaN, NaN, NaN, NaN, NaN, NaN, …
+\end{Soutput}
+\end{Schunk}
+% 
+If a more detailed analysis is desired, the level of aggregation can of course be changed to show for example separate scores for the different locations as well.
+This can be achieved using the \code{summarise_by} argument.
+To additionally stratify by location, we could specify \code{summarise_by = c("mode", "value_type", "geography")}.
+If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
+This can be useful if we, for example, want to analyse what proportion of true values are covered by certain interval ranges, or if we want to analyse the accuracy of the tails of the forecasts.
+When aggregating, \fct{eval\_forecasts} takes the mean according to the group defined in \code{summarise_by}.
+In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
+In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.
+
+The user must, however, still exercise some caution when aggregating scores, as many of the metrics are absolute and scale with the magnitude of the quantity to forecast.
+Looking at one score per model (i.e.
+specifying \code{summarise_by = c("model")}) may not be so useful in this instance, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
+
+In the above example, we did not have to explicitly specify the \code{by} argument, but this may be necessary if additional columns are present in the data that do not indicate a grouping of forecasts.
+The \code{by} argument must then be used to denote the unit of a single forecast.
+In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")}.
+Quantiles should not be included, as several quantiles make up one forecast (and similarly for samples).
+If we had additional columns that do not serve to group forecasts (like for example the number of inhabitants in a certain location over time), these should also not be included.
+By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
+
+\subsection{Pairwise comparisons}
+
+Pairwise comparisons between models [CITATION] can be obtained in two different ways.
+First, relative skill scores based on pairwise comparisons are by default returned from \fct{eval\_forecasts}.
+These will be computed separately for the categories defined in the \code{summarise_by} argument (excluding the category 'model').
+Alternatively, a set of scores can be post-processed using the separate function \fct{pairwise\_comparison}.
+This approach is to be used for visualisation and if p-values for the pairwise comparisons are needed, as those are not returned from \fct{eval\_forecasts}.
+Usually, one would compute scores without specifying a \code{summarise_by} argument, but sometimes it may be sensible to average over certain scores, for example for predictions generated at a certain date.
+This allows to reduce the correlation between observations that enter the computation of p-values, which in turn makes the test less liberal.
+Using the function \fct{plot\_pairwise\_comparison} we can visualise the mean score ratios between all models as well as the
+
+\begin{Schunk}
+\begin{Sinput}
+R> # unsummarised scores
+R> unsummarised_scores <- eval_forecasts(combined)
+R> pairwise <- pairwise_comparison(unsummarised_scores, 
++                                  summarise_by = "value_desc")
+\end{Sinput}
+\end{Schunk}
+
+The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
+
+\section{Visualisation and interpretation of evaluation results}
+
+\subsection{Visualising aggregate scores and rankings}
+A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
+
+\begin{figure}[h]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> score_table(scores, y = "model", facet_formula = ~ value_desc)
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-score-table}
+\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
+\end{figure}
+
+The most informative metric in terms of model ranking is the relative\_skill.
+However, interpretation is not always straightforward and has to be done carefully.
+We can see that performance varied quite a bit across different metrics, where some models did well on one target, but poorly on another.
+Especially the Exponential growth/decline model stands out as it received the lowest relative skill score for hospital admissions, but the highest for the total number of beds occupied.
+Looking back at Figure \ref{fig:avail-forecasts}, we see that the model has only submitted very few forecasts over all.
+It may therefore be sensible to require all models to have submitted forecasts for at least 50\% of all forecast targets in order to enter the pairwise comparisons.
+For similar reasons, the interval score may be deceiving if looked at in isolation.
+As can be seen, the DetSEIRwithNB MLE model received a lower relative skill score, but a higher interval score than the DetSEIRwithNB MCMC model.
+This, again, can be explained by the fact that they forecasted different targets.
+The interval score, as an absolute metric, is highly influenced by the absolute value of the quantity that is forecasted.
+For the same reason, one should be careful when summarising interval scores from different locations or forecast targets, as the average score will be dominated by outliers as well as differences in the absolute level.
+Assuming a large enough set of available overlapping forecasts, the relative skill score is more robust.
+It therefore is reasonable to assume that the DetSEIRwithNB MLE forecasted quantities with a higher absolute value, but tended to perform worse than the DetSEIRwithNB MCMC model as far as we can tell based on the set of all pariwise comparisons.
+This can be confirmed for the direct comparison between the two by looking at the mean score ratios from the pairwise comparisons.
+These can be obtained by calling
+
+\begin{figure}[h!]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> plot_pairwise_comparison(pairwise) +
++    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-pairwise-plot}
+\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
+If a tile is blue, then the model on the y-axis performed better.
+If it is red, the model on the x-axis performed better in direct comparison.
+}
+\end{figure}
+
+we can also look at p-values in Figure \ref{fig:pairwise-comparison-pval} PROBABLY REMOVE THAT FROM THE PAPER
+\begin{figure}[h!]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> plot_pairwise_comparison(pairwise, type = "pval") +
++    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-pairwise-plot-ratio}
+\caption{\label{fig:pairwise-comparison-pval} XXXX}
+\end{figure}
+
+Both plots can be combined in one: PLOT NEEDS FIXING AND ALSO PROBABLY REMOVE THAT FROM THE PAPER? Figure \ref{fig:pairwise-comparison-both}
+\begin{figure}[h!]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> plot_pairwise_comparison(pairwise, type = "together") +
++    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-pairwise-plot-pval}
+\caption{\label{fig:pairwise-comparison-both} p-values and ratios together.}
+\end{figure}
+
+In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
+
+\subsection{Visual model diagnostics}
+
+For forecasts in an interval format, looking at the components of the weighted interval score separately is a natural next step.
+We can see in Figure \ref{fig:wis-components} that the majority of penalties come from over-and underprediction, instead of the sharpness component.
+We also see that most models tended to either over- or underpredict actual numbers.
+
+\begin{figure}[h!]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> wis_components(scores, 
++                 facet_formula = ~ value_desc, 
++                 scales = "free_x") + 
++    ggplot2::coord_flip() + 
++    ggplot2::theme(legend.position = "bottom")
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-WIS-components}
+\caption{\label{fig:wis-components} p-values and ratios together.}
+\end{figure}
+
+We can have a closer look at calibration using the functions \fct{interval\_coverage} and \fct{quantile\_coverage}.
+The interval coverage plot shows the proportion of all true values that fall within all the different prediction intervals.
+This gives a visual impression of probabilistic calibration [GNEITING].
+Ideally, $x$ percent of true values should be covered by the $x$\%-prediction intervals, resulting in a 45° line.
+Areas shaded in green indicate that the model is covering more true values than it actually should, while areas in white indicate that the model fails to cover the desired proportion of true values with its prediction intervals.
+The majority of the models were too confident in their predictions, while some showed showed good calibration.
+The quantile coverage plot shows the proportion of all true values below certain predictive quantiles.
+While this plot is slightly harder to interpret, it also includes information about bias as and allows to separate the lower and upper boundaries of the prediction intervals.
+We can see, for example, that the Exponential growth/decline model was consistently biased downwards.
+Figure \ref{fig:coverage}
+
+\begin{figure}[h!]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> cov_scores <- eval_forecasts(combined, 
++                               summarise_by = c("model", 
++                                                "range", "quantile"))
+R> scoringutils::interval_coverage(cov_scores)
+R> scoringutils::quantile_coverage(cov_scores)
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-coverage}
+
+\begin{Schunk}
+\begin{Sinput}
+R> cov_scores <- eval_forecasts(combined, 
++                               summarise_by = c("model", 
++                                                "range", "quantile"))
+R> scoringutils::quantile_coverage(cov_scores)
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-quantile-coverage}
+
+\caption{\label{fig:coverage} quantile and interval coverage}
+\end{figure}
+
+DO I WANT TO INCLUDE PIT PLOTS AS WELL? I GUESS? NEED TO LOOK AT THE IMPLEMENTATION FOR QUANTILE FORECASTS
+
+% \begin{figure}[h]
+% \centering
+% <<
+% pit-plots, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+% # out <- eval_forecasts(combined, 
+% #                       summarise_by = c("model"), 
+% #                       pit_plots = TRUE)
+% 
+% @
+% \caption{\label{fig:calibration-pit} pit plots}
+% \end{figure}
+
+Look at e.g. bias by location? Figure \ref{fig:bias-heatmap}
+
+\begin{figure}[h]
+\centering
+\begin{Schunk}
+\begin{Sinput}
+R> scores <- eval_forecasts(combined, 
++                           summarise_by = c("model",
++                                            "value_desc", 
++                                            "geography"), 
++                           compute_relative_skill = FALSE)
+R> scoringutils::score_heatmap(scores, metric = "bias", 
++                              x = "geography", facet_formula = ~ value_desc, 
++                              scale = "free_x")
+\end{Sinput}
+\end{Schunk}
+\includegraphics{plots/plot-calibration}
+\caption{\label{fig:bias-heatmap} bias by location}
+\end{figure}
+
+WHAT IS NEEDED HERE IS A BIT OF THINKING WITH REGARDS TO WHAT VISUALISATION I WANT TO SHOW AND IN HOW MUCH DETAIL I WANT TO ANALYSE THE MODELS.
 
@@
 
+%% -- Summary/conclusions/discussion -------------------------------------------
+
+\section{Summary and discussion} \label{sec:summary}
+
+COMING SOON.
+
+% The results in this paper were obtained using
+% \proglang{R}~4.0.3 with the
+% \pkg{MASS}~7.3.53 package. \proglang{R} itself
+% and all packages used are available from the Comprehensive
+% \proglang{R} Archive Network (CRAN) at
+% \url{https://CRAN.R-project.org/}.
+
+\section*{Acknowledgments}
+
+% All acknowledgments (note the AE spelling) should be collected in this
+% unnumbered section before the references. It may contain the usual information
+% about funding and feedback from colleagues/reviewers/etc. Furthermore,
+% information such as relative contributions of the authors may be added here
+% (if any).
+
+%% -- Bibliography -------------------------------------------------------------
+%% - References need to be provided in a .bib BibTeX database.
+%% - All references should be made with \cite, \citet, \citep, \citealp etc.
+%%   (and never hard-coded). See the FAQ for details.
+%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
+%% - Titles in the .bib should be in title case.
+%% - DOIs should be included where available.
+
+\bibliography{refs}
+
+%% -- Appendix (if any) --------------------------------------------------------
+%% - After the bibliography with page break.
+%% - With proper section titles and _not_ just "Appendix".
+
+\newpage
+
+\begin{appendix}
+
+\section{Appendix section} \label{app:technical}
+
+\end{appendix}
+
+%% -----------------------------------------------------------------------------
+
 \end{document}

### 2021-03-02 article.Rnw
update paper with scores table and update plots

@@
 %% another package (only for this demo article)
-\usepackage{framed}
+\usepackage{booktabs}
+\usepackage{adjustbox}
+\usepackage{tabularx}
+\usepackage{rotating}
 
@@
 \section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}
+
+Table:
+
+\input{tables/scores-table.tex}
 
@@
 \centering
-<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12>>=
+<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12, eval = FALSE>>=
 show_avail_forecasts(combined, 
@@
 \centering
-<<show-forecasts, echo=TRUE, fig=TRUE, height=5.2, width=7>>=
+<<show-forecasts, echo=TRUE, fig=TRUE, height=5.2, width=7, eval = FALSE>>=
 locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
@@
 % 
-<<scoring>>=
+<<scoring, eval = FALSE>>=
 scores <- eval_forecasts(combined, 
@@
 
-<<pairwise-comparison>>=
+<<pairwise-comparison, eval = FALSE>>=
 # unsummarised scores
@@
 \centering
-<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=14>>=
+<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
 score_table(scores, y = "model", facet_formula = ~ value_desc)
@@
 \centering
-<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14>>=
+<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
 plot_pairwise_comparison(pairwise) +
@@
 \centering
-<<pairwise-plot-ratio, echo=TRUE, fig=TRUE, height=7.2, width=14>>=
+<<pairwise-plot-ratio, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
 plot_pairwise_comparison(pairwise, type = "pval") +
@@
 \caption{\label{fig:pairwise-comparison-pval} XXXX}
-\end{figure}
-
-Both plots can be combined in one: PLOT NEEDS FIXING AND ALSO PROBABLY REMOVE THAT FROM THE PAPER? Figure \ref{fig:pairwise-comparison-both}
-\begin{figure}[h!]
-\centering
-<<pairwise-plot-pval, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
-plot_pairwise_comparison(pairwise, type = "together") +
-  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-@
-\caption{\label{fig:pairwise-comparison-both} p-values and ratios together.}
 \end{figure}
@@
 \centering
-<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
+<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12, eval = FALSE>>=
 wis_components(scores, 
@@
 \centering
-<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold'>>=
+<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = FALSE>>=
 cov_scores <- eval_forecasts(combined, 
@@
 
-<<quantile-coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold'>>=
+<<quantile-coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = FALSE>>=
 cov_scores <- eval_forecasts(combined, 
@@
 \centering
-<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15>>=
+<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15, eval = FALSE>>=
 scores <- eval_forecasts(combined, 

### 2021-03-02 article.tex
update paper with scores table and update plots

@@
 \end{Schunk}
-\includegraphics{plots/plot-show-availability}
 \caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
@@
 \end{Schunk}
-\includegraphics{plots/plot-show-forecasts}
 \caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
@@
 \end{Sinput}
-\begin{Soutput}
-Rows: 29
-Columns: 11
-$ model              <chr> "EpiSoon", "StructuredODE", "Microsimula…
-$ value_desc         <fct> Hospital admissions, Hospital admissions…
-$ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876,…
-$ sharpness          <dbl> 22.886153, 6.133974, 33.985553, 10.36508…
-$ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.…
-$ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.5…
-$ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174,…
-$ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0…
-$ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563,…
-$ relative_skill     <dbl> 0.9085917, 1.2479010, 0.7405995, 0.82232…
-$ scaled_rel_skill   <dbl> NaN, NaN, NaN, NaN, NaN, NaN, NaN, NaN, …
-\end{Soutput}
 \end{Schunk}
@@
 \end{Schunk}
-\includegraphics{plots/plot-score-table}
 \caption{\label{fig:score-table} Coloured table to visualise the computed scores}
@@
 \end{Schunk}
-\includegraphics{plots/plot-pairwise-plot}
 \caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
@@
 \end{Schunk}
-\includegraphics{plots/plot-pairwise-plot-ratio}
 \caption{\label{fig:pairwise-comparison-pval} XXXX}
-\end{figure}
-
-Both plots can be combined in one: PLOT NEEDS FIXING AND ALSO PROBABLY REMOVE THAT FROM THE PAPER? Figure \ref{fig:pairwise-comparison-both}
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise, type = "together") +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-pairwise-plot-pval}
-\caption{\label{fig:pairwise-comparison-both} p-values and ratios together.}
 \end{figure}
@@
 \end{Schunk}
-\includegraphics{plots/plot-WIS-components}
 \caption{\label{fig:wis-components} p-values and ratios together.}
@@
 \end{Schunk}
-\includegraphics{plots/plot-coverage}
 
@@
 \end{Schunk}
-\includegraphics{plots/plot-quantile-coverage}
 
@@
 \end{Schunk}
-\includegraphics{plots/plot-calibration}
 \caption{\label{fig:bias-heatmap} bias by location}

### 2021-03-02 scores-table.Rnw
update paper with scores table and update plots

\documentclass{article}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{float}
\usepackage{longtable}
\begin{document}

<<table, echo=FALSE, results=tex>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")

crps <- list(
  `Metric` = "CRPS (Continuous) ranked probability score", 
  `Explanation` = r"(Proper scoring rule for continuous and integer forecasts that measures the distance between the CDF of the predictive distribution and the data-generating distribution. The CRPS is given as 
  $$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
  where y is the true observed value an F the predictive distribution. Often An alternative representation is used:
  $$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value. In this represenation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.  For integer-valued forecasts, the RPS is given as 
  $$ \text{RPS}(F, y) = \sum_{x = 0}^\infty (F(x) - 1(x \geq y))^2. $$

  \textbf{Usage}: Proper scoring rule, recommended in most instances, smaller values are better)", 
  `Caveats` = "",
  `References` = ""
)

log_score <- list(
  `Metric` = "Log score", 
  `Explanation` = r"(The Log score is a proper scoring rule and is simply the log of the predictive density evaluated at the true observed value. It is given as 
  $$ \text{log score} = \log f(y), $$ where $f$ is the predictive density function and y is the true value.

  \textbf{Usage and caveats}: Larger values are better, but sometimes the sign is reversed. Sensitive to outliers, as individual negative log score contributions can become very large if $f(y)$ is close to zero. In practice the log score is also hard to use for integer values, as a predictive density is required.)",
  `Caveats` = "",
  `References` = ""
)

wis <- list(
  Metric = "WIS (Weighted) interval score", 
  `Explanation` = r"(Proper scoring rule for quantile forecasts. converges to crps for increasing number of interval. The score can be decomposed into a sharpness contribution and penalties for over- and underprediction. For a single interval, the score is computed as $$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u)$$, where 1() is the indicator function, y is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantile of $F$, i.e. the lower and upper bound of a single prediction interval. For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, $WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y))$. $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.

  \textbf{Caveat}: The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.)", 
  `Caveats` = "The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.",
  `References` = ""
)

dss <- list(
  `Metric` = "DSS Dawid-Sebastiani score", 
  `Explanation` = r"(proper scoring rule for continuous and integer forecasts. The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution. If in doubt, we would recommend the crps, but the difference should not be large in practice.)",
  `Caveats` = "",
  `References` = ""
)

brier_score <- list(
  `Metric` = "Brier score", 
  `Explanation` = r"(Proper scoring rule for binary forecasts. $\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (\text{prediction}_n - \text{outcome}_n)^2$, where prediction)",
  `Caveats` = "",
  `References` = ""
)

interval_coverage <- list(
  `Metric` = "interval coverage", 
  `Explanation` = r"((Interval) coverage for a single prediction interval can be calculated as $IC_{\alpha} = \text{nominal coverage} - \text{actual empirical coverage}$, where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by the $1 - \alpha$ prediction intervals. Interval coverage can then be aggregated over all interval levels: $\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$)",
  `Caveats` = "",
  `References` = ""
)

quantile_coverage <- list(
  `Metric` = "Quantile coverage", 
  `Explanation` = r"(Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level. )",
  `Caveats` = "",
  `References` = ""
)

coverage_deviation <- list(
  `Metric` = "", 
  `Explanation` = r"()",
  `Caveats` = "",
  `References` = ""
)

bias <- list(
  `Metric` = "Bias", 
  `Explanation` = r"(For continuous forecasts, bias is given as $B (F, y) = 1 - 2 \cdot (F (y))$. For integer-valued forecasts, bias can be calculated as$B (F, y) = 1 - (F (y) + P_t (x_t + 1))$ and for quantile forecasts as $B =$ the maximum percentile rank that satisfies prediction smaller than y, if the true value is smaller than the median of the predictive distribution. If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value. If the true value is exactly the median, both terms cancel out and $B_t$ is zero. For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.)",
  `Caveats` = "",
  `References` = ""
)

mean_score_ratio <- list(
  `Metric` = "Mean score ratio", 
  `Explanation` = r"(The mean score ratio is used to compare two models on the overlapping set of forecast targets for which both models have made a prediction. It is calculated as the mean score achieved by the first model over the mean score achieved by the second model.)",
  `Caveats` = "",
  `References` = ""
)

relative_skill <- list(
  `Metric` = "Mean score ratio", 
  `Explanation` = r"(Relative skill is used to create a ranking between models based on pairwise comparisons between all models. To compute the relative skill of model $m$, we take the geometric mean of all mean score ratios that involve model $m$. )",
  `Caveats` = "",
  `References` = ""
)

data <- rbind(as.data.frame(crps), 
              as.data.frame(log_score), 
              as.data.frame(wis), 
              as.data.frame(dss), 
              as.data.frame(brier_score), 
              as.data.frame(interval_coverage),
              as.data.frame(quantile_coverage), 
              as.data.frame(bias), 
              as.data.frame(mean_score_ratio), 
              as.data.frame(relative_skill))

library(kableExtra)
library(magrittr)
library(knitr)
data[, 1:2] %>%
  kableExtra::kbl(format = "latex", booktabs = TRUE, 
                  escape = FALSE,
                  longtable = TRUE,
                  linesep = c('\\addlinespace \\addlinespace')) %>%
  kableExtra::column_spec(1, width = "2.5cm") %>%
  kableExtra::column_spec(2, width = "16.5cm") %>%
  # kableExtra::column_spec(3, width = "4.5cm") %>%
  kableExtra::kable_styling(latex_options = c("striped", "repeat_header")) %>%
  kableExtra::landscape()

@

\end{document}

### 2021-03-02 scores-table.tex
update paper with scores table and update plots

--- 2021-03-02 scores-table.Rnw
@@
 \usepackage{longtable}
+\usepackage{Sweave}
 \begin{document}
 
-<<table, echo=FALSE, results=tex>>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-library("MASS")
+\begin{landscape}
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{16.5cm}}
+\toprule
+Metric & Explanation\\
+\midrule
+\endfirsthead
+\multicolumn{2}{@{}l}{\textit{(continued)}}\\
+\toprule
+Metric & Explanation\\
+\midrule
+\endhead
 
-crps <- list(
-  `Metric` = "CRPS (Continuous) ranked probability score", 
-  `Explanation` = r"(Proper scoring rule for continuous and integer forecasts that measures the distance between the CDF of the predictive distribution and the data-generating distribution.
+\endfoot
+\bottomrule
+\endlastfoot
+CRPS (Continuous) ranked probability score & Proper scoring rule for continuous and integer forecasts that measures the distance between the CDF of the predictive distribution and the data-generating distribution.
 The CRPS is given as
@@
 
-  \textbf{Usage}: Proper scoring rule, recommended in most instances, smaller values are better)", 
-  `Caveats` = "",
-  `References` = ""
-)
-
-log_score <- list(
-  `Metric` = "Log score", 
-  `Explanation` = r"(The Log score is a proper scoring rule and is simply the log of the predictive density evaluated at the true observed value. It is given as 
+\cellcolor{gray!6}{  \textbf{Usage}: Proper scoring rule, recommended in most instances, smaller values are better}\\
+\addlinespace \addlinespace
+Log score & The Log score is a proper scoring rule and is simply the log of the predictive density evaluated at the true observed value. It is given as 
   $$ \text{log score} = \log f(y), $$ where $f$ is the predictive density function and y is the true value.
@@
 Sensitive to outliers, as individual negative log score contributions can become very large if $f(y)$ is close to zero.
-In practice the log score is also hard to use for integer values, as a predictive density is required.)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-wis <- list(
-  Metric = "WIS (Weighted) interval score", 
-  `Explanation` = r"(Proper scoring rule for quantile forecasts.
+In practice the log score is also hard to use for integer values, as a predictive density is required.\\
+\addlinespace \addlinespace
+WIS (Weighted) interval score & Proper scoring rule for quantile forecasts.
 converges to crps for increasing number of interval.
@@
 
-  \textbf{Caveat}: The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.)", 
-  `Caveats` = "The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.",
-  `References` = ""
-)
-
-dss <- list(
-  `Metric` = "DSS Dawid-Sebastiani score", 
-  `Explanation` = r"(proper scoring rule for continuous and integer forecasts.
+\cellcolor{gray!6}{  \textbf{Caveat}: The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.}\\
+\addlinespace \addlinespace
+DSS Dawid-Sebastiani score & proper scoring rule for continuous and integer forecasts.
 The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution.
-If in doubt, we would recommend the crps, but the difference should not be large in practice.)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-brier_score <- list(
-  `Metric` = "Brier score", 
-  `Explanation` = r"(Proper scoring rule for binary forecasts. $\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (\text{prediction}_n - \text{outcome}_n)^2$, where prediction)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-interval_coverage <- list(
-  `Metric` = "interval coverage", 
-  `Explanation` = r"((Interval) coverage for a single prediction interval can be calculated as $IC_{\alpha} = \text{nominal coverage} - \text{actual empirical coverage}$, where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by the $1 - \alpha$ prediction intervals.
-Interval coverage can then be aggregated over all interval levels: $\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-quantile_coverage <- list(
-  `Metric` = "Quantile coverage", 
-  `Explanation` = r"(Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level. )",
-  `Caveats` = "",
-  `References` = ""
-)
-
-coverage_deviation <- list(
-  `Metric` = "", 
-  `Explanation` = r"()",
-  `Caveats` = "",
-  `References` = ""
-)
-
-bias <- list(
-  `Metric` = "Bias", 
-  `Explanation` = r"(For continuous forecasts, bias is given as $B (F, y) = 1 - 2 \cdot (F (y))$.
+If in doubt, we would recommend the crps, but the difference should not be large in practice.\\
+\addlinespace \addlinespace
+\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{Proper scoring rule for binary forecasts.
+$\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (\text{prediction}_n - \text{outcome}_n)^2$, where prediction}\\
+\addlinespace \addlinespace
+interval coverage & (Interval) coverage for a single prediction interval can be calculated as $IC_{\alpha} = \text{nominal coverage} - \text{actual empirical coverage}$, where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by the $1 - \alpha$ prediction intervals.
+Interval coverage can then be aggregated over all interval levels: $\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$\\
+\addlinespace \addlinespace
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level.}\\
+\addlinespace \addlinespace
+Bias & For continuous forecasts, bias is given as $B (F, y) = 1 - 2 \cdot (F (y))$.
 For integer-valued forecasts, bias can be calculated as$B (F, y) = 1 - (F (y) + P_t (x_t + 1))$ and for quantile forecasts as $B =$ the maximum percentile rank that satisfies prediction smaller than y, if the true value is smaller than the median of the predictive distribution.
@@
 If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
-For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-mean_score_ratio <- list(
-  `Metric` = "Mean score ratio", 
-  `Explanation` = r"(The mean score ratio is used to compare two models on the overlapping set of forecast targets for which both models have made a prediction.
-It is calculated as the mean score achieved by the first model over the mean score achieved by the second model.)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-relative_skill <- list(
-  `Metric` = "Mean score ratio", 
-  `Explanation` = r"(Relative skill is used to create a ranking between models based on pairwise comparisons between all models.
-To compute the relative skill of model $m$, we take the geometric mean of all mean score ratios that involve model $m$.
-)",
-  `Caveats` = "",
-  `References` = ""
-)
-
-data <- rbind(as.data.frame(crps), 
-              as.data.frame(log_score), 
-              as.data.frame(wis), 
-              as.data.frame(dss), 
-              as.data.frame(brier_score), 
-              as.data.frame(interval_coverage),
-              as.data.frame(quantile_coverage), 
-              as.data.frame(bias), 
-              as.data.frame(mean_score_ratio), 
-              as.data.frame(relative_skill))
-
-library(kableExtra)
-library(magrittr)
-library(knitr)
-data[, 1:2] %>%
-  kableExtra::kbl(format = "latex", booktabs = TRUE, 
-                  escape = FALSE,
-                  longtable = TRUE,
-                  linesep = c('\\addlinespace \\addlinespace')) %>%
-  kableExtra::column_spec(1, width = "2.5cm") %>%
-  kableExtra::column_spec(2, width = "16.5cm") %>%
-  # kableExtra::column_spec(3, width = "4.5cm") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "repeat_header")) %>%
-  kableExtra::landscape()
-
-@
+For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.\\
+\addlinespace \addlinespace
+\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{The mean score ratio is used to compare two models on the overlapping set of forecast targets for which both models have made a prediction.
+It is calculated as the mean score achieved by the first model over the mean score achieved by the second model.}\\
+\addlinespace \addlinespace
+Mean score ratio & Relative skill is used to create a ranking between models based on pairwise comparisons between all models.
+To compute the relative skill of model $m$, we take the geometric mean of all mean score ratios that involve model $m$.\\*
+\end{longtable}
+\end{landscape}
 

### 2021-03-02 tables/scores-table.Rnw
update paper with scores table and update plots

% !TeX root = ../article.Rnw
\begin{table}[]
\centering
\begin{tabular}{@{}llll@{}}
\toprule
Metric                                                           & Formula                           & Application and Explanation                                                                                                                                                                                                                                                                                                             & Caveats                                                                                                                                                                                                     \\ \midrule
\multicolumn{1}{|l|}{(weighted) interval score (wis)}            & \multicolumn{1}{l|}{$\alpha$}     & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}proper scoring rule for quantile forecasts. \\ converges to crps for increasing number of interval. \\ The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.\end{tabular}}                                                              & \begin{tabular}[c]{@{}l@{}}The wis is based on measures of absolute error. \\ When averaging across multiple targets, it will therefore be dominated\\ by targets with higher absolute values.\end{tabular} \\ \cmidrule(r){1-3}
\multicolumn{1}{|l|}{continuous ranked probability score (crps)} & \multicolumn{1}{l|}{4.892523e+02} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}proper scoring rule for continuous and integer forecasts. \\ Implemented for sample-based forecasts.\end{tabular}}                                                                                                                                                                       & Also sensitive to the absolute value of the predicted quantity                                                                                                                                              \\ \cmidrule(r){1-3}
\multicolumn{1}{|l|}{Dawid-Sebastiani score (dss)}               & \multicolumn{1}{l|}{2.257546e+02} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}proper scoring rule for continuous and integer forecasts. \\ The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution. \\ If in doubt, we would recommend the crps, but the difference should not be large in practice.\end{tabular}} &                                                                                                                                                                                                             \\ \cmidrule(r){1-3}
log score                                                        &                                   & The log score is a proper scoring rule best suited for continuous forecasts.                                                                                                                                                                                                                                                            & Sensitive to outliers. Individual log score contributions can go to infinity                                                                                                                                \\
Brier score                                                      &                                   & Proper scoring rule for binary forecasts                                                                                                                                                                                                                                                                                                &                                                                                                                                                                                                             \\
interval coverage                                                &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
quantile coverage                                                &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
Coverage deviation                                               &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
Bias                                                             &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\ \cmidrule(r){1-3}
Mean score ratio                                                 &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
Relative skill                                                   &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
Absolute error of the median                                     &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
Mean squared error                                               &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\ \bottomrule
\end{tabular}
\end{table}

### 2021-03-02 tables/scores-table.tex
update paper with scores table and update plots

% !TeX root = ../article.Rnw
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{sidewaystable}[]

% \begin{adjustbox}{angle=90}

\centering
\begin{tabularx}{\linewidth}{|X|X|X|X|}

\hline
Metric                                                                  & Formula                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         & Application and Explanation                                                                                                                                                                                                                                                                                                                                                                                                                    & Caveats                                                                                                                                                        \\ \hline
WIS (Weighted) interval score                                           & For a single interval, the score is computed as $IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u)$, where 1() is the indicator function, y is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantile of $F$, i.e. the lower and upper bound of a single prediction interval. For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, $WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y))$. $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$                                                                                                                                                                                                                   & proper scoring rule for quantile forecasts. converges to crps for increasing number of interval. The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.                                                                                                                                                                                                                                        & The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values. \\
CRPS (Continuous) ranked probability score                              & $\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \mathbb{1}(x \geq y) \right)^2 dx$, where y is the true observed value an F the predictive distribution. Another represenation allows for easier computation using samples. Take $ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|$, where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value. We can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.  For integer-valued forecasts, the RPS is given as $ \text{RPS}(F, y) = \sum_{x = 0}^\infty (F(x) - \mathbb{1} (x \geq y))^2. $                                                                                                                                                                                                     & proper scoring rule for continuous and integer forecasts. Implemented for sample-based forecasts.                                                                                                                                                                                                                                                                                                                                              & Also sensitive to the absolute value of the predicted quantity                                                                                                 \\
\begin{tabular}[c]{@{}l@{}}DSS\\ \\ Dawid-Sebastiani score\end{tabular} & 2.257546e+02                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    & proper scoring rule for continuous and integer forecasts. The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution. If in doubt, we would recommend the crps, but the difference should not be large in practice.                                                                                                                                                                           &                                                                                                                                                                \\
log score                                                               & The Log score is given by $ \text{log score} = \log f(y)$, where $f$ is the predictive density function and y is the true value.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                & \begin{tabular}[c]{@{}l@{}}The log score is a proper scoring rule best suited for continuous forecasts. While in principle applicable to integer valued forecasts, its computation requires a density function which is in practice difficult for integer valued forecasts. \\ One concern with the log score is that it can assign extreme scores if the predictive distribution gives little probability to the observed event.\end{tabular} & Sensitive to outliers. Individual log score contributions can go to infinity                                                                                   \\
Brier score                                                             & $\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (\text{prediction}_n - \text{outcome}_n)^2$, where prediction                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                & Proper scoring rule for binary forecasts                                                                                                                                                                                                                                                                                                                                                                                                       &                                                                                                                                                                \\
interval coverage                                                       & (Interval) coverage for a single prediction interval can be calculated as $IC_\alpha = \text{nominal coverage} - \text{actual empirical coverage}$, where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by the $1 - \alpha$ prediction intervals. Interval coverage can then be aggregated over all interval levels: $\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} IC_{\alpha_k}$                                                                                                                                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                                                                                                                                                                                                                                                                &                                                                                                                                                                \\

\end{tabularx}

% \end{adjustbox}

\end{sidewaystable}

### 2021-03-02 scores-table.R

@@ -1,0 +1,14 @@
+scores_table <- data.frame(
+  
+)
+
+wis <- list(
+  Metric = "WIS (Weighted) interval score", 
+  `Formula` = r"(For a single interval, the score is computed as $IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u)$, where 1() is the indicator function, y is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantile of $F$, i.e. the lower and upper bound of a single prediction interval. For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, $WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y))$. $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$)", 
+  `Application and Explanation` = "proper scoring rule for quantile forecasts. converges to crps for increasing number of interval. The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.", 
+  `Caveats` = "The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values."
+)
+
+library(magrittr)
+as.data.frame(wis) %>%
+  kableExtra::kable()

### 2021-03-05 scores-table.Rnw
update scores table

@@
   `Metric` = "CRPS (Continuous) ranked probability score", 
-  `Explanation` = r"(Proper scoring rule for continuous and integer forecasts that measures the distance between the CDF of the predictive distribution and the data-generating distribution.
+  `Explanation` = r"(The crps is a proper scoring rule that generalises the absolute error to probabilistic forecasts.
+It measures the 'distance' of the predictive distribution to the observed data-generating distribution.
 The CRPS is given as
   $$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
-  where y is the true observed value an F the predictive distribution. Often An alternative representation is used:
+  where y is the true observed value and F the CDF of predictive distribution. Often An alternative representation is used:
   $$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
@@
 
-  \textbf{Usage}: Proper scoring rule, recommended in most instances, smaller values are better)", 
+\textbf{Usage and caveats} Smaller values are better.
+The crps is a good choice for most practical purposes that involve decision making, as it takes the entire predictive distribution into account.
+If two forecasters assign the same probability to the true event $y$, then the forecaster who assigned high probability to events far away from $y$ will still get a worse score.
+The crps (in contrast to the log score) can at times be quite lenient towards extreme mispredictions.
+Also, due to it's similarity to the absolute error, the level of scores depend a lot on the absolute value of what is predicted, which makes it hard to compare scores of forecasts for quantities that are orders of magnitude apart.)",
   `Caveats` = "",
@@
   `Metric` = "Log score", 
-  `Explanation` = r"(The Log score is a proper scoring rule and is simply the log of the predictive density evaluated at the true observed value. It is given as 
-  $$ \text{log score} = \log f(y), $$ where $f$ is the predictive density function and y is the true value.
+  `Explanation` = r"(The Log score is a proper scoring rule that is simply compuated as the log of the predictive density evaluated at the true observed value. It is given as 
+  $$ \text{log score} = \log f(y), $$ 
+  where $f$ is the predictive density function and y is the true value. For integer-valued forecasts, the log score can be computed as 
+  $$ \text{log score} = \log p_y, $$
+  where $p_y$ is the probability assigned to outcome p by the forecast F.
 
 \textbf{Usage and caveats}: Larger values are better, but sometimes the sign is reversed.
-Sensitive to outliers, as individual negative log score contributions can become very large if $f(y)$ is close to zero.
-In practice the log score is also hard to use for integer values, as a predictive density is required.)",
+The log score is ensitive to outliers, as individual negative log score contributions quickly can become very large if the event falls in the tails of the predictive distribution, where $f(y)$ (or $p_y$) is close to zero.
+Whether or not that is desirable depends ont the application.
+In scoringutils, the log score cannot be used for integer-valued forecasts, as the implementation requires a predictive density.
+In contrast to the crps, the log score is a local scoring rule: it's value only depends only on the probability that was assigned to the actual outcome.
+This property is desirable for inferential purposes, for example in a Bayesian context (@winklerScoringRulesEvaluation1996).
+In settings where forecasts inform decision making, it may be more appropriate to score forecasts based on the entire predictive distribution.)",
   `Caveats` = "",
@@
   Metric = "WIS (Weighted) interval score", 
-  `Explanation` = r"(Proper scoring rule for quantile forecasts.
-converges to crps for increasing number of interval.
-The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.
-For a single interval, the score is computed as $$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u)$$, where 1() is the indicator function, y is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantile of $F$, i.e.
+  `Explanation` = r"(The (weighted) interval score is a proper scoring rule for quantile forecasts that converges to the crps for an increasing number of intervals.
+The score can be decomposed into a sharpness (uncertainty) component and penalties for over- and underprediction.
+For a single interval, the score is computed as
+  $$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
+  where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
 the lower and upper bound of a single prediction interval.
-For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, $WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y))$.
-$w_k$ is a weight for every interval.
-Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum,
+  $$WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),$$ 
+  where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
 
-  \textbf{Caveat}: The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.)", 
+\textbf{Usage and caveats}:
+  Smaller scores are better.
+Applicable to all quantile forecasts, takes the entire predictive distribution into account.
+Just as the crps, the wis is based on measures of absolute error.
+When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.
+The decomposition into sharpness, over- and underprediction make it easy to interpret scores and use them for model improvement.
+)",
   `Caveats` = "The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.",
@@
 
+quantile_score <- "yet to come"
+
 dss <- list(
   `Metric` = "DSS Dawid-Sebastiani score", 
-  `Explanation` = r"(proper scoring rule for continuous and integer forecasts.
-The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution.
-If in doubt, we would recommend the crps, but the difference should not be large in practice.)",
+  `Explanation` = r"(The Dawid-Sebastiani-Score is a proper scoring rule proposed by Gneiting and Raftery in [@gneitingStrictlyProperScoring2007] that only relies on the first moments of the predictive distribution and is therefore easy to compute.
+It is given as
+
+$$\text{dss}(F, y) = \left( \frac{y - \mu}{\sigma} \right) + 2 \cdot \log \sigma,$$
+where $F$ is the predictive distribution with mean $\mu$ and standard deviation $\sigma$ and $y$ is the true observed value.
+
+textbf{Usage and caveats} The dss is applicable to continuous and integer forecasts and easy to compute.
+Apart from the ease of computation we see little advantage in using it.
+MAYBE SCRATCH IT ALTOGETHER FROM THE PAPER?)",
   `Caveats` = "",
@@
   `Metric` = "Brier score", 
-  `Explanation` = r"(Proper scoring rule for binary forecasts. $\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (\text{prediction}_n - \text{outcome}_n)^2$, where prediction)",
+  `Explanation` = r"(Proper scoring rule for binary forecasts. The Brier score is computed as 
+$$\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (f_n - y_n), $$
+where $f_n$, with $n = 1, \dots, N$ are the predicted probablities that the corresponding events, $y_n \in (0, 1)$ will be equal to one.)
+
+\textbf{Usage}: Applicable to all binary forecasts.)",
   `Caveats` = "",
@@
 interval_coverage <- list(
-  `Metric` = "interval coverage", 
-  `Explanation` = r"((Interval) coverage for a single prediction interval can be calculated as $IC_{\alpha} = \text{nominal coverage} - \text{actual empirical coverage}$, where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by the $1 - \alpha$ prediction intervals.
-Interval coverage can then be aggregated over all interval levels: $\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$)",
+  `Metric` = "Interval coverage", 
+  `Explanation` = r"(Interval coverage measures the proportion of observed values that fall in a given prediction interval range.
+Interval coverage for a single prediction interval range can be calculated as $$IC_{\alpha} = \text{nominal coverage} - \text{empirical coverage},$$
+where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by all $1 - \alpha$ prediction intervals.
+
+To summarise interval coverage over different over multiple interval ranges, we can compute coverage deviation defined as the mean interval coverage over all $K$ interval ranges $\alpha_k$ with $k = 1, \dots, K$:
+$$\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$$
+
+\textbf{Usage}: Interval coverage for a set of chosen intervals, (e.g.
+50\% and 90\%) gives a good indication of marginal calibration and is easy to interpret.
+Reporting coverage deviation has the advantage of summarising calibration in a single number, but loses some of the nuance.)",
   `Caveats` = "",
@@
   `Metric` = "Quantile coverage", 
-  `Explanation` = r"(Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level. )",
-  `Caveats` = "",
-  `References` = ""
-)
+  `Explanation` = r"(Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level.
 
-coverage_deviation <- list(
-  `Metric` = "", 
-  `Explanation` = r"()",
+\textbf{Usage}: Quantile coverage is similar to interval coverage, but conveys more information.
+For example, it allows us to look at the 5\% and 95\% quantile separately, instead of jointly at the 90\% prediction interval).
+This helps to diagnose whether it is the upper or lower end of a prediction interval that is causing problems.
+Plots of quantile coverage are conceptually very similar to PIT histograms.)",
   `Caveats` = "",
@@
   `Metric` = "Bias", 
-  `Explanation` = r"(For continuous forecasts, bias is given as $B (F, y) = 1 - 2 \cdot (F (y))$.
-For integer-valued forecasts, bias can be calculated as$B (F, y) = 1 - (F (y) + P_t (x_t + 1))$ and for quantile forecasts as $B =$ the maximum percentile rank that satisfies prediction smaller than y, if the true value is smaller than the median of the predictive distribution.
-If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
-If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
-For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.)",
+  `Explanation` = r"(Bias is a measure of the tendency of a forecaster to over- or underpredict. For continuous forecasts, bias is given as 
+$$B(F, y) = 1 - 2 \cdot (F (y)), $$ 
+where $F$ is the CDF of the predictive distribution and $y$ is the observed value.
+
+For integer-valued forecasts, bias can be calculated as 
+$$B(P, y) = 1 - (P(y) + P(y + 1)), $$ 
+where $P(y)$ is the cumulative probability assigned to all outcomes smaller or equal to $y$.
+
+For quantile forecasts, Bias can be calculated as the maximum percentile rank for which the prediction is smaller than $y$, if the true value is smaller than the median of the predictive distribution.
+If the true value is above the median of the predictive distribution, then bias is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
+If the true value is exactly the median, bias is zero.
+For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
+
+\textbf{Usage}:
+In contrast to the over- and underprediction penalties of the interval score it is bound between 0 and 1 and represents the tendency of forecasts to be biased rather than the absolute amount of over- and underprediction.
+It is therefore a more robust measurement, but harder to interpet.
+It largely depends on the application whether one is more interested in the tendency to be biased or in the absolute value of over- and underpredictions.)",
   `Caveats` = "",
@@
   `Explanation` = r"(The mean score ratio is used to compare two models on the overlapping set of forecast targets for which both models have made a prediction.
-It is calculated as the mean score achieved by the first model over the mean score achieved by the second model.)",
+The mean score ratio is calculated as the mean score achieved by the first model over the mean score achieved by the second model.
+More precisely, for two models $i, j$, we determine the set of overlapping forecasts, denoted by $\mathcal{A}_{ij}$ and compute the mean score ratio $\theta_{ij}$ as
+$$\theta_{ij} =\frac{\text{mean score model } i \text{ on } \mathcal{A}_{ij}}{\text{mean score model } j \text{ on } \mathcal{A}_{ij}}.$$
+The mean score ratio can in principle be computed for any arbitrary score.
+
+\textbf{Usage}:
+Mean scores ratios are usually calculated in the context of pairwise comparisons, where a set of models is compared by looking at mean score ratios of all possible parings.
+Whether smaller or larger values are better depends on the orientation of the original score used)",
   `Caveats` = "",
@@
 relative_skill <- list(
-  `Metric` = "Mean score ratio", 
-  `Explanation` = r"(Relative skill is used to create a ranking between models based on pairwise comparisons between all models.
-To compute the relative skill of model $m$, we take the geometric mean of all mean score ratios that involve model $m$.
+  `Metric` = "Relative skill", 
+  `Explanation` = r"(Relative skill scores can be used to obtain a ranking of models based on pairwise comparisons between all models.
+To compute the relative skill $\theta_i$ of model $i$, we take the geometric mean of all [mean score ratios](mean%20score%20ratio.md) that involve model $i$, i.e.
+$$ \theta_{i} = \left(\prod_{m = 1}^M \theta_{im}\right)^{1/M}, $$
+where M is the number of models.
+
+\textbf{Usage and caveats}: 
+Relative skill is a helpful way to obtain a model ranking. Whether smaller or larger values are better depends on the orientation of the original score used. 
+It is in principle relatively robust against biases that arise when models only forecast some of the available targets and is a reasonable way to handle missing forecasts.
+One possible precautionary measure to reduces issues with missing forecasts is to only compare models that have forecasted at least half of all possible targets (this ensures that there is always an overlap between models).
+If there is no overlap between models, the relative skill implicitly estimates how a model would have forecasted on those missing targets.
 )",

### 2021-03-06 article.Rnw
delete some unneccessary text

@@
 
-A third way to look at calibration is to look a the probability integral tranform.
-As explained previously, the CDF of predictice distribution $P_t$ should ideally be equal to the CDF of the true unknown distribution $F_t$ that generated the observed value $x_t$.
-In order to assess whether there are substantial deviations between the two, @dawidPresentPositionPotential1984 suggested to transform the observed values using the probability integral transform (PIT).
-Agreement between the forecasts and the observed values can then be examined by observing whether or not the transformed values follow a uniform distribution.
-The PIT is given by
-$$u_t = P_t (x_t),$$
-where $u_t$ is the transformed variable and $P_t(x_t)$ is the predictive distribution evaluated at the true observed value $x_t$.
-If $P_t = F_t$ at all times $t$, then $u_t, t = 1 \dots T$ follows a uniform distribution (for a proof see e.g.
-@angusProbabilityIntegralTransform1994).
-
-In the case of discrete outcomes, the PIT is no longer uniform even when forecasts are ideal. As @funkAssessingPerformanceRealtime2019 suggest, we use a randomised PIT instead by redefining 
-$$u_t = P_t(x_t) + vt \cdot (P_t(x_t) - P_t(x_t - 1) ),$$
-where $x_t$ is again the observed value at time $t$, $P_t()$ is the CDF of the predictive distribution function, $P_t (-1) = 0$ by definition, and $v_t$ is a standard uniform variable independent of $x_t$.
-If $P_t$ is equal to the true data-generating distribution function, then $u_t$ is standard uniform.
-@czadoPredictiveModelAssessment2009 also propose a non-randomised version of the PIT for count data that could be used alternatively.
-
-One can then plot a histogram of $u_t$ values to look for deviations from uniformity.
-U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
-Biased predictions will usually result in a triangle-shaped histogram.
-
 \subsubsection{Assessing sharpness}
-
-Sharpness is the second property central to model evaluation.
-The ability to produce narrow forecasts is a quality of the forecasts only and does not depend on the observations.
-Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
-Again we need to take slightly different approaches for continuous, integer and quantile forecasts.
-For continuous and integer forecasts we follow the suggestion from @funkAssessingPerformanceRealtime2019.
-For quantile forecasts we propose a novel metric.
-
-For continuous and integer forecasts, @funkAssessingPerformanceRealtime2019 suggest to measure sharpness as the normalised median absolute deviation about the median (MADN), i.e. 
-$$ S_t(P_t) = \frac{1}{0.675} \cdot \text{median}(|y - \text{median(y)}|), $$ 
-where $y$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant that ensures that sharpness will equal the standard deviation of the predictive distribution if $P_t$ is the CDF of a normal distribution.
-
-MAKE THE FOLLOWING PARAGRAPH SHORTER AND JUST SAY WE USE THE SHARPNESS COMPONENT OF THE WIS.
-
-For quantile forecasts, we propose to measure sharpness as a weighted mean of the width of the interval ranges.
-Let $Q_t$ be a set of predicted quantiles for a true $x_t$ at time $t$.
-This set of quantiles is assumed to be symmetric, such that there exist $K$ corresponding pairs of elements $q_{t, \frac{\alpha}{2}}$ and $q_{t, 1-\frac{\alpha}{2}}$.
-These $K$ corresponding pairs of quantiles cover a $(1 - \alpha) \cdot 100$ prediction interval.
-We can, accordingly, also denote $q_{t, \frac{\alpha}{2}}$ as $l_t$ the lower bound of the prediction interval at time $t$ and $q_{t, 1-\frac{\alpha}{2}}$ as $u_t$, the upper bound.
-We measure the sharpness of a quantile forecast at time $t$ as
-\begin{align*}
-\text{sharpness}_t &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (q_{t, 1 - \frac{\alpha}{2}} - q_{t, \frac{\alpha}{2}}) \\
-                   &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (u_t - l_t).
-\end{align*}
-Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
-We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score.
 

### 2021-03-06 article.Rnw
shorten section about forecast types and add figure

@@
 
-\subsection{Forecast types}
+\subsection{Forecast types and forecast formats}
 
-In its most general sense, a forecast is the forecaster’s stated belief about the future \citep{gneitingStrictlyProperScoring2007}.
-For conceptual clarity, it is however useful to distinguish between different types of forecasts.
-In terms of the quantitative forecasts that can be evaluated with \pkg{soringutils}, we can distinguish point forecasts from probabilistic forecasts.
-A point forecast states a single number and is the simplest form of a forecast.
-It is limited in its usefulness, as it does not give information about the uncertainty of the forecast.
-A very certain forecast may, for example, warrant a very different course of actions than does a very uncertain one.
-A probabilistic forecast, in contrast, is a forecast made in terms of the entire predictive distribution.
-Providing the entire predictive distribution allows the forecaster to express their belief about all aspects of the underlying data-generating distribution (including e.g.
+In its most general sense, a forecast is the forecaster’s stated belief about the future \citep{gneitingStrictlyProperScoring2007} that can come in many different forms.
+Quantitative forecasts are either point forecasts or probabilistic in nature and can make statements about continuous, discrete or binary outcome variables.
+Point forecasts only give one single number for the most likely outcome, but do not quantify the forecaster's uncertainty.
+This limits their usefulness, as a very certain forecast may, for example, warrant a very different course of actions than does a very uncertain one.
+Probabilistic forecasts, in contrast, by definition provide a full predictive distribution.
+This makes them much more useful in any applied setting, as we learn about the forecaster's uncertainty and their belief about all aspects of the underlying data-generating distribution (including e.g.
 skewness or the width of its tails).
@@
 
-\subsection{Forecast formats}
-
-Not only can we distinguish different forecast types, but also different formats in which predictions may be reported.
-This distinction is important, as different formats and prediction types require slightly different evaluation approaches.
-Forecasts can be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
+The predictive distribution of a probabilistic forecast can be represented in different ways with implications for the appropriate evaluation approach.
+Forecasts can wither be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
 Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
 They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
-For that reason, often quantiles are reported instead [citation FORECAST HUBS].
-As an alternative representation of a quantile forecast, one can also express predictions using central prediction intervals.
-In this case, the forecaster reports a range to denote the interval as well as a value for the lower and upper bound.
-A forecaster could also in principle state their forecasts in a binary way by defining an outcome and assigning a probability that the outcome will come true.
-This type of forecasting is common in many classification problems.
-If we think of the outcome of some number being larger or smaller than a certain value, we can immediately see the connection between binary predictions and the concept of a cumulative distribution functions (CDF).
-All of these forecast types can be evaluated using \pkg{scoringutils}.
-The general forecasting paradigm [GNEITING] that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+For that reason, often quantiles or central prediction intervals are reported instead [citation FORECAST HUBS].
+For binary prediction targets, common in many classification problems, a probabilistic forecasts is represented by the probability that an outcome will come true.
+DO I NEED TO TALK ABOUT MULTINOMIAL CLASSIFICATION PROBLEMS?
+Figure \ref{fig:overview-forecast-types} summarises the different forecast types and formats.
+The general forecasting paradigm [GNEITING], however, that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+
+\begin{figure}[h]
+\centering
+\includegraphics{plots/overview-forecast-types.png}
+\caption{\label{fig:overview-forecast-types} Overview of forecast types}
+\end{figure}
 
@@
 
-\input{tables/scores-table.tex}
+% \input{tables/scores-table.tex}
 

### 2021-03-06 tables/scores-table.Rnw
move scores-table file to correct folder

@@
-% !TeX root = ../article.Rnw
-\begin{table}[]
-\centering
-\begin{tabular}{@{}llll@{}}
-\toprule
-Metric                                                           & Formula                           & Application and Explanation                                                                                                                                                                                                                                                                                                             & Caveats                                                                                                                                                                                                     \\ \midrule
-\multicolumn{1}{|l|}{(weighted) interval score (wis)}            & \multicolumn{1}{l|}{$\alpha$}     & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}proper scoring rule for quantile forecasts.
-\\ converges to crps for increasing number of interval.
-\\ The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.\end{tabular}}                                                              & \begin{tabular}[c]{@{}l@{}}The wis is based on measures of absolute error.
-\\ When averaging across multiple targets, it will therefore be dominated\\ by targets with higher absolute values.\end{tabular} \\ \cmidrule(r){1-3}
-\multicolumn{1}{|l|}{continuous ranked probability score (crps)} & \multicolumn{1}{l|}{4.892523e+02} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}proper scoring rule for continuous and integer forecasts.
-\\ Implemented for sample-based forecasts.\end{tabular}}                                                                                                                                                                       & Also sensitive to the absolute value of the predicted quantity                                                                                                                                              \\ \cmidrule(r){1-3}
-\multicolumn{1}{|l|}{Dawid-Sebastiani score (dss)}               & \multicolumn{1}{l|}{2.257546e+02} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}proper scoring rule for continuous and integer forecasts.
-\\ The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution.
-\\ If in doubt, we would recommend the crps, but the difference should not be large in practice.\end{tabular}} &                                                                                                                                                                                                             \\ \cmidrule(r){1-3}
-log score                                                        &                                   & The log score is a proper scoring rule best suited for continuous forecasts.
-& Sensitive to outliers.
-Individual log score contributions can go to infinity                                                                                                                                \\
-Brier score                                                      &                                   & Proper scoring rule for binary forecasts                                                                                                                                                                                                                                                                                                &                                                                                                                                                                                                             \\
-interval coverage                                                &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
-quantile coverage                                                &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
-Coverage deviation                                               &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
-Bias                                                             &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\ \cmidrule(r){1-3}
-Mean score ratio                                                 &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
-Relative skill                                                   &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
-Absolute error of the median                                     &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\
-Mean squared error                                               &                                   &                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                             \\ \bottomrule
-\end{tabular}
-\end{table}
+\documentclass{article}
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{float}
+\usepackage{longtable}
+\begin{document}
+
+<<table, echo=FALSE, results=tex>>=
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+library("MASS")
+
+crps <- list(
+  `Metric` = "CRPS (Continuous) ranked probability score", 
+  `Explanation` = r"(The crps is a proper scoring rule that generalises the absolute error to probabilistic forecasts.
+It measures the 'distance' of the predictive distribution to the observed data-generating distribution.
+The CRPS is given as
+  $$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
+  where y is the true observed value and F the CDF of predictive distribution. Often An alternative representation is used:
+  $$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
+In this represenation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
+  For integer-valued forecasts, the RPS is given as 
+  $$ \text{RPS}(F, y) = \sum_{x = 0}^\infty (F(x) - 1(x \geq y))^2. $$
+
+\textbf{Usage and caveats} Smaller values are better.
+The crps is a good choice for most practical purposes that involve decision making, as it takes the entire predictive distribution into account.
+If two forecasters assign the same probability to the true event $y$, then the forecaster who assigned high probability to events far away from $y$ will still get a worse score.
+The crps (in contrast to the log score) can at times be quite lenient towards extreme mispredictions.
+Also, due to it's similarity to the absolute error, the level of scores depend a lot on the absolute value of what is predicted, which makes it hard to compare scores of forecasts for quantities that are orders of magnitude apart.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+log_score <- list(
+  `Metric` = "Log score", 
+  `Explanation` = r"(The Log score is a proper scoring rule that is simply compuated as the log of the predictive density evaluated at the true observed value. It is given as 
+  $$ \text{log score} = \log f(y), $$ 
+  where $f$ is the predictive density function and y is the true value. For integer-valued forecasts, the log score can be computed as 
+  $$ \text{log score} = \log p_y, $$
+  where $p_y$ is the probability assigned to outcome p by the forecast F.
+
+\textbf{Usage and caveats}: Larger values are better, but sometimes the sign is reversed.
+The log score is ensitive to outliers, as individual negative log score contributions quickly can become very large if the event falls in the tails of the predictive distribution, where $f(y)$ (or $p_y$) is close to zero.
+Whether or not that is desirable depends ont the application.
+In scoringutils, the log score cannot be used for integer-valued forecasts, as the implementation requires a predictive density.
+In contrast to the crps, the log score is a local scoring rule: it's value only depends only on the probability that was assigned to the actual outcome.
+This property is desirable for inferential purposes, for example in a Bayesian context (@winklerScoringRulesEvaluation1996).
+In settings where forecasts inform decision making, it may be more appropriate to score forecasts based on the entire predictive distribution.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+wis <- list(
+  Metric = "WIS (Weighted) interval score", 
+  `Explanation` = r"(The (weighted) interval score is a proper scoring rule for quantile forecasts that converges to the crps for an increasing number of intervals.
+The score can be decomposed into a sharpness (uncertainty) component and penalties for over- and underprediction.
+For a single interval, the score is computed as
+  $$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
+  where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
+the lower and upper bound of a single prediction interval.
+For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum,
+  $$WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),$$ 
+  where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+
+\textbf{Usage and caveats}:
+  Smaller scores are better.
+Applicable to all quantile forecasts, takes the entire predictive distribution into account.
+Just as the crps, the wis is based on measures of absolute error.
+When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.
+The decomposition into sharpness, over- and underprediction make it easy to interpret scores and use them for model improvement.
+)",
+  `Caveats` = "The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.",
+  `References` = ""
+)
+
+quantile_score <- "yet to come"
+
+dss <- list(
+  `Metric` = "DSS Dawid-Sebastiani score", 
+  `Explanation` = r"(The Dawid-Sebastiani-Score is a proper scoring rule proposed by Gneiting and Raftery in [@gneitingStrictlyProperScoring2007] that only relies on the first moments of the predictive distribution and is therefore easy to compute.
+It is given as
+
+$$\text{dss}(F, y) = \left( \frac{y - \mu}{\sigma} \right)^2 + 2 \cdot \log \sigma,$$
+where $F$ is the predictive distribution with mean $\mu$ and standard deviation $\sigma$ and $y$ is the true observed value.
+
+textbf{Usage and caveats} The dss is applicable to continuous and integer forecasts and easy to compute.
+Apart from the ease of computation we see little advantage in using it.
+MAYBE SCRATCH IT ALTOGETHER FROM THE PAPER?)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+brier_score <- list(
+  `Metric` = "Brier score", 
+  `Explanation` = r"(Proper scoring rule for binary forecasts. The Brier score is computed as 
+$$\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (f_n - y_n), $$
+where $f_n$, with $n = 1, \dots, N$ are the predicted probablities that the corresponding events, $y_n \in (0, 1)$ will be equal to one.)
+
+\textbf{Usage}: Applicable to all binary forecasts.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+interval_coverage <- list(
+  `Metric` = "Interval coverage", 
+  `Explanation` = r"(Interval coverage measures the proportion of observed values that fall in a given prediction interval range.
+Interval coverage for a single prediction interval range can be calculated as $$IC_{\alpha} = \text{nominal coverage} - \text{empirical coverage},$$
+where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by all $1 - \alpha$ prediction intervals.
+
+To summarise interval coverage over different over multiple interval ranges, we can compute coverage deviation defined as the mean interval coverage over all $K$ interval ranges $\alpha_k$ with $k = 1, \dots, K$:
+$$\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$$
+
+\textbf{Usage}: Interval coverage for a set of chosen intervals, (e.g.
+50\% and 90\%) gives a good indication of marginal calibration and is easy to interpret.
+Reporting coverage deviation has the advantage of summarising calibration in a single number, but loses some of the nuance.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+quantile_coverage <- list(
+  `Metric` = "Quantile coverage", 
+  `Explanation` = r"(Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level.
+
+\textbf{Usage}: Quantile coverage is similar to interval coverage, but conveys more information.
+For example, it allows us to look at the 5\% and 95\% quantile separately, instead of jointly at the 90\% prediction interval).
+This helps to diagnose whether it is the upper or lower end of a prediction interval that is causing problems.
+Plots of quantile coverage are conceptually very similar to PIT histograms.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+sharpness <- list(
+  `Metric` = "Sharpness", 
+  `Explanation` = r"(Sharpness is the ability to produce narrow forecasts and is a feature of the forecasts only and does not depend on the observations.
+Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
+
+As suggested by @funkAssessingPerformanceRealtime2019, we measure sharpness for continuous and integer forecasts represented by predictive samples as the normalised median absolute deviation about the median (MADN) ), i.e.
+$$ S(F) = \frac{1}{0.675} \cdot \text{median}(|x - \text{median(x)}|), $$ 
+where $x$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant.
+If the predictive distribution $F$ is the CDF of a normal distribution, then sharpness will equal the standard deviation of $F$.
+
+For quantile forecasts we can directly use the sharpness component of the weighted interval score. Sharpness is then simply the weighted mean of the widths of the central prediction intervals.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+bias <- list(
+  `Metric` = "Bias", 
+  `Explanation` = r"(Bias is a measure of the tendency of a forecaster to over- or underpredict. For continuous forecasts, bias is given as 
+$$B(F, y) = 1 - 2 \cdot (F (y)), $$ 
+where $F$ is the CDF of the predictive distribution and $y$ is the observed value.
+
+For integer-valued forecasts, bias can be calculated as 
+$$B(P, y) = 1 - (P(y) + P(y + 1)), $$ 
+where $P(y)$ is the cumulative probability assigned to all outcomes smaller or equal to $y$.
+
+For quantile forecasts, Bias can be calculated as the maximum percentile rank for which the prediction is smaller than $y$, if the true value is smaller than the median of the predictive distribution.
+If the true value is above the median of the predictive distribution, then bias is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
+If the true value is exactly the median, bias is zero.
+For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
+
+\textbf{Usage}:
+In contrast to the over- and underprediction penalties of the interval score it is bound between 0 and 1 and represents the tendency of forecasts to be biased rather than the absolute amount of over- and underprediction.
+It is therefore a more robust measurement, but harder to interpet.
+It largely depends on the application whether one is more interested in the tendency to be biased or in the absolute value of over- and underpredictions.)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+pit <- list(
+  `Metric` = "Probability integral transform (PIT)",
+  `Explanation` = r"(The probability integral transform (PIT) @dawidPresentPositionPotential1984 represents a succinct way to visualise deviations between the predictive distribution $F$ and the true data-generating distribution $G$.
+The idea is to transform the observed values such that agreement between forecasts and data can then be examined by observing whether or not the transformed values follow a uniform distribution.
+The PIT is given by
+$$u = F (y),$$
+where $u$ is the transformed variable and $F(y)$ is the predictive distribution $F$ evaluated at the true observed value $y$.
+If $F = G$, then $u$ follows a uniform distribution (for a proof see e.g.
+@angusProbabilityIntegralTransform1994).
+
+For integer outcomes, the PIT is no longer uniform even when forecasts are ideal. Instead, a randomised PIT (@funkAssessingPerformanceRealtime2019) can be used: 
+$$u = P(y) + v \cdot (P(y) - P(y - 1) ),$$
+where $y$ is again the observed value $P()$ is the cumulative probability assigned to all values smaller or equal to $y$ (where $P(-1) = 0$ by definition, and $v$ is a standard uniform variable independent of $y$.
+If $P$ is equal to the true data-generating distribution function, then $u$ is standard uniform.
+also propose a non-randomised version of the PIT for count data that could be used alternatively.
+
+\textbf{Usage}: 
+One can plot a histogram of $u$ values to look for deviations from uniformity.
+U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
+Biased predictions will usually result in a triangle-shaped histogram.
+One can also test for deviations from normality, using for example an Anderson-Darling test.
+This, however, proves to be overly strict in practice and even slight deviations from perfect calibration are punished in a way that makes it very hard to compare models at all.
+I HAVE MADE A SIMULATION TO TEST THAT IN MY MASTER THESIS AND COULD ADD IT TO THE PAPER?)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+mean_score_ratio <- list(
+  `Metric` = "Mean score ratio", 
+  `Explanation` = r"(The mean score ratio is used to compare two models on the overlapping set of forecast targets for which both models have made a prediction.
+The mean score ratio is calculated as the mean score achieved by the first model over the mean score achieved by the second model.
+More precisely, for two models $i, j$, we determine the set of overlapping forecasts, denoted by $\mathcal{A}_{ij}$ and compute the mean score ratio $\theta_{ij}$ as
+$$\theta_{ij} =\frac{\text{mean score model } i \text{ on } \mathcal{A}_{ij}}{\text{mean score model } j \text{ on } \mathcal{A}_{ij}}.$$
+The mean score ratio can in principle be computed for any arbitrary score.
+
+\textbf{Usage}:
+Mean scores ratios are usually calculated in the context of pairwise comparisons, where a set of models is compared by looking at mean score ratios of all possible parings.
+Whether smaller or larger values are better depends on the orientation of the original score used)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+relative_skill <- list(
+  `Metric` = "Relative skill", 
+  `Explanation` = r"(Relative skill scores can be used to obtain a ranking of models based on pairwise comparisons between all models.
+To compute the relative skill $\theta_i$ of model $i$, we take the geometric mean of all [mean score ratios](mean%20score%20ratio.md) that involve model $i$, i.e.
+$$ \theta_{i} = \left(\prod_{m = 1}^M \theta_{im}\right)^{1/M}, $$
+where M is the number of models.
+
+\textbf{Usage and caveats}: 
+Relative skill is a helpful way to obtain a model ranking. Whether smaller or larger values are better depends on the orientation of the original score used. 
+It is in principle relatively robust against biases that arise when models only forecast some of the available targets and is a reasonable way to handle missing forecasts.
+One possible precautionary measure to reduces issues with missing forecasts is to only compare models that have forecasted at least half of all possible targets (this ensures that there is always an overlap between models).
+If there is no overlap between models, the relative skill implicitly estimates how a model would have forecasted on those missing targets.
+)",
+  `Caveats` = "",
+  `References` = ""
+)
+
+data <- rbind(as.data.frame(crps), 
+              as.data.frame(log_score), 
+              as.data.frame(wis), 
+              as.data.frame(dss), 
+              as.data.frame(brier_score), 
+              as.data.frame(interval_coverage),
+              as.data.frame(quantile_coverage), 
+              as.data.frame(pit),
+              as.data.frame(sharpness),
+              as.data.frame(bias), 
+              as.data.frame(mean_score_ratio), 
+              as.data.frame(relative_skill))
+
+library(kableExtra)
+library(magrittr)
+library(knitr)
+data[, 1:2] %>%
+  kableExtra::kbl(format = "latex", booktabs = TRUE, 
+                  escape = FALSE,
+                  longtable = TRUE,
+                  linesep = c('\\addlinespace \\addlinespace')) %>%
+  kableExtra::column_spec(1, width = "2.5cm") %>%
+  kableExtra::column_spec(2, width = "16.5cm") %>%
+  # kableExtra::column_spec(3, width = "4.5cm") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "repeat_header")) %>%
+  kableExtra::landscape()
+
+@
+
+\end{document}

### 2021-03-06 scores-table.R

@@ -1,14 +1,0 @@
-scores_table <- data.frame(
-  
-)
-
-wis <- list(
-  Metric = "WIS (Weighted) interval score", 
-  `Formula` = r"(For a single interval, the score is computed as $IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u)$, where 1() is the indicator function, y is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantile of $F$, i.e. the lower and upper bound of a single prediction interval. For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, $WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y))$. $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$)", 
-  `Application and Explanation` = "proper scoring rule for quantile forecasts. converges to crps for increasing number of interval. The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.", 
-  `Caveats` = "The wis is based on measures of absolute error. When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values."
-)
-
-library(magrittr)
-as.data.frame(wis) %>%
-  kableExtra::kable()

### 2021-03-30 article.Rnw
update scores sections of the paper

@@
-\documentclass[article]{jss}
+\documentclass[article,shortnames]{jss}
 
@@
 
+%% packages for the table
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{multirow}
+
+\usepackage{xcolor}
+\usepackage{colortbl}
+
 %% new custom commands
@@
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+
+# knitr::knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed = TRUE)})
+
 library("MASS")
+library(patchwork)
+library(ggplot2)
+library(kableExtra)
 
@@
 Its role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
-Even with a plethora of tools and models available to make forecasts, evaluating the predictions is not trivial.
+Evaluating forecasts, however, is not trivial in practice.
 Even though scoring methods such as proper scoring rules have been widely studied in the past, there is a lack of software implementation that allows a forecaster to conveniently evaluate their forecasts.
@@
 It gives the user access to a wide range of scoring metrics for various types of forecasts as well as a variety of ways to visualise the evaluation.
-We give an overview of the rationale behind forecast evaluation and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
+We give an overview of the evaluation process and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
 }
@@
 For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
-\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{FUNK}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
+\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{funkAssessingPerformanceRealtime2019}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
 Yet even with this rich body of research available, implementing a complete forecast evaluation in \proglang{R} is not trivial.
@@
 The predictive distribution of a probabilistic forecast can be represented in different ways with implications for the appropriate evaluation approach.
-Forecasts can wither be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
-Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
-They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
+For most forecasting problems, predictive distributions are not readily available in a closed form (and the \pkg{scoringutils} package therefore does not support scoring them directly).
+Instead, predictive distributions are often represented by a set of quantiles or predictive samples.
+Predictive samples require a lot of storage space and also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
 For that reason, often quantiles or central prediction intervals are reported instead [citation FORECAST HUBS].
@@
 DO I NEED TO TALK ABOUT MULTINOMIAL CLASSIFICATION PROBLEMS?
-Figure \ref{fig:overview-forecast-types} summarises the different forecast types and formats.
-The general forecasting paradigm [GNEITING], however, that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+Table \ref{tab:forecast-types} summarises the different forecast types and formats.
+The general forecasting paradigm \cite{gneitingProbabilisticForecastsCalibration2007}, however, that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+
+\begin{table}[h!]
+\centering
+<<forecast-types-table, echo=FALSE, results=tex>>=
+
+data <- readRDS("tables/forecast-types.rds")
+
+data %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kbl(format = "latex", 
+                  booktabs = TRUE, 
+                  escape = FALSE,
+                  longtable = TRUE,
+                  linesep = c('\\addlinespace')) %>%
+  kableExtra::column_spec(1, width = "3.5cm") %>%
+  kableExtra::column_spec(2, width = "2.5cm") %>%
+  kableExtra::column_spec(3, width = "4.5cm") %>%
+  kableExtra::row_spec(0, bold = TRUE) %>%
+  kableExtra::kable_styling() %>% 
+  kableExtra::collapse_rows(valign = "middle")
+
+@
+\caption{\label{tab:forecast-types} Ssummary of the different forceast types and forecast formats}
+\end{table}
+
+\subsection{The forecasting paradigm}
+
+Any forecaster should aim to minimise the difference between the predictive distribution $F$ and the unknown true data-generating distribution $G$ \citep{gneitingProbabilisticForecastsCalibration2007}.
+For an ideal forecast, we therefore have
+%
+$$ F = G. $$
+%
+As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
+\cite{gneitingProbabilisticForecastsCalibration2007} instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
+Calibration refers to the statistical consistency between the predictive distribution and the observations.
+A well calibrated forecast does not systematically deviate from the observed values.
+For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to \cite{gneitingProbabilisticForecastsCalibration2007}.
+Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
+how precise the forecasts are.
+The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
+A model that made very precise forecasts would be at best useless if the forecasts were wrong most of the time.
+On the other hand, a model may be well calibrated, but not sharp enough to be useful.
+Take a weather forecast that would assign 30 percent rain probability for every single day.
+It may be well calibrated over the course of a year (it would be marginally calibrated according to \cite{gneitingProbabilisticForecastsCalibration2007}), but we would of course prefer a more precise forecast.
+Figure \ref{fig:forecast-paradigm} illustrates the concepts of calibration and sharpness.
 
@@
 \centering
-\includegraphics{plots/overview-forecast-types.png}
-\caption{\label{fig:overview-forecast-types} Overview of forecast types}
+\includegraphics{plots/forecast-paradigm.png}
+\caption{\label{fig:forecast-paradigm} Schematic illustration of calibration and sharpness. True value are represented in red, the predictive distribution is shown in black}
 \end{figure}
-
-\subsection{The forecasting paradigm}
-
-Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007].
-For an ideal forecast, we therefore have
-%
-$$ P_t = F_t, $$
-%
-where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution.
-As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
-[CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
-Calibration refers to the statistical consistency between the predictive distribution and the observations.
-A well calibrated forecast does not systematically differ deviate from the observed values.
-For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007].
-Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
-how precise the forecasts are.
-The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
-Take for example the task of predicting rainfall in a city like London.
-A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time.
-On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
-%Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
-DO I WANT THAT FIGURE?
 
@@
 
-Table:
+Some of the metrics in \pkg{scoringuitls} focus only on sharpness or on calibration.
+Others, called proper scoring rules, combine both aspects into a single number.
+The former can be helpful to learn about specific model aspects and improve them, the latter are especially useful to assess and rank predictive performance of a forecaster.
+The following gives an introduction to how these metrics can be used to evaluate forecasts.
+Table \ref{tab:table-summary-scores} shows an overview of the metrics implemented in \pkg{scoringutils}.
+Table \ref{tab:score-table} in the Appendix gives a thorough explanation of all the metrics.
 
-% \input{tables/scores-table.tex}
+\begin{table}[h!]
+\centering
+<<summary-table, echo=FALSE, results=tex>>=
 
-The metrics included in the \pkg{scoringuitls} package can be divided into two categories.
-The first consists of metrics that aim to capture different aspects of sharpness and calibration, the second comprises various proper scoring rules.
-We begin with the latter.
+data <- readRDS("tables/scores-table-summary.rds")
+
+library(kableExtra)
+library(magrittr)
+library(knitr)
+data[, 1:4] %>%
+  kableExtra::kbl(format = "latex", booktabs = TRUE, 
+                  escape = FALSE,
+                  longtable = TRUE,
+                  linesep = c('\\addlinespace')) %>%
+  kableExtra::column_spec(1, width = "2.5cm") %>%
+  kableExtra::column_spec(2, width = "2cm") %>%
+  kableExtra::column_spec(3, width = "4.5cm") %>%
+  kableExtra::column_spec(4, width = "4cm") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))
+
+@
+\caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
+\end{table}
 
@@
 
-Proper scoring rules [CITATION] jointly assess sharpness and calibration and assign a single numeric value to a forecast.
+Proper scoring rules \citep{gneitingStrictlyProperScoring2007} jointly assess sharpness and calibration and assign a single numeric value to a forecast.
 A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
-This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state state their true best belief.
-The following scoring rules are implemented in \pkg{scoringutils}: The (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], the Dawid-Sebastiani-score (dss) [CITATION], the (weighted) interval score (wis), and the Brier score (bs).
-The first three proper scoring rules are implemented as wrappers around functions from the \pkg{scoringRules} package.
-They are suitable for any sample-based prediction format.
-The Log Score, however, is not applied if integer-valued forecasts are supplied, as the implementation in \pkg{scoringRules} requires a kernel density estimation that may be inappropriate for integer values.
-The interval score is appropriate for forecasts in a quantile formats.
-For every central prediction interval, the interval score is computed as the sum of separate penalties for overprediction, underprediction and sharpness (i.e.
-width) of the forecast.
-By default, the weighted interval score is returned.
-The overall score for a forecast is obtained as the mean of all scores for the individual central prediction intervals, weighted according to the range of the prediction interval.
-This ensures that the weighted interval score converges to the continuous ranked probability score for an increasing number of available prediction intervals.
-DO I WANT FORMULAS FOR ALL THE SCORES? The Brier score is used for binary predictions.
+This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state their best estimate.
 
-TALK ABOUT CAVEATS (LATER?) E.G. THAT THE MEAN WIS IS INFLUENCED BY THE ABSOLUTE VALUE OF THE TARGET?
+For sample-based forecasts, the (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], and the Dawid-Sebastiani-score (dss) [CITATION] are available in \pkg{scoringutils}.
+These are implemented as wrappers around functions from the \pkg{scoringRules} package (which also has closed-form versions of the scoring rules available).
+They are in principle applicable to continuous as well as integer forecasts.
+The \pkg{scoringRules} implementation of the log score, however, requires a kernel density estimation that may be inappropriate for integer values (see also Table \ref{tab:table-summary-scores}) and is therefore not available for discrete predictions in \pkg{scoringutils}.
+For forecasts in an interval or quantile format, the weighted interval score (wis) \citep{bracherEvaluatingEpidemicForecasts2020a} is available.
+For an increasing number of equally-spaced prediction intervals, the wis converges to the crps and therefore has very similar properties.
+Binary forecasts can be scored using the Brier score (bs) [CITATION].
+
+%sample-based proper scoring rules\usepackage{multirow}
+When scoring forecasts in a sample-based format, the choice is usually between the log score and the crps.
+The dss is much less commonly used.
+It is easier to compute, but apart from that does not have immediate advantages over the former two.
+Crps and log score differ in two important aspects: the first is sensitivity to distance \cite{winklerScoringRulesEvaluation1996}, the second is how harshly far-off predictions are punished.
+
+The crps is a so-called global scoring rule, which means that the entire predictive distribution is taken into account when scoring a single forecast.
+The log score, on the other hand is local.
+The resulting score does not depend on the overall distance between the observed value and the distribution, but only on the probality density assigend to the actual outcome.
+Imagine two forecasters, A and B, who forecast the number of points scored in a basketball game.
+If both forecasters assigned the same probability to the true outcome (100 points), but A assigned higher probability to extreme outcomes far away from the actually observed outcome, then A will receive a worse score than B.
+The log score, in contrast, is a local scoring rule that only scores the probability assigned to the actual outcome and ignores the rest of the predictive distribution.
+Judged by the log score, A and B would receive exactly the same score.
+Sensitivity to distance (taking the entire predictive distribution into account) is arguably an advantage in most settings that involve decision making.
+Forecaster A's prediction that assigns high probability to results far away from the observed value is arguably less useful than B's forecast that assigns higher probability to values closer to it (the probability assigned to the actual outcome being equal for both forecasts).
+The log score is only implicitly sensitive to distance if we assume that values close to the observed value are actually more likely to occur.
+It may, however, be more appropriate for inferential purposes (see \cite{winklerScoringRulesEvaluation1996}).
+
+% stable vs. non-stable
+A second important difference is how forecasts are treated that deviate strongly from the observered outcome.
+The crps can be thought of as a generalisation of the absolute error to a predictive distribution.
+It therefore scales linearly with the distance between forecast distribution and true value.
+The log score, however, is the log of the predictive density evaluated at the observed value.
+It can therfore quickly go to negative infinity if the probability assigned to the observed outcome is close to zero.
+The crps is therefore considered more stable than the log score.
+The behaviour of the dss is in between the two.
+Whether or not harsh punishment of bad predictions is desirable or not depends of course on the setting.
+\cite{bracherEvaluatingEpidemicForecasts2020a} exemplify that in practice there may indeed be substantial differences between how the crps and log score judge the same forecast.
+
+% properties of the WIS + decomposition
+For quantile forecasts the wis is the best available option and can be understood as an approximation to the crps.
+One additional benefit of the wis is that it can easily be decomposed into three additive components: an uncertainty penalty (sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval).
+This can be very helpful in diagnosing model problems.
+It may therefore even be useful to convert samples into quantiles and use the wis instead of the crps to make use of this decomposition for the purpose of model diagnostics.
+% fact that scores scale with the overall level of the prediction
+Both crps and wis, as generalisations of the absolute value, suffer from the fact that overall scores depend on the order of magnitude of the quantity we try to forecast.
+It can therefore be difficult to compare forecasts for very different targets.
+One possibility to address this is the use of pairwise comparisons introduced later.
+In general, this is much less of a problem for the log score and the dss.
+IS IT ONE AT ALL?
 
@@
 Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g.
-@funkAssessingPerformanceRealtime2019; @gneitingProbabilisticForecastsCalibration2007; @gneitingStrictlyProperScoring2007).
+\cite{funkAssessingPerformanceRealtime2019}; \cite{gneitingProbabilisticForecastsCalibration2007}; \cite{gneitingStrictlyProperScoring2007}).
 Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
 
+% Bias
 Bias, i.e.
-systematic over- or underprediction, is a very common form of miscalibration.
-For continuous forecasts, assessing whether a predictive distribution has a tendency to over- or underpredict can be very easily achieved by simply evaluating the predictive distribution at the true observed value.
-This metric is a generalisation of the integer-valued one @funkAssessingPerformanceRealtime2019 have proposed.
-It is also closely related to the probability integral transform (PIT) discussed later in this chapter.
-To improve the interpretability of the score we can transform it to a value between -1 (under-prediction) and 1 (over-prediction).
-Consequently, we measure bias as
-$$B_t (P_t, x_t) = 1 - 2 \cdot (P_t (x_t)),$$
-where $P_t$ is the cumulative distribution function of the predictive distribution for the true value $x_t$.
-When using predictive samples, $P_t (x_t)$ is simply the fraction of predictive samples for $x_t$ that are smaller than the true observed $x_t$.
+systematic over- or underprediction, is a very common form of miscalibration which therefore deserves separate attention.
+The bias metric (with slightly different versions for the various forecast types and formats) captures a general tendency to over- and underpredict that is bound to be between minus one (underpredicton) and one (overprediction), where zero is ideal.
+It is derived by looking at how much of the probability mass of the predictive distribution is below or above the true observed value.
+For quantile forecasts we have second alternative approach available to assess over- and underprediction - by simply looking at the corresponding components of the weighted interval score.
+What is different between the over- and underprediction components and bias as described above is its sensitivity to outliers.
+The former are derived from absolute differences, while the latter is bound and rather captures a general tendency to be biased.
 
-For integer valued forecasts, we use the metric proposed by @funkAssessingPerformanceRealtime2019: 
-$$B_t (P_t, x_t) = 1 - (P_t (x_t) + P_t (x_t + 1)).$$
-Bias can again assume values between -1 (under-prediction) and 1 (over-prediction) and is 0 ideally.
-
-For quantile forecasts, we propose the following metric to assess bias: 
-\begin{align*}
-  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t \}) \mathbbm{1}( x_t \leq q_{t, 0.5}) \\
-  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbbm{1}( x_t \geq q_{t, 0.5}),
-\end{align*}
-where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$.
-They represent our belief about what the true value $x_t$ will be.
-For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$.
-$\mathbbm{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
-In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution.
-If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
-If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
-For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
-For quantile forecasts, an alternative approach is to look at the over- and underprediction components of the weighted interval score.
-These however, capture the bias in absolute terms, while the above proposed metric captures a tendency to over- and underpredict that is less sensitive to outliers.
-
-I ASSUME ALL OF THIS IS A BIT TOO LONG?
-
-Another way to look at calibration[precisely: probabilistic calibration in @gneitingProbabilisticForecastsCalibration2007] is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
-This is most easily understood in the context of quantile forecasts, but can in principle be transferred to continuous and integer forecasts as well.
-
+% interval coverage and quantile coverage
+Another way to look at calibration (precisely: probabilistic calibration in \cite{gneitingProbabilisticForecastsCalibration2007}) is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
+This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and integer forecasts as well.
+%
 To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution.
@@
 We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
+%
+To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
+This allows to diagnose issues in the lower and upper tails of the prediction intervals separately.
+A similar way to visualise the same information is a PIT histogram.
+In order to conveniently assess deviations between the predictive distribution and the true data-generating distribution we can transform the observed values using the probability integral transformation (PIT) \citep{dawidPresentPositionPotential1984} (see more details in Table \ref{tab:score-table}).
+If both distributions are equal, the transformed values will follow a uniform distribution.
+A histogram of the transformed values can help to diagnose systematic differences between the predictions and the observed values.
+Figure \ref{fig:calibration-plots} exemplifies the characteristic shape of certain systematic deviations of the predicitve distribution from the true data-generating distribution.
+In the PIT histograms, bias leads to a triangular shape, overdispersion results in a hump shaped form and underdispersion in a U-shape.
+ADD INTERPRETATION FOR QUANTILE AND INTERVAL COVERAGE PLOTS HERE.
 
-To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
+\begin{figure}[h]
+\centering
+\includegraphics{plots/calibration-diagnostic-examples.png}
+\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
+\end{figure}
+
+<<example-calibration, echo=FALSE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
+n_truth = 1000
+n_samples = 100
+true_values <- rnorm(n_truth, 0, 1)
+predictions1 <- rnorm(n_truth * n_samples)
+predictions2 <- rnorm(n_truth * n_samples, mean = 0.5)
+predictions3 <- rnorm(n_truth * n_samples, sd = 1.4)
+predictions4 <- rnorm(n_truth * n_samples, sd =  0.7)
+predictions5 <- rnorm(n_truth * n_samples, mean = 0.2, sd =  0.7)
+
+df <- data.table::data.table(true_value = rep(true_values, each = n_samples), 
+                             id = rep(1:n_truth, each = n_samples),
+                             prediction = c(predictions1, predictions2, 
+                                            predictions3, predictions4, 
+                                            predictions5), 
+                             sample = 1:n_samples,
+                             model = rep(c("Normal(0,1)", "Normal(0.5, 1)", 
+                                       "Normal(0, 1.4)", "Normal(0, 0.7)", 
+                                       "Normal(0.2, 0.7)"), each = n_truth * n_samples))
+
+res <- eval_forecasts(df, 
+                      by = c("model", "id"), 
+                      summarise_by = c("model", "range"), 
+                      pit_plots = TRUE)
+
+# plot with observations
+true_value_plot <- ggplot2::ggplot(data = data.frame(x = true_values),
+                ggplot2::aes(x = x)) +
+  ggplot2::geom_histogram(ggplot2::aes(y = ..density..),
+                          fill = "grey",
+                          colour = "dark grey") +
+  cowplot::theme_cowplot() +
+  ggplot2::labs(x = "True values",
+                y = "Density") +
+  ggplot2::theme(legend.position = "bottom")
+
+# plot with standard normal distribution
+standard_normal <- true_value_plot +
+  ggplot2::geom_function(fun = dnorm, colour = "black") +
+  ggplot2::ggtitle("Normal(0, 1)")
+
+# plot with shifted mean
+shifted_mean <- true_value_plot +
+  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(mean = 0.5)) +
+  ggplot2::ggtitle("Normal(0.5, 1)")
+
+# plot with overdispersion
+overdispersion <- true_value_plot +
+  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 1.4)) +
+  ggplot2::ggtitle("Normal(0, 1.4)")
+
+# plot with underdispersion
+underdispersion <- true_value_plot +
+  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 0.7)) +
+  ggplot2::ggtitle("Normal(0, 0.7)")
+
+(standard_normal | shifted_mean | overdispersion | underdispersion) /
+  (res$pit_plots$`Normal(0,1)` | res$pit_plots$`Normal(0.5, 1)` | res$pit_plots$`Normal(0, 1.4)` | res$pit_plots$`Normal(0, 0.7)`)
+
+ggsave("plots/calibration-diagnostic-examples.png", width = 12.5, height = 6)
+@
 
@@
 
+Sharpness is the ability to produce narrow forecasts.
+It does not depend on the actual observations and is a quality of the forecast only \ref{gneitingProbabilisticForecastsCalibration2007}.
+Sharpness is therefore only useful subject to calibration, as exemplified above in Figure \ref{fig:calibration-example}.
+We may be willing to trade off a little calibration for a lot more sharpness, but usually not much.
+For sample-based forecasts, \pkg{scoringutils} calculates sharpness as the normalised median absolute deviation about the median (MADN) [funkAssessingPerformanceRealtime2019] (for details see Table \ref{tab:score-overview}).
+For quantile forecasts, we take the sharpness component of the wis which corresponds to a weighted average of the individual interval widths.
+
+WHAT WOULD BE REALLY COOL IS TO DETERMINE A WAY TO FIND THE OPTIMAL SHARPNESS OF A FORECAST. AS IS, I FEEL THIS PARAGRAPH IS SOMEWHAT USELESS.
+
 \subsection{Pairwise comparisons}
 
-If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION].
-In turn, each pair of models is evaluated based on the targets that that both models have predicted.
-The mean score by one model is divided by the mean score of the other model to obtain a measure of relative performance.
-Relative skill scores [citation (not sure there is something published, preprint fron Estee Cramer and Nick Reich)] for every model can then be obtained by taking the geometric mean of all pairwise ratios calculated for each model (omitting comparisons where there is no overlapping set of forecasts).
-ADD FORMULAS.
-A skill score smaller than 1 indicates that a model is performing better than the average model.
-One can also scale the skill by a reference baseline.
+If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION CRAMER et al.].
+In turn, each pair of models is evaluated based on the targets that both models have predicted.
+The mean score by one model is divided by the mean score of the other model to obtain the mean score ratio (see Table \ref{tab:score-overview}, a measure of relative performance.
+To obtain an overall relative skill score for a model, we take the geomatric mean of all mean score ratios that involve that model  (omitting comparisons where there is no overlapping set of forecasts).
+This gives us an indicator of performance relative to all other models.
+The orientation depends on the score used.
+For the proper scoring rules described above, smaller is better and a relative skill score smaller than 1 indicates that a model is performing better than the average model.
+We can obtain a scaled relative skill score by dividing a model's relative skill by the relative skill of a baseline model.
 A scaled relative skill smaller than one then means that the model in question performed better than the baseline.
 
-By applying a permutation test or non-parametric test like the Wilcoxon rank sum test we can also obtain p-values that help determine whether or not models perform significantly differently.
-THERE IS AN ISSUE WITH CORRELATION: THE TEST WOULD TREAT ALL OBSERVATIONS AS INDEPENDENT, WHICH THEY ARENT'T, SO I WOULD ASSUME THAT IT IS TOO LIBERAL.
-CURRENTLY, NO ADJUSTMENTS ARE IMPLEMENTED, I'M NOT REALLY SURE WHETHER TO DISCUSS THIS HERE OR NOT.
+It is in principle possible to obtain p-values that help determine whether two models perform significantly differently.
+\pkg{scoringutils} allows to compute these using eitab:score-tablether the Wilcoxon rank sum test or a permutation test.
+In practice, this is slightly complicated by the fact that both tests assume independent observations.
+In reality, however, forecasts by a model may be correlated across time or another dimension (e.g.
+if a forecaster has a bad day, they will likely perform badly across different targets for a given forecast date).
+P-values may therefore be too quick to suggest significant differences where there aren't any.
+One way to mitigate this is to aggregate observations over a category where one suspects correlation.
+A test that is performed on aggregate scores will likely be more conservative.
 
@@
 \centering
-<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12, eval = FALSE>>=
+<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12, eval = TRUE>>=
 show_avail_forecasts(combined, 
@@
 As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
-% 
-<<scoring, eval = FALSE>>=
+% ion G
+<<scoring, eval = TRUE>>=
 scores <- eval_forecasts(combined, 
@@
 
-<<pairwise-comparison, eval = FALSE>>=
+<<pairwise-comparison, eval = TRUE>>=
 # unsummarised scores
@@
 \centering
-<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
+<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
 score_table(scores, y = "model", facet_formula = ~ value_desc)
@@
 \centering
-<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
+<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
 plot_pairwise_comparison(pairwise) +
@@
 \centering
-<<pairwise-plot-ratio, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
+<<pairwise-plot-ratio, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
 plot_pairwise_comparison(pairwise, type = "pval") +
@@
 \centering
-<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12, eval = FALSE>>=
+<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12, eval = TRUE>>=
 wis_components(scores, 
@@
 The interval coverage plot shows the proportion of all true values that fall within all the different prediction intervals.
-This gives a visual impression of probabilistic calibration [GNEITING].
+This gives a visual impression of probabilistic calibration \ref{gneitingProbabilisticForecastsCalibration2007}.
 Ideally, $x$ percent of true values should be covered by the $x$\%-prediction intervals, resulting in a 45° line.
@@
 \centering
-<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = FALSE>>=
+<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = TRUE>>=
 cov_scores <- eval_forecasts(combined, 
@@
 
-<<quantile-coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = FALSE>>=
+<<quantile-coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = TRUE>>=
 cov_scores <- eval_forecasts(combined, 
@@
 \centering
-<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15, eval = FALSE>>=
+<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15, eval = TRUE>>=
 scores <- eval_forecasts(combined, 
@@
 
-\bibliography{refs}
+\bibliography{references}
 
@@
 
+\begin{table}[h!]
+\centering
+<<scores-table, echo=FALSE, results=tex>>=
+
+data <- readRDS("tables/scores-table.rds")
+
+# data[, 1:4] %>%
+#   kableExtra::kbl(format = "latex", booktabs = TRUE, 
+#                   escape = FALSE,
+#                   longtable = TRUE,
+#                   linesep = c('\\addlinespace')) %>%
+#   kableExtra::column_spec(1, width = "2.5cm") %>%
+#   kableExtra::column_spec(2, width = "2cm") %>%
+#   kableExtra::column_spec(3, width = "4.5cm") %>%
+#   kableExtra::column_spec(4, width = "4cm") %>%
+#   kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))
+
+@
+\caption{\label{tab:score-table} Explanation of all the scores}
+\end{table}
+
 \end{appendix}
@@
 \end{document}
+
+%BIAS
+% For continuous forecasts, assessing whether a predictive distribution has a tendency to over- or underpredict can be very easily achieved by simply evaluating the predictive distribution at the true observed value.
+This metric is a generalisation of the integer-valued one @funkAssessingPerformanceRealtime2019 have proposed.
+It is also closely related to the probability integral transform (PIT) discussed later in this chapter.
+To improve the interpretability of the score we can transform it to a value between -1 (under-prediction) and 1 (over-prediction).
+Consequently, we measure bias as
+% $$B_t (P_t, x_t) = 1 - 2 \cdot (P_t (x_t)),$$
+% where $P_t$ is the cumulative distribution function of the predictive distribution for the true value $x_t$.
+When using predictive samples, $P_t (x_t)$ is simply the fraction of predictive samples for $x_t$ that are smaller than the true observed $x_t$.
+% 
+% For integer valued forecasts, we use the metric proposed by @funkAssessingPerformanceRealtime2019: 
+% $$B_t (P_t, x_t) = 1 - (P_t (x_t) + P_t (x_t + 1)).$$
+% Bias can again assume values between -1 (under-prediction) and 1 (over-prediction) and is 0 ideally. 
+% 
+% For quantile forecasts, we propose the following metric to assess bias: 
+% \begin{align*}
+%   B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t \}) \mathbbm{1}( x_t \leq q_{t, 0.5}) \\
+%   &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbbm{1}( x_t \geq q_{t, 0.5}),
+% \end{align*}
+% where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$.
+They represent our belief about what the true value $x_t$ will be.
+For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$.
+$\mathbbm{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
+In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution.
+If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
+If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
+For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.

### 2021-03-30 article.tex
update scores sections of the paper

@@
 \usepackage{rotating}
+
+%% packages for the table
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{multirow}
+
+\usepackage{xcolor}
+\usepackage{colortbl}
 
@@
 
-\subsection{Forecast formats}
+The predictive distribution of a probabilistic forecast can be represented in different ways with implications for the appropriate evaluation approach.
+For most forecasting problems, predictive distributions are not readily available in a closed form (and the \pkg{scoringutils} package therefore does not support scoring them directly).
+Instead, predictive distributions are often represented by a set of quantiles or predictive samples.
+Predictive samples require a lot of storage space and also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
+For that reason, often quantiles or central prediction intervals are reported instead [citation FORECAST HUBS].
+For binary prediction targets, common in many classification problems, a probabilistic forecasts is represented by the probability that an outcome will come true.
+DO I NEED TO TALK ABOUT MULTINOMIAL CLASSIFICATION PROBLEMS?
+Table \ref{tab:forecast-types} summarises the different forecast types and formats.
+The general forecasting paradigm \cite{gneitingProbabilisticForecastsCalibration2007}, however, that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
 
-Not only can we distinguish different forecast types, but also different formats in which predictions may be reported.
-This distinction is important, as different formats and prediction types require slightly different evaluation approaches.
-Forecasts can be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
-Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
-They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
-For that reason, often quantiles are reported instead [citation FORECAST HUBS].
-As an alternative representation of a quantile forecast, one can also express predictions using central prediction intervals.
-In this case, the forecaster reports a range to denote the interval as well as a value for the lower and upper bound.
-A forecaster could also in principle state their forecasts in a binary way by defining an outcome and assigning a probability that the outcome will come true.
-This type of forecasting is common in many classification problems.
-If we think of the outcome of some number being larger or smaller than a certain value, we can immediately see the connection between binary predictions and the concept of a cumulative distribution functions (CDF).
-All of these forecast types can be evaluated using \pkg{scoringutils}.
-The general forecasting paradigm [GNEITING] that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+\begin{table}[h!]
+\centering
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{4.5cm}}
+\toprule
+\textbf{Forecast type} & \textbf{Target type} & \textbf{Representation of the predictive distribution}\\
+\midrule
+Point forecast & continuous \newline  integer \newline  binary & one single number for the predicted outcome\\
+\cmidrule{1-3}\pagebreak[0]
+ & continuous \newline  integer & predictive samples \newline   quantiles \newline   closed analytical form\\
+\cmidrule{2-3}\nopagebreak
+\multirow{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
+\bottomrule
+\end{longtable}\caption{\label{tab:forecast-types} Ssummary of the different forceast types and forecast formats}
+\end{table}
 
@@
 
-Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007].
+Any forecaster should aim to minimise the difference between the predictive distribution $F$ and the unknown true data-generating distribution $G$ \citep{gneitingProbabilisticForecastsCalibration2007}.
 For an ideal forecast, we therefore have
 %
-$$ P_t = F_t, $$
+$$ F = G. $$
 %
-where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution.
 As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
-[CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
+\cite{gneitingProbabilisticForecastsCalibration2007} instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
 Calibration refers to the statistical consistency between the predictive distribution and the observations.
-A well calibrated forecast does not systematically differ deviate from the observed values.
-For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007].
+A well calibrated forecast does not systematically deviate from the observed values.
+For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to \cite{gneitingProbabilisticForecastsCalibration2007}.
 Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
@@
 The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
-Take for example the task of predicting rainfall in a city like London.
-A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time.
-On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
-%Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
-DO I WANT THAT FIGURE?
+A model that made very precise forecasts would be at best useless if the forecasts were wrong most of the time.
+On the other hand, a model may be well calibrated, but not sharp enough to be useful.
+Take a weather forecast that would assign 30 percent rain probability for every single day.
+It may be well calibrated over the course of a year (it would be marginally calibrated according to \cite{gneitingProbabilisticForecastsCalibration2007}), but we would of course prefer a more precise forecast.
+Figure \ref{fig:forecast-paradigm} illustrates the concepts of calibration and sharpness.
+
+\begin{figure}[h]
+\centering
+\includegraphics{plots/forecast-paradigm.png}
+\caption{\label{fig:forecast-paradigm} Schematic illustration of calibration and sharpness. True value are represented in red, the predictive distribution is shown in black}
+\end{figure}
 
@@
 
-Table:
+Some of the metrics in \pkg{scoringuitls} focus only on sharpness or on calibration.
+Others, called proper scoring rules, combine both aspects into a single number.
+The former can be helpful to learn about specific model aspects and improve them, the latter are especially useful to assess and rank predictive performance of a forecaster.
+The following gives an introduction to how these metrics can be used to evaluate forecasts.
+Table \ref{tab:table-summary-scores} shows an overview of the metrics implemented in \pkg{scoringutils}.
+Table \ref{tab:score-table} in the Appendix gives a thorough explanation of all the metrics.
 
-\input{tables/scores-table.tex}
+\begin{table}[h!]
+\centering
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{4cm}}
+\toprule
+Metric & Target types & Forecast formats & Properties\\
+\midrule
+\endfirsthead
+\multicolumn{4}{@{}l}{\textit{(continued)}}\\
+\toprule
+Metric & Target types & Forecast formats & Properties\\
+\midrule
+\endhead
 
-The metrics included in the \pkg{scoringuitls} package can be divided into two categories.
-The first consists of metrics that aim to capture different aspects of sharpness and calibration, the second comprises various proper scoring rules.
-We begin with the latter.
+\endfoot
+\bottomrule
+\endlastfoot
+\cellcolor{gray!6}{CRPS (Continuous) ranked probability score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
+\addlinespace
+Log score & continuous, (integer not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
+\addlinespace
+\cellcolor{gray!6}{WIS (Weighted) interval score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
+\addlinespace
+DSS Dawid-Sebastiani score & continuous, integer & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
+\addlinespace
+\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
+\addlinespace
+Interval coverage & continuous, integer & interval forecasts (needs matching quantiles) & measure for calibration\\
+\addlinespace
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
+\addlinespace
+Probability integral transform (PIT) & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
+\addlinespace
+\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
+\addlinespace
+Bias & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
+\addlinespace
+\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{compares performance of two models}\\
+\addlinespace
+Relative skill & depends on scored & depends on score & Ranks models based on pairwise comparisons\\*
+\end{longtable}\caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
+\end{table}
 
@@
 We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
+%
+To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
+This allows to diagnose issues in the lower and upper tails of the prediction intervals separately.
+A similar way to visualise the same information is a PIT histogram.
+In order to conveniently assess deviations between the predictive distribution and the true data-generating distribution we can transform the observed values using the probability integral transformation (PIT) \citep{dawidPresentPositionPotential1984} (see more details in Table \ref{tab:score-table}).
+If both distributions are equal, the transformed values will follow a uniform distribution.
+A histogram of the transformed values can help to diagnose systematic differences between the predictions and the observed values.
+Figure \ref{fig:calibration-plots} exemplifies the characteristic shape of certain systematic deviations of the predicitve distribution from the true data-generating distribution.
+In the PIT histograms, bias leads to a triangular shape, overdispersion results in a hump shaped form and underdispersion in a U-shape.
+ADD INTERPRETATION FOR QUANTILE AND INTERVAL COVERAGE PLOTS HERE.
 
-To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
-
-A third way to look at calibration is to look a the probability integral tranform.
-As explained previously, the CDF of predictice distribution $P_t$ should ideally be equal to the CDF of the true unknown distribution $F_t$ that generated the observed value $x_t$.
-In order to assess whether there are substantial deviations between the two, @dawidPresentPositionPotential1984 suggested to transform the observed values using the probability integral transform (PIT).
-Agreement between the forecasts and the observed values can then be examined by observing whether or not the transformed values follow a uniform distribution.
-The PIT is given by
-$$u_t = P_t (x_t),$$
-where $u_t$ is the transformed variable and $P_t(x_t)$ is the predictive distribution evaluated at the true observed value $x_t$.
-If $P_t = F_t$ at all times $t$, then $u_t, t = 1 \dots T$ follows a uniform distribution (for a proof see e.g.
-@angusProbabilityIntegralTransform1994).
-
-In the case of discrete outcomes, the PIT is no longer uniform even when forecasts are ideal. As @funkAssessingPerformanceRealtime2019 suggest, we use a randomised PIT instead by redefining 
-$$u_t = P_t(x_t) + vt \cdot (P_t(x_t) - P_t(x_t - 1) ),$$
-where $x_t$ is again the observed value at time $t$, $P_t()$ is the CDF of the predictive distribution function, $P_t (-1) = 0$ by definition, and $v_t$ is a standard uniform variable independent of $x_t$.
-If $P_t$ is equal to the true data-generating distribution function, then $u_t$ is standard uniform.
-@czadoPredictiveModelAssessment2009 also propose a non-randomised version of the PIT for count data that could be used alternatively.
-
-One can then plot a histogram of $u_t$ values to look for deviations from uniformity.
-U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
-Biased predictions will usually result in a triangle-shaped histogram.
+\begin{figure}[h]
+\centering
+\includegraphics{plots/calibration-diagnostic-examples.png}
+\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
+\end{figure}
 
@@
 
-Sharpness is the second property central to model evaluation.
-The ability to produce narrow forecasts is a quality of the forecasts only and does not depend on the observations.
-Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
-Again we need to take slightly different approaches for continuous, integer and quantile forecasts.
-For continuous and integer forecasts we follow the suggestion from @funkAssessingPerformanceRealtime2019.
-For quantile forecasts we propose a novel metric.
+Sharpness is the ability to produce narrow forecasts.
+It does not depend on the actual observations and is a quality of the forecast only \ref{gneitingProbabilisticForecastsCalibration2007}.
+Sharpness is therefore only useful subject to calibration, as exemplified above in Figure \ref{fig:calibration-example}.
+We may be willing to trade off a little calibration for a lot more sharpness, but usually not much.
+For sample-based forecasts, \pkg{scoringutils} calculates sharpness as the normalised median absolute deviation about the median (MADN) [funkAssessingPerformanceRealtime2019] (for details see Table \ref{tab:score-overview}).
+For quantile forecasts, we take the sharpness component of the wis which corresponds to a weighted average of the individual interval widths.
 
-For continuous and integer forecasts, @funkAssessingPerformanceRealtime2019 suggest to measure sharpness as the normalised median absolute deviation about the median (MADN), i.e. 
-$$ S_t(P_t) = \frac{1}{0.675} \cdot \text{median}(|y - \text{median(y)}|), $$ 
-where $y$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant that ensures that sharpness will equal the standard deviation of the predictive distribution if $P_t$ is the CDF of a normal distribution.
-
-MAKE THE FOLLOWING PARAGRAPH SHORTER AND JUST SAY WE USE THE SHARPNESS COMPONENT OF THE WIS.
-
-For quantile forecasts, we propose to measure sharpness as a weighted mean of the width of the interval ranges.
-Let $Q_t$ be a set of predicted quantiles for a true $x_t$ at time $t$.
-This set of quantiles is assumed to be symmetric, such that there exist $K$ corresponding pairs of elements $q_{t, \frac{\alpha}{2}}$ and $q_{t, 1-\frac{\alpha}{2}}$.
-These $K$ corresponding pairs of quantiles cover a $(1 - \alpha) \cdot 100$ prediction interval.
-We can, accordingly, also denote $q_{t, \frac{\alpha}{2}}$ as $l_t$ the lower bound of the prediction interval at time $t$ and $q_{t, 1-\frac{\alpha}{2}}$ as $u_t$, the upper bound.
-We measure the sharpness of a quantile forecast at time $t$ as
-\begin{align*}
-\text{sharpness}_t &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (q_{t, 1 - \frac{\alpha}{2}} - q_{t, \frac{\alpha}{2}}) \\
-                   &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (u_t - l_t).
-\end{align*}
-Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
-We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score.
+WHAT WOULD BE REALLY COOL IS TO DETERMINE A WAY TO FIND THE OPTIMAL SHARPNESS OF A FORECAST. AS IS, I FEEL THIS PARAGRAPH IS SOMEWHAT USELESS.
 
@@
 
-If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION].
-In turn, each pair of models is evaluated based on the targets that that both models have predicted.
-The mean score by one model is divided by the mean score of the other model to obtain a measure of relative performance.
-Relative skill scores [citation (not sure there is something published, preprint fron Estee Cramer and Nick Reich)] for every model can then be obtained by taking the geometric mean of all pairwise ratios calculated for each model (omitting comparisons where there is no overlapping set of forecasts).
-ADD FORMULAS.
-A skill score smaller than 1 indicates that a model is performing better than the average model.
-One can also scale the skill by a reference baseline.
+If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION CRAMER et al.].
+In turn, each pair of models is evaluated based on the targets that both models have predicted.
+The mean score by one model is divided by the mean score of the other model to obtain the mean score ratio (see Table \ref{tab:score-overview}, a measure of relative performance.
+To obtain an overall relative skill score for a model, we take the geomatric mean of all mean score ratios that involve that model  (omitting comparisons where there is no overlapping set of forecasts).
+This gives us an indicator of performance relative to all other models.
+The orientation depends on the score used.
+For the proper scoring rules described above, smaller is better and a relative skill score smaller than 1 indicates that a model is performing better than the average model.
+We can obtain a scaled relative skill score by dividing a model's relative skill by the relative skill of a baseline model.
 A scaled relative skill smaller than one then means that the model in question performed better than the baseline.
 
-By applying a permutation test or non-parametric test like the Wilcoxon rank sum test we can also obtain p-values that help determine whether or not models perform significantly differently.
-THERE IS AN ISSUE WITH CORRELATION: THE TEST WOULD TREAT ALL OBSERVATIONS AS INDEPENDENT, WHICH THEY ARENT'T, SO I WOULD ASSUME THAT IT IS TOO LIBERAL.
-CURRENTLY, NO ADJUSTMENTS ARE IMPLEMENTED, I'M NOT REALLY SURE WHETHER TO DISCUSS THIS HERE OR NOT.
+It is in principle possible to obtain p-values that help determine whether two models perform significantly differently.
+\pkg{scoringutils} allows to compute these using eitab:score-tablether the Wilcoxon rank sum test or a permutation test.
+In practice, this is slightly complicated by the fact that both tests assume independent observations.
+In reality, however, forecasts by a model may be correlated across time or another dimension (e.g.
+if a forecaster has a bad day, they will likely perform badly across different targets for a given forecast date).
+P-values may therefore be too quick to suggest significant differences where there aren't any.
+One way to mitigate this is to aggregate observations over a category where one suspects correlation.
+A test that is performed on aggregate scores will likely be more conservative.
 
@@
 Columns: 6
-$ geography  <fct> London, London, London, London, London, London, …
-$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospit…
-$ value_desc <fct> Hospital admissions, Hospital admissions, Hospit…
-$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …
-$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23,…
-$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546,…
+$ geography  <fct> London, London, London, London, London, London, L…
+$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita…
+$ value_desc <fct> Hospital admissions, Hospital admissions, Hospita…
+$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0…
+$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, …
+$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, …
 \end{Soutput}
@@
 Columns: 8
-$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, …
-$ geography     <chr> "East of England", "East of England", "East o…
-$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hos…
-$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-…
-$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-…
-$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.0…
-$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, …
-$ value_desc    <fct> Hospital admissions, Hospital admissions, Hos…
+$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E…
+$ geography     <chr> "East of England", "East of England", "East of…
+$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp…
+$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3…
+$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0…
+$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05…
+$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0…
+$ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp…
 \end{Soutput}
@@
 \end{Schunk}
+\includegraphics{plots/plot-show-availability}
 \caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
@@
 As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
-% 
+% ion G
 \begin{Schunk}
@@
 \end{Sinput}
+\begin{Soutput}
+Rows: 29
+Columns: 9
+$ model              <fct> EpiSoon, StructuredODE, Microsimulation, …
+$ value_desc         <fct> Hospital admissions, Hospital admissions,…
+$ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876, …
+$ sharpness          <dbl> 22.886153, 6.133974, 33.985553, 10.365082…
+$ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.5…
+$ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.55…
+$ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174, …
+$ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0.…
+$ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563, …
+\end{Soutput}
 \end{Schunk}
@@
 \end{Schunk}
+\includegraphics{plots/plot-score-table}
 \caption{\label{fig:score-table} Coloured table to visualise the computed scores}
@@
 \end{Schunk}
+\includegraphics{plots/plot-pairwise-plot}
 \caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
@@
 \end{Schunk}
+\includegraphics{plots/plot-pairwise-plot-ratio}
 \caption{\label{fig:pairwise-comparison-pval} XXXX}
@@
 \end{Schunk}
+\includegraphics{plots/plot-WIS-components}
 \caption{\label{fig:wis-components} p-values and ratios together.}
@@
 \end{Schunk}
+\includegraphics{plots/plot-coverage}
 
@@
 \end{Schunk}
+\includegraphics{plots/plot-quantile-coverage}
 
@@
 \end{Schunk}
+\includegraphics{plots/plot-calibration}
 \caption{\label{fig:bias-heatmap} bias by location}
@@
 % The results in this paper were obtained using
-% \proglang{R}~4.0.3 with the
+% \proglang{R}~4.0.4 with the
 % \pkg{MASS}~7.3.53 package. \proglang{R} itself
@@
 
+\begin{table}[h!]
+\centering
+\caption{\label{tab:score-table} Explanation of all the scores}
+\end{table}
+
 \end{appendix}

### 2021-03-30 tables/forecast-types.tex
update scores sections of the paper

--- 2021-03-02 article.tex
@@
-\documentclass[article]{jss}
-
-%% -- LaTeX packages and custom commands ---------------------------------------
-
-%% recommended packages
-\usepackage{thumbpdf,lmodern}
-\usepackage{amsmath, amsfonts}
-\usepackage{bbm}% to get \mathbbmm{1}
-
-%% another package (only for this demo article)
+\documentclass{article}
 \usepackage{booktabs}
-\usepackage{adjustbox}
-\usepackage{tabularx}
-\usepackage{rotating}
-
-%% new custom commands
-\newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\code{#1()}}
-
-%% For Sweave-based articles about R packages:
-%% need no \usepackage{Sweave}
-
-%% -- Article metainformation (author, title, ...) -----------------------------
-
-%% - \author{} with primary affiliation
-%% - \Plainauthor{} without affiliations
-%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
-%% - \AND starts a new line, \And does not.
-\author{Nikos I. Bosse\\London School of Hygiene and Tropical Medicine
-   \And Second Author\\Plus Affiliation}
-\Plainauthor{Nikos I. Bosse, Second Author}
-
-%% - \title{} in title case
-%% - \Plaintitle{} without LaTeX markup (if any)
-%% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}
-\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using scoringutils in R}
-\Shorttitle{Evaluating Forecasts with \pkg{scoringutils} in \proglang{R}}
-
-%% - \Abstract{} almost as usual
-\Abstract{
-  Forecasts play an important role in a variety of fields.
-Its role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
-Even with a plethora of tools and models available to make forecasts, evaluating the predictions is not trivial.
-Even though scoring methods such as proper scoring rules have been widely studied in the past, there is a lack of software implementation that allows a forecaster to conveniently evaluate their forecasts.
-In this paper we introduce \pkg{scoringutils}, an \proglang{R} package that facilitates automated forecast evaluation.
-It gives the user access to a wide range of scoring metrics for various types of forecasts as well as a variety of ways to visualise the evaluation.
-We give an overview of the rationale behind forecast evaluation and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
-}
-
-%% - \Keywords{} with LaTeX markup, at least one required
-%% - \Plainkeywords{} without LaTeX markup (if necessary)
-%% - Should be comma-separated and in sentence case.
-\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
-\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
-
-%% - \Address{} of at least one author
-%% - May contain multiple affiliations for each author
-%%   (in extra lines, separated by \emph{and}\\).
-%% - May contain multiple authors for the same affiliation
-%%   (in the same first line, separated by comma).
-\Address{
-  Nikos Bosse\\
-  Centre for Mathematical Modelling of Infectious Diseases\\
-  London School of Hygiene and Tropical Medicine\\
-  Keppel Street \\ 
-  London WC1E 7HT \\
-  E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
-  % URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
-}
-
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{multirow}
+\usepackage{Sweave}
 \begin{document}
 
-%% -- Introduction -------------------------------------------------------------
-
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
-\section[Introduction]{Introduction}
-Good forecasts are of great interest to decision makers in various fields like finance (), weather predictions or infectious disease modeling \citep{funkShorttermForecastsInform2020}.
-An integral part of assessing and improving their usefulness is forecast evaluation.
-For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
-\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{FUNK}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
-Yet even with this rich body of research available, implementing a complete forecast evaluation in \proglang{R} is not trivial.
-We therefore present the \pkg{scoringutils} package.
-The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts.
-In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK \citep{funkShorttermForecastsInform2020}.
-
-\subsection{Forecast types}
-
-In its most general sense, a forecast is the forecaster’s stated belief about the future \citep{gneitingStrictlyProperScoring2007}.
-For conceptual clarity, it is however useful to distinguish between different types of forecasts.
-In terms of the quantitative forecasts that can be evaluated with \pkg{soringutils}, we can distinguish point forecasts from probabilistic forecasts.
-A point forecast states a single number and is the simplest form of a forecast.
-It is limited in its usefulness, as it does not give information about the uncertainty of the forecast.
-A very certain forecast may, for example, warrant a very different course of actions than does a very uncertain one.
-A probabilistic forecast, in contrast, is a forecast made in terms of the entire predictive distribution.
-Providing the entire predictive distribution allows the forecaster to express their belief about all aspects of the underlying data-generating distribution (including e.g.
-skewness or the width of its tails).
-Probabilistic forecasts are therefore the focus of this paper as well as the \pkg{scoringutils} package.
-
-\subsection{Forecast formats}
-
-Not only can we distinguish different forecast types, but also different formats in which predictions may be reported.
-This distinction is important, as different formats and prediction types require slightly different evaluation approaches.
-Forecasts can be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
-Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
-They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
-For that reason, often quantiles are reported instead [citation FORECAST HUBS].
-As an alternative representation of a quantile forecast, one can also express predictions using central prediction intervals.
-In this case, the forecaster reports a range to denote the interval as well as a value for the lower and upper bound.
-A forecaster could also in principle state their forecasts in a binary way by defining an outcome and assigning a probability that the outcome will come true.
-This type of forecasting is common in many classification problems.
-If we think of the outcome of some number being larger or smaller than a certain value, we can immediately see the connection between binary predictions and the concept of a cumulative distribution functions (CDF).
-All of these forecast types can be evaluated using \pkg{scoringutils}.
-The general forecasting paradigm [GNEITING] that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
-
-\subsection{The forecasting paradigm}
-
-Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007].
-For an ideal forecast, we therefore have
-%
-$$ P_t = F_t, $$
-%
-where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution.
-As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
-[CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
-Calibration refers to the statistical consistency between the predictive distribution and the observations.
-A well calibrated forecast does not systematically differ deviate from the observed values.
-For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007].
-Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
-how precise the forecasts are.
-The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
-Take for example the task of predicting rainfall in a city like London.
-A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time.
-On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
-%Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
-DO I WANT THAT FIGURE?
-
-% ==============================================================================
-\section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}
-
-Table:
-
-\input{tables/scores-table.tex}
-
-The metrics included in the \pkg{scoringuitls} package can be divided into two categories.
-The first consists of metrics that aim to capture different aspects of sharpness and calibration, the second comprises various proper scoring rules.
-We begin with the latter.
-
-\subsection{Proper scoring rules}
-
-Proper scoring rules [CITATION] jointly assess sharpness and calibration and assign a single numeric value to a forecast.
-A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
-This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state state their true best belief.
-The following scoring rules are implemented in \pkg{scoringutils}: The (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], the Dawid-Sebastiani-score (dss) [CITATION], the (weighted) interval score (wis), and the Brier score (bs).
-The first three proper scoring rules are implemented as wrappers around functions from the \pkg{scoringRules} package.
-They are suitable for any sample-based prediction format.
-The Log Score, however, is not applied if integer-valued forecasts are supplied, as the implementation in \pkg{scoringRules} requires a kernel density estimation that may be inappropriate for integer values.
-The interval score is appropriate for forecasts in a quantile formats.
-For every central prediction interval, the interval score is computed as the sum of separate penalties for overprediction, underprediction and sharpness (i.e.
-width) of the forecast.
-By default, the weighted interval score is returned.
-The overall score for a forecast is obtained as the mean of all scores for the individual central prediction intervals, weighted according to the range of the prediction interval.
-This ensures that the weighted interval score converges to the continuous ranked probability score for an increasing number of available prediction intervals.
-DO I WANT FORMULAS FOR ALL THE SCORES? The Brier score is used for binary predictions.
-
-TALK ABOUT CAVEATS (LATER?) E.G. THAT THE MEAN WIS IS INFLUENCED BY THE ABSOLUTE VALUE OF THE TARGET?
-
-\subsection{Evaluating calibration and sharpness independently}
-
-In addition to the proper scoring rules outlined above, \pkg{scoringutils} makes numerous metrics available to evaluate calibration and sharpness independently.
-This is especially helpful for model diagnostics.
-
-\subsubsection{Assessing calibration}
-
-Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g.
-@funkAssessingPerformanceRealtime2019; @gneitingProbabilisticForecastsCalibration2007; @gneitingStrictlyProperScoring2007).
-Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
-
-Bias, i.e.
-systematic over- or underprediction, is a very common form of miscalibration.
-For continuous forecasts, assessing whether a predictive distribution has a tendency to over- or underpredict can be very easily achieved by simply evaluating the predictive distribution at the true observed value.
-This metric is a generalisation of the integer-valued one @funkAssessingPerformanceRealtime2019 have proposed.
-It is also closely related to the probability integral transform (PIT) discussed later in this chapter.
-To improve the interpretability of the score we can transform it to a value between -1 (under-prediction) and 1 (over-prediction).
-Consequently, we measure bias as
-$$B_t (P_t, x_t) = 1 - 2 \cdot (P_t (x_t)),$$
-where $P_t$ is the cumulative distribution function of the predictive distribution for the true value $x_t$.
-When using predictive samples, $P_t (x_t)$ is simply the fraction of predictive samples for $x_t$ that are smaller than the true observed $x_t$.
-
-For integer valued forecasts, we use the metric proposed by @funkAssessingPerformanceRealtime2019: 
-$$B_t (P_t, x_t) = 1 - (P_t (x_t) + P_t (x_t + 1)).$$
-Bias can again assume values between -1 (under-prediction) and 1 (over-prediction) and is 0 ideally.
-
-For quantile forecasts, we propose the following metric to assess bias: 
-\begin{align*}
-  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t \}) \mathbbm{1}( x_t \leq q_{t, 0.5}) \\
-  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbbm{1}( x_t \geq q_{t, 0.5}),
-\end{align*}
-where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$.
-They represent our belief about what the true value $x_t$ will be.
-For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$.
-$\mathbbm{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
-In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution.
-If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
-If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
-For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
-For quantile forecasts, an alternative approach is to look at the over- and underprediction components of the weighted interval score.
-These however, capture the bias in absolute terms, while the above proposed metric captures a tendency to over- and underpredict that is less sensitive to outliers.
-
-I ASSUME ALL OF THIS IS A BIT TOO LONG?
-
-Another way to look at calibration[precisely: probabilistic calibration in @gneitingProbabilisticForecastsCalibration2007] is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
-This is most easily understood in the context of quantile forecasts, but can in principle be transferred to continuous and integer forecasts as well.
-
-To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution.
-If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles.
-We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
-
-To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
-
-A third way to look at calibration is to look a the probability integral tranform.
-As explained previously, the CDF of predictice distribution $P_t$ should ideally be equal to the CDF of the true unknown distribution $F_t$ that generated the observed value $x_t$.
-In order to assess whether there are substantial deviations between the two, @dawidPresentPositionPotential1984 suggested to transform the observed values using the probability integral transform (PIT).
-Agreement between the forecasts and the observed values can then be examined by observing whether or not the transformed values follow a uniform distribution.
-The PIT is given by
-$$u_t = P_t (x_t),$$
-where $u_t$ is the transformed variable and $P_t(x_t)$ is the predictive distribution evaluated at the true observed value $x_t$.
-If $P_t = F_t$ at all times $t$, then $u_t, t = 1 \dots T$ follows a uniform distribution (for a proof see e.g.
-@angusProbabilityIntegralTransform1994).
-
-In the case of discrete outcomes, the PIT is no longer uniform even when forecasts are ideal. As @funkAssessingPerformanceRealtime2019 suggest, we use a randomised PIT instead by redefining 
-$$u_t = P_t(x_t) + vt \cdot (P_t(x_t) - P_t(x_t - 1) ),$$
-where $x_t$ is again the observed value at time $t$, $P_t()$ is the CDF of the predictive distribution function, $P_t (-1) = 0$ by definition, and $v_t$ is a standard uniform variable independent of $x_t$.
-If $P_t$ is equal to the true data-generating distribution function, then $u_t$ is standard uniform.
-@czadoPredictiveModelAssessment2009 also propose a non-randomised version of the PIT for count data that could be used alternatively.
-
-One can then plot a histogram of $u_t$ values to look for deviations from uniformity.
-U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
-Biased predictions will usually result in a triangle-shaped histogram.
-
-\subsubsection{Assessing sharpness}
-
-Sharpness is the second property central to model evaluation.
-The ability to produce narrow forecasts is a quality of the forecasts only and does not depend on the observations.
-Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
-Again we need to take slightly different approaches for continuous, integer and quantile forecasts.
-For continuous and integer forecasts we follow the suggestion from @funkAssessingPerformanceRealtime2019.
-For quantile forecasts we propose a novel metric.
-
-For continuous and integer forecasts, @funkAssessingPerformanceRealtime2019 suggest to measure sharpness as the normalised median absolute deviation about the median (MADN), i.e. 
-$$ S_t(P_t) = \frac{1}{0.675} \cdot \text{median}(|y - \text{median(y)}|), $$ 
-where $y$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant that ensures that sharpness will equal the standard deviation of the predictive distribution if $P_t$ is the CDF of a normal distribution.
-
-MAKE THE FOLLOWING PARAGRAPH SHORTER AND JUST SAY WE USE THE SHARPNESS COMPONENT OF THE WIS.
-
-For quantile forecasts, we propose to measure sharpness as a weighted mean of the width of the interval ranges.
-Let $Q_t$ be a set of predicted quantiles for a true $x_t$ at time $t$.
-This set of quantiles is assumed to be symmetric, such that there exist $K$ corresponding pairs of elements $q_{t, \frac{\alpha}{2}}$ and $q_{t, 1-\frac{\alpha}{2}}$.
-These $K$ corresponding pairs of quantiles cover a $(1 - \alpha) \cdot 100$ prediction interval.
-We can, accordingly, also denote $q_{t, \frac{\alpha}{2}}$ as $l_t$ the lower bound of the prediction interval at time $t$ and $q_{t, 1-\frac{\alpha}{2}}$ as $u_t$, the upper bound.
-We measure the sharpness of a quantile forecast at time $t$ as
-\begin{align*}
-\text{sharpness}_t &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (q_{t, 1 - \frac{\alpha}{2}} - q_{t, \frac{\alpha}{2}}) \\
-                   &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (u_t - l_t).
-\end{align*}
-Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
-We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score.
-
-\subsection{Pairwise comparisons}
-
-If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION].
-In turn, each pair of models is evaluated based on the targets that that both models have predicted.
-The mean score by one model is divided by the mean score of the other model to obtain a measure of relative performance.
-Relative skill scores [citation (not sure there is something published, preprint fron Estee Cramer and Nick Reich)] for every model can then be obtained by taking the geometric mean of all pairwise ratios calculated for each model (omitting comparisons where there is no overlapping set of forecasts).
-ADD FORMULAS.
-A skill score smaller than 1 indicates that a model is performing better than the average model.
-One can also scale the skill by a reference baseline.
-A scaled relative skill smaller than one then means that the model in question performed better than the baseline.
-
-By applying a permutation test or non-parametric test like the Wilcoxon rank sum test we can also obtain p-values that help determine whether or not models perform significantly differently.
-THERE IS AN ISSUE WITH CORRELATION: THE TEST WOULD TREAT ALL OBSERVATIONS AS INDEPENDENT, WHICH THEY ARENT'T, SO I WOULD ASSUME THAT IT IS TOO LIBERAL.
-CURRENTLY, NO ADJUSTMENTS ARE IMPLEMENTED, I'M NOT REALLY SURE WHETHER TO DISCUSS THIS HERE OR NOT.
-
-% ==============================================================================
-\section{Evaluating UK short-term forecasts}
-\subsection{The data}
-
-To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
-Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
-The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
-Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
-The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL].
-
-- timing (weekly?) and number of forecast dates
-- the four targets
-- the models. 
-- is the set complete?
-
-We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> # install and load data pacakge from external repository
-R> # remotes::install_github("sbfnk/covid19.forecasts.uk")
-R> 
-R> # load packages
-R> library(covid19.forecasts.uk)
-R> library(dplyr)
-R> library(scoringutils)
-R> # load truth data
-R> data(covid_uk_data)
-R> # head(covid_uk_data)
-R> 
-R> # load forecasts
-R> data(uk_forecasts)
-R> # head(uk_forecasts)
-\end{Sinput}
-\end{Schunk}
-% 
-Let us take a first look a the data:
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> glimpse(covid_uk_data)
-\end{Sinput}
-\begin{Soutput}
-Rows: 4,174
-Columns: 6
-$ geography  <fct> London, London, London, London, London, London, …
-$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospit…
-$ value_desc <fct> Hospital admissions, Hospital admissions, Hospit…
-$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …
-$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23,…
-$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546,…
-\end{Soutput}
-\begin{Sinput}
-R> glimpse(uk_forecasts)
-\end{Sinput}
-\begin{Soutput}
-Rows: 1,254,513
-Columns: 8
-$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, …
-$ geography     <chr> "East of England", "East of England", "East o…
-$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hos…
-$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-…
-$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-…
-$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.0…
-$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, …
-$ value_desc    <fct> Hospital admissions, Hospital admissions, Hos…
-\end{Soutput}
-\end{Schunk}
-% 
-To bring the forecasts into the format needed for the evaluation, some minor changes need to be made to the data.
-The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
-While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public.
-The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way.
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> uk_forecasts <- rename(uk_forecasts, prediction = value)
-R> covid_uk_data <-rename(covid_uk_data, true_value = value)
-R> combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
-\end{Sinput}
-\end{Schunk}
-% 
-Before we start with scoring the forecasts, it makes sense to start the evaluation process by visualising the data.
-To get a feeling for how complete the data set is, we can run the following code to obtain a heatmap with the number of available forecasts:
-% 
-\begin{figure}[h]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> show_avail_forecasts(combined, 
-+                       x = "creation_date", 
-+                       show_numbers = FALSE,
-+                       legend_position = "bottom",
-+                       facet_formula = ~ value_desc)
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
-\end{figure}
-%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
-% 
-Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
-By default, the function treats a set of different quantiles or samples as one forecast.
-However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.
-
-Forecasts can be visualised using the \fct{plot\_predictions} function.
-The function accepts either a single combined data set or two separate truth data sets that can be merged together.
-Forecasts and observed values can be filtered independently, to show for example only predictions from a certain model or a specific number of weeks of true data before the forecast.
-Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
-To obtain, for example, a specific forecast from a model we are interest in, we can call:
-% 
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
-R> plot_predictions(truth = covid_uk_data,
-+                   forecasts = uk_forecasts,
-+                   filter_both = list(paste("geography %in%", locations, collapse = "")),
-+                   filter_truth = list('value_date <= "2020-06-22"',
-+                                       'value_date > "2020-06-01"'), 
-+                   filter_forecasts = list('model == "SIRCOVID"', 
-+                                           'creation_date == "2020-06-22"'),
-+                   x = "value_date",
-+                   facet_formula = geography ~ value_type) + 
-+    ggplot2::theme(legend.position = "bottom")
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
-\end{figure}
-% 
-The output is shown in Figure \ref{fig:forecast-visualisation}.
-
-\subsection[Scoring forecasts with eval_forecasts]{Scoring forecasts with \fct{eval\_forecasts}}
-
-A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
-This requires a \code{data.frame} or similar which has at least a column called "prediction" and one called "true\_value".
-Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
-Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
-We will not discuss all possible input formats here, but instead refer to the example data for each format that is provided with the package.
-Where possible, \pkg{scoringutils} also provides functionality to transform between various formats, e.g.
-from a sample based format to a quantile format.
-The \fct{eval\_forecasts} function automatically recognises the prediction type and input format, applies the appropriate scoring metrics and aggregates results as desired by the user.
-Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
-
-As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> scores <- eval_forecasts(combined, 
-+                           summarise_by = c("model", "value_desc"))
-R> glimpse(scores)
-\end{Sinput}
-\end{Schunk}
-% 
-If a more detailed analysis is desired, the level of aggregation can of course be changed to show for example separate scores for the different locations as well.
-This can be achieved using the \code{summarise_by} argument.
-To additionally stratify by location, we could specify \code{summarise_by = c("mode", "value_type", "geography")}.
-If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
-This can be useful if we, for example, want to analyse what proportion of true values are covered by certain interval ranges, or if we want to analyse the accuracy of the tails of the forecasts.
-When aggregating, \fct{eval\_forecasts} takes the mean according to the group defined in \code{summarise_by}.
-In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
-In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.
-
-The user must, however, still exercise some caution when aggregating scores, as many of the metrics are absolute and scale with the magnitude of the quantity to forecast.
-Looking at one score per model (i.e.
-specifying \code{summarise_by = c("model")}) may not be so useful in this instance, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
-
-In the above example, we did not have to explicitly specify the \code{by} argument, but this may be necessary if additional columns are present in the data that do not indicate a grouping of forecasts.
-The \code{by} argument must then be used to denote the unit of a single forecast.
-In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")}.
-Quantiles should not be included, as several quantiles make up one forecast (and similarly for samples).
-If we had additional columns that do not serve to group forecasts (like for example the number of inhabitants in a certain location over time), these should also not be included.
-By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
-
-\subsection{Pairwise comparisons}
-
-Pairwise comparisons between models [CITATION] can be obtained in two different ways.
-First, relative skill scores based on pairwise comparisons are by default returned from \fct{eval\_forecasts}.
-These will be computed separately for the categories defined in the \code{summarise_by} argument (excluding the category 'model').
-Alternatively, a set of scores can be post-processed using the separate function \fct{pairwise\_comparison}.
-This approach is to be used for visualisation and if p-values for the pairwise comparisons are needed, as those are not returned from \fct{eval\_forecasts}.
-Usually, one would compute scores without specifying a \code{summarise_by} argument, but sometimes it may be sensible to average over certain scores, for example for predictions generated at a certain date.
-This allows to reduce the correlation between observations that enter the computation of p-values, which in turn makes the test less liberal.
-Using the function \fct{plot\_pairwise\_comparison} we can visualise the mean score ratios between all models as well as the
-
-\begin{Schunk}
-\begin{Sinput}
-R> # unsummarised scores
-R> unsummarised_scores <- eval_forecasts(combined)
-R> pairwise <- pairwise_comparison(unsummarised_scores, 
-+                                  summarise_by = "value_desc")
-\end{Sinput}
-\end{Schunk}
-
-The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
-
-\section{Visualisation and interpretation of evaluation results}
-
-\subsection{Visualising aggregate scores and rankings}
-A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
-
-\begin{figure}[h]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> score_table(scores, y = "model", facet_formula = ~ value_desc)
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
-\end{figure}
-
-The most informative metric in terms of model ranking is the relative\_skill.
-However, interpretation is not always straightforward and has to be done carefully.
-We can see that performance varied quite a bit across different metrics, where some models did well on one target, but poorly on another.
-Especially the Exponential growth/decline model stands out as it received the lowest relative skill score for hospital admissions, but the highest for the total number of beds occupied.
-Looking back at Figure \ref{fig:avail-forecasts}, we see that the model has only submitted very few forecasts over all.
-It may therefore be sensible to require all models to have submitted forecasts for at least 50\% of all forecast targets in order to enter the pairwise comparisons.
-For similar reasons, the interval score may be deceiving if looked at in isolation.
-As can be seen, the DetSEIRwithNB MLE model received a lower relative skill score, but a higher interval score than the DetSEIRwithNB MCMC model.
-This, again, can be explained by the fact that they forecasted different targets.
-The interval score, as an absolute metric, is highly influenced by the absolute value of the quantity that is forecasted.
-For the same reason, one should be careful when summarising interval scores from different locations or forecast targets, as the average score will be dominated by outliers as well as differences in the absolute level.
-Assuming a large enough set of available overlapping forecasts, the relative skill score is more robust.
-It therefore is reasonable to assume that the DetSEIRwithNB MLE forecasted quantities with a higher absolute value, but tended to perform worse than the DetSEIRwithNB MCMC model as far as we can tell based on the set of all pariwise comparisons.
-This can be confirmed for the direct comparison between the two by looking at the mean score ratios from the pairwise comparisons.
-These can be obtained by calling
-
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise) +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
-If a tile is blue, then the model on the y-axis performed better.
-If it is red, the model on the x-axis performed better in direct comparison.
-}
-\end{figure}
-
-we can also look at p-values in Figure \ref{fig:pairwise-comparison-pval} PROBABLY REMOVE THAT FROM THE PAPER
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise, type = "pval") +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:pairwise-comparison-pval} XXXX}
-\end{figure}
-
-In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
-
-\subsection{Visual model diagnostics}
-
-For forecasts in an interval format, looking at the components of the weighted interval score separately is a natural next step.
-We can see in Figure \ref{fig:wis-components} that the majority of penalties come from over-and underprediction, instead of the sharpness component.
-We also see that most models tended to either over- or underpredict actual numbers.
-
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> wis_components(scores, 
-+                 facet_formula = ~ value_desc, 
-+                 scales = "free_x") + 
-+    ggplot2::coord_flip() + 
-+    ggplot2::theme(legend.position = "bottom")
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:wis-components} p-values and ratios together.}
-\end{figure}
-
-We can have a closer look at calibration using the functions \fct{interval\_coverage} and \fct{quantile\_coverage}.
-The interval coverage plot shows the proportion of all true values that fall within all the different prediction intervals.
-This gives a visual impression of probabilistic calibration [GNEITING].
-Ideally, $x$ percent of true values should be covered by the $x$\%-prediction intervals, resulting in a 45° line.
-Areas shaded in green indicate that the model is covering more true values than it actually should, while areas in white indicate that the model fails to cover the desired proportion of true values with its prediction intervals.
-The majority of the models were too confident in their predictions, while some showed showed good calibration.
-The quantile coverage plot shows the proportion of all true values below certain predictive quantiles.
-While this plot is slightly harder to interpret, it also includes information about bias as and allows to separate the lower and upper boundaries of the prediction intervals.
-We can see, for example, that the Exponential growth/decline model was consistently biased downwards.
-Figure \ref{fig:coverage}
-
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> cov_scores <- eval_forecasts(combined, 
-+                               summarise_by = c("model", 
-+                                                "range", "quantile"))
-R> scoringutils::interval_coverage(cov_scores)
-R> scoringutils::quantile_coverage(cov_scores)
-\end{Sinput}
-\end{Schunk}
-
-\begin{Schunk}
-\begin{Sinput}
-R> cov_scores <- eval_forecasts(combined, 
-+                               summarise_by = c("model", 
-+                                                "range", "quantile"))
-R> scoringutils::quantile_coverage(cov_scores)
-\end{Sinput}
-\end{Schunk}
-
-\caption{\label{fig:coverage} quantile and interval coverage}
-\end{figure}
-
-DO I WANT TO INCLUDE PIT PLOTS AS WELL? I GUESS? NEED TO LOOK AT THE IMPLEMENTATION FOR QUANTILE FORECASTS
-
-% \begin{figure}[h]
-% \centering
-% <<
-% pit-plots, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
-% # out <- eval_forecasts(combined, 
-% #                       summarise_by = c("model"), 
-% #                       pit_plots = TRUE)
-% 
-% @
-% \caption{\label{fig:calibration-pit} pit plots}
-% \end{figure}
-
-Look at e.g. bias by location? Figure \ref{fig:bias-heatmap}
-
-\begin{figure}[h]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> scores <- eval_forecasts(combined, 
-+                           summarise_by = c("model",
-+                                            "value_desc", 
-+                                            "geography"), 
-+                           compute_relative_skill = FALSE)
-R> scoringutils::score_heatmap(scores, metric = "bias", 
-+                              x = "geography", facet_formula = ~ value_desc, 
-+                              scale = "free_x")
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:bias-heatmap} bias by location}
-\end{figure}
-
-WHAT IS NEEDED HERE IS A BIT OF THINKING WITH REGARDS TO WHAT VISUALISATION I WANT TO SHOW AND IN HOW MUCH DETAIL I WANT TO ANALYSE THE MODELS.
-
-\begin{CodeInput}
-.
-\end{CodeInput}
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-
-\section{Summary and discussion} \label{sec:summary}
-
-COMING SOON.
-
-% The results in this paper were obtained using
-% \proglang{R}~4.0.3 with the
-% \pkg{MASS}~7.3.53 package. \proglang{R} itself
-% and all packages used are available from the Comprehensive
-% \proglang{R} Archive Network (CRAN) at
-% \url{https://CRAN.R-project.org/}.
-
-\section*{Acknowledgments}
-
-% All acknowledgments (note the AE spelling) should be collected in this
-% unnumbered section before the references. It may contain the usual information
-% about funding and feedback from colleagues/reviewers/etc. Furthermore,
-% information such as relative contributions of the authors may be added here
-% (if any).
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
-\bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-\newpage
-
-\begin{appendix}
-
-\section{Appendix section} \label{app:technical}
-
-\end{appendix}
-
-%% -----------------------------------------------------------------------------
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{4.5cm}}
+\toprule
+\textbf{Forecast type} & \textbf{Target type} & \textbf{Representation of the predictive distribution}\\
+\midrule
+Point forecast & continuous \newline  integer \newline  binary & one single number for the predicted outcome\\
+\cmidrule{1-3}\pagebreak[0]
+ & continuous \newline  integer & predictive samples \newline   quantiles \newline   closed analytical form\\
+\cmidrule{2-3}\nopagebreak
+\multirow[t]{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
+\bottomrule
+\end{longtable}
 

### 2021-03-30 tables/scores-table-summary.Rnw
update scores sections of the paper

\documentclass{article}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{float}
\usepackage{longtable}
\begin{document}

<<table, echo=FALSE, results=tex>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
library(data.table)

crps <- list(
  `Metric` = "CRPS (Continuous) ranked probability score", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(closed-form, samples (approximation))",
  `Properties` = "proper scoring rule, global, stable handling of outliers",
  `References` = ""
)

log_score <- list(
  `Metric` = "Log score", 
  `Target types` = r"(continuous, (integer not in scoringutils))", 
  `Forecast formats` = r"(closed-form, samples (approximation))",
  `Properties` = "proper scoring rule, local, unstable for outliers",
  `References` = ""
)

wis <- list(
  Metric = "WIS (Weighted) interval score", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(quantile or interval predictions)",
  `Properties` = "proper scoring rule, global, stable handling of outliers, converges to crps",
  `References` = ""
)

dss <- list(
  `Metric` = "DSS Dawid-Sebastiani score", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(closed-form, samples (approximation))",
  `Properties` = "proper scoring rule, somewhat global, somewhat stable handling of outliers",
  `References` = ""
)

brier_score <- list(
  `Metric` = "Brier score", 
  `Target types` = r"(binary)", 
  `Forecast formats` = r"(binary probabilities)",
  `Properties` = "proper scoring rule",
  `References` = ""
)

interval_coverage <- list(
  `Metric` = "Interval coverage", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(interval forecasts (needs matching quantiles))",
  `Properties` = "measure for calibration",
  `References` = ""
)

quantile_coverage <- list(
  `Metric` = "Quantile coverage", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(quantile or interval forecasts)",
  `Properties` = "measure for calibration",
  `References` = ""
)

sharpness <- list(
  `Metric` = "Sharpness", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
  `Properties` = "measures sharpness, slightly different depending on forecast format",
  `References` = ""
)

bias <- list(
  `Metric` = "Bias", 
  `Target types` = r"(continuous, integer, quantile)", 
  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
  `Properties` = "captures tendency to over-or underpredict (aspect of calibration)",
  `References` = ""
)

pit <- list(
  `Metric` = "Probability integral transform (PIT)",
  `Target types` = r"(continuous, integer, quantile)", 
  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
  `Properties` = "assesses calibration",
  `References` = ""
)

mean_score_ratio <- list(
  `Metric` = "Mean score ratio", 
  `Target types` = r"(depends on score)", 
  `Forecast formats` = r"(depends on score)",
  `Properties` = "compares performance of two models",
  `References` = ""
)

relative_skill <- list(
  `Metric` = "Relative skill", 
  `Target types` = r"(depends on scored)", 
  `Forecast formats` = r"(depends on score)",
  `Properties` = "Ranks models based on pairwise comparisons",
  `References` = ""
)

data <- rbind(as.data.table(crps), 
              as.data.table(log_score), 
              as.data.table(wis), 
              as.data.table(dss), 
              as.data.table(brier_score), 
              as.data.table(interval_coverage),
              as.data.table(quantile_coverage), 
              as.data.table(pit),
              as.data.table(sharpness),
              as.data.table(bias), 
              as.data.table(mean_score_ratio), 
              as.data.table(relative_skill))

saveRDS(data, file = "tables/scores-table-summary.rds")

library(kableExtra)
library(magrittr)
library(knitr)
data[, 1:4] %>%
  kableExtra::kbl(format = "latex", booktabs = TRUE, 
                  escape = FALSE,
                  longtable = TRUE,
                  linesep = c('\\addlinespace')) %>%
  kableExtra::column_spec(1, width = "2.5cm") %>%
  kableExtra::column_spec(2, width = "2cm") %>%
  kableExtra::column_spec(3, width = "4.5cm") %>%
  kableExtra::column_spec(4, width = "4cm") %>%
  kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))

@

\end{document}

### 2021-03-30 tables/scores-table-summary.tex
update scores sections of the paper

--- 2021-03-02 article.tex
@@
-\documentclass[article]{jss}
-
-%% -- LaTeX packages and custom commands ---------------------------------------
-
-%% recommended packages
-\usepackage{thumbpdf,lmodern}
-\usepackage{amsmath, amsfonts}
-\usepackage{bbm}% to get \mathbbmm{1}
-
-%% another package (only for this demo article)
+\documentclass{article}
 \usepackage{booktabs}
-\usepackage{adjustbox}
-\usepackage{tabularx}
-\usepackage{rotating}
-
-%% new custom commands
-\newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\code{#1()}}
-
-%% For Sweave-based articles about R packages:
-%% need no \usepackage{Sweave}
-
-%% -- Article metainformation (author, title, ...) -----------------------------
-
-%% - \author{} with primary affiliation
-%% - \Plainauthor{} without affiliations
-%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
-%% - \AND starts a new line, \And does not.
-\author{Nikos I. Bosse\\London School of Hygiene and Tropical Medicine
-   \And Second Author\\Plus Affiliation}
-\Plainauthor{Nikos I. Bosse, Second Author}
-
-%% - \title{} in title case
-%% - \Plaintitle{} without LaTeX markup (if any)
-%% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}}
-\Plaintitle{Evaluating Covid-19 Short-Term Forecasts using scoringutils in R}
-\Shorttitle{Evaluating Forecasts with \pkg{scoringutils} in \proglang{R}}
-
-%% - \Abstract{} almost as usual
-\Abstract{
-  Forecasts play an important role in a variety of fields.
-Its role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
-Even with a plethora of tools and models available to make forecasts, evaluating the predictions is not trivial.
-Even though scoring methods such as proper scoring rules have been widely studied in the past, there is a lack of software implementation that allows a forecaster to conveniently evaluate their forecasts.
-In this paper we introduce \pkg{scoringutils}, an \proglang{R} package that facilitates automated forecast evaluation.
-It gives the user access to a wide range of scoring metrics for various types of forecasts as well as a variety of ways to visualise the evaluation.
-We give an overview of the rationale behind forecast evaluation and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
-}
-
-%% - \Keywords{} with LaTeX markup, at least one required
-%% - \Plainkeywords{} without LaTeX markup (if necessary)
-%% - Should be comma-separated and in sentence case.
-\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
-\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
-
-%% - \Address{} of at least one author
-%% - May contain multiple affiliations for each author
-%%   (in extra lines, separated by \emph{and}\\).
-%% - May contain multiple authors for the same affiliation
-%%   (in the same first line, separated by comma).
-\Address{
-  Nikos Bosse\\
-  Centre for Mathematical Modelling of Infectious Diseases\\
-  London School of Hygiene and Tropical Medicine\\
-  Keppel Street \\ 
-  London WC1E 7HT \\
-  E-mail: \email{nikos.bosse@lshtm.ac.uk}\\
-  % URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
-}
-
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{xcolor}
+% \usepackage{colortbl}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{Sweave}
 \begin{document}
 
-%% -- Introduction -------------------------------------------------------------
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{4cm}}
+\toprule
+Metric & Target types & Forecast formats & Properties\\
+\midrule
+\endfirsthead
+\multicolumn{4}{@{}l}{\textit{(continued)}}\\
+\toprule
+Metric & Target types & Forecast formats & Properties\\
+\midrule
+\endhead
 
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
-\section[Introduction]{Introduction}
-Good forecasts are of great interest to decision makers in various fields like finance (), weather predictions or infectious disease modeling \citep{funkShorttermForecastsInform2020}.
-An integral part of assessing and improving their usefulness is forecast evaluation.
-For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
-\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{FUNK}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
-Yet even with this rich body of research available, implementing a complete forecast evaluation in \proglang{R} is not trivial.
-We therefore present the \pkg{scoringutils} package.
-The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts.
-In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK \citep{funkShorttermForecastsInform2020}.
-
-\subsection{Forecast types}
-
-In its most general sense, a forecast is the forecaster’s stated belief about the future \citep{gneitingStrictlyProperScoring2007}.
-For conceptual clarity, it is however useful to distinguish between different types of forecasts.
-In terms of the quantitative forecasts that can be evaluated with \pkg{soringutils}, we can distinguish point forecasts from probabilistic forecasts.
-A point forecast states a single number and is the simplest form of a forecast.
-It is limited in its usefulness, as it does not give information about the uncertainty of the forecast.
-A very certain forecast may, for example, warrant a very different course of actions than does a very uncertain one.
-A probabilistic forecast, in contrast, is a forecast made in terms of the entire predictive distribution.
-Providing the entire predictive distribution allows the forecaster to express their belief about all aspects of the underlying data-generating distribution (including e.g.
-skewness or the width of its tails).
-Probabilistic forecasts are therefore the focus of this paper as well as the \pkg{scoringutils} package.
-
-\subsection{Forecast formats}
-
-Not only can we distinguish different forecast types, but also different formats in which predictions may be reported.
-This distinction is important, as different formats and prediction types require slightly different evaluation approaches.
-Forecasts can be expressed analytically in a closed form, or more commonly through either predictive samples or quantiles.
-Predictive samples are useful if no closed-form predictive distribution is available, but require a lot of storage space.
-They may also come with a loss of precision that is especially pronounced in the tails of the predictive distribution, where quite a lot of samples are needed to accurately characterise the distribution.
-For that reason, often quantiles are reported instead [citation FORECAST HUBS].
-As an alternative representation of a quantile forecast, one can also express predictions using central prediction intervals.
-In this case, the forecaster reports a range to denote the interval as well as a value for the lower and upper bound.
-A forecaster could also in principle state their forecasts in a binary way by defining an outcome and assigning a probability that the outcome will come true.
-This type of forecasting is common in many classification problems.
-If we think of the outcome of some number being larger or smaller than a certain value, we can immediately see the connection between binary predictions and the concept of a cumulative distribution functions (CDF).
-All of these forecast types can be evaluated using \pkg{scoringutils}.
-The general forecasting paradigm [GNEITING] that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
-
-\subsection{The forecasting paradigm}
-
-Any forecaster should aim to minimise the difference between the predictive distribution and the unknown true data-generating distribution [@gneitingProbabilisticForecastsCalibration2007].
-For an ideal forecast, we therefore have
-%
-$$ P_t = F_t, $$
-%
-where $P_t$ is the the cumulative density function (CDF) of the predictive distribution at time $t$ and $F_t$ is the CDF of the true, unknown data-generating distribution.
-As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
-[CITATIONS] instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
-Calibration refers to the statistical consistency between the predictive distribution and the observations.
-A well calibrated forecast does not systematically differ deviate from the observed values.
-For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to [GneitingProbabilisticForecastsCalibration2007].
-Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
-how precise the forecasts are.
-The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
-Take for example the task of predicting rainfall in a city like London.
-A model that made very precise forecasts would not be useful if the forecasts were wrong most of the time.
-On the other hand, a model that predicts the same rainfall probability for every day can be correct on average [To be precise, this model would be marginally calibrated according to @gneitingProbabilisticForecastsCalibration2007], but is also less useful than a model that were able to accurately predict the weather every single day.
-%Figure \@ref(fig:forecast-paradigm) illustrates the concepts of calibration and sharpness once again.
-DO I WANT THAT FIGURE?
-
-% ==============================================================================
-\section[metrics]{Scoring metrics implemented in \pkg{scoringutils}}
-
-Table:
-
-\input{tables/scores-table.tex}
-
-The metrics included in the \pkg{scoringuitls} package can be divided into two categories.
-The first consists of metrics that aim to capture different aspects of sharpness and calibration, the second comprises various proper scoring rules.
-We begin with the latter.
-
-\subsection{Proper scoring rules}
-
-Proper scoring rules [CITATION] jointly assess sharpness and calibration and assign a single numeric value to a forecast.
-A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
-This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state state their true best belief.
-The following scoring rules are implemented in \pkg{scoringutils}: The (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], the Dawid-Sebastiani-score (dss) [CITATION], the (weighted) interval score (wis), and the Brier score (bs).
-The first three proper scoring rules are implemented as wrappers around functions from the \pkg{scoringRules} package.
-They are suitable for any sample-based prediction format.
-The Log Score, however, is not applied if integer-valued forecasts are supplied, as the implementation in \pkg{scoringRules} requires a kernel density estimation that may be inappropriate for integer values.
-The interval score is appropriate for forecasts in a quantile formats.
-For every central prediction interval, the interval score is computed as the sum of separate penalties for overprediction, underprediction and sharpness (i.e.
-width) of the forecast.
-By default, the weighted interval score is returned.
-The overall score for a forecast is obtained as the mean of all scores for the individual central prediction intervals, weighted according to the range of the prediction interval.
-This ensures that the weighted interval score converges to the continuous ranked probability score for an increasing number of available prediction intervals.
-DO I WANT FORMULAS FOR ALL THE SCORES? The Brier score is used for binary predictions.
-
-TALK ABOUT CAVEATS (LATER?) E.G. THAT THE MEAN WIS IS INFLUENCED BY THE ABSOLUTE VALUE OF THE TARGET?
-
-\subsection{Evaluating calibration and sharpness independently}
-
-In addition to the proper scoring rules outlined above, \pkg{scoringutils} makes numerous metrics available to evaluate calibration and sharpness independently.
-This is especially helpful for model diagnostics.
-
-\subsubsection{Assessing calibration}
-
-Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g.
-@funkAssessingPerformanceRealtime2019; @gneitingProbabilisticForecastsCalibration2007; @gneitingStrictlyProperScoring2007).
-Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
-
-Bias, i.e.
-systematic over- or underprediction, is a very common form of miscalibration.
-For continuous forecasts, assessing whether a predictive distribution has a tendency to over- or underpredict can be very easily achieved by simply evaluating the predictive distribution at the true observed value.
-This metric is a generalisation of the integer-valued one @funkAssessingPerformanceRealtime2019 have proposed.
-It is also closely related to the probability integral transform (PIT) discussed later in this chapter.
-To improve the interpretability of the score we can transform it to a value between -1 (under-prediction) and 1 (over-prediction).
-Consequently, we measure bias as
-$$B_t (P_t, x_t) = 1 - 2 \cdot (P_t (x_t)),$$
-where $P_t$ is the cumulative distribution function of the predictive distribution for the true value $x_t$.
-When using predictive samples, $P_t (x_t)$ is simply the fraction of predictive samples for $x_t$ that are smaller than the true observed $x_t$.
-
-For integer valued forecasts, we use the metric proposed by @funkAssessingPerformanceRealtime2019: 
-$$B_t (P_t, x_t) = 1 - (P_t (x_t) + P_t (x_t + 1)).$$
-Bias can again assume values between -1 (under-prediction) and 1 (over-prediction) and is 0 ideally.
-
-For quantile forecasts, we propose the following metric to assess bias: 
-\begin{align*}
-  B_t =& (1 - 2 \cdot \max \{i | q_{t,i} \in Q_t \land q_{t,i} \leq x_t \}) \mathbbm{1}( x_t \leq q_{t, 0.5}) \\
-  &+ (1 - 2 \cdot \min \{i | q_{t,i} \in Q_t \land q_{t,i} \geq x_t\}) \mathbbm{1}( x_t \geq q_{t, 0.5}),
-\end{align*}
-where $Q_t$ is the set of quantiles that form the predictive distribution at time $t$.
-They represent our belief about what the true value $x_t$ will be.
-For consistency, we define $Q_t$ such that it always includes the element $q_{t, 0} = - \infty$ and $q_{t,1} = \infty$.
-$\mathbbm{1}()$ is the indicator function that is $1$ if the condition is satisfied and $0$ otherwise.
-In clearer terms, $B_t$ is defined as the maximum percentile rank for which the corresponding quantile is still below the true value, if the true value is smaller than the median of the predictive distribution.
-If the true value is above the median of the predictive distribution, then $B_t$ is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
-If the true value is exactly the median, both terms cancel out and $B_t$ is zero.
-For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
-For quantile forecasts, an alternative approach is to look at the over- and underprediction components of the weighted interval score.
-These however, capture the bias in absolute terms, while the above proposed metric captures a tendency to over- and underpredict that is less sensitive to outliers.
-
-I ASSUME ALL OF THIS IS A BIT TOO LONG?
-
-Another way to look at calibration[precisely: probabilistic calibration in @gneitingProbabilisticForecastsCalibration2007] is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
-This is most easily understood in the context of quantile forecasts, but can in principle be transferred to continuous and integer forecasts as well.
-
-To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution.
-If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles.
-We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
-
-To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
-
-A third way to look at calibration is to look a the probability integral tranform.
-As explained previously, the CDF of predictice distribution $P_t$ should ideally be equal to the CDF of the true unknown distribution $F_t$ that generated the observed value $x_t$.
-In order to assess whether there are substantial deviations between the two, @dawidPresentPositionPotential1984 suggested to transform the observed values using the probability integral transform (PIT).
-Agreement between the forecasts and the observed values can then be examined by observing whether or not the transformed values follow a uniform distribution.
-The PIT is given by
-$$u_t = P_t (x_t),$$
-where $u_t$ is the transformed variable and $P_t(x_t)$ is the predictive distribution evaluated at the true observed value $x_t$.
-If $P_t = F_t$ at all times $t$, then $u_t, t = 1 \dots T$ follows a uniform distribution (for a proof see e.g.
-@angusProbabilityIntegralTransform1994).
-
-In the case of discrete outcomes, the PIT is no longer uniform even when forecasts are ideal. As @funkAssessingPerformanceRealtime2019 suggest, we use a randomised PIT instead by redefining 
-$$u_t = P_t(x_t) + vt \cdot (P_t(x_t) - P_t(x_t - 1) ),$$
-where $x_t$ is again the observed value at time $t$, $P_t()$ is the CDF of the predictive distribution function, $P_t (-1) = 0$ by definition, and $v_t$ is a standard uniform variable independent of $x_t$.
-If $P_t$ is equal to the true data-generating distribution function, then $u_t$ is standard uniform.
-@czadoPredictiveModelAssessment2009 also propose a non-randomised version of the PIT for count data that could be used alternatively.
-
-One can then plot a histogram of $u_t$ values to look for deviations from uniformity.
-U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
-Biased predictions will usually result in a triangle-shaped histogram.
-
-\subsubsection{Assessing sharpness}
-
-Sharpness is the second property central to model evaluation.
-The ability to produce narrow forecasts is a quality of the forecasts only and does not depend on the observations.
-Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
-Again we need to take slightly different approaches for continuous, integer and quantile forecasts.
-For continuous and integer forecasts we follow the suggestion from @funkAssessingPerformanceRealtime2019.
-For quantile forecasts we propose a novel metric.
-
-For continuous and integer forecasts, @funkAssessingPerformanceRealtime2019 suggest to measure sharpness as the normalised median absolute deviation about the median (MADN), i.e. 
-$$ S_t(P_t) = \frac{1}{0.675} \cdot \text{median}(|y - \text{median(y)}|), $$ 
-where $y$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant that ensures that sharpness will equal the standard deviation of the predictive distribution if $P_t$ is the CDF of a normal distribution.
-
-MAKE THE FOLLOWING PARAGRAPH SHORTER AND JUST SAY WE USE THE SHARPNESS COMPONENT OF THE WIS.
-
-For quantile forecasts, we propose to measure sharpness as a weighted mean of the width of the interval ranges.
-Let $Q_t$ be a set of predicted quantiles for a true $x_t$ at time $t$.
-This set of quantiles is assumed to be symmetric, such that there exist $K$ corresponding pairs of elements $q_{t, \frac{\alpha}{2}}$ and $q_{t, 1-\frac{\alpha}{2}}$.
-These $K$ corresponding pairs of quantiles cover a $(1 - \alpha) \cdot 100$ prediction interval.
-We can, accordingly, also denote $q_{t, \frac{\alpha}{2}}$ as $l_t$ the lower bound of the prediction interval at time $t$ and $q_{t, 1-\frac{\alpha}{2}}$ as $u_t$, the upper bound.
-We measure the sharpness of a quantile forecast at time $t$ as
-\begin{align*}
-\text{sharpness}_t &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (q_{t, 1 - \frac{\alpha}{2}} - q_{t, \frac{\alpha}{2}}) \\
-                   &= \frac{1}{K} \sum_{\alpha} \frac{\alpha}{2} (u_t - l_t).
-\end{align*}
-Weighting the width of different intervals with $\frac{\alpha}{2}$ ensures that the score does not grow indefinitely for very large prediction intervals, and correspondingly, very small $\alpha$.
-We also argue that this sharpness metric for quantile forecasts is a natural choice for quantile forecasts as it corresponds to the sharpness component of the Weighted Interval Score.
-
-\subsection{Pairwise comparisons}
-
-If what we care about is to determine which model performs best, pairwise comparisons between models are a suitable approach [CITATION].
-In turn, each pair of models is evaluated based on the targets that that both models have predicted.
-The mean score by one model is divided by the mean score of the other model to obtain a measure of relative performance.
-Relative skill scores [citation (not sure there is something published, preprint fron Estee Cramer and Nick Reich)] for every model can then be obtained by taking the geometric mean of all pairwise ratios calculated for each model (omitting comparisons where there is no overlapping set of forecasts).
-ADD FORMULAS.
-A skill score smaller than 1 indicates that a model is performing better than the average model.
-One can also scale the skill by a reference baseline.
-A scaled relative skill smaller than one then means that the model in question performed better than the baseline.
-
-By applying a permutation test or non-parametric test like the Wilcoxon rank sum test we can also obtain p-values that help determine whether or not models perform significantly differently.
-THERE IS AN ISSUE WITH CORRELATION: THE TEST WOULD TREAT ALL OBSERVATIONS AS INDEPENDENT, WHICH THEY ARENT'T, SO I WOULD ASSUME THAT IT IS TOO LIBERAL.
-CURRENTLY, NO ADJUSTMENTS ARE IMPLEMENTED, I'M NOT REALLY SURE WHETHER TO DISCUSS THIS HERE OR NOT.
-
-% ==============================================================================
-\section{Evaluating UK short-term forecasts}
-\subsection{The data}
-
-To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
-Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
-The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
-Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
-The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL].
-
-- timing (weekly?) and number of forecast dates
-- the four targets
-- the models. 
-- is the set complete?
-
-We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> # install and load data pacakge from external repository
-R> # remotes::install_github("sbfnk/covid19.forecasts.uk")
-R> 
-R> # load packages
-R> library(covid19.forecasts.uk)
-R> library(dplyr)
-R> library(scoringutils)
-R> # load truth data
-R> data(covid_uk_data)
-R> # head(covid_uk_data)
-R> 
-R> # load forecasts
-R> data(uk_forecasts)
-R> # head(uk_forecasts)
-\end{Sinput}
-\end{Schunk}
-% 
-Let us take a first look a the data:
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> glimpse(covid_uk_data)
-\end{Sinput}
-\begin{Soutput}
-Rows: 4,174
-Columns: 6
-$ geography  <fct> London, London, London, London, London, London, …
-$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospit…
-$ value_desc <fct> Hospital admissions, Hospital admissions, Hospit…
-$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, …
-$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23,…
-$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546,…
-\end{Soutput}
-\begin{Sinput}
-R> glimpse(uk_forecasts)
-\end{Sinput}
-\begin{Soutput}
-Rows: 1,254,513
-Columns: 8
-$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, …
-$ geography     <chr> "East of England", "East of England", "East o…
-$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hos…
-$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-…
-$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-…
-$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.0…
-$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, …
-$ value_desc    <fct> Hospital admissions, Hospital admissions, Hos…
-\end{Soutput}
-\end{Schunk}
-% 
-To bring the forecasts into the format needed for the evaluation, some minor changes need to be made to the data.
-The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
-While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public.
-The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way.
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> uk_forecasts <- rename(uk_forecasts, prediction = value)
-R> covid_uk_data <-rename(covid_uk_data, true_value = value)
-R> combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
-\end{Sinput}
-\end{Schunk}
-% 
-Before we start with scoring the forecasts, it makes sense to start the evaluation process by visualising the data.
-To get a feeling for how complete the data set is, we can run the following code to obtain a heatmap with the number of available forecasts:
-% 
-\begin{figure}[h]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> show_avail_forecasts(combined, 
-+                       x = "creation_date", 
-+                       show_numbers = FALSE,
-+                       legend_position = "bottom",
-+                       facet_formula = ~ value_desc)
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
-\end{figure}
-%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
-% 
-Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
-By default, the function treats a set of different quantiles or samples as one forecast.
-However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.
-
-Forecasts can be visualised using the \fct{plot\_predictions} function.
-The function accepts either a single combined data set or two separate truth data sets that can be merged together.
-Forecasts and observed values can be filtered independently, to show for example only predictions from a certain model or a specific number of weeks of true data before the forecast.
-Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
-To obtain, for example, a specific forecast from a model we are interest in, we can call:
-% 
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
-R> plot_predictions(truth = covid_uk_data,
-+                   forecasts = uk_forecasts,
-+                   filter_both = list(paste("geography %in%", locations, collapse = "")),
-+                   filter_truth = list('value_date <= "2020-06-22"',
-+                                       'value_date > "2020-06-01"'), 
-+                   filter_forecasts = list('model == "SIRCOVID"', 
-+                                           'creation_date == "2020-06-22"'),
-+                   x = "value_date",
-+                   facet_formula = geography ~ value_type) + 
-+    ggplot2::theme(legend.position = "bottom")
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
-\end{figure}
-% 
-The output is shown in Figure \ref{fig:forecast-visualisation}.
-
-\subsection[Scoring forecasts with eval_forecasts]{Scoring forecasts with \fct{eval\_forecasts}}
-
-A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
-This requires a \code{data.frame} or similar which has at least a column called "prediction" and one called "true\_value".
-Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
-Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
-We will not discuss all possible input formats here, but instead refer to the example data for each format that is provided with the package.
-Where possible, \pkg{scoringutils} also provides functionality to transform between various formats, e.g.
-from a sample based format to a quantile format.
-The \fct{eval\_forecasts} function automatically recognises the prediction type and input format, applies the appropriate scoring metrics and aggregates results as desired by the user.
-Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
-
-As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
-% 
-\begin{Schunk}
-\begin{Sinput}
-R> scores <- eval_forecasts(combined, 
-+                           summarise_by = c("model", "value_desc"))
-R> glimpse(scores)
-\end{Sinput}
-\end{Schunk}
-% 
-If a more detailed analysis is desired, the level of aggregation can of course be changed to show for example separate scores for the different locations as well.
-This can be achieved using the \code{summarise_by} argument.
-To additionally stratify by location, we could specify \code{summarise_by = c("mode", "value_type", "geography")}.
-If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
-This can be useful if we, for example, want to analyse what proportion of true values are covered by certain interval ranges, or if we want to analyse the accuracy of the tails of the forecasts.
-When aggregating, \fct{eval\_forecasts} takes the mean according to the group defined in \code{summarise_by}.
-In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
-In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.
-
-The user must, however, still exercise some caution when aggregating scores, as many of the metrics are absolute and scale with the magnitude of the quantity to forecast.
-Looking at one score per model (i.e.
-specifying \code{summarise_by = c("model")}) may not be so useful in this instance, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
-
-In the above example, we did not have to explicitly specify the \code{by} argument, but this may be necessary if additional columns are present in the data that do not indicate a grouping of forecasts.
-The \code{by} argument must then be used to denote the unit of a single forecast.
-In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")}.
-Quantiles should not be included, as several quantiles make up one forecast (and similarly for samples).
-If we had additional columns that do not serve to group forecasts (like for example the number of inhabitants in a certain location over time), these should also not be included.
-By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
-
-\subsection{Pairwise comparisons}
-
-Pairwise comparisons between models [CITATION] can be obtained in two different ways.
-First, relative skill scores based on pairwise comparisons are by default returned from \fct{eval\_forecasts}.
-These will be computed separately for the categories defined in the \code{summarise_by} argument (excluding the category 'model').
-Alternatively, a set of scores can be post-processed using the separate function \fct{pairwise\_comparison}.
-This approach is to be used for visualisation and if p-values for the pairwise comparisons are needed, as those are not returned from \fct{eval\_forecasts}.
-Usually, one would compute scores without specifying a \code{summarise_by} argument, but sometimes it may be sensible to average over certain scores, for example for predictions generated at a certain date.
-This allows to reduce the correlation between observations that enter the computation of p-values, which in turn makes the test less liberal.
-Using the function \fct{plot\_pairwise\_comparison} we can visualise the mean score ratios between all models as well as the
-
-\begin{Schunk}
-\begin{Sinput}
-R> # unsummarised scores
-R> unsummarised_scores <- eval_forecasts(combined)
-R> pairwise <- pairwise_comparison(unsummarised_scores, 
-+                                  summarise_by = "value_desc")
-\end{Sinput}
-\end{Schunk}
-
-The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
-
-\section{Visualisation and interpretation of evaluation results}
-
-\subsection{Visualising aggregate scores and rankings}
-A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
-
-\begin{figure}[h]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> score_table(scores, y = "model", facet_formula = ~ value_desc)
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
-\end{figure}
-
-The most informative metric in terms of model ranking is the relative\_skill.
-However, interpretation is not always straightforward and has to be done carefully.
-We can see that performance varied quite a bit across different metrics, where some models did well on one target, but poorly on another.
-Especially the Exponential growth/decline model stands out as it received the lowest relative skill score for hospital admissions, but the highest for the total number of beds occupied.
-Looking back at Figure \ref{fig:avail-forecasts}, we see that the model has only submitted very few forecasts over all.
-It may therefore be sensible to require all models to have submitted forecasts for at least 50\% of all forecast targets in order to enter the pairwise comparisons.
-For similar reasons, the interval score may be deceiving if looked at in isolation.
-As can be seen, the DetSEIRwithNB MLE model received a lower relative skill score, but a higher interval score than the DetSEIRwithNB MCMC model.
-This, again, can be explained by the fact that they forecasted different targets.
-The interval score, as an absolute metric, is highly influenced by the absolute value of the quantity that is forecasted.
-For the same reason, one should be careful when summarising interval scores from different locations or forecast targets, as the average score will be dominated by outliers as well as differences in the absolute level.
-Assuming a large enough set of available overlapping forecasts, the relative skill score is more robust.
-It therefore is reasonable to assume that the DetSEIRwithNB MLE forecasted quantities with a higher absolute value, but tended to perform worse than the DetSEIRwithNB MCMC model as far as we can tell based on the set of all pariwise comparisons.
-This can be confirmed for the direct comparison between the two by looking at the mean score ratios from the pairwise comparisons.
-These can be obtained by calling
-
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise) +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
-If a tile is blue, then the model on the y-axis performed better.
-If it is red, the model on the x-axis performed better in direct comparison.
-}
-\end{figure}
-
-we can also look at p-values in Figure \ref{fig:pairwise-comparison-pval} PROBABLY REMOVE THAT FROM THE PAPER
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise, type = "pval") +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:pairwise-comparison-pval} XXXX}
-\end{figure}
-
-In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
-
-\subsection{Visual model diagnostics}
-
-For forecasts in an interval format, looking at the components of the weighted interval score separately is a natural next step.
-We can see in Figure \ref{fig:wis-components} that the majority of penalties come from over-and underprediction, instead of the sharpness component.
-We also see that most models tended to either over- or underpredict actual numbers.
-
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> wis_components(scores, 
-+                 facet_formula = ~ value_desc, 
-+                 scales = "free_x") + 
-+    ggplot2::coord_flip() + 
-+    ggplot2::theme(legend.position = "bottom")
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:wis-components} p-values and ratios together.}
-\end{figure}
-
-We can have a closer look at calibration using the functions \fct{interval\_coverage} and \fct{quantile\_coverage}.
-The interval coverage plot shows the proportion of all true values that fall within all the different prediction intervals.
-This gives a visual impression of probabilistic calibration [GNEITING].
-Ideally, $x$ percent of true values should be covered by the $x$\%-prediction intervals, resulting in a 45° line.
-Areas shaded in green indicate that the model is covering more true values than it actually should, while areas in white indicate that the model fails to cover the desired proportion of true values with its prediction intervals.
-The majority of the models were too confident in their predictions, while some showed showed good calibration.
-The quantile coverage plot shows the proportion of all true values below certain predictive quantiles.
-While this plot is slightly harder to interpret, it also includes information about bias as and allows to separate the lower and upper boundaries of the prediction intervals.
-We can see, for example, that the Exponential growth/decline model was consistently biased downwards.
-Figure \ref{fig:coverage}
-
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> cov_scores <- eval_forecasts(combined, 
-+                               summarise_by = c("model", 
-+                                                "range", "quantile"))
-R> scoringutils::interval_coverage(cov_scores)
-R> scoringutils::quantile_coverage(cov_scores)
-\end{Sinput}
-\end{Schunk}
-
-\begin{Schunk}
-\begin{Sinput}
-R> cov_scores <- eval_forecasts(combined, 
-+                               summarise_by = c("model", 
-+                                                "range", "quantile"))
-R> scoringutils::quantile_coverage(cov_scores)
-\end{Sinput}
-\end{Schunk}
-
-\caption{\label{fig:coverage} quantile and interval coverage}
-\end{figure}
-
-DO I WANT TO INCLUDE PIT PLOTS AS WELL? I GUESS? NEED TO LOOK AT THE IMPLEMENTATION FOR QUANTILE FORECASTS
-
-% \begin{figure}[h]
-% \centering
-% <<
-% pit-plots, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
-% # out <- eval_forecasts(combined, 
-% #                       summarise_by = c("model"), 
-% #                       pit_plots = TRUE)
-% 
-% @
-% \caption{\label{fig:calibration-pit} pit plots}
-% \end{figure}
-
-Look at e.g. bias by location? Figure \ref{fig:bias-heatmap}
-
-\begin{figure}[h]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> scores <- eval_forecasts(combined, 
-+                           summarise_by = c("model",
-+                                            "value_desc", 
-+                                            "geography"), 
-+                           compute_relative_skill = FALSE)
-R> scoringutils::score_heatmap(scores, metric = "bias", 
-+                              x = "geography", facet_formula = ~ value_desc, 
-+                              scale = "free_x")
-\end{Sinput}
-\end{Schunk}
-\caption{\label{fig:bias-heatmap} bias by location}
-\end{figure}
-
-WHAT IS NEEDED HERE IS A BIT OF THINKING WITH REGARDS TO WHAT VISUALISATION I WANT TO SHOW AND IN HOW MUCH DETAIL I WANT TO ANALYSE THE MODELS.
-
-\begin{CodeInput}
-.
-\end{CodeInput}
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-
-\section{Summary and discussion} \label{sec:summary}
-
-COMING SOON.
-
-% The results in this paper were obtained using
-% \proglang{R}~4.0.3 with the
-% \pkg{MASS}~7.3.53 package. \proglang{R} itself
-% and all packages used are available from the Comprehensive
-% \proglang{R} Archive Network (CRAN) at
-% \url{https://CRAN.R-project.org/}.
-
-\section*{Acknowledgments}
-
-% All acknowledgments (note the AE spelling) should be collected in this
-% unnumbered section before the references. It may contain the usual information
-% about funding and feedback from colleagues/reviewers/etc. Furthermore,
-% information such as relative contributions of the authors may be added here
-% (if any).
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
-\bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-\newpage
-
-\begin{appendix}
-
-\section{Appendix section} \label{app:technical}
-
-\end{appendix}
-
-%% -----------------------------------------------------------------------------
+\endfoot
+\bottomrule
+\endlastfoot
+\cellcolor{gray!6}{CRPS (Continuous) ranked probability score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
+\addlinespace
+Log score & continuous, (integer not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
+\addlinespace
+\cellcolor{gray!6}{WIS (Weighted) interval score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
+\addlinespace
+DSS Dawid-Sebastiani score & continuous, integer & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
+\addlinespace
+\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
+\addlinespace
+Interval coverage & continuous, integer & interval forecasts (needs matching quantiles) & measure for calibration\\
+\addlinespace
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
+\addlinespace
+Probability integral transform (PIT) & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
+\addlinespace
+\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
+\addlinespace
+Bias & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
+\addlinespace
+\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{compares performance of two models}\\
+\addlinespace
+Relative skill & depends on scored & depends on score & Ranks models based on pairwise comparisons\\*
+\end{longtable}
 

### 2021-03-30 tables/scores-table.tex
update scores sections of the paper

@@
-% !TeX root = ../article.Rnw
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{sidewaystable}[]
+\documentclass{article}
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{float}
+\usepackage{longtable}
+\usepackage{Sweave}
+\begin{document}
 
-% \begin{adjustbox}{angle=90}
+\begin{landscape}
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{16.5cm}}
+\toprule
+Metric & Explanation\\
+\midrule
+\endfirsthead
+\multicolumn{2}{@{}l}{\textit{(continued)}}\\
+\toprule
+Metric & Explanation\\
+\midrule
+\endhead
 
-\centering
-\begin{tabularx}{\linewidth}{|X|X|X|X|}
+\endfoot
+\bottomrule
+\endlastfoot
+CRPS (Continuous) ranked probability score & The crps is a proper scoring rule that generalises the absolute error to probabilistic forecasts.
+It measures the 'distance' of the predictive distribution to the observed data-generating distribution.
+The CRPS is given as
+  $$\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - 1(x \geq y) \right)^2 dx,$$
+  where y is the true observed value and F the CDF of predictive distribution. Often An alternative representation is used:
+  $$ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|,$$ where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
+In this represenation we can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
+  For integer-valued forecasts, the RPS is given as 
+  $$ \text{RPS}(F, y) = \sum_{x = 0}^\infty (F(x) - 1(x \geq y))^2. $$
 
-\hline
-Metric                                                                  & Formula                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         & Application and Explanation                                                                                                                                                                                                                                                                                                                                                                                                                    & Caveats                                                                                                                                                        \\ \hline
-WIS (Weighted) interval score                                           & For a single interval, the score is computed as $IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u)$, where 1() is the indicator function, y is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantile of $F$, i.e.
+\cellcolor{gray!6}{\textbf{Usage and caveats} Smaller values are better.
+The crps is a good choice for most practical purposes that involve decision making, as it takes the entire predictive distribution into account.
+If two forecasters assign the same probability to the true event $y$, then the forecaster who assigned high probability to events far away from $y$ will still get a worse score.
+The crps (in contrast to the log score) can at times be quite lenient towards extreme mispredictions.
+Also, due to it's similarity to the absolute error, the level of scores depend a lot on the absolute value of what is predicted, which makes it hard to compare scores of forecasts for quantities that are orders of magnitude apart.}\\
+\addlinespace \addlinespace
+Log score & The Log score is a proper scoring rule that is simply compuated as the log of the predictive density evaluated at the true observed value. It is given as 
+  $$ \text{log score} = \log f(y), $$ 
+  where $f$ is the predictive density function and y is the true value. For integer-valued forecasts, the log score can be computed as 
+  $$ \text{log score} = \log p_y, $$
+  where $p_y$ is the probability assigned to outcome p by the forecast F.
+
+\textbf{Usage and caveats}: Larger values are better, but sometimes the sign is reversed.
+The log score is ensitive to outliers, as individual negative log score contributions quickly can become very large if the event falls in the tails of the predictive distribution, where $f(y)$ (or $p_y$) is close to zero.
+Whether or not that is desirable depends ont the application.
+In scoringutils, the log score cannot be used for integer-valued forecasts, as the implementation requires a predictive density.
+In contrast to the crps, the log score is a local scoring rule: it's value only depends only on the probability that was assigned to the actual outcome.
+This property is desirable for inferential purposes, for example in a Bayesian context (@winklerScoringRulesEvaluation1996).
+In settings where forecasts inform decision making, it may be more appropriate to score forecasts based on the entire predictive distribution.\\
+\addlinespace \addlinespace
+WIS (Weighted) interval score & The (weighted) interval score is a proper scoring rule for quantile forecasts that converges to the crps for an increasing number of intervals.
+The score can be decomposed into a sharpness (uncertainty) component and penalties for over- and underprediction.
+For a single interval, the score is computed as
+  $$IS_\alpha(F,y) = (u-l) + \frac{2}{\alpha} \cdot (l-y) \cdot 1(y \leq l) + \frac{2}{\alpha} \cdot (y-u) \cdot 1(y \geq u), $$ 
+  where $1()$ is the indicator function, $y$ is the true value, and $l$ and $u$ are the $\frac{\alpha}{2}$ and $1 - \frac{\alpha}{2}$ quantiles of the predictive distribution $F$, i.e.
 the lower and upper bound of a single prediction interval.
-For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum, $WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y))$.
-$w_k$ is a weight for every interval.
-Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$                                                                                                                                                                                                                   & proper scoring rule for quantile forecasts.
-converges to crps for increasing number of interval.
-The score can be decomposed into a sharpness contribution and penalties for over- and underprediction.
-& The wis is based on measures of absolute error.
+For a set of $K$ prediction intervals and the median $m$, the score is computed as a weighted sum,
+  $$WIS = \frac{1}{K + 0.5} \cdot (w_0 \cdot |y - m| + \sum_{k = 1}^{K} w_k \cdot IS_{\alpha}(F, y)),$$ 
+  where $w_k$ is a weight for every interval. Usually, $w_k = \frac{\alpha_k}{2}$ and $w_0 = 0.5$.
+
+\textbf{Usage and caveats}:
+\cellcolor{gray!6}{  Smaller scores are better.
+Applicable to all quantile forecasts, takes the entire predictive distribution into account.
+Just as the crps, the wis is based on measures of absolute error.
 When averaging across multiple targets, it will therefore be dominated by targets with higher absolute values.
-\\
-CRPS (Continuous) ranked probability score                              & $\text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \mathbb{1}(x \geq y) \right)^2 dx$, where y is the true observed value an F the predictive distribution.
-Another represenation allows for easier computation using samples.
-Take $ \text{CRPS}(F, y) = \frac{1}{2} \mathbb{E}_{F} |X - X'| - \mathbb{E}_P |X - y|$, where $X$ and $X'$ are independent realisations from the predictive distributions $F$ with finite first moment and $y$ is the true value.
-We can simply replace $X$ and $X'$ by samples sum over all possible combinations to obtain the CRPS.
-For integer-valued forecasts, the RPS is given as $ \text{RPS}(F, y) = \sum_{x = 0}^\infty (F(x) - \mathbb{1} (x \geq y))^2.
-$                                                                                                                                                                                                     & proper scoring rule for continuous and integer forecasts.
-Implemented for sample-based forecasts.
-& Also sensitive to the absolute value of the predicted quantity                                                                                                 \\
-\begin{tabular}[c]{@{}l@{}}DSS\\ \\ Dawid-Sebastiani score\end{tabular} & 2.257546e+02                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    & proper scoring rule for continuous and integer forecasts.
-The dss has a slightly simpler formula that only relies on the first moments of the predictive distribution.
-If in doubt, we would recommend the crps, but the difference should not be large in practice.
-&                                                                                                                                                                \\
-log score                                                               & The Log score is given by $ \text{log score} = \log f(y)$, where $f$ is the predictive density function and y is the true value.
-& \begin{tabular}[c]{@{}l@{}}The log score is a proper scoring rule best suited for continuous forecasts.
-While in principle applicable to integer valued forecasts, its computation requires a density function which is in practice difficult for integer valued forecasts.
-\\ One concern with the log score is that it can assign extreme scores if the predictive distribution gives little probability to the observed event.\end{tabular} & Sensitive to outliers.
-Individual log score contributions can go to infinity                                                                                   \\
-Brier score                                                             & $\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (\text{prediction}_n - \text{outcome}_n)^2$, where prediction                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                & Proper scoring rule for binary forecasts                                                                                                                                                                                                                                                                                                                                                                                                       &                                                                                                                                                                \\
-interval coverage                                                       & (Interval) coverage for a single prediction interval can be calculated as $IC_\alpha = \text{nominal coverage} - \text{actual empirical coverage}$, where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by the $1 - \alpha$ prediction intervals.
-Interval coverage can then be aggregated over all interval levels: $\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} IC_{\alpha_k}$                                                                                                                                                                                                                                                                                                                                                                                                                                                         &                                                                                                                                                                                                                                                                                                                                                                                                                                                &                                                                                                                                                                \\
+The decomposition into sharpness, over- and underprediction make it easy to interpret scores and use them for model improvement.}\\
+\addlinespace \addlinespace
+DSS Dawid-Sebastiani score & The Dawid-Sebastiani-Score is a proper scoring rule proposed by Gneiting and Raftery in [@gneitingStrictlyProperScoring2007] that only relies on the first moments of the predictive distribution and is therefore easy to compute.
+It is given as
 
-\end{tabularx}
+$$\text{dss}(F, y) = \left( \frac{y - \mu}{\sigma} \right)^2 + 2 \cdot \log \sigma,$$
+where $F$ is the predictive distribution with mean $\mu$ and standard deviation $\sigma$ and $y$ is the true observed value.
 
-% \end{adjustbox}
+textbf{Usage and caveats} The dss is applicable to continuous and integer forecasts and easy to compute.
+Apart from the ease of computation we see little advantage in using it.
+MAYBE SCRATCH IT ALTOGETHER FROM THE PAPER?\\
+\addlinespace \addlinespace
+Brier score & Proper scoring rule for binary forecasts. The Brier score is computed as 
+$$\text{Brier Score} = \frac{1}{N} \sum_{n = 1}^{N} (f_n - y_n), $$
+where $f_n$, with $n = 1, \dots, N$ are the predicted probablities that the corresponding events, $y_n \in (0, 1)$ will be equal to one.)
 
-\end{sidewaystable}
+\cellcolor{gray!6}{\textbf{Usage}: Applicable to all binary forecasts.}\\
+\addlinespace \addlinespace
+Interval coverage & Interval coverage measures the proportion of observed values that fall in a given prediction interval range.
+Interval coverage for a single prediction interval range can be calculated as $$IC_{\alpha} = \text{nominal coverage} - \text{empirical coverage},$$
+where nominal coverage is $1 - \alpha$ and empirical coverage is the percentage of true values actually covered by all $1 - \alpha$ prediction intervals.
+
+To summarise interval coverage over different over multiple interval ranges, we can compute coverage deviation defined as the mean interval coverage over all $K$ interval ranges $\alpha_k$ with $k = 1, \dots, K$:
+$$\text{Coverage deviation} = \frac{1}{K} \sum_{k = 1}^{K} \text{IC}_{\alpha_k}$$
+
+\textbf{Usage}: Interval coverage for a set of chosen intervals, (e.g.
+50\% and 90\%) gives a good indication of marginal calibration and is easy to interpret.
+Reporting coverage deviation has the advantage of summarising calibration in a single number, but loses some of the nuance.\\
+\addlinespace \addlinespace
+Quantile coverage & Quantile coverage for a given quantile level is the percentage of true values smaller than the predictions corresponding to that quantile level.
+
+\cellcolor{gray!6}{\textbf{Usage}: Quantile coverage is similar to interval coverage, but conveys more information.
+For example, it allows us to look at the 5\% and 95\% quantile separately, instead of jointly at the 90\% prediction interval).
+This helps to diagnose whether it is the upper or lower end of a prediction interval that is causing problems.
+Plots of quantile coverage are conceptually very similar to PIT histograms.}\\
+\addlinespace \addlinespace
+Probability integral transform (PIT) & The probability integral transform (PIT) @dawidPresentPositionPotential1984 represents a succinct way to visualise deviations between the predictive distribution $F$ and the true data-generating distribution $G$.
+The idea is to transform the observed values such that agreement between forecasts and data can then be examined by observing whether or not the transformed values follow a uniform distribution.
+The PIT is given by
+$$u = F (y),$$
+where $u$ is the transformed variable and $F(y)$ is the predictive distribution $F$ evaluated at the true observed value $y$.
+If $F = G$, then $u$ follows a uniform distribution (for a proof see e.g.
+@angusProbabilityIntegralTransform1994).
+
+For integer outcomes, the PIT is no longer uniform even when forecasts are ideal. Instead, a randomised PIT (@funkAssessingPerformanceRealtime2019) can be used: 
+$$u = P(y) + v \cdot (P(y) - P(y - 1) ),$$
+where $y$ is again the observed value $P()$ is the cumulative probability assigned to all values smaller or equal to $y$ (where $P(-1) = 0$ by definition, and $v$ is a standard uniform variable independent of $y$.
+If $P$ is equal to the true data-generating distribution function, then $u$ is standard uniform.
+also propose a non-randomised version of the PIT for count data that could be used alternatively.
+
+\textbf{Usage}: 
+One can plot a histogram of $u$ values to look for deviations from uniformity.
+U-shaped histograms often result from predictions that are too narrow, while hump-shaped histograms indicate that predictions may be too wide.
+Biased predictions will usually result in a triangle-shaped histogram.
+One can also test for deviations from normality, using for example an Anderson-Darling test.
+This, however, proves to be overly strict in practice and even slight deviations from perfect calibration are punished in a way that makes it very hard to compare models at all.
+I HAVE MADE A SIMULATION TO TEST THAT IN MY MASTER THESIS AND COULD ADD IT TO THE PAPER?\\
+\addlinespace \addlinespace
+Sharpness & Sharpness is the ability to produce narrow forecasts and is a feature of the forecasts only and does not depend on the observations.
+Sharpness is therefore only of interest conditional on calibration: a very precise forecast is not useful if it is clearly wrong.
+
+As suggested by @funkAssessingPerformanceRealtime2019, we measure sharpness for continuous and integer forecasts represented by predictive samples as the normalised median absolute deviation about the median (MADN) ), i.e.
+$$ S(F) = \frac{1}{0.675} \cdot \text{median}(|x - \text{median(x)}|), $$ 
+where $x$ is the vector of all predictive samples and $\frac{1}{0.675}$ is a normalising constant.
+If the predictive distribution $F$ is the CDF of a normal distribution, then sharpness will equal the standard deviation of $F$.
+
+\cellcolor{gray!6}{For quantile forecasts we can directly use the sharpness component of the weighted interval score.
+Sharpness is then simply the weighted mean of the widths of the central prediction intervals.}\\
+\addlinespace \addlinespace
+Bias & Bias is a measure of the tendency of a forecaster to over- or underpredict. For continuous forecasts, bias is given as 
+$$B(F, y) = 1 - 2 \cdot (F (y)), $$ 
+where $F$ is the CDF of the predictive distribution and $y$ is the observed value.
+
+For integer-valued forecasts, bias can be calculated as 
+$$B(P, y) = 1 - (P(y) + P(y + 1)), $$ 
+where $P(y)$ is the cumulative probability assigned to all outcomes smaller or equal to $y$.
+
+For quantile forecasts, Bias can be calculated as the maximum percentile rank for which the prediction is smaller than $y$, if the true value is smaller than the median of the predictive distribution.
+If the true value is above the median of the predictive distribution, then bias is the minimum percentile rank for which the corresponding quantile is still larger than the true value.
+If the true value is exactly the median, bias is zero.
+For a large enough number of quantiles, the percentile rank will equal the proportion of predictive samples below the observed true value, and this metric coincides with the one for continuous forecasts.
+
+\textbf{Usage}:
+In contrast to the over- and underprediction penalties of the interval score it is bound between 0 and 1 and represents the tendency of forecasts to be biased rather than the absolute amount of over- and underprediction.
+It is therefore a more robust measurement, but harder to interpet.
+It largely depends on the application whether one is more interested in the tendency to be biased or in the absolute value of over- and underpredictions.\\
+\addlinespace \addlinespace
+Mean score ratio & The mean score ratio is used to compare two models on the overlapping set of forecast targets for which both models have made a prediction.
+The mean score ratio is calculated as the mean score achieved by the first model over the mean score achieved by the second model.
+More precisely, for two models $i, j$, we determine the set of overlapping forecasts, denoted by $\mathcal{A}_{ij}$ and compute the mean score ratio $\theta_{ij}$ as
+$$\theta_{ij} =\frac{\text{mean score model } i \text{ on } \mathcal{A}_{ij}}{\text{mean score model } j \text{ on } \mathcal{A}_{ij}}.$$
+The mean score ratio can in principle be computed for any arbitrary score.
+
+\textbf{Usage}:
+\cellcolor{gray!6}{Mean scores ratios are usually calculated in the context of pairwise comparisons, where a set of models is compared by looking at mean score ratios of all possible parings.
+Whether smaller or larger values are better depends on the orientation of the original score used}\\
+\addlinespace \addlinespace
+Relative skill & Relative skill scores can be used to obtain a ranking of models based on pairwise comparisons between all models.
+To compute the relative skill $\theta_i$ of model $i$, we take the geometric mean of all [mean score ratios](mean%20score%20ratio.md) that involve model $i$, i.e.
+$$ \theta_{i} = \left(\prod_{m = 1}^M \theta_{im}\right)^{1/M}, $$
+where M is the number of models.
+
+\textbf{Usage and caveats}: 
+Relative skill is a helpful way to obtain a model ranking. Whether smaller or larger values are better depends on the orientation of the original score used. 
+It is in principle relatively robust against biases that arise when models only forecast some of the available targets and is a reasonable way to handle missing forecasts.
+One possible precautionary measure to reduces issues with missing forecasts is to only compare models that have forecasted at least half of all possible targets (this ensures that there is always an overlap between models).
+If there is no overlap between models, the relative skill implicitly estimates how a model would have forecasted on those missing targets.\\*
+\end{longtable}
+\end{landscape}
+
+\end{document}

### 2021-11-22 article.Rnw
update paper

@@
 
+%% packages for landscape
+%% there are probably duplicates
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{float}
+\usepackage{longtable}
+
 %% new custom commands
@@
 library(kableExtra)
+library(scoringutils)
+library(data.table)
+library(kableExtra)
+library(magrittr)
+library(knitr)
 
@@
 knitr::opts_chunk$set(
-  out.width = "0.9\\textwidth"
+  out.width = "0.98\\textwidth"
 )
@@
   Forecasts play an important role in a variety of fields.
-Its role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
-Evaluating forecasts, however, is not trivial in practice.
-Even though scoring methods such as proper scoring rules have been widely studied in the past, there is a lack of software implementation that allows a forecaster to conveniently evaluate their forecasts.
+Their role in informing public policy has attracted increased attention from the general public with the emergence of the Covid-19 pandemic.
+Much theoretical work has been done on the development of proper scoring rules and other scoring metrics that can help evaluate these forecasts.
+Making these theoretical insights available to end users, however, has been less of a focus.
 In this paper we introduce \pkg{scoringutils}, an \proglang{R} package that facilitates automated forecast evaluation.
@@
 \section[Introduction]{Introduction}
-Good forecasts are of great interest to decision makers in various fields like finance (), weather predictions or infectious disease modeling \citep{funkShorttermForecastsInform2020}.
+Good forecasts are of great interest to decision makers in various fields like finance \citep{timmermannForecastingMethodsFinance2018} [CITATION], weather predictions [CITATION] or infectious disease modeling \citep{funkShorttermForecastsInform2020} [CITATIONS].
 An integral part of assessing and improving their usefulness is forecast evaluation.
-For decades, researchers therefore have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
-\cite{bracherEvaluatingEpidemicForecasts2020a}, \cite{funkAssessingPerformanceRealtime2019}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
+For decades, researchers have developed and refined an arsenal of techniques not only to forecast, but also to evaluate these forecasts (see e.g.
+\cite{bracherEvaluatingEpidemicForecasts2021}, \cite{funkAssessingPerformanceRealtime2019}, \cite{gneitingProbabilisticForecastsCalibration2007}, and \cite{gneitingStrictlyProperScoring2007}).
 Yet even with this rich body of research available, implementing a complete forecast evaluation in \proglang{R} is not trivial.
-We therefore present the \pkg{scoringutils} package.
-The goal of the scoringuitls package is to facilitate the evaluation process and to allow even inexperienced users to perform a thorough evaluation of their forecasts.
-In this paper we give a quick introduction of the fundamental ideas behind forecast evaluation, explain the evaluation metrics implemented in \pkg{scoringutils} and present a full example evaluation of Covid-19 related short-term forecasts in the UK \citep{funkShorttermForecastsInform2020}.
 
+Some packages exist that bundle different scoring metrics together, but none offer the user a standalone solution to forecast evaluation.
+The \pkg{scoringRules} package \citep{scoringRules} offers a very extensive collection of proper scoring rules.
+Its exclusive focus on proper scoring rules and the fact that all functions are implemented using vectors and matrices instead of data.frames, however, make it more suitable for experienced users or as a building block in larger application.
+It also lacks features like pairwise comparisons between forecast models and plotting functionality that are invaluable in the evaluation process.
+Other packages like \pkg{Metrics} [https://cran.r-project.org/web/packages/Metrics/Metrics.pdf] \pkg{MLmetrics} [https://cran.r-project.org/web/packages/MLmetrics/MLmetrics.pdf] are geared towards machine learning problems and don't implement the set of metrics and scoring rules desired for forecast evaluation.
+The \pkg{scoringutils} package aims to bring forth a standardised and tested toolkit.
+It offers convenient automated forecast evaluation in a data.table format, but also provides experienced users with a set of reliable lower-level scoring metrics they can build upon in other applications.
+In addition it implements a wide range of flexible plots that are able to cover most day-to-day use cases.
+
+This paper provides an overview of the fundamental ideas behind forecast evaluation, gives a detailed explanation of the evaluation metrics in \pkg{scoringutils} and discusses what to think about when applying them in practice.
+It then presents a case study based on the evaluation of Covid-19 related short-term forecasts in the UK \citep{funkShorttermForecastsInform2020}.
 \subsection{Forecast types and forecast formats}
@@
 For that reason, often quantiles or central prediction intervals are reported instead [citation FORECAST HUBS].
-For binary prediction targets, common in many classification problems, a probabilistic forecasts is represented by the probability that an outcome will come true.
-DO I NEED TO TALK ABOUT MULTINOMIAL CLASSIFICATION PROBLEMS?
+For binary or multinomial prediction targets, common in many classification problems, a probabilistic forecasts is represented by the probability that an outcome will come true.
 Table \ref{tab:forecast-types} summarises the different forecast types and formats.
-The general forecasting paradigm \cite{gneitingProbabilisticForecastsCalibration2007}, however, that guides the evaluation process is the same irrespective of the reporting format, even though specific scoring metrics differ.
+While specific metrics may differ depending on the forecast type or format, the general forecasting paradigm \cite{gneitingProbabilisticForecastsCalibration2007} that guides the evaluation process is the same.
 
@@
 @
-\caption{\label{tab:forecast-types} Ssummary of the different forceast types and forecast formats}
+\caption{\label{tab:forecast-types} Forecasts can be distinguished by whether they are probabilistic in nature, or a point forecast only.
+Depending on the type of the target (discrete, continuous or binary) different representation of the predictive distribution are possible.}
 \end{table}
@@
 
-Any forecaster should aim to minimise the difference between the predictive distribution $F$ and the unknown true data-generating distribution $G$ \citep{gneitingProbabilisticForecastsCalibration2007}.
+Any forecaster should aim to minimise the difference between the (cumulative) predictive distribution $F$ and the unknown true data-generating distribution $G$ \citep{gneitingProbabilisticForecastsCalibration2007}.
 For an ideal forecast, we therefore have
 %
-$$ F = G. $$
+$$ F = G, $$
 %
+where $F$ and $G$ are both cumulative distribution functions.
 As we don't know the true data-generating distribution, we cannot assess the difference between the two distributions directly.
-\cite{gneitingProbabilisticForecastsCalibration2007} instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness.
-Calibration refers to the statistical consistency between the predictive distribution and the observations.
-A well calibrated forecast does not systematically deviate from the observed values.
-For an in-depth discussion of different ways in which a forecast can be miscalibrated, we refer to \cite{gneitingProbabilisticForecastsCalibration2007}.
+\cite{gneitingProbabilisticForecastsCalibration2007} instead suggest to focus on two central aspects of the predictive distribution, calibration and sharpness (illustrated in Figure \ref{fig:forecast-paradigm}.
+Calibration refers to the statistical consistency (i.e.
+absence of systematic deviations) between the predictive distribution and the observations.
 Sharpness is a feature of the forecast only and describes how concentrated the predictive distribution is, i.e.
@@
 The general forecasting paradigm states that we should maximise sharpness of the predictive distribution subject to calibration.
-A model that made very precise forecasts would be at best useless if the forecasts were wrong most of the time.
+A model that made very precise forecasts would at best be useless if the forecasts were wrong most of the time.
 On the other hand, a model may be well calibrated, but not sharp enough to be useful.
 Take a weather forecast that would assign 30 percent rain probability for every single day.
-It may be well calibrated over the course of a year (it would be marginally calibrated according to \cite{gneitingProbabilisticForecastsCalibration2007}), but we would of course prefer a more precise forecast.
-Figure \ref{fig:forecast-paradigm} illustrates the concepts of calibration and sharpness.
+It may be (marginally) calibrated when looking at the average rainfall over the course of a year, but it doesn't give much guidance on a day to day basis.
+\cite{gneitingProbabilisticForecastsCalibration2007} discuss different forms of calibration in more detail.
 
@@
 Table \ref{tab:table-summary-scores} shows an overview of the metrics implemented in \pkg{scoringutils}.
-Table \ref{tab:score-table} in the Appendix gives a thorough explanation of all the metrics.
+Table \ref{tab:score-table} in the Appendix gives a mathematical definition and thorough explanation of all the metrics.
+Figure \ref{fig:calibration-plots} gives an applied example of different scoring metrics and visualisations for simple toy data.
 
-\begin{table}[h!]
-\centering
-<<summary-table, echo=FALSE, results=tex>>=
+% \begin{table}[h!]
+% \centering
+
+<<summary-table, echo=FALSE, eval = FALSE, results=tex>>=
 
@@
 
-library(kableExtra)
-library(magrittr)
-library(knitr)
 data[, 1:4] %>%
@@
                   linesep = c('\\addlinespace')) %>%
-  kableExtra::column_spec(1, width = "2.5cm") %>%
-  kableExtra::column_spec(2, width = "2cm") %>%
-  kableExtra::column_spec(3, width = "4.5cm") %>%
-  kableExtra::column_spec(4, width = "4cm") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))
+  kableExtra::column_spec(1, width = "12.5cm") %>%
+  # kableExtra::column_spec(2, width = "2cm") %>%
+  # kableExtra::column_spec(3, width = "4.5cm") %>%
+  # kableExtra::column_spec(4, width = "4cm") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "repeat_header", "scale_down"), font_size = 2)
 
 @
-\caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
+% \caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
+% \end{table}
+
+\newpage
+\begin{table}[h!]
+\centering
+<<summary-table, echo=FALSE, results=tex>>=
+
+data <- readRDS("tables/scores-table-summary2.rds")
+
+data[, 1:8] %>%
+  kableExtra::kbl(format = "latex", booktabs = TRUE, 
+                  escape = FALSE,
+                  longtable = TRUE,
+                  align = c("lccccccl"),
+                  linesep = c('\\addlinespace')) %>%
+  kableExtra::column_spec(1, width = "2.0 cm") %>%
+  kableExtra::column_spec(2, width = "1.4cm") %>%
+  kableExtra::column_spec(3, width = "1.4cm") %>%
+  kableExtra::column_spec(4, width = "1.2cm") %>%
+  kableExtra::column_spec(5, width = "1.2cm") %>%
+  kableExtra::column_spec(6, width = "1.4cm") %>%
+  kableExtra::column_spec(7, width = "1.2cm") %>%
+  kableExtra::column_spec(8, width = "4.5cm") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "repeat_header, scale_down"), font_size = 10)
+
+@
+\caption{\label{tab:table-summary-scores2} Summary table of scores available in scoringutils}
 \end{table}
 
-\subsection{Proper scoring rules}
-
-Proper scoring rules \citep{gneitingStrictlyProperScoring2007} jointly assess sharpness and calibration and assign a single numeric value to a forecast.
-A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
-This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state their best estimate.
-
-For sample-based forecasts, the (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], and the Dawid-Sebastiani-score (dss) [CITATION] are available in \pkg{scoringutils}.
-These are implemented as wrappers around functions from the \pkg{scoringRules} package (which also has closed-form versions of the scoring rules available).
-They are in principle applicable to continuous as well as integer forecasts.
-The \pkg{scoringRules} implementation of the log score, however, requires a kernel density estimation that may be inappropriate for integer values (see also Table \ref{tab:table-summary-scores}) and is therefore not available for discrete predictions in \pkg{scoringutils}.
-For forecasts in an interval or quantile format, the weighted interval score (wis) \citep{bracherEvaluatingEpidemicForecasts2020a} is available.
-For an increasing number of equally-spaced prediction intervals, the wis converges to the crps and therefore has very similar properties.
-Binary forecasts can be scored using the Brier score (bs) [CITATION].
-
-%sample-based proper scoring rules\usepackage{multirow}
-When scoring forecasts in a sample-based format, the choice is usually between the log score and the crps.
-The dss is much less commonly used.
-It is easier to compute, but apart from that does not have immediate advantages over the former two.
-Crps and log score differ in two important aspects: the first is sensitivity to distance \cite{winklerScoringRulesEvaluation1996}, the second is how harshly far-off predictions are punished.
-
-The crps is a so-called global scoring rule, which means that the entire predictive distribution is taken into account when scoring a single forecast.
-The log score, on the other hand is local.
-The resulting score does not depend on the overall distance between the observed value and the distribution, but only on the probality density assigend to the actual outcome.
-Imagine two forecasters, A and B, who forecast the number of points scored in a basketball game.
-If both forecasters assigned the same probability to the true outcome (100 points), but A assigned higher probability to extreme outcomes far away from the actually observed outcome, then A will receive a worse score than B.
-The log score, in contrast, is a local scoring rule that only scores the probability assigned to the actual outcome and ignores the rest of the predictive distribution.
-Judged by the log score, A and B would receive exactly the same score.
-Sensitivity to distance (taking the entire predictive distribution into account) is arguably an advantage in most settings that involve decision making.
-Forecaster A's prediction that assigns high probability to results far away from the observed value is arguably less useful than B's forecast that assigns higher probability to values closer to it (the probability assigned to the actual outcome being equal for both forecasts).
-The log score is only implicitly sensitive to distance if we assume that values close to the observed value are actually more likely to occur.
-It may, however, be more appropriate for inferential purposes (see \cite{winklerScoringRulesEvaluation1996}).
-
-% stable vs. non-stable
-A second important difference is how forecasts are treated that deviate strongly from the observered outcome.
-The crps can be thought of as a generalisation of the absolute error to a predictive distribution.
-It therefore scales linearly with the distance between forecast distribution and true value.
-The log score, however, is the log of the predictive density evaluated at the observed value.
-It can therfore quickly go to negative infinity if the probability assigned to the observed outcome is close to zero.
-The crps is therefore considered more stable than the log score.
-The behaviour of the dss is in between the two.
-Whether or not harsh punishment of bad predictions is desirable or not depends of course on the setting.
-\cite{bracherEvaluatingEpidemicForecasts2020a} exemplify that in practice there may indeed be substantial differences between how the crps and log score judge the same forecast.
-
-% properties of the WIS + decomposition
-For quantile forecasts the wis is the best available option and can be understood as an approximation to the crps.
-One additional benefit of the wis is that it can easily be decomposed into three additive components: an uncertainty penalty (sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval).
-This can be very helpful in diagnosing model problems.
-It may therefore even be useful to convert samples into quantiles and use the wis instead of the crps to make use of this decomposition for the purpose of model diagnostics.
-% fact that scores scale with the overall level of the prediction
-Both crps and wis, as generalisations of the absolute value, suffer from the fact that overall scores depend on the order of magnitude of the quantity we try to forecast.
-It can therefore be difficult to compare forecasts for very different targets.
-One possibility to address this is the use of pairwise comparisons introduced later.
-In general, this is much less of a problem for the log score and the dss.
-IS IT ONE AT ALL?
-
-\subsection{Evaluating calibration and sharpness independently}
-
-In addition to the proper scoring rules outlined above, \pkg{scoringutils} makes numerous metrics available to evaluate calibration and sharpness independently.
-This is especially helpful for model diagnostics.
-
-\subsubsection{Assessing calibration}
-
-Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g.
-\cite{funkAssessingPerformanceRealtime2019}; \cite{gneitingProbabilisticForecastsCalibration2007}; \cite{gneitingStrictlyProperScoring2007}).
-Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
-
-% Bias
-Bias, i.e.
-systematic over- or underprediction, is a very common form of miscalibration which therefore deserves separate attention.
-The bias metric (with slightly different versions for the various forecast types and formats) captures a general tendency to over- and underpredict that is bound to be between minus one (underpredicton) and one (overprediction), where zero is ideal.
-It is derived by looking at how much of the probability mass of the predictive distribution is below or above the true observed value.
-For quantile forecasts we have second alternative approach available to assess over- and underprediction - by simply looking at the corresponding components of the weighted interval score.
-What is different between the over- and underprediction components and bias as described above is its sensitivity to outliers.
-The former are derived from absolute differences, while the latter is bound and rather captures a general tendency to be biased.
-
-% interval coverage and quantile coverage
-Another way to look at calibration (precisely: probabilistic calibration in \cite{gneitingProbabilisticForecastsCalibration2007}) is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
-This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and integer forecasts as well.
-%
-To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution.
-If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles.
-We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
-%
-To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
-This allows to diagnose issues in the lower and upper tails of the prediction intervals separately.
-A similar way to visualise the same information is a PIT histogram.
-In order to conveniently assess deviations between the predictive distribution and the true data-generating distribution we can transform the observed values using the probability integral transformation (PIT) \citep{dawidPresentPositionPotential1984} (see more details in Table \ref{tab:score-table}).
-If both distributions are equal, the transformed values will follow a uniform distribution.
-A histogram of the transformed values can help to diagnose systematic differences between the predictions and the observed values.
-Figure \ref{fig:calibration-plots} exemplifies the characteristic shape of certain systematic deviations of the predicitve distribution from the true data-generating distribution.
-In the PIT histograms, bias leads to a triangular shape, overdispersion results in a hump shaped form and underdispersion in a U-shape.
-ADD INTERPRETATION FOR QUANTILE AND INTERVAL COVERAGE PLOTS HERE.
+\newpage
 
@@
 \centering
-\includegraphics{plots/calibration-diagnostic-examples.png}
-\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
-\end{figure}
-
 <<example-calibration, echo=FALSE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
@@
 predictions4 <- rnorm(n_truth * n_samples, sd =  0.7)
-predictions5 <- rnorm(n_truth * n_samples, mean = 0.2, sd =  0.7)
+# predictions5 <- rnorm(n_truth * n_samples, mean = 0.2, sd =  0.7)
 
@@
                              prediction = c(predictions1, predictions2, 
-                                            predictions3, predictions4, 
-                                            predictions5), 
+                                            predictions3, predictions4), 
                              sample = 1:n_samples,
-                             model = rep(c("Normal(0,1)", "Normal(0.5, 1)", 
-                                       "Normal(0, 1.4)", "Normal(0, 0.7)", 
-                                       "Normal(0.2, 0.7)"), each = n_truth * n_samples))
+                             model = rep(c("Normal(0, 1)", "Normal(0.5, 1)", 
+                                       "Normal(0, 1.4)", "Normal(0, 0.7)"), 
+                                       each = n_truth * n_samples))
 
@@
                       by = c("model", "id"), 
-                      summarise_by = c("model", "range"), 
+                      summarise_by = c("model"), 
                       pit_plots = TRUE)
@@
 
+scores_table <- dcast(melt(res$scores, id.vars = "model", 
+                           variable.name = "score"), 
+                      score ~ model)
+setcolorder(
+  scores_table, 
+  c("score", "Normal(0, 1)", "Normal(0.5, 1)", "Normal(0, 1.4)", "Normal(0, 0.7)")
+)
+
 (standard_normal | shifted_mean | overdispersion | underdispersion) /
-  (res$pit_plots$`Normal(0,1)` | res$pit_plots$`Normal(0.5, 1)` | res$pit_plots$`Normal(0, 1.4)` | res$pit_plots$`Normal(0, 0.7)`)
+  (res$pit_plots$`Normal(0, 1)` | res$pit_plots$`Normal(0.5, 1)` | res$pit_plots$`Normal(0, 1.4)` | res$pit_plots$`Normal(0, 0.7)`) / 
+  gridExtra::tableGrob(scores_table)
 
@@
 @
+
+\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
+\end{figure}
+
+\subsection{Evaluating calibration and sharpness independently}
+
+Evaluating calibration and sharpness independently is helpful for model diagnostics.
+To that end \pkg{scoringutils} makes numerous metrics available that aim to capture different aspects of sharpness and calibration.
+
+\subsubsection{Assessing calibration}
+
+Calibration means consistency between forecasts and observed values, but there are various ways in which a forecast can systematically deviate from the observations (see \cite{gneitingProbabilisticForecastsCalibration2007} for a discussion of different forms of calibration MAYBE IT WOULD BE A GOOD IDEA TO INCLUDE THESE DIFFERENT FORMS OF CALIBRATION HERE).
+\pkg{scoringutils} allows the user to examine three different sub-aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
+
+% Bias
+Bias, i.e.
+systematic over- or underprediction, is a very common form of miscalibration which therefore deserves separate attention.
+The bias metric (with slightly different versions for the various forecast types and formats) captures a general tendency to over- and underpredict that is bound to be between minus one (underpredicton) and one (overprediction), where zero is ideal.
+It is derived by looking at how much of the probability mass of the predictive distribution is below or above the true observed value.
+For quantile forecasts we have second alternative approach available to assess over- and underprediction - by simply looking at the corresponding components of the weighted interval score.
+What is different between the over- and underprediction components and bias as described above is its sensitivity to outliers.
+The former are derived from absolute differences, while the latter is bound and rather captures a general tendency to be biased.
+
+% interval coverage and quantile coverage
+Another way to look at calibration (precisely: probabilistic calibration in \cite{gneitingProbabilisticForecastsCalibration2007}) is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
+This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and discrete forecasts as well.
+%
+To assess empirical coverage at a certain interval range, we simply measure the proportion of true observed values that fall into corresponding range of the predictive distribution.
+If the 0.05, 0.25, 0.75, and 0.95 quantiles are given, then 50\% of the true values should fall between the 0.25 and 0.75 quantiles and 90\% should fall between the 0.05 and 0.95 quantiles.
+We can calculate and plot these values to inspect how well different parts of the forecast distribution are calibrated.
+%
+To get an even more precise picture, we can also look at the percentage of true values below every single quantile of the predictive distribution.
+This allows to diagnose issues in the lower and upper tails of the prediction intervals separately.
+A similar way to visualise the same information is a PIT histogram.
+In order to conveniently assess deviations between the predictive distribution and the true data-generating distribution we can transform the observed values using the probability integral transformation (PIT) \citep{dawidPresentPositionPotential1984} (see more details in Table \ref{tab:score-table}).
+If both distributions are equal, the transformed values will follow a uniform distribution.
+A histogram of the transformed values can help to diagnose systematic differences between the predictions and the observed values.
+Figure \ref{fig:calibration-plots} exemplifies the characteristic shape of certain systematic deviations of the predicitve distribution from the true data-generating distribution.
+In the PIT histograms, bias leads to a triangular shape, overdispersion results in a hump shaped form and underdispersion in a U-shape.
+ADD INTERPRETATION FOR QUANTILE AND INTERVAL COVERAGE PLOTS HERE.
 
@@
 For sample-based forecasts, \pkg{scoringutils} calculates sharpness as the normalised median absolute deviation about the median (MADN) [funkAssessingPerformanceRealtime2019] (for details see Table \ref{tab:score-overview}).
-For quantile forecasts, we take the sharpness component of the wis which corresponds to a weighted average of the individual interval widths.
+For quantile forecasts, we take the sharpness component of the WIS which corresponds to a weighted average of the individual interval widths.
 
 WHAT WOULD BE REALLY COOL IS TO DETERMINE A WAY TO FIND THE OPTIMAL SHARPNESS OF A FORECAST. AS IS, I FEEL THIS PARAGRAPH IS SOMEWHAT USELESS.
+
+\subsection{Proper scoring rules}
+Proper scoring rules \citep{gneitingStrictlyProperScoring2007} jointly assess sharpness and calibration and assign a single numeric value to a forecast.
+A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
+This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state their best estimate.
+As summarised in Table \ref{tab:table-summary-scores}, not all proper scoring rules are suitable for all forecast formats.
+
+\subsection{Proper scoring rules for sample-based forecasts (CRPS, logS and DSS)}
+For sample-based forecasts, the \pkg{scoringutils} provides the following proper scoring rules: the (continuous) ranked probability score (CRPS) [CITATION], the log score (logS) [CITATION], and the Dawid-Sebastiani-score (DSS) [CITATION] (formal definitions are given in Table \ref{tab:score-table}).
+The corresponding functions are imported from the \pkg{scoringRules} package and exposed to the user through a slightly adapted interface.
+Other, non-sample-based variants of the CRPS, logS and DSS are available directly in the \pkg{scoringRules} package, but not in \pkg{scoringutils}.
+All three scores are in principle applicable to continuous as well as discrete forecasts.
+The \pkg{scoringRules} implementation of the log score, however, requires a kernel density estimation that may be inappropriate for discrete values (see also Table \ref{tab:table-summary-scores}).
+The logS is therefore not computed for discrete predictions in \pkg{scoringutils}.
+
+When scoring forecasts in a sample-based format, the choice is usually between the logS and the CRPS.
+The DSS is much less commonly used.
+It is easier to compute, but apart from that does not have immediate advantages over the former two.
+CRPS and logS differ in three important aspects: sensitivity to distance \cite{winklerScoringRulesEvaluation1996}, sensitivity to outlier predictions, and sensitivity to the order of magnitude of the forecasted quantity.
+
+The CRPS is a so-called global scoring rule, which means that the entire predictive distribution is taken into account when scoring a single forecast.
+The log score, on the other hand is local.
+The resulting score does not depend on the overall distance between the observed value and the distribution, but only on the probability density assigned to the actual outcome.
+Imagine two forecasters, A and B, who forecast the number of goals scored by a team in a football match.
+If both forecasters assigned the same probability to the true outcome (e.g.
+2 points), but A assigned higher probability to extreme outcomes far away from the actually observed outcome (e.g.
+A stated 10 goals may be quite likely, while B's predictions are concentrated around 1, 2 or 3 points), then A will receive a worse score than B.
+The log score, in contrast, is a local scoring rule that only scores the probability assigned to the actual outcome and ignores the rest of the predictive distribution.
+Judged by the log score, A and B would receive exactly the same score.
+Sensitivity to distance (taking the entire predictive distribution into account) may be an advantage in most settings that involve decision making.
+Forecaster A's prediction that assigns high probability to results far away from the observed value is arguably less useful than B's forecast that assigns higher probability to values closer to it (the probability assigned to the actual outcome being equal for both forecasts).
+The log score is only implicitly sensitive to distance if we assume that values close to the observed value are actually more likely to occur.
+The logS may, however, be more appropriate for inferential purposes (see \cite{winklerScoringRulesEvaluation1996}) and is commonly used in Bayesian statistics [CITATION].
+
+% stable vs. non-stable
+A second important difference is how forecasts are treated that deviate strongly from the observed outcome.
+The CRPS can be thought of as a generalisation of the absolute error to a predictive distribution.
+It therefore scales linearly with the distance between forecast distribution and true value.
+The log score, however, is the log of the predictive density evaluated at the observed value.
+It can quickly go to negative infinity if the probability assigned to the observed outcome is close to zero.
+The CRPS is therefore considered more stable than the log score.
+The behaviour of the DSS is in between the two.
+Whether or not harsh punishment of bad predictions is desirable or not depends of course on the setting.
+\cite{bracherEvaluatingEpidemicForecasts2021} exemplify that in practice there may indeed be substantial differences between how the CRPS and log score judge the same forecast.
+
+As the CRPS is a generalisation of the absolute value, overall scores depend on the order of magnitude of the quantity we try to forecast.
+This makes it harder to compare forecasts for very different targets, or assess average performance if the quantity of interest varies substantially over time.
+Average scores are dominated by forecasts for targets with high absolute numbers.
+This may be desirable,if we care most about forecasts in situations where numbers are high, but usually it is not.
+LogS and DSS are more robust against this effect.
+Another way to address this issue by using pairwise comparisons will be introduced later.
+
+\subsection{Proper scoring rules for non-sample-based forecasts (WIS and BS)}
+For forecasts in an interval or quantile format, \pkg{scoringutils} offers the weighted interval score (WIS) \citep{bracherEvaluatingEpidemicForecasts2021}. 
+% properties of the WIS + decomposition
+The WIS has very similar properties to the CRPS and can be thought of as a quantile-based approximation.
+For an increasing number of equally-spaced prediction intervals the WIS converges to the CRPS.
+One additional benefit of the WIS is that it can easily be decomposed into three additive components: an uncertainty penalty (sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval).
+This can be very helpful in diagnosing model problems.
+It may even be useful to convert samples into quantiles and use the WIS instead of the CRPS to make use of this decomposition for the purpose of model diagnostics.
+% Brier Score
+Binary forecasts can be scored using the Brier score (BS) [CITATION], which corresponds to the squared difference between the given probability and the outcome (either 0 or 1).
 
@@
 \section{Evaluating UK short-term forecasts}
-\subsection{The data}
 
-To illustrate the evaluation process with \pkg{scoringutils} we use short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
-Forecasts were produced by six groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
+The following section shows an example evaluation of short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
+Forecasts were produced by six research groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
 The forecasts aimed to assess the likely future burden the UK healthcare system would face from the Covid-19 pandemic.
-Predictions were then aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
-The data, as well as the individual forecast models are discussed in more depth in [FUNK ET AL].
+Predictions submitted to SPI-M were aggregated and used to inform UK government health policy through the Strategic Advisory Group of Experts (SAGE).
+All predictions as well as most of the observed values are now publicly available and discussed in more depth in [FUNK ET AL].
+% - timing (weekly?) and number of forecast dates
+% - the four targets
+% - the models.
 
-- timing (weekly?) and number of forecast dates
-- the four targets
-- the models. 
-- is the set complete?
+The evaluation process looks as follows: First we load, prepare and visualise (some of) the data.
+Then we obtain forecast scores and a model ranking based on pairwise-comparisons, followed by a more detailed analysis of calibration and sharpness.
+
+\subsection{Preparing and visualising the data}
 
@@
 % 
-To bring the forecasts into the format needed for the evaluation, some minor changes need to be made to the data.
+Some minor changes need to be made to the data before we can evaluate the forecasts.
 The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
@@
 % 
-Before we start with scoring the forecasts, it makes sense to start the evaluation process by visualising the data.
-To get a feeling for how complete the data set is, we can run the following code to obtain a heatmap with the number of available forecasts:
+It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance.
+The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts.
+By default, the function treats a set of different quantiles or samples as one forecast.
+However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.
 % 
@@
 <<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12, eval = TRUE>>=
-show_avail_forecasts(combined, 
+show_avail_forecasts(data = combined, 
                      x = "creation_date", 
@@
 @
-\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
+\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
 \end{figure}
@@
 % 
-Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
-By default, the function treats a set of different quantiles or samples as one forecast.
-However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts.
-
-Forecasts can be visualised using the \fct{plot\_predictions} function.
-The function accepts either a single combined data set or two separate truth data sets that can be merged together.
-Forecasts and observed values can be filtered independently, to show for example only predictions from a certain model or a specific number of weeks of true data before the forecast.
+The forecasts and observed values themselves can be visualised using the \fct{plot\_predictions} function.
+Data visualisation is of course highly context-dependent, but the function tries to accommodate most practical use-cases.
+In addition to basic plotting functionality it offers the user an optional ad hoc way to filter both forecasts and observed values.
+This makes it possible to tweak the plot in a beginner-friendly way without having to manipulate the data separately.
+Forecasts and observed values can be passed in separately (and are merged internally) or as a single data.frame.
 Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
-To obtain, for example, a specific forecast from a model we are interest in, we can call:
-% 
+To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: %
 \begin{figure}[h!]
@@
                  forecasts = uk_forecasts,
-                 filter_both = list(paste("geography %in%", locations, collapse = "")),
+                 filter_both = list(paste("geography %in%", locations, collapse = ""), 
+                                    'value_type == "death_inc_line"'),
                  filter_truth = list('value_date <= "2020-06-22"',
@@
                  x = "value_date",
-                 facet_formula = geography ~ value_type) + 
+                 facet_formula = ~ geography) + 
   ggplot2::theme(legend.position = "bottom")
 @
-\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
+\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
 \end{figure}
@@
 
-\subsection[Scoring forecasts with eval_forecasts]{Scoring forecasts with \fct{eval\_forecasts}}
+\subsection[Scoring forecasts and obtaining a model ranking]{Scoring forecasts with \fct{eval\_forecasts}}
 
-A full evaluation of all forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
-This requires a \code{data.frame} or similar which has at least a column called "prediction" and one called "true\_value".
-Depending on the exact input format, additional columns like "sample", "quantile", or "range" and "boundary" are needed.
+The actual scoring of forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
+The function automatically detects the forecast type and format, applies the appropriate scoring metrics and aggregates results as desired by the user.
+Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
+The function requires a \code{data.frame} or similar with some pre-specified columns (e.g.
+a column called "prediction", one called "true\_value" and more depending on the forecast format).
 Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
-We will not discuss all possible input formats here, but instead refer to the example data for each format that is provided with the package.
-Where possible, \pkg{scoringutils} also provides functionality to transform between various formats, e.g.
+To illustrate the requirements, \pkg{scoringutils} provides example data for all possible forecast formats.
+It also offers functions to to transform between various formats, e.g.
 from a sample based format to a quantile format.
-The \fct{eval\_forecasts} function automatically recognises the prediction type and input format, applies the appropriate scoring metrics and aggregates results as desired by the user.
-Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
 
-As a first start for the evaluation of UK short-term forecasts, it makes sense to look at the scores achieved by every model separate for all prediction targets. This can be achieved by calling
+When conducting an evaluation, we need to decide on the level of grouping over which we want to aggregate scores.
+For the UK short-term forecasts it makes sense to look at the scores achieved by every model separate for all prediction targets.
+This can be achieved by calling
 % ion G
@@
 % 
-If a more detailed analysis is desired, the level of aggregation can of course be changed to show for example separate scores for the different locations as well.
-This can be achieved using the \code{summarise_by} argument.
-To additionally stratify by location, we could specify \code{summarise_by = c("mode", "value_type", "geography")}.
-If we wanted to have one score per quantile or one per prediction interval range, we could specify something like \code{summarise_by = c("model", "quantile")} or \code{summarise_by = c("model", "quantile", "range")}.
-This can be useful if we, for example, want to analyse what proportion of true values are covered by certain interval ranges, or if we want to analyse the accuracy of the tails of the forecasts.
-When aggregating, \fct{eval\_forecasts} takes the mean according to the group defined in \code{summarise_by}.
+\fct{eval\_forecasts} aggregates scores by taking the mean over according to the grouping specified in \code{summarise_by}.
 In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
+For a more detailed analysis we could for example specifiy \code{summarise_by = c("mode", "value_type", "geography")} to additionally stratify by location.
+If we wanted to have one score per quantile or one per prediction interval range, we could use something like \code{summarise_by = c("model", "quantile", "range")}.
+This would allow us to examine interval or quantile coverage or makes it possible to analyse the accuracy of the tails of the forecasts.
 In addition to the mean, we can also obtain the standard deviation of the scores over which we average, as well as any desired quantile, by specifying \code{sd = TRUE} and for example \code{quantiles = c(0.5)} for the median.
 
-The user must, however, still exercise some caution when aggregating scores, as many of the metrics are absolute and scale with the magnitude of the quantity to forecast.
-Looking at one score per model (i.e.
-specifying \code{summarise_by = c("model")}) may not be so useful in this instance, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
+The user must, however, exercise some caution when aggregating scores.
+As explained above, many of the metrics are absolute and scale with the magnitude of the quantity to forecast.
+This makes it sometimes ill-advised to average over them.
+In the given example, looking at one score per model (i.e.
+specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
+Similarly, aggregating over different forecast horizons is often not a good idea as the mean will be dominated by further ahead forecast horizons.
 
-In the above example, we did not have to explicitly specify the \code{by} argument, but this may be necessary if additional columns are present in the data that do not indicate a grouping of forecasts.
-The \code{by} argument must then be used to denote the unit of a single forecast.
-In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")}.
-Quantiles should not be included, as several quantiles make up one forecast (and similarly for samples).
-If we had additional columns that do not serve to group forecasts (like for example the number of inhabitants in a certain location over time), these should also not be included.
+The second important argument of the \fct{eval\_forecasts} function is the \code{by} argument that denotes the unit of a single forecast.
+In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")} as there is one distinct forecast for each combination of these categories.
+Quantiles do not need to be included, as several quantiles make up one forecast (and similarly for samples).
 By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
+In the above example, we did not have to explicitly specify the \code{by} argument, as all available columns were relevant to the unit of a single forecast.
+If there are additional columns that do not correspond to the unit of a single forecast, like for example the observed temperature in a given location, \code{by} must be specified and these columns must be excluded.
 
-\subsection{Pairwise comparisons}
+In order to obtain a model ranking, we recommend to look at the relative skill in terms of an appropriate proper scoring rule instead of the raw score.
+Relative skill scores can be aggregated more easily across different forecast targets as they are less influenced by the order of magnitude of the quantity to forecast than e.g.
+the WIS or the CRPS.
 
@@
 \centering
-<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
+<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=15, eval = TRUE>>=
 score_table(scores, y = "model", facet_formula = ~ value_desc)
@@
 
-we can also look at p-values in Figure \ref{fig:pairwise-comparison-pval} PROBABLY REMOVE THAT FROM THE PAPER
-\begin{figure}[h!]
-\centering
-<<pairwise-plot-ratio, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
-plot_pairwise_comparison(pairwise, type = "pval") +
-  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-@
-\caption{\label{fig:pairwise-comparison-pval} XXXX}
-\end{figure}
-
 In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
@@
                facet_formula = ~ value_desc, 
-               scales = "free_x") + 
+               scales = "free_x", 
+               relative_contributions = TRUE, 
+               x_text_angle = 0) + 
   ggplot2::coord_flip() + 
@@
 scoringutils::interval_coverage(cov_scores)
-scoringutils::quantile_coverage(cov_scores)
-@
-
-<<quantile-coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = TRUE>>=
-cov_scores <- eval_forecasts(combined, 
-                             summarise_by = c("model", 
-                                              "range", "quantile"))
 scoringutils::quantile_coverage(cov_scores)
@@
 
-\begin{CodeInput}
-.
-\end{CodeInput}
+% \begin{CodeInput}
+% .
+% \end{CodeInput}
 
@@
 
-\bibliography{references}
+\bibliography{scoringutils-paper, references}
 
@@
 
+%Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
+
 %BIAS

### 2021-11-22 article.tex
update paper

@@
 \usepackage{colortbl}
+
+%% packages for landscape
+%% there are probably duplicates
+\usepackage{array}
+\usepackage{pdflscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{xcolor}
+\usepackage{colortbl}
+\usepackage{float}
+\usepackage{longtable}
 
@@
 \midrule
-Point forecast & continuous \newline  integer \newline  binary & one single number for the predicted outcome\\
+Point forecast & continuous \newline  discrete \newline  binary & one single number for the predicted outcome\\
 \cmidrule{1-3}\pagebreak[0]
- & continuous \newline  integer & predictive samples \newline   quantiles \newline   closed analytical form\\
+ & continuous \newline  discrete & predictive samples \newline   quantiles \newline   closed analytical form\\
 \cmidrule{2-3}\nopagebreak
@@
 \bottomrule
-\end{longtable}\caption{\label{tab:forecast-types} Ssummary of the different forceast types and forecast formats}
+\end{longtable}\caption{\label{tab:forecast-types} Forecasts can be distinguished by whether they are probabilistic in nature, or a point forecast only.
+Depending on the type of the target (discrete, continuous or binary) different representation of the predictive distribution are possible.}
 \end{table}
@@
 Table \ref{tab:table-summary-scores} shows an overview of the metrics implemented in \pkg{scoringutils}.
-Table \ref{tab:score-table} in the Appendix gives a thorough explanation of all the metrics.
+Table \ref{tab:score-table} in the Appendix gives a mathematical definition and thorough explanation of all the metrics.
+Figure \ref{fig:calibration-plots} gives an applied example of different scoring metrics and visualisations for simple toy data.
 
+% \begin{table}[h!]
+% \centering
+
+% \caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
+% \end{table}
+
+\newpage
 \begin{table}[h!]
 \centering
-\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{4cm}}
+\begingroup\fontsize{10}{12}\selectfont
+
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.0 cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.2cm}>{\raggedright\arraybackslash}p{4.5cm}}
 \toprule
-Metric & Target types & Forecast formats & Properties\\
+Metric & Discrete & Continuous & Binary & Closed-form & Samples (approx.) & Quantiles & Properties\\
 \midrule
-\endfirsthead
-\multicolumn{4}{@{}l}{\textit{(continued)}}\\
-\toprule
-Metric & Target types & Forecast formats & Properties\\
-\midrule
-\endhead
-
-\endfoot
+\cellcolor{gray!6}{(Continuous) ranked probability score (CRPS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
+\addlinespace
+Log score (logS) & ($\checkmark$) & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, local, unstable for outliers\\
+\addlinespace
+\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK AGAIN} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
+\addlinespace
+Dawid-Sebastiani score (DSS) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
+\addlinespace
+\cellcolor{gray!6}{Brier score (BS)} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule}\\
+\addlinespace
+Interval coverage & $\checkmark$ & $\checkmark$ &  &  & THINK AGAIN & $\checkmark$ & measure for calibration\\
+\addlinespace
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK AGAIN} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measure for calibration}\\
+\addlinespace
+Probability integral transform (PIT) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & assesses calibration\\
+\addlinespace
+\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
+\addlinespace
+Bias & $\checkmark$ & $\checkmark$ & $\checkmark$??? & $\checkmark$ & $\checkmark$ & $\checkmark$ & captures tendency to over-or underpredict (aspect of calibration)\\
+\addlinespace
+\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{compares performance of two models}\\
+\addlinespace
+Relative skill & depends & depends & depends & depends & depends & depends & Ranks models based on pairwise comparisons\\
 \bottomrule
-\endlastfoot
-\cellcolor{gray!6}{CRPS (Continuous) ranked probability score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
-\addlinespace
-Log score & continuous, (integer not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
-\addlinespace
-\cellcolor{gray!6}{WIS (Weighted) interval score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
-\addlinespace
-DSS Dawid-Sebastiani score & continuous, integer & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
-\addlinespace
-\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
-\addlinespace
-Interval coverage & continuous, integer & interval forecasts (needs matching quantiles) & measure for calibration\\
-\addlinespace
-\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
-\addlinespace
-Probability integral transform (PIT) & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
-\addlinespace
-\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
-\addlinespace
-Bias & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
-\addlinespace
-\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{compares performance of two models}\\
-\addlinespace
-Relative skill & depends on scored & depends on score & Ranks models based on pairwise comparisons\\*
-\end{longtable}\caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
+\end{longtable}
+\endgroup{}\caption{\label{tab:table-summary-scores2} Summary table of scores available in scoringutils}
 \end{table}
 
-\subsection{Proper scoring rules}
+\newpage
 
-Proper scoring rules \citep{gneitingStrictlyProperScoring2007} jointly assess sharpness and calibration and assign a single numeric value to a forecast.
-A scoring rule is proper if a perfect forecaster (the predictive distribution equals the data-generating distribution) receives the lowest score on average.
-This makes sure that a forecaster evaluated by a proper scoring rule is always incentivised to state their best estimate.
+\begin{figure}[h]
+\centering
 
-For sample-based forecasts, the (continuous) ranked probability score (crps) [CITATION], the log score (logs) [CITATION], and the Dawid-Sebastiani-score (dss) [CITATION] are available in \pkg{scoringutils}.
-These are implemented as wrappers around functions from the \pkg{scoringRules} package (which also has closed-form versions of the scoring rules available).
-They are in principle applicable to continuous as well as integer forecasts.
-The \pkg{scoringRules} implementation of the log score, however, requires a kernel density estimation that may be inappropriate for integer values (see also Table \ref{tab:table-summary-scores}) and is therefore not available for discrete predictions in \pkg{scoringutils}.
-For forecasts in an interval or quantile format, the weighted interval score (wis) \citep{bracherEvaluatingEpidemicForecasts2020a} is available.
-For an increasing number of equally-spaced prediction intervals, the wis converges to the crps and therefore has very similar properties.
-Binary forecasts can be scored using the Brier score (bs) [CITATION].
-
-%sample-based proper scoring rules\usepackage{multirow}
-When scoring forecasts in a sample-based format, the choice is usually between the log score and the crps.
-The dss is much less commonly used.
-It is easier to compute, but apart from that does not have immediate advantages over the former two.
-Crps and log score differ in two important aspects: the first is sensitivity to distance \cite{winklerScoringRulesEvaluation1996}, the second is how harshly far-off predictions are punished.
-
-The crps is a so-called global scoring rule, which means that the entire predictive distribution is taken into account when scoring a single forecast.
-The log score, on the other hand is local.
-The resulting score does not depend on the overall distance between the observed value and the distribution, but only on the probality density assigend to the actual outcome.
-Imagine two forecasters, A and B, who forecast the number of points scored in a basketball game.
-If both forecasters assigned the same probability to the true outcome (100 points), but A assigned higher probability to extreme outcomes far away from the actually observed outcome, then A will receive a worse score than B.
-The log score, in contrast, is a local scoring rule that only scores the probability assigned to the actual outcome and ignores the rest of the predictive distribution.
-Judged by the log score, A and B would receive exactly the same score.
-Sensitivity to distance (taking the entire predictive distribution into account) is arguably an advantage in most settings that involve decision making.
-Forecaster A's prediction that assigns high probability to results far away from the observed value is arguably less useful than B's forecast that assigns higher probability to values closer to it (the probability assigned to the actual outcome being equal for both forecasts).
-The log score is only implicitly sensitive to distance if we assume that values close to the observed value are actually more likely to occur.
-It may, however, be more appropriate for inferential purposes (see \cite{winklerScoringRulesEvaluation1996}).
-
-% stable vs. non-stable
-A second important difference is how forecasts are treated that deviate strongly from the observered outcome.
-The crps can be thought of as a generalisation of the absolute error to a predictive distribution.
-It therefore scales linearly with the distance between forecast distribution and true value.
-The log score, however, is the log of the predictive density evaluated at the observed value.
-It can therfore quickly go to negative infinity if the probability assigned to the observed outcome is close to zero.
-The crps is therefore considered more stable than the log score.
-The behaviour of the dss is in between the two.
-Whether or not harsh punishment of bad predictions is desirable or not depends of course on the setting.
-\cite{bracherEvaluatingEpidemicForecasts2020a} exemplify that in practice there may indeed be substantial differences between how the crps and log score judge the same forecast.
-
-% properties of the WIS + decomposition
-For quantile forecasts the wis is the best available option and can be understood as an approximation to the crps.
-One additional benefit of the wis is that it can easily be decomposed into three additive components: an uncertainty penalty (sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval).
-This can be very helpful in diagnosing model problems.
-It may therefore even be useful to convert samples into quantiles and use the wis instead of the crps to make use of this decomposition for the purpose of model diagnostics.
-% fact that scores scale with the overall level of the prediction
-Both crps and wis, as generalisations of the absolute value, suffer from the fact that overall scores depend on the order of magnitude of the quantity we try to forecast.
-It can therefore be difficult to compare forecasts for very different targets.
-One possibility to address this is the use of pairwise comparisons introduced later.
-In general, this is much less of a problem for the log score and the dss.
-IS IT ONE AT ALL?
+\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
+\end{figure}
 
@@
 
-In addition to the proper scoring rules outlined above, \pkg{scoringutils} makes numerous metrics available to evaluate calibration and sharpness independently.
-This is especially helpful for model diagnostics.
+Evaluating calibration and sharpness independently is helpful for model diagnostics.
+To that end \pkg{scoringutils} makes numerous metrics available that aim to capture different aspects of sharpness and calibration.
 
@@
 
-Several strategies have been proposed to detect systematic deviations of the predictive distributions from the observations (see e.g.
-\cite{funkAssessingPerformanceRealtime2019}; \cite{gneitingProbabilisticForecastsCalibration2007}; \cite{gneitingStrictlyProperScoring2007}).
-Using \pkg{scoringutils}, we can look at three different aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
+Calibration means consistency between forecasts and observed values, but there are various ways in which a forecast can systematically deviate from the observations (see \cite{gneitingProbabilisticForecastsCalibration2007} for a discussion of different forms of calibration MAYBE IT WOULD BE A GOOD IDEA TO INCLUDE THESE DIFFERENT FORMS OF CALIBRATION HERE).
+\pkg{scoringutils} allows the user to examine three different sub-aspects of calibration: bias, empirical coverage, and the probability integral transform (PIT).
 
@@
 Another way to look at calibration (precisely: probabilistic calibration in \cite{gneitingProbabilisticForecastsCalibration2007}) is to compare the proportion of observed values covered by different parts of the predictive distribution with the nominal coverage implied by the CDF of the distribution.
-This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and integer forecasts as well.
+This is most easily understood in the context of quantile forecasts, but can easily be transferred to sample-based continuous and discrete forecasts as well.
 %
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics{plots/calibration-diagnostic-examples.png}
-\caption{\label{fig:calibration-plots} Calibration plots for different forecast distributions. I plan to add interval and quantile coverage plots, but these still have to be made.}
-\end{figure}
-
 \subsubsection{Assessing sharpness}
@@
 \begin{Sinput}
-R> show_avail_forecasts(combined, 
+R> show_avail_forecasts(data = combined, 
 +                       x = "creation_date", 
@@
 \includegraphics{plots/plot-show-availability}
-\caption{\label{fig:avail-forecasts} Overview of the number of forecasts available}
+\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
 \end{figure}
@@
 +                   forecasts = uk_forecasts,
-+                   filter_both = list(paste("geography %in%", locations, collapse = "")),
++                   filter_both = list(paste("geography %in%", locations, collapse = ""), 
++                                      'value_type == "death_inc_line"'),
 +                   filter_truth = list('value_date <= "2020-06-22"',
@@
 +                   x = "value_date",
-+                   facet_formula = geography ~ value_type) + 
++                   facet_formula = ~ geography) + 
 +    ggplot2::theme(legend.position = "bottom")
@@
 \end{Schunk}
-\caption{\label{fig:forecast-visualisation} Short-term forecasts made by the SIRCOVID model on June 22 2020.}
+\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
 \end{figure}
@@
 
-we can also look at p-values in Figure \ref{fig:pairwise-comparison-pval} PROBABLY REMOVE THAT FROM THE PAPER
-\begin{figure}[h!]
-\centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise, type = "pval") +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-pairwise-plot-ratio}
-\caption{\label{fig:pairwise-comparison-pval} XXXX}
-\end{figure}
-
 In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
@@
 +                 facet_formula = ~ value_desc, 
-+                 scales = "free_x") + 
++                 scales = "free_x", 
++                 relative_contributions = TRUE, 
++                 x_text_angle = 0) + 
 +    ggplot2::coord_flip() + 
@@
 \includegraphics{plots/plot-coverage}
-
-\begin{Schunk}
-\begin{Sinput}
-R> cov_scores <- eval_forecasts(combined, 
-+                               summarise_by = c("model", 
-+                                                "range", "quantile"))
-R> scoringutils::quantile_coverage(cov_scores)
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-quantile-coverage}
 
@@
 % The results in this paper were obtained using
-% \proglang{R}~4.0.4 with the
-% \pkg{MASS}~7.3.53 package. \proglang{R} itself
+% \proglang{R}~4.1.2 with the
+% \pkg{MASS}~7.3.54 package. \proglang{R} itself
 % and all packages used are available from the Comprehensive
@@
 \end{document}
+
+%Missing forecasts can have a large impact on the forecast evaluation, if forecasts are not missing at random, but instead missingness correlates with performance.
 

### 2021-11-22 tables/forecast-types.tex
update paper

@@
 \midrule
-Point forecast & continuous \newline  integer \newline  binary & one single number for the predicted outcome\\
+Point forecast & continuous \newline  discrete \newline  binary & one single number for the predicted outcome\\
 \cmidrule{1-3}\pagebreak[0]
- & continuous \newline  integer & predictive samples \newline   quantiles \newline   closed analytical form\\
+\midrule
+ & continuous \newline  discrete & predictive samples \newline   quantiles \newline   closed analytical form\\
 \cmidrule{2-3}\nopagebreak
-\multirow[t]{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
+\multirow{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
 \bottomrule

### 2021-11-22 tables/scores-table-summary.Rnw
update paper

@@
 
+# create checkmark with $\checkmark$
+
 crps <- list(
-  `Metric` = "CRPS (Continuous) ranked probability score", 
-  `Target types` = r"(continuous, integer)", 
+  `Metric` = "(Continuous) ranked probability score (CRPS)", 
+  `Target types` = r"(continuous, discrete)", 
   `Forecast formats` = r"(closed-form, samples (approximation))",
@@
 log_score <- list(
-  `Metric` = "Log score", 
-  `Target types` = r"(continuous, (integer not in scoringutils))", 
+  `Metric` = "Log score (logS)", 
+  `Target types` = r"(continuous, (discrete not in scoringutils))", 
   `Forecast formats` = r"(closed-form, samples (approximation))",
@@
 wis <- list(
-  Metric = "WIS (Weighted) interval score", 
-  `Target types` = r"(continuous, integer)", 
+  Metric = "(Weighted) interval score (WIS)", 
+  `Target types` = r"(continuous, discrete)", 
   `Forecast formats` = r"(quantile or interval predictions)",
@@
 dss <- list(
-  `Metric` = "DSS Dawid-Sebastiani score", 
-  `Target types` = r"(continuous, integer)", 
+  `Metric` = "Dawid-Sebastiani score (DSS)", 
+  `Target types` = r"(continuous, discrete)", 
   `Forecast formats` = r"(closed-form, samples (approximation))",
@@
 brier_score <- list(
-  `Metric` = "Brier score", 
+  `Metric` = "Brier score (BS)", 
   `Target types` = r"(binary)", 
@@
   `Metric` = "Interval coverage", 
-  `Target types` = r"(continuous, integer)", 
+  `Target types` = r"(continuous, discrete)", 
   `Forecast formats` = r"(interval forecasts (needs matching quantiles))",
@@
   `Metric` = "Quantile coverage", 
-  `Target types` = r"(continuous, integer)", 
+  `Target types` = r"(continuous, discrete)", 
   `Forecast formats` = r"(quantile or interval forecasts)",
@@
   `Metric` = "Sharpness", 
-  `Target types` = r"(continuous, integer)", 
+  `Target types` = r"(continuous, discrete)", 
   `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
@@
   `Metric` = "Bias", 
-  `Target types` = r"(continuous, integer, quantile)", 
+  `Target types` = r"(continuous, discrete, quantile)", 
   `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
@@
   `Metric` = "Probability integral transform (PIT)",
-  `Target types` = r"(continuous, integer, quantile)", 
+  `Target types` = r"(continuous, discrete, quantile)", 
   `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
@@
 
-saveRDS(data, file = "tables/scores-table-summary.rds")
+# saveRDS(data, file = "tables/scores-table-summary.rds")
 
@@
 
+<<table2, echo=FALSE, results=tex>>=
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+library("MASS")
+library(data.table)
+
+# create checkmark with $\checkmark$
+
+crps <- list(
+  `Metric` = "(Continuous) ranked probability score (CRPS)", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"($\checkmark$)",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"()",
+  `Properties` = "proper scoring rule, global, stable handling of outliers",
+  `References` = ""
+)
+
+log_score <- list(
+  `Metric` = "Log score (logS)", 
+  `Discrete` = r"(($\checkmark$))",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"($\checkmark$)",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"()",
+  `Properties` = "proper scoring rule, local, unstable for outliers",
+  `References` = ""
+)
+
+wis <- list(
+  Metric = "(Weighted) interval score (WIS)", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"()",
+  `Samples (approx.)` = r"(THINK )",
+  `Quantiles` = r"($\checkmark$)",
+  `Properties` = "proper scoring rule, global, stable handling of outliers, converges to crps",
+  `References` = ""
+)
+
+dss <- list(
+  `Metric` = "Dawid-Sebastiani score (DSS)", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"($\checkmark$)",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"()",
+  `Properties` = "proper scoring rule, somewhat global, somewhat stable handling of outliers",
+  `References` = ""
+)
+
+brier_score <- list(
+  `Metric` = "Brier score (BS)", 
+  `Discrete` = r"()",
+  `Continuous` = r"()",
+  `Binary` = r"($\checkmark$)",
+  `Closed-form` = r"()",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"()",
+  #`Forecast formats` = r"(binary probabilities)",
+  `Properties` = "proper scoring rule",
+  `References` = ""
+)
+
+interval_coverage <- list(
+  `Metric` = "Interval coverage", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"()",
+  `Samples (approx.)` = r"(THINK )",
+  `Quantiles` = r"($\checkmark$)",
+  `Properties` = "measure for calibration",
+  `References` = ""
+)
+
+quantile_coverage <- list(
+  `Metric` = "Quantile coverage", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"()",
+  `Samples (approx.)` = r"(THINK )",
+  `Quantiles` = r"($\checkmark$)",
+  `Properties` = "measure for calibration",
+  `References` = ""
+)
+
+sharpness <- list(
+  `Metric` = "Sharpness", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"($\checkmark$)",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"($\checkmark$)",
+  `Properties` = "measures forecast dispersions",
+  `References` = ""
+)
+
+bias <- list(
+  `Metric` = "Bias", 
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"($\checkmark$???)",
+  `Closed-form` = r"($\checkmark$)",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"($\checkmark$)",
+  `Properties` = "captures tendency to over-or underpredict (aspect of calibration)",
+  `References` = ""
+)
+
+pit <- list(
+  `Metric` = "Probability integral transform (PIT)",
+  `Discrete` = r"($\checkmark$)",
+  `Continuous` = r"($\checkmark$)",
+  `Binary` = r"()",
+  `Closed-form` = r"($\checkmark$)",
+  `Samples (approx.)` = r"($\checkmark$)",
+  `Quantiles` = r"($\checkmark$)",
+  `Properties` = "assesses calibration",
+  `References` = ""
+)
+
+mean_score_ratio <- list(
+  `Metric` = "Mean score ratio", 
+  `Discrete` = r"(depends)",
+  `Continuous` = r"(depends)",
+  `Binary` = r"(depends)",
+  `Closed-form` = r"(depends)",
+  `Samples (approx.)` = r"(depends)",
+  `Quantiles` = r"(depends)",
+  `Properties` = "compares performance of two models",
+  `References` = ""
+)
+
+relative_skill <- list(
+  `Metric` = "Relative skill", 
+  `Discrete` = r"(depends)",
+  `Continuous` = r"(depends)",
+  `Binary` = r"(depends)",
+  `Closed-form` = r"(depends)",
+  `Samples (approx.)` = r"(depends)",
+  `Quantiles` = r"(depends)",
+  `Properties` = "Ranks models based on pairwise comparisons",
+  `References` = ""
+)
+
+data <- rbind(as.data.table(crps), 
+              as.data.table(log_score), 
+              as.data.table(wis), 
+              as.data.table(dss), 
+              as.data.table(brier_score), 
+              as.data.table(interval_coverage),
+              as.data.table(quantile_coverage), 
+              as.data.table(pit),
+              as.data.table(sharpness),
+              as.data.table(bias), 
+              as.data.table(mean_score_ratio), 
+              as.data.table(relative_skill))
+
+# saveRDS(data, file = "tables/scores-table-summary2.rds")
+
+library(kableExtra)
+library(magrittr)
+library(knitr)
+data[, 1:8] %>%
+  kableExtra::kbl(format = "latex", booktabs = TRUE, 
+                  escape = FALSE,
+                  longtable = TRUE,
+                  align = c("lccccccl"),
+                  linesep = c('\\addlinespace')) %>%
+  kableExtra::column_spec(1, width = "3.2cm") %>%
+  kableExtra::column_spec(2, width = "1.5cm") %>%
+  kableExtra::column_spec(3, width = "1.5cm") %>%
+  kableExtra::column_spec(4, width = "1.3cm") %>%
+  kableExtra::column_spec(5, width = "1.3cm") %>%
+  kableExtra::column_spec(6, width = "1.5cm") %>%
+  kableExtra::column_spec(7, width = "1.3cm") %>%
+  kableExtra::column_spec(8, width = "6.0cm") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "repeat_header")) %>%
+  kableExtra::landscape()
+
+@
+
 \end{document}

### 2021-11-22 tables/scores-table-summary.tex
update paper

@@
 \usepackage{xcolor}
-% \usepackage{colortbl}
+\usepackage{colortbl}
 \usepackage{float}
@@
 \endlastfoot
-\cellcolor{gray!6}{CRPS (Continuous) ranked probability score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
+\cellcolor{gray!6}{(Continuous) ranked probability score (CRPS)} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
 \addlinespace
-Log score & continuous, (integer not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
+Log score (logS) & continuous, (discrete not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
 \addlinespace
-\cellcolor{gray!6}{WIS (Weighted) interval score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
+\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
 \addlinespace
-DSS Dawid-Sebastiani score & continuous, integer & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
+Dawid-Sebastiani score (DSS) & continuous, discrete & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
 \addlinespace
-\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
+\cellcolor{gray!6}{Brier score (BS)} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
 \addlinespace
-Interval coverage & continuous, integer & interval forecasts (needs matching quantiles) & measure for calibration\\
+Interval coverage & continuous, discrete & interval forecasts (needs matching quantiles) & measure for calibration\\
 \addlinespace
-\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
 \addlinespace
-Probability integral transform (PIT) & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
+Probability integral transform (PIT) & continuous, discrete, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
 \addlinespace
-\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
+\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
 \addlinespace
-Bias & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
+Bias & continuous, discrete, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
 \addlinespace
@@
 \end{longtable}
+\begin{landscape}
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.2cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.3cm}>{\centering\arraybackslash}p{1.3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.3cm}>{\raggedright\arraybackslash}p{6.0cm}}
+\toprule
+Metric & Discrete & Continuous & Binary & Closed-form & Samples (approx.) & Quantiles & Properties\\
+\midrule
+\endfirsthead
+\multicolumn{8}{@{}l}{\textit{(continued)}}\\
+\toprule
+Metric & Discrete & Continuous & Binary & Closed-form & Samples (approx.) & Quantiles & Properties\\
+\midrule
+\endhead
+
+\endfoot
+\bottomrule
+\endlastfoot
+\cellcolor{gray!6}{(Continuous) ranked probability score (CRPS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
+\addlinespace
+Log score (logS) & ($\checkmark$) & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, local, unstable for outliers\\
+\addlinespace
+\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
+\addlinespace
+Dawid-Sebastiani score (DSS) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
+\addlinespace
+\cellcolor{gray!6}{Brier score (BS)} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule}\\
+\addlinespace
+Interval coverage & $\checkmark$ & $\checkmark$ &  &  & THINK & $\checkmark$ & measure for calibration\\
+\addlinespace
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measure for calibration}\\
+\addlinespace
+Probability integral transform (PIT) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & assesses calibration\\
+\addlinespace
+\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measures forecast dispersions}\\
+\addlinespace
+Bias & $\checkmark$ & $\checkmark$ & $\checkmark$??? & $\checkmark$ & $\checkmark$ & $\checkmark$ & captures tendency to over-or underpredict (aspect of calibration)\\
+\addlinespace
+\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{compares performance of two models}\\
+\addlinespace
+Relative skill & depends & depends & depends & depends & depends & depends & Ranks models based on pairwise comparisons\\*
+\end{longtable}
+\end{landscape}
 

### 2021-12-10 article-knitr.Rnw
switch to knitr, update a few plots

--- 2021-11-22 article.Rnw
@@
 %% need no \usepackage{Sweave}
-\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE, prefix.string=plots/plot}
-<<preliminaries, echo=FALSE, results=hide>>=
+
+<<include=FALSE>>=
+library(knitr)
+opts_chunk$set(
+  engine='R', 
+  tidy=TRUE, 
+  fig.path='plots/plot-', 
+  message = FALSE, 
+  warning = FALSE
+)
+@
+
+<<preliminaries, echo=FALSE, results='hide'>>=
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@@
 library(knitr)
+library(here)
 
@@
 \begin{document}
-\SweaveOpts{concordance=FALSE}
+
+<<include=FALSE>>=
+library(knitr)
+opts_chunk$set(
+concordance=FALSE
+)
+@
 
@@
 \centering
-<<forecast-types-table, echo=FALSE, results=tex>>=
+<<forecast-types-table, echo=FALSE, results='asis'>>=
 
@@
 
-<<summary-table, echo=FALSE, eval = FALSE, results=tex>>=
+<<summary-table, echo=FALSE, eval = FALSE, results='asis'>>=
 
@@
 \centering
-<<summary-table, echo=FALSE, results=tex>>=
+<<summary-table2, echo=FALSE, results='asis'>>=
 
@@
 \centering
-<<example-calibration, echo=FALSE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
-n_truth = 1000
-n_samples = 100
-true_values <- rnorm(n_truth, 0, 1)
-predictions1 <- rnorm(n_truth * n_samples)
-predictions2 <- rnorm(n_truth * n_samples, mean = 0.5)
-predictions3 <- rnorm(n_truth * n_samples, sd = 1.4)
-predictions4 <- rnorm(n_truth * n_samples, sd =  0.7)
-# predictions5 <- rnorm(n_truth * n_samples, mean = 0.2, sd =  0.7)
-
-df <- data.table::data.table(true_value = rep(true_values, each = n_samples), 
-                             id = rep(1:n_truth, each = n_samples),
-                             prediction = c(predictions1, predictions2, 
-                                            predictions3, predictions4), 
-                             sample = 1:n_samples,
-                             model = rep(c("Normal(0, 1)", "Normal(0.5, 1)", 
-                                       "Normal(0, 1.4)", "Normal(0, 0.7)"), 
-                                       each = n_truth * n_samples))
-
-res <- eval_forecasts(df, 
-                      by = c("model", "id"), 
-                      summarise_by = c("model"), 
-                      pit_plots = TRUE)
-
-# plot with observations
-true_value_plot <- ggplot2::ggplot(data = data.frame(x = true_values),
-                ggplot2::aes(x = x)) +
-  ggplot2::geom_histogram(ggplot2::aes(y = ..density..),
-                          fill = "grey",
-                          colour = "dark grey") +
-  cowplot::theme_cowplot() +
-  ggplot2::labs(x = "True values",
-                y = "Density") +
-  ggplot2::theme(legend.position = "bottom")
-
-# plot with standard normal distribution
-standard_normal <- true_value_plot +
-  ggplot2::geom_function(fun = dnorm, colour = "black") +
-  ggplot2::ggtitle("Normal(0, 1)")
-
-# plot with shifted mean
-shifted_mean <- true_value_plot +
-  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(mean = 0.5)) +
-  ggplot2::ggtitle("Normal(0.5, 1)")
-
-# plot with overdispersion
-overdispersion <- true_value_plot +
-  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 1.4)) +
-  ggplot2::ggtitle("Normal(0, 1.4)")
-
-# plot with underdispersion
-underdispersion <- true_value_plot +
-  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 0.7)) +
-  ggplot2::ggtitle("Normal(0, 0.7)")
-
-scores_table <- dcast(melt(res$scores, id.vars = "model", 
-                           variable.name = "score"), 
-                      score ~ model)
-setcolorder(
-  scores_table, 
-  c("score", "Normal(0, 1)", "Normal(0.5, 1)", "Normal(0, 1.4)", "Normal(0, 0.7)")
-)
-
-(standard_normal | shifted_mean | overdispersion | underdispersion) /
-  (res$pit_plots$`Normal(0, 1)` | res$pit_plots$`Normal(0.5, 1)` | res$pit_plots$`Normal(0, 1.4)` | res$pit_plots$`Normal(0, 0.7)`) / 
-  gridExtra::tableGrob(scores_table)
-
-ggsave("plots/calibration-diagnostic-examples.png", width = 12.5, height = 6)
+<<example-calibration2, echo=FALSE, fig.height=7.2, fig.width=14, eval = TRUE>>=
+knitr::include_graphics(here("plots", "calibration-diagnostic-examples.png"))
 @
@@
 \centering
-<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12, eval = TRUE>>=
+<<show-availability, echo=TRUE, fig.height=6.7, fig.width=12, eval = TRUE>>=
 show_avail_forecasts(data = combined, 
@@
 \centering
-<<show-forecasts, echo=TRUE, fig=TRUE, height=5.2, width=7, eval = FALSE>>=
+<<show-forecasts, echo=TRUE, fig.height=5.2, fig.width=7, eval = FALSE>>=
 locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
@@
 \centering
-<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=15, eval = TRUE>>=
+<<score-table, echo=TRUE, fig.height=7.2, fig.width=15, eval = TRUE>>=
 score_table(scores, y = "model", facet_formula = ~ value_desc)
@@
 \centering
-<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
+<<pairwise-plot, echo=TRUE, fig.height=7.2, fig.width=14, eval = TRUE>>=
 plot_pairwise_comparison(pairwise) +
@@
 \centering
-<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12, eval = TRUE>>=
+<<WIS-components, echo=TRUE, fig.height=5.2, fig.width=12, eval = TRUE>>=
 wis_components(scores, 
@@
 \centering
-<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = TRUE>>=
+<<coverage, echo=TRUE, fig.height=5.2, fig.width=14, out.width="40%", fig.show='hold', eval = TRUE>>=
 cov_scores <- eval_forecasts(combined, 
@@
 \centering
-<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15, eval = TRUE>>=
+<<calibration, echo=TRUE, fig.height=5.2, fig.width=15, eval = TRUE>>=
 scores <- eval_forecasts(combined, 
@@
 \centering
-<<scores-table, echo=FALSE, results=tex>>=
+<<scores-table, echo=FALSE, results='asis'>>=
 

### 2021-12-10 article-knitr.tex
switch to knitr, update a few plots

--- 2021-11-22 article.tex
@@
-\documentclass[article,shortnames]{jss}
+\documentclass[article,shortnames]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 }
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \centering
+
 \begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{4.5cm}}
@@
 \bottomrule
-\end{longtable}\caption{\label{tab:forecast-types} Forecasts can be distinguished by whether they are probabilistic in nature, or a point forecast only.
+\end{longtable}
+
+\caption{\label{tab:forecast-types} Forecasts can be distinguished by whether they are probabilistic in nature, or a point forecast only.
 Depending on the type of the target (discrete, continuous or binary) different representation of the predictive distribution are possible.}
@@
 \end{longtable}
-\endgroup{}\caption{\label{tab:table-summary-scores2} Summary table of scores available in scoringutils}
+\endgroup{}
+
+\caption{\label{tab:table-summary-scores2} Summary table of scores available in scoringutils}
 \end{table}
@@
 \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=0.98\textwidth]{/mnt/data/GoogleDrive/Uni/PhD/scoringutils-paper/plots/calibration-diagnostic-examples} 
+\end{knitrout}
 
@@
 % 
-\begin{Schunk}
-\begin{Sinput}
-R> # install and load data pacakge from external repository
-R> # remotes::install_github("sbfnk/covid19.forecasts.uk")
-R> 
-R> # load packages
-R> library(covid19.forecasts.uk)
-R> library(dplyr)
-R> library(scoringutils)
-R> # load truth data
-R> data(covid_uk_data)
-R> # head(covid_uk_data)
-R> 
-R> # load forecasts
-R> data(uk_forecasts)
-R> # head(uk_forecasts)
-\end{Sinput}
-\end{Schunk}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{# install and load data pacakge from external repository}
+\hlcom{# remotes::install_github('sbfnk/covid19.forecasts.uk')}
+
+\hlcom{# load packages}
+\hlkwd{library}\hlstd{(covid19.forecasts.uk)}
+\hlkwd{library}\hlstd{(dplyr)}
+\hlkwd{library}\hlstd{(scoringutils)}
+
+\hlcom{# load truth data}
+\hlkwd{data}\hlstd{(covid_uk_data)}
+\hlcom{# head(covid_uk_data)}
+
+\hlcom{# load forecasts}
+\hlkwd{data}\hlstd{(uk_forecasts)}
+\hlcom{# head(uk_forecasts)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 % 
@@
 % 
-\begin{Schunk}
-\begin{Sinput}
-R> glimpse(covid_uk_data)
-\end{Sinput}
-\begin{Soutput}
-Rows: 4,174
-Columns: 6
-$ geography  <fct> London, London, London, London, London, London, L…
-$ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita…
-$ value_desc <fct> Hospital admissions, Hospital admissions, Hospita…
-$ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0…
-$ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, …
-$ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, …
-\end{Soutput}
-\begin{Sinput}
-R> glimpse(uk_forecasts)
-\end{Sinput}
-\begin{Soutput}
-Rows: 1,254,513
-Columns: 8
-$ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E…
-$ geography     <chr> "East of England", "East of England", "East of…
-$ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp…
-$ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3…
-$ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0…
-$ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05…
-$ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0…
-$ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp…
-\end{Soutput}
-\end{Schunk}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{glimpse}\hlstd{(covid_uk_data)}
+\end{alltt}
+\begin{verbatim}
+## Rows: 4,174
+## Columns: 6
+## $ geography  <fct> London, London, London, London, London, London, L~
+## $ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita~
+## $ value_desc <fct> Hospital admissions, Hospital admissions, Hospita~
+## $ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0~
+## $ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, ~
+## $ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, ~
+\end{verbatim}
+\begin{alltt}
+\hlkwd{glimpse}\hlstd{(uk_forecasts)}
+\end{alltt}
+\begin{verbatim}
+## Rows: 1,254,513
+## Columns: 8
+## $ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E~
+## $ geography     <chr> "East of England", "East of England", "East of~
+## $ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp~
+## $ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3~
+## $ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0~
+## $ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05~
+## $ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0~
+## $ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp~
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 % 
@@
 % 
-\begin{Schunk}
-\begin{Sinput}
-R> uk_forecasts <- rename(uk_forecasts, prediction = value)
-R> covid_uk_data <-rename(covid_uk_data, true_value = value)
-R> combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
-\end{Sinput}
-\end{Schunk}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{uk_forecasts} \hlkwb{<-} \hlkwd{rename}\hlstd{(uk_forecasts,} \hlkwc{prediction} \hlstd{= value)}
+\hlstd{covid_uk_data} \hlkwb{<-} \hlkwd{rename}\hlstd{(covid_uk_data,} \hlkwc{true_value} \hlstd{= value)}
+
+\hlstd{combined} \hlkwb{<-} \hlkwd{merge_pred_and_obs}\hlstd{(uk_forecasts, covid_uk_data)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 % 
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> show_avail_forecasts(data = combined, 
-+                       x = "creation_date", 
-+                       show_numbers = FALSE,
-+                       legend_position = "bottom",
-+                       facet_formula = ~ value_desc)
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-show-availability}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{show_avail_forecasts}\hlstd{(}\hlkwc{data} \hlstd{= combined,} \hlkwc{x} \hlstd{=} \hlstr{"creation_date"}\hlstd{,} \hlkwc{show_numbers} \hlstd{=} \hlnum{FALSE}\hlstd{,}
+    \hlkwc{legend_position} \hlstd{=} \hlstr{"bottom"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=0.98\textwidth]{plots/plot-show-availability-1} 
+\end{knitrout}
 \caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
-R> plot_predictions(truth = covid_uk_data,
-+                   forecasts = uk_forecasts,
-+                   filter_both = list(paste("geography %in%", locations, collapse = ""), 
-+                                      'value_type == "death_inc_line"'),
-+                   filter_truth = list('value_date <= "2020-06-22"',
-+                                       'value_date > "2020-06-01"'), 
-+                   filter_forecasts = list('model == "SIRCOVID"', 
-+                                           'creation_date == "2020-06-22"'),
-+                   x = "value_date",
-+                   facet_formula = ~ geography) + 
-+    ggplot2::theme(legend.position = "bottom")
-\end{Sinput}
-\end{Schunk}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{locations} \hlkwb{<-} \hlstr{"c(\textbackslash{}"Scotland\textbackslash{}", \textbackslash{}"Wales\textbackslash{}", \textbackslash{}"Northern Ireland\textbackslash{}", \textbackslash{}"England\textbackslash{}")"}
+\hlkwd{plot_predictions}\hlstd{(}\hlkwc{truth} \hlstd{= covid_uk_data,} \hlkwc{forecasts} \hlstd{= uk_forecasts,} \hlkwc{filter_both} \hlstd{=} \hlkwd{list}\hlstd{(}\hlkwd{paste}\hlstd{(}\hlstr{"geography %in%"}\hlstd{,}
+    \hlstd{locations,} \hlkwc{collapse} \hlstd{=} \hlstr{""}\hlstd{),} \hlstr{"value_type == \textbackslash{}"death_inc_line\textbackslash{}""}\hlstd{),} \hlkwc{filter_truth} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"value_date <= \textbackslash{}"2020-06-22\textbackslash{}""}\hlstd{,}
+    \hlstr{"value_date > \textbackslash{}"2020-06-01\textbackslash{}""}\hlstd{),} \hlkwc{filter_forecasts} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"model == \textbackslash{}"SIRCOVID\textbackslash{}""}\hlstd{,}
+    \hlstr{"creation_date == \textbackslash{}"2020-06-22\textbackslash{}""}\hlstd{),} \hlkwc{x} \hlstd{=} \hlstr{"value_date"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{geography)} \hlopt{+}
+    \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 \caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
@@
 % ion G
-\begin{Schunk}
-\begin{Sinput}
-R> scores <- eval_forecasts(combined, 
-+                           summarise_by = c("model", "value_desc"))
-R> glimpse(scores)
-\end{Sinput}
-\begin{Soutput}
-Rows: 29
-Columns: 9
-$ model              <fct> EpiSoon, StructuredODE, Microsimulation, …
-$ value_desc         <fct> Hospital admissions, Hospital admissions,…
-$ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876, …
-$ sharpness          <dbl> 22.886153, 6.133974, 33.985553, 10.365082…
-$ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.5…
-$ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.55…
-$ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174, …
-$ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0.…
-$ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563, …
-\end{Soutput}
-\end{Schunk}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"value_desc"}\hlstd{))}
+
+\hlkwd{glimpse}\hlstd{(scores)}
+\end{alltt}
+\begin{verbatim}
+## Rows: 29
+## Columns: 9
+## $ model              <fct> EpiSoon, StructuredODE, Microsimulation, ~
+## $ value_desc         <fct> Hospital admissions, Hospital admissions,~
+## $ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876, ~
+## $ dispersion         <dbl> 22.886153, 6.133974, 33.985553, 10.365082~
+## $ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.5~
+## $ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.55~
+## $ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174, ~
+## $ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0.~
+## $ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563, ~
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 % 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> # unsummarised scores
-R> unsummarised_scores <- eval_forecasts(combined)
-R> pairwise <- pairwise_comparison(unsummarised_scores, 
-+                                  summarise_by = "value_desc")
-\end{Sinput}
-\end{Schunk}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{# unsummarised scores}
+\hlstd{unsummarised_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined)}
+\hlstd{pairwise} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(unsummarised_scores,} \hlkwc{summarise_by} \hlstd{=} \hlstr{"value_desc"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> score_table(scores, y = "model", facet_formula = ~ value_desc)
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-score-table}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{score_table}\hlstd{(scores,} \hlkwc{y} \hlstd{=} \hlstr{"model"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=0.98\textwidth]{plots/plot-score-table-1} 
+\end{knitrout}
 \caption{\label{fig:score-table} Coloured table to visualise the computed scores}
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> plot_pairwise_comparison(pairwise) +
-+    ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-pairwise-plot}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot_pairwise_comparison}\hlstd{(pairwise)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{value_desc,} \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=0.98\textwidth]{plots/plot-pairwise-plot-1} 
+\end{knitrout}
 \caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> wis_components(scores, 
-+                 facet_formula = ~ value_desc, 
-+                 scales = "free_x", 
-+                 relative_contributions = TRUE, 
-+                 x_text_angle = 0) + 
-+    ggplot2::coord_flip() + 
-+    ggplot2::theme(legend.position = "bottom")
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-WIS-components}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{wis_components}\hlstd{(scores,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc,} \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{,}
+    \hlkwc{relative_contributions} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{x_text_angle} \hlstd{=} \hlnum{0}\hlstd{)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{coord_flip}\hlstd{()} \hlopt{+}
+    \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=0.98\textwidth]{plots/plot-WIS-components-1} 
+\end{knitrout}
 \caption{\label{fig:wis-components} p-values and ratios together.}
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> cov_scores <- eval_forecasts(combined, 
-+                               summarise_by = c("model", 
-+                                                "range", "quantile"))
-R> scoringutils::interval_coverage(cov_scores)
-R> scoringutils::quantile_coverage(cov_scores)
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-coverage}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{cov_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"range"}\hlstd{,}
+    \hlstr{"quantile"}\hlstd{))}
+\hlstd{scoringutils}\hlopt{::}\hlkwd{interval_coverage}\hlstd{(cov_scores)}
+\hlstd{scoringutils}\hlopt{::}\hlkwd{quantile_coverage}\hlstd{(cov_scores)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=0.4\linewidth]{plots/plot-coverage-1} 
+\includegraphics[width=0.4\linewidth]{plots/plot-coverage-2} 
+\end{knitrout}
 
@@
 \centering
-\begin{Schunk}
-\begin{Sinput}
-R> scores <- eval_forecasts(combined, 
-+                           summarise_by = c("model",
-+                                            "value_desc", 
-+                                            "geography"), 
-+                           compute_relative_skill = FALSE)
-R> scoringutils::score_heatmap(scores, metric = "bias", 
-+                              x = "geography", facet_formula = ~ value_desc, 
-+                              scale = "free_x")
-\end{Sinput}
-\end{Schunk}
-\includegraphics{plots/plot-calibration}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"value_desc"}\hlstd{,}
+    \hlstr{"geography"}\hlstd{),} \hlkwc{compute_relative_skill} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+\hlstd{scoringutils}\hlopt{::}\hlkwd{score_heatmap}\hlstd{(scores,} \hlkwc{metric} \hlstd{=} \hlstr{"bias"}\hlstd{,} \hlkwc{x} \hlstd{=} \hlstr{"geography"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc,}
+    \hlkwc{scale} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=0.98\textwidth]{plots/plot-calibration-1} 
+\end{knitrout}
 \caption{\label{fig:bias-heatmap} bias by location}
@@
 % The results in this paper were obtained using
-% \proglang{R}~4.1.2 with the
-% \pkg{MASS}~7.3.54 package. \proglang{R} itself
+% \proglang{R}~paste(R.Version()[6:7], collapse = ".") with the
+% \pkg{MASS}~packageVersion("MASS") package. \proglang{R} itself
 % and all packages used are available from the Comprehensive
@@
 \centering
+
 \caption{\label{tab:score-table} Explanation of all the scores}

### 2021-12-10 plots/plots.R

@@ -1,0 +1,78 @@
+
+library(scoringutils)
+library(patchwork)
+library(ggplot2)
+library(data.table)
+
+
+# Calibration Diagnostic plot --------------------------------------------------
+n_truth = 1000
+n_samples = 100
+true_values <- rnorm(n_truth, 0, 1)
+predictions1 <- rnorm(n_truth * n_samples)
+predictions2 <- rnorm(n_truth * n_samples, mean = 0.5)
+predictions3 <- rnorm(n_truth * n_samples, sd = 1.4)
+predictions4 <- rnorm(n_truth * n_samples, sd =  0.7)
+# predictions5 <- rnorm(n_truth * n_samples, mean = 0.2, sd =  0.7)
+
+df <- data.table::data.table(true_value = rep(true_values, each = n_samples), 
+                             id = rep(1:n_truth, each = n_samples),
+                             prediction = c(predictions1, predictions2, 
+                                            predictions3, predictions4), 
+                             sample = 1:n_samples,
+                             model = rep(c("Normal(0, 1)", "Normal(0.5, 1)", 
+                                           "Normal(0, 1.4)", "Normal(0, 0.7)"), 
+                                         each = n_truth * n_samples))
+
+res <- eval_forecasts(df, 
+                      by = c("model", "id"), 
+                      summarise_by = c("model"), 
+                      pit_plots = TRUE)
+
+# plot with observations
+true_value_plot <- ggplot2::ggplot(data = data.frame(x = true_values),
+                                   ggplot2::aes(x = x)) +
+  ggplot2::geom_histogram(ggplot2::aes(y = ..density..),
+                          fill = "grey",
+                          colour = "dark grey") +
+  cowplot::theme_cowplot() +
+  ggplot2::labs(x = "True values",
+                y = "Density") +
+  ggplot2::theme(legend.position = "bottom") 
+
+# plot with standard normal distribution
+standard_normal <- true_value_plot +
+  ggplot2::geom_function(fun = dnorm, colour = "black") +
+  ggplot2::ggtitle("Normal(0, 1)")
+
+# plot with shifted mean
+shifted_mean <- true_value_plot +
+  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(mean = 0.5)) +
[…]

### 2021-12-10 plots/plots.R
update calibration plot

@@ -1,4 +1,3 @@
-
 library(scoringutils)
 library(patchwork)
 library(ggplot2)
@@ -25,9 +24,11 @@ df <- data.table::data.table(true_value = rep(true_values, each = n_samples),
                                          each = n_truth * n_samples))
 
 res <- eval_forecasts(df, 
-                      by = c("model", "id"), 
-                      summarise_by = c("model"), 
-                      pit_plots = TRUE)
+                      summarise_by = c("model"))
+
+pit <- scoringutils::pit_df(df, summarise_by = "model")
+pit_plots <- scoringutils::hist_PIT(pit)
+
 
 # plot with observations
 true_value_plot <- ggplot2::ggplot(data = data.frame(x = true_values),
@@ -35,7 +36,7 @@ true_value_plot <- ggplot2::ggplot(data = data.frame(x = true_values),
   ggplot2::geom_histogram(ggplot2::aes(y = ..density..),
                           fill = "grey",
                           colour = "dark grey") +
-  cowplot::theme_cowplot() +
+  theme_minimal() +
   ggplot2::labs(x = "True values",
                 y = "Density") +
   ggplot2::theme(legend.position = "bottom") 
@@ -60,7 +61,7 @@ underdispersion <- true_value_plot +
   ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 0.7)) +
   ggplot2::ggtitle("Normal(0, 0.7)")
 
-scores_table <- dcast(melt(res$scores, id.vars = "model", 
+scores_table <- dcast(melt(res, id.vars = "model", 
                            variable.name = "score"), 
                       score ~ model)
 scores_table <- scores_table[, lapply(.SD, round, 2), by = score]
@@ -71,7 +72,7 @@ setcolorder(
 )
 
 (standard_normal | shifted_mean | overdispersion | underdispersion) /
-  (res$pit_plots$`Normal(0, 1)` | res$pit_plots$`Normal(0.5, 1)` | res$pit_plots$`Normal(0, 1.4)` | res$pit_plots$`Normal(0, 0.7)`) / 
+  (pit_plots$`Normal(0, 1)` | pit_plots$`Normal(0.5, 1)` | pit_plots$`Normal(0, 1.4)` | pit_plots$`Normal(0, 0.7)`) / 
   gridExtra::tableGrob(scores_table)
 
 

### 2021-12-10 article-knitr.Rnw
update sumamry table and calibration plot

@@
 
-<<summary-table, echo=FALSE, eval = FALSE, results='asis'>>=
-
-data <- readRDS("tables/scores-table-summary.rds")
-
-data[, 1:4] %>%
-  kableExtra::kbl(format = "latex", booktabs = TRUE, 
-                  escape = FALSE,
-                  longtable = TRUE,
-                  linesep = c('\\addlinespace')) %>%
-  kableExtra::column_spec(1, width = "12.5cm") %>%
-  # kableExtra::column_spec(2, width = "2cm") %>%
-  # kableExtra::column_spec(3, width = "4.5cm") %>%
-  # kableExtra::column_spec(4, width = "4cm") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "repeat_header", "scale_down"), font_size = 2)
-
-@
-% \caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
-% \end{table}
-
 \newpage
@@
                   linesep = c('\\addlinespace')) %>%
-  kableExtra::column_spec(1, width = "2.0 cm") %>%
-  kableExtra::column_spec(2, width = "1.4cm") %>%
-  kableExtra::column_spec(3, width = "1.4cm") %>%
-  kableExtra::column_spec(4, width = "1.2cm") %>%
-  kableExtra::column_spec(5, width = "1.2cm") %>%
-  kableExtra::column_spec(6, width = "1.4cm") %>%
-  kableExtra::column_spec(7, width = "1.2cm") %>%
-  kableExtra::column_spec(8, width = "4.5cm") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "repeat_header, scale_down"), font_size = 10)
+  kableExtra::column_spec(1, width = "1.8 cm") %>%
+  kableExtra::column_spec(2, width = "1.2cm") %>%
+  kableExtra::column_spec(3, width = "1.2cm") %>%
+  kableExtra::column_spec(4, width = "1.0cm") %>%
+  kableExtra::column_spec(5, width = "1.0cm") %>%
+  kableExtra::column_spec(6, width = "1.3cm") %>%
+  kableExtra::column_spec(7, width = "1.1cm") %>%
+  kableExtra::column_spec(8, width = "3.4cm") %>%
+  kableExtra::kable_styling(latex_options = c("striped", "repeat_header, scale_down"), font_size = 8)
 

### 2021-12-10 article-knitr.tex
update sumamry table and calibration plot

@@
 
-% \caption{\label{tab:table-summary-scores} Summary table of scores available in scoringutils}
-% \end{table}
-
 \newpage
@@
 \centering
-\begingroup\fontsize{10}{12}\selectfont
+\begingroup\fontsize{8}{10}\selectfont
 
-\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.0 cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.2cm}>{\raggedright\arraybackslash}p{4.5cm}}
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{1.8 cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.2cm}>{\centering\arraybackslash}p{1.0cm}>{\centering\arraybackslash}p{1.0cm}>{\centering\arraybackslash}p{1.3cm}>{\centering\arraybackslash}p{1.1cm}>{\raggedright\arraybackslash}p{3.4cm}}
 \toprule
@@
 \addlinespace
-Log score (logS) & ($\checkmark$) & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, local, unstable for outliers\\
+Log score (logS) &  & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, log of predictive density evaluated at observed value, local, unstable for outliers\\
 \addlinespace
-\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK AGAIN} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
+\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps for an increasing numbre of equally spaced intervals}\\
 \addlinespace
@@
 \addlinespace
-Interval coverage & $\checkmark$ & $\checkmark$ &  &  & THINK AGAIN & $\checkmark$ & measure for calibration\\
+Interval coverage & $\checkmark$ & $\checkmark$ &  &  & THINK & $\checkmark$ & measure for calibration\\
 \addlinespace
-\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK AGAIN} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measure for calibration}\\
+\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measure for calibration}\\
 \addlinespace
@@
 \addlinespace
-\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
+\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measures forecast dispersions}\\
 \addlinespace
@@
 \addlinespace
-\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{compares performance of two models}\\
+\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{$\sim$} & \cellcolor{gray!6}{$\sim$} & \cellcolor{gray!6}{$\sim$} & \cellcolor{gray!6}{$\sim$} & \cellcolor{gray!6}{$\sim$} & \cellcolor{gray!6}{$\sim$} & \cellcolor{gray!6}{compares performance of two models.
+Properties depend on the metric chosen for the comparison.}\\
 \addlinespace
-Relative skill & depends & depends & depends & depends & depends & depends & Ranks models based on pairwise comparisons\\
+Relative skill & $\sim$ & $\sim$ & $\sim$ & $\sim$ & $\sim$ & $\sim$ & Ranks models based on pairwise comparisons. Properties depend on the metric chosen for the comparison.\\
 \bottomrule

### 2021-12-10 tables/scores-table-summary.Rnw
update sumamry table and calibration plot

@@
 \begin{document}
-
-<<table, echo=FALSE, results=tex>>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-library("MASS")
-library(data.table)
-
-# create checkmark with $\checkmark$
-
-crps <- list(
-  `Metric` = "(Continuous) ranked probability score (CRPS)", 
-  `Target types` = r"(continuous, discrete)", 
-  `Forecast formats` = r"(closed-form, samples (approximation))",
-  `Properties` = "proper scoring rule, global, stable handling of outliers",
-  `References` = ""
-)
-
-log_score <- list(
-  `Metric` = "Log score (logS)", 
-  `Target types` = r"(continuous, (discrete not in scoringutils))", 
-  `Forecast formats` = r"(closed-form, samples (approximation))",
-  `Properties` = "proper scoring rule, local, unstable for outliers",
-  `References` = ""
-)
-
-wis <- list(
-  Metric = "(Weighted) interval score (WIS)", 
-  `Target types` = r"(continuous, discrete)", 
-  `Forecast formats` = r"(quantile or interval predictions)",
-  `Properties` = "proper scoring rule, global, stable handling of outliers, converges to crps",
-  `References` = ""
-)
-
-dss <- list(
-  `Metric` = "Dawid-Sebastiani score (DSS)", 
-  `Target types` = r"(continuous, discrete)", 
-  `Forecast formats` = r"(closed-form, samples (approximation))",
-  `Properties` = "proper scoring rule, somewhat global, somewhat stable handling of outliers",
-  `References` = ""
-)
-
-brier_score <- list(
-  `Metric` = "Brier score (BS)", 
-  `Target types` = r"(binary)", 
-  `Forecast formats` = r"(binary probabilities)",
-  `Properties` = "proper scoring rule",
-  `References` = ""
-)
-
-interval_coverage <- list(
-  `Metric` = "Interval coverage", 
-  `Target types` = r"(continuous, discrete)", 
-  `Forecast formats` = r"(interval forecasts (needs matching quantiles))",
-  `Properties` = "measure for calibration",
-  `References` = ""
-)
-
-quantile_coverage <- list(
-  `Metric` = "Quantile coverage", 
-  `Target types` = r"(continuous, discrete)", 
-  `Forecast formats` = r"(quantile or interval forecasts)",
-  `Properties` = "measure for calibration",
-  `References` = ""
-)
-
-sharpness <- list(
-  `Metric` = "Sharpness", 
-  `Target types` = r"(continuous, discrete)", 
-  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
-  `Properties` = "measures sharpness, slightly different depending on forecast format",
-  `References` = ""
-)
-
-bias <- list(
-  `Metric` = "Bias", 
-  `Target types` = r"(continuous, discrete, quantile)", 
-  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
-  `Properties` = "captures tendency to over-or underpredict (aspect of calibration)",
-  `References` = ""
-)
-
-pit <- list(
-  `Metric` = "Probability integral transform (PIT)",
-  `Target types` = r"(continuous, discrete, quantile)", 
-  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
-  `Properties` = "assesses calibration",
-  `References` = ""
-)
-
-mean_score_ratio <- list(
-  `Metric` = "Mean score ratio", 
-  `Target types` = r"(depends on score)", 
-  `Forecast formats` = r"(depends on score)",
-  `Properties` = "compares performance of two models",
-  `References` = ""
-)
-
-relative_skill <- list(
-  `Metric` = "Relative skill", 
-  `Target types` = r"(depends on scored)", 
-  `Forecast formats` = r"(depends on score)",
-  `Properties` = "Ranks models based on pairwise comparisons",
-  `References` = ""
-)
-
-data <- rbind(as.data.table(crps), 
-              as.data.table(log_score), 
-              as.data.table(wis), 
-              as.data.table(dss), 
-              as.data.table(brier_score), 
-              as.data.table(interval_coverage),
-              as.data.table(quantile_coverage), 
-              as.data.table(pit),
-              as.data.table(sharpness),
-              as.data.table(bias), 
-              as.data.table(mean_score_ratio), 
-              as.data.table(relative_skill))
-
-# saveRDS(data, file = "tables/scores-table-summary.rds")
-
-library(kableExtra)
-library(magrittr)
-library(knitr)
-data[, 1:4] %>%
-  kableExtra::kbl(format = "latex", booktabs = TRUE, 
-                  escape = FALSE,
-                  longtable = TRUE,
-                  linesep = c('\\addlinespace')) %>%
-  kableExtra::column_spec(1, width = "2.5cm") %>%
-  kableExtra::column_spec(2, width = "2cm") %>%
-  kableExtra::column_spec(3, width = "4.5cm") %>%
-  kableExtra::column_spec(4, width = "4cm") %>%
-  kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))
-
-@
 
@@
   `Metric` = "Log score (logS)", 
-  `Discrete` = r"(($\checkmark$))",
+  `Discrete` = "",
   `Continuous` = r"($\checkmark$)",
@@
   `Quantiles` = r"()",
-  `Properties` = "proper scoring rule, local, unstable for outliers",
+  `Properties` = "proper scoring rule, log of predictive density evaluated at observed value, local, unstable for outliers",
   `References` = ""
@@
   `Quantiles` = r"($\checkmark$)",
-  `Properties` = "proper scoring rule, global, stable handling of outliers, converges to crps",
+  `Properties` = "proper scoring rule, global, stable handling of outliers, converges to crps for an increasing numbre of equally spaced intervals",
   `References` = ""
@@
   `Metric` = "Mean score ratio", 
-  `Discrete` = r"(depends)",
-  `Continuous` = r"(depends)",
-  `Binary` = r"(depends)",
-  `Closed-form` = r"(depends)",
-  `Samples (approx.)` = r"(depends)",
-  `Quantiles` = r"(depends)",
-  `Properties` = "compares performance of two models",
+  `Discrete` = r"($\sim$)",
+  `Continuous` = r"($\sim$)",
+  `Binary` = r"($\sim$)",
+  `Closed-form` = r"($\sim$)",
+  `Samples (approx.)` = r"($\sim$)",
+  `Quantiles` = r"($\sim$)",
+  `Properties` = "compares performance of two models. Properties depend on the metric chosen for the comparison.",
   `References` = ""
@@
   `Metric` = "Relative skill", 
-  `Discrete` = r"(depends)",
-  `Continuous` = r"(depends)",
-  `Binary` = r"(depends)",
-  `Closed-form` = r"(depends)",
-  `Samples (approx.)` = r"(depends)",
-  `Quantiles` = r"(depends)",
-  `Properties` = "Ranks models based on pairwise comparisons",
+  `Discrete` = r"($\sim$)",
+  `Continuous` = r"($\sim$)",
+  `Binary` = r"($\sim$)",
+  `Closed-form` = r"($\sim$)",
+  `Samples (approx.)` = r"($\sim$)",
+  `Quantiles` = r"($\sim$)",
+  `Properties` = "Ranks models based on pairwise comparisons. Properties depend on the metric chosen for the comparison.",
   `References` = ""
@@
 
-# saveRDS(data, file = "tables/scores-table-summary2.rds")
+saveRDS(data, file = "tables/scores-table-summary2.rds")
 

### 2021-12-10 plots/plots.R

@@ -23,11 +23,29 @@ df <- data.table::data.table(true_value = rep(true_values, each = n_samples),
                                            "Normal(0, 1.4)", "Normal(0, 0.7)"), 
                                          each = n_truth * n_samples))
 
+# get scores
 res <- eval_forecasts(df, 
                       summarise_by = c("model"))
 
+# create pit plots
 pit <- scoringutils::pit_df(df, summarise_by = "model")
 pit_plots <- scoringutils::hist_PIT(pit)
+
+# create coverage plots by transforming to quantile format first
+quantiles <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)
+df_quantile <- scoringutils::sample_to_quantile(df, 
+                                                quantiles = quantiles)
+
+res_quantile <- eval_forecasts(df_quantile, 
+                               summarise_by = c("model", "range", "quantile"))
+
+res_quantile[, model := factor(model, levels = c("Normal(0, 1)", "Normal(0.5, 1)", "Normal(0, 1.4)", "Normal(0, 0.7)"))]
+
+interval_coverage <- scoringutils::interval_coverage(res_quantile) + 
+  facet_wrap(~ model, nrow = 1)
+
+quantile_coverage <- scoringutils::quantile_coverage(res_quantile) + 
+  facet_wrap(~ model, nrow = 1)
 
 
 # plot with observations
@@ -73,7 +91,9 @@ setcolorder(
 
 (standard_normal | shifted_mean | overdispersion | underdispersion) /
   (pit_plots$`Normal(0, 1)` | pit_plots$`Normal(0.5, 1)` | pit_plots$`Normal(0, 1.4)` | pit_plots$`Normal(0, 0.7)`) / 
+  interval_coverage / 
+  quantile_coverage /
   gridExtra::tableGrob(scores_table)
 
 
-ggsave("plots/calibration-diagnostic-examples.png", width = 12.5, height = 6)
+ggsave("plots/calibration-diagnostic-examples.png", width = 12.5, height = 10)

### 2021-12-12 article-knitr.Rnw
update paper and plots

@@
   engine='R', 
+  cache=TRUE,
   tidy=TRUE, 
+  tidy.opts=list(width.cutoff=60),
   fig.path='plots/plot-', 
   message = FALSE, 
-  warning = FALSE
+  warning = FALSE, 
+  prompt = TRUE,
+  continue = "+  ", 
+  width = 70, 
+  useFancyQuotes = FALSE
 )
@@
 It gives the user access to a wide range of scoring metrics for various types of forecasts as well as a variety of ways to visualise the evaluation.
-We give an overview of the evaluation process and the metrics implemented in \pkg{scoringutils} and show a full evaluation of a set of short-term forecasts of public health related targets made by SPI-M during the 2020 Covid-19 epidemic in the United Kingdom.
+We give an overview of the evaluation process and the metrics implemented in \pkg{scoringutils} and show an example evaluation of forecasts for COVID-19 cases and deaths submitted to the European Forecast Hub between May und September 2021.
 }
@@
 For an increasing number of equally-spaced prediction intervals the WIS converges to the CRPS.
-One additional benefit of the WIS is that it can easily be decomposed into three additive components: an uncertainty penalty (sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval).
+One additional benefit of the WIS is that it can easily be decomposed into three additive components: an uncertainty penalty (called dispersion or sharpness) for the width of a prediction interval and penalties for over- and underprediction (if a value falls outside of a prediction interval).
 This can be very helpful in diagnosing model problems.
 It may even be useful to convert samples into quantiles and use the WIS instead of the CRPS to make use of this decomposition for the purpose of model diagnostics.
+
 % Brier Score
@@
 
-The following section shows an example evaluation of short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
+% The following section shows an example evaluation of short-term predictions of four different Covid-19 related targets in the UK made between March 31 and July 13 2020.
 Forecasts were produced by six research groups in the UK, and submitted to the Scientific Pandemic Influenza Group on Modelling (SPI-M).
@@
 
+The following section shows an example evaluation of short-term predictions for COVID-19 cases and deaths submitted to the European Forecast Hub [CITATION].
+The full oficial hub evaluations, which also use \pkg{scoringutils}, can be seen at https://covid19forecasthub.eu/.
+The European Forecast Hub each week collates, aggregates and evaluates one to four week ahead predictions of different COVID-19 related targets submitted by different research groups.
+Forecasts are submitted in a quantile-based format with a set of 22 quantiles plus the median ($0.01, 0.025, 0.05, ..., 0.5, ...
+0.95, 0.975, 0.99$).
+The following example evaluation is based on a subset of forecasts from the European Forecast Hub which includes one to three week ahead forecasts made between May and September 2021 for COVID-19 cases and deaths from four different models.
+
+This example data follows a quantile-based format, but the evaluation process would work analogously for forecasts in a different format.
+We will point out differences to scoring different forecast formats where appropriate.
+The example data set (as well as versions of it in different formats) is also included in the \pkg{scoringutils} package.
+
 The evaluation process looks as follows: First we load, prepare and visualise (some of) the data.
@@
 
-\subsection{Preparing and visualising the data}
+\subsection{Checking the data}
 
-We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
 % 
 <<load-data>>=
-# install and load data pacakge from external repository
-# remotes::install_github("sbfnk/covid19.forecasts.uk")
-
 # load packages
-library(covid19.forecasts.uk)
+library(scoringutils)
 library(dplyr)
-library(scoringutils)
-
-# load truth data
-data(covid_uk_data)
-# head(covid_uk_data)
+library(data.table)
+library(kableExtra)
 
 # load forecasts
-data(uk_forecasts)
-# head(uk_forecasts)
+data(example_quantile)
 @
 % 
-Let us take a first look a the data:
+All higher level functions that allow for convenient evaluation of forecasts expect a \code{data.frame} or similar.
+Lower-level functions (discussed in a later section) accept inputs as vectors and matrices.
+
+We omit \code{NA} values as the data set also contains entries for which we have an observed value, but no forecast. 
 % 
 <<glimpse-data>>=
-glimpse(covid_uk_data)
-glimpse(uk_forecasts)
+example_quantile |>
+  na.omit() |>
+  glimpse()
 @
 %
-Some minor changes need to be made to the data before we can evaluate the forecasts.
-The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
-While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public.
-The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way.
-% 
-<<modify-data>>=
-uk_forecasts <- rename(uk_forecasts, prediction = value)
-covid_uk_data <-rename(covid_uk_data, true_value = value)
 
-combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
+For forecasts in a quantile-based format it is necessary to have columns called \code{true_value}, \code{prediction} and \code{quantile}.
+It is recommended (although not necessary except in order to do pairwise comparisons) to have a column called "model" with an identifier for the forecaster.
+Table \ref{tab:column-requirements} shows the columns expected for different input formats.
+
+<<column-requirements>>=
+requirements <- 
+  data.table(
+    "Format" = c("quantile-based", "sample-based", "binary", "pairwise-comparisons"), 
+    `Required columns` = c("'true_value', 'prediction', 'quantile'", 
+                           "'true_value', 'prediction', 'sample'", 
+                           "'true_value', 'prediction'", 
+                           "additionally a column 'model'")
+  )
+
+requirements |>
+  kable(format = "latex")
 @
-% 
-It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance.
+
+We can check whether the data conforms to the requirements by running \fct{check\_forecasts}.
+
+<<>>=
+check_forecasts(example_quantile)
+@
+
+This returns a list with different entries giving information about what \pkg{scoringutils} infers from the data.
+\code{target_type} \code{prediction_type} refer to the forecast format.
+\code{forecast_unit} contains a vector of the columns which \pkg{scoringutils} thinks denote the unit of a single forecast.
+This means that in this instance a single forecast (with a set of 23 quantiles) can uniquely be identified by the values in the columns "location", "target\_end\_date", "target\_type", "location\_name", "forecast\_date", "model", "horizon".
+In this example, having "location" as well as "location\_name" included does not make a difference, as they contain duplicated information.
+In general, however, it is strongly advised to remove all unnecessary columns that do not help identify a single forecast.
+\code{unique_values} gives an overview of the number of unique values per column across the entire data set, providing a first hint as to whether the forecast set is complete.
+\code{warnings} shows potential warnings about the data.
+In this example, \pkg{scoringutils} warns that there are observed values present for which there is no corresponding forecast.
+These warnings can often be ignored, but may provide important information.
+If there are errors that cannot be ignored, a list entry \code{errors} will appear.
+
+It is helpful to start the evaluation process by visualising data availability, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance.
 The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts.
@@
 <<show-availability, echo=TRUE, fig.height=6.7, fig.width=12, eval = TRUE>>=
-show_avail_forecasts(data = combined, 
-                     x = "creation_date", 
+show_avail_forecasts(data = example_quantile, 
+                     x = "target_end_date", 
                      show_numbers = FALSE,
                      legend_position = "bottom",
-                     facet_formula = ~ value_desc)
+                     facet_formula = ~ target_type)
 @
@@
 Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
-To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: %
+To display, for example, short-term forecasts for COVID-19 cases and deaths made by the EuroCOVIDhub-ensemble model on June 28 2021 as well as 5 weeks of prior data, we can call: %
 \begin{figure}[h!]
 \centering
-<<show-forecasts, echo=TRUE, fig.height=5.2, fig.width=7, eval = FALSE>>=
-locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
-plot_predictions(truth = covid_uk_data,
-                 forecasts = uk_forecasts,
-                 filter_both = list(paste("geography %in%", locations, collapse = ""), 
-                                    'value_type == "death_inc_line"'),
-                 filter_truth = list('value_date <= "2020-06-22"',
-                                     'value_date > "2020-06-01"'), 
-                 filter_forecasts = list('model == "SIRCOVID"', 
-                                         'creation_date == "2020-06-22"'),
-                 x = "value_date",
-                 facet_formula = ~ geography) + 
+<<show-forecasts, echo=TRUE, fig.height=5.2, fig.width=7, eval = TRUE>>=
+plot_predictions(data = example_quantile,
+                 x = "target_end_date",
+                 filter_truth = list('target_end_date <= "2021-07-15"',
+                                     'target_end_date > "2021-05-22"'),
+                 filter_forecasts = list("model == 'EuroCOVIDhub-ensemble'",
+                                         'forecast_date == "2021-06-28"'),
+                 facet_formula =  target_type ~ location, 
+                 ncol = 4) + 
   ggplot2::theme(legend.position = "bottom")
 @
-\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
+\caption{\label{fig:forecast-visualisation} Short-term forecasts for COVID-19 cases and deaths made by the EuroCOVIDhub-ensemble model on June 28 2021.}
 \end{figure}
@@
 The actual scoring of forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
-The function automatically detects the forecast type and format, applies the appropriate scoring metrics and aggregates results as desired by the user.
+The function automatically detects the forecast type and format (as shown in the output of \fct{check\_forecasts}), applies the appropriate scoring metrics and can in addition also aggregate results as well as perform pairwise comparisons between models.
 Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
-The function requires a \code{data.frame} or similar with some pre-specified columns (e.g.
-a column called "prediction", one called "true\_value" and more depending on the forecast format).
-Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
-To illustrate the requirements, \pkg{scoringutils} provides example data for all possible forecast formats.
-It also offers functions to to transform between various formats, e.g.
-from a sample based format to a quantile format.
 
-When conducting an evaluation, we need to decide on the level of grouping over which we want to aggregate scores.
-For the UK short-term forecasts it makes sense to look at the scores achieved by every model separate for all prediction targets.
-This can be achieved by calling
-% ion G
-<<scoring, eval = TRUE>>=
-scores <- eval_forecasts(combined, 
-                         summarise_by = c("model", "value_desc"))
-
+<<>>=
+scores <- eval_forecasts(example_quantile)
 glimpse(scores)
 @
+
+The above produces one score for every forecast. However, we usually like to summarise scores to learn about average performance.
+
+We can either run a separate function to summarise the scores, 
+<<>>=
+summarised_scores <- 
+  summarise_scores(scores = scores, 
+                   summarise_by = c("model", "target_type"))
+@
+
+or we score and summarise in one function call:
+
+<<>>=
+summarised_scores <- 
+  eval_forecasts(example_quantile, 
+                 summarise_by = c("model", "target_type"))
+@
+
+<<>>=
+summarised_scores
+@
+
 % 
-\fct{eval\_forecasts} aggregates scores by taking the mean over according to the grouping specified in \code{summarise_by}.
-In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
-For a more detailed analysis we could for example specifiy \code{summarise_by = c("mode", "value_type", "geography")} to additionally stratify by location.
+\fct{eval\_forecasts} and \fct{summarise\_scores} aggregate scores by taking the mean over according to the grouping specified in \code{summarise_by}.
+In the above example, if \code{summarise_by = c("model", "target_type")}, then scores would be averaged over all other categories to obtain one score per model and forecast target type.
+For a more detailed analysis we could for example specify \code{summarise_by = c("mode", "target_type", "location")} to additionally stratify by country.
+Summarised scores can be visualised using the function \fct{scores\_table} as is shown in Figure \ref{fig:score-table}.
+
 If we wanted to have one score per quantile or one per prediction interval range, we could use something like \code{summarise_by = c("model", "quantile", "range")}.
@@
 In the given example, looking at one score per model (i.e.
-specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
+specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by case forecasts, while errors on death forecasts would have little influence.
 Similarly, aggregating over different forecast horizons is often not a good idea as the mean will be dominated by further ahead forecast horizons.
 
-The second important argument of the \fct{eval\_forecasts} function is the \code{by} argument that denotes the unit of a single forecast.
-In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")} as there is one distinct forecast for each combination of these categories.
-Quantiles do not need to be included, as several quantiles make up one forecast (and similarly for samples).
-By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
-In the above example, we did not have to explicitly specify the \code{by} argument, as all available columns were relevant to the unit of a single forecast.
-If there are additional columns that do not correspond to the unit of a single forecast, like for example the observed temperature in a given location, \code{by} must be specified and these columns must be excluded.
+\begin{figure}[h]
+\centering
+<<score-table, echo=TRUE, fig.height=7.2, fig.width=15, eval = TRUE>>=
+score_table(summarised_scores, y = "model", facet_formula = ~ target_type)
+@
+\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
+\end{figure}
 
@@
 the WIS or the CRPS.
+
+Again there are two possible ways to do pairwise comparisons.
+First, the function \fct{pairiwse\_comparison} can be called to compare models against each other.
+Here \code{summarise_by} denotes what relative skill scores shall be computed for.
+\code{summarise_by = "model"} means there will be one relative skill score per model, while \code{summarise_by = c("target_type", "target_type")} would mean there is one relative skill score per model, computed completely separately for different target types.
+
+When using \fct{pairwise\_comparison}, it is important that the scores used as input have not been summarised in any way.
+
+<<>>=
+pairwise <- pairwise_comparison(scores, 
+                                baseline = "EuroCOVIDhub-baseline")
+
+pairwise2 <- pairwise_comparison(scores, 
+                                summarise_by = c("model", "target_type"), 
+                                baseline = "EuroCOVIDhub-baseline")
+@
 
@@
 
-<<pairwise-comparison, eval = TRUE>>=
-# unsummarised scores
-unsummarised_scores <- eval_forecasts(combined)
-pairwise <- pairwise_comparison(unsummarised_scores, 
-                                summarise_by = "value_desc")
+The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
+
+\begin{figure}[h!]
+\centering
+<<pairwise-plot, echo=TRUE, fig.height=7.2, fig.width=14, eval = TRUE>>=
+plot_pairwise_comparison(pairwise2) +
+  ggplot2::facet_wrap(~ target_type, scales = "free_x") 
 @
-
-The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
+\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
+If a tile is blue, then the model on the y-axis performed better.
+If it is red, the model on the x-axis performed better in direct comparison.
+}
+\end{figure}
 
@@
 A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
-
-\begin{figure}[h]
-\centering
-<<score-table, echo=TRUE, fig.height=7.2, fig.width=15, eval = TRUE>>=
-score_table(scores, y = "model", facet_formula = ~ value_desc)
-@
-\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
-\end{figure}
 
@@
 
-\begin{figure}[h!]
-\centering
-<<pairwise-plot, echo=TRUE, fig.height=7.2, fig.width=14, eval = TRUE>>=
-plot_pairwise_comparison(pairwise) +
-  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
-@
-\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
-If a tile is blue, then the model on the y-axis performed better.
-If it is red, the model on the x-axis performed better in direct comparison.
-}
-\end{figure}
-
 In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
@@
 <<WIS-components, echo=TRUE, fig.height=5.2, fig.width=12, eval = TRUE>>=
-wis_components(scores, 
-               facet_formula = ~ value_desc, 
+wis_components(summarised_scores, 
+               facet_formula = ~ target_type, 
                scales = "free_x", 
@@
 
+<<pit-plots, echo=TRUE, fig.height=5.2, fig.width=14, out.width="49%", fig.show='hold', eval = TRUE>>=
+pit <- pit_df(example_quantile, 
+              summarise_by = "model", "target")
+
+hist_PIT(pit)
+@
+
 \begin{figure}[h!]
 \centering
-<<coverage, echo=TRUE, fig.height=5.2, fig.width=14, out.width="40%", fig.show='hold', eval = TRUE>>=
-cov_scores <- eval_forecasts(combined, 
-                             summarise_by = c("model", 
+<<coverage, echo=TRUE, fig.height=5.2, fig.width=14, fig.show='hold', eval = TRUE>>=
+cov_scores <- eval_forecasts(example_quantile, 
+                             summarise_by = c("model", "target_type",
                                               "range", "quantile"))
-scoringutils::interval_coverage(cov_scores)
-scoringutils::quantile_coverage(cov_scores)
+
+scoringutils::interval_coverage(cov_scores, facet_formula = ~ target_type)
+@
+
+<<quantile-coverage, fig.height=5.2, fig.width=14, fig.show='hold', eval = TRUE>>=
+scoringutils::quantile_coverage(cov_scores, facet_formula = ~ target_type)
 @
@@
 % pit-plots, echo=TRUE, fig=TRUE, height=5.2, width=12>>=
-% # out <- eval_forecasts(combined, 
+% # out <- eval_forecasts(example_quantile, 
 % #                       summarise_by = c("model"), 
@@
 <<calibration, echo=TRUE, fig.height=5.2, fig.width=15, eval = TRUE>>=
-scores <- eval_forecasts(combined, 
+scores <- eval_forecasts(example_quantile, 
                          summarise_by = c("model",
-                                          "value_desc", 
-                                          "geography"), 
+                                          "target_type", 
+                                          "location_name"), 
                          compute_relative_skill = FALSE)
-scoringutils::score_heatmap(scores, metric = "bias", 
-                            x = "geography", facet_formula = ~ value_desc, 
+scoringutils::score_heatmap(scores, metric = "interval_score", 
+                            x = "location_name", facet_formula = ~ target_type, 
                             scale = "free_x")

### 2021-12-12 article-knitr.tex
update paper and plots

@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=0.98\textwidth]{/mnt/data/GoogleDrive/Uni/PhD/scoringutils-paper/plots/calibration-diagnostic-examples} 
+\includegraphics[width=\maxwidth]{/mnt/data/GoogleDrive/Uni/PhD/scoringutils-paper/plots/calibration-diagnostic-examples} 
 \end{knitrout}
@@
 
-\subsection{Preparing and visualising the data}
+\subsection{Checking the data}
 
-We first need to obtain the data by installing and loading the \pkg{covid19.forecasts.uk} using the following commands: 
 % 
@@
 \begin{alltt}
-\hlcom{# install and load data pacakge from external repository}
-\hlcom{# remotes::install_github('sbfnk/covid19.forecasts.uk')}
-
-\hlcom{# load packages}
-\hlkwd{library}\hlstd{(covid19.forecasts.uk)}
-\hlkwd{library}\hlstd{(dplyr)}
-\hlkwd{library}\hlstd{(scoringutils)}
-
-\hlcom{# load truth data}
-\hlkwd{data}\hlstd{(covid_uk_data)}
-\hlcom{# head(covid_uk_data)}
-
-\hlcom{# load forecasts}
-\hlkwd{data}\hlstd{(uk_forecasts)}
-\hlcom{# head(uk_forecasts)}
+\hlstd{> }\hlcom{# load packages}
+\hlstd{> }\hlkwd{library}\hlstd{(scoringutils)}
+\hlstd{> }\hlkwd{library}\hlstd{(dplyr)}
+\hlstd{> }\hlkwd{library}\hlstd{(data.table)}
+\hlstd{> }\hlkwd{library}\hlstd{(kableExtra)}
+\hlstd{> }
+\hlstd{> }\hlcom{# load forecasts}
+\hlstd{> }\hlkwd{data}\hlstd{(example_quantile)}
 \end{alltt}
@@
 % 
-Let us take a first look a the data:
+All higher level functions that allow for convenient evaluation of forecasts expect a \code{data.frame} or similar.
+Lower-level functions (discussed in a later section) accept inputs as vectors and matrices.
+
+We omit \code{NA} values as the data set also contains entries for which we have an observed value, but no forecast. 
 % 
@@
 \begin{alltt}
-\hlkwd{glimpse}\hlstd{(covid_uk_data)}
+\hlstd{> }\hlstd{example_quantile |>}
+\hlstd{+ }    \hlkwd{na.omit}\hlstd{() |>}
+\hlstd{+ }    \hlkwd{glimpse}\hlstd{()}
 \end{alltt}
 \begin{verbatim}
-## Rows: 4,174
-## Columns: 6
-## $ geography  <fct> London, London, London, London, London, London, L~
-## $ value_type <fct> hospital_inc, hospital_inc, hospital_inc, hospita~
-## $ value_desc <fct> Hospital admissions, Hospital admissions, Hospita~
-## $ truncation <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0~
-## $ value_date <date> 2020-03-20, 2020-03-21, 2020-03-22, 2020-03-23, ~
-## $ value      <dbl> 18, 232, 280, 241, 313, 353, 516, 637, 677, 546, ~
-\end{verbatim}
-\begin{alltt}
-\hlkwd{glimpse}\hlstd{(uk_forecasts)}
-\end{alltt}
-\begin{verbatim}
-## Rows: 1,254,513
-## Columns: 8
-## $ model         <fct> EpiSoon, EpiSoon, EpiSoon, EpiSoon, EpiSoon, E~
-## $ geography     <chr> "East of England", "East of England", "East of~
-## $ value_type    <fct> hospital_inc, hospital_inc, hospital_inc, hosp~
-## $ creation_date <date> 2020-03-31, 2020-03-31, 2020-03-31, 2020-03-3~
-## $ value_date    <date> 2020-04-01, 2020-04-02, 2020-04-03, 2020-04-0~
-## $ quantile      <dbl> 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05~
-## $ value         <dbl> 70, 64, 53, 42, 30, 19, 4, 0, 0, 0, 0, 0, 0, 0~
-## $ value_desc    <fct> Hospital admissions, Hospital admissions, Hosp~
+## Rows: 20,401
+## Columns: 10
+## $ location        <chr> "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE", ~
+## $ target_end_date <date> 2021-05-08, 2021-05-08, 2021-05-08, 2021-05-08, 2021-~
+## $ target_type     <chr> "Cases", "Cases", "Cases", "Cases", "Cases", "Cases", ~
+## $ true_value      <dbl> 106987, 106987, 106987, 106987, 106987, 106987, 106987~
+## $ location_name   <chr> "Germany", "Germany", "Germany", "Germany", "Germany",~
+## $ forecast_date   <chr> "2021-05-03", "2021-05-03", "2021-05-03", "2021-05-03"~
+## $ quantile        <dbl> 0.010, 0.025, 0.050, 0.100, 0.150, 0.200, 0.250, 0.300~
+## $ prediction      <int> 82466, 86669, 90285, 95341, 99171, 102990, 105962, 108~
+## $ model           <chr> "EuroCOVIDhub-ensemble", "EuroCOVIDhub-ensemble", "Eur~
+## $ horizon         <dbl> 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, ~
 \end{verbatim}
@@
 %
-Some minor changes need to be made to the data before we can evaluate the forecasts.
-The names of the columns that hold the forecasts and the true observed values need to be changed to \code{prediction} and \code{true_value}.
-While we could also proceed with separate data sets for the evaluation, we merge the two data sets in order to remove all instances where the forecasts, but not the true observations were made public.
-The \pkg{scoringutils} package provides a function that attempts to merge the data sets in a sensible way.
-% 
+
+For forecasts in a quantile-based format it is necessary to have columns called \code{true_value}, \code{prediction} and \code{quantile}.
+It is recommended (although not necessary except in order to do pairwise comparisons) to have a column called "model" with an identifier for the forecaster.
+Table \ref{tab:column-requirements} shows the columns expected for different input formats.
+
 \begin{knitrout}
@@
 \begin{alltt}
-\hlstd{uk_forecasts} \hlkwb{<-} \hlkwd{rename}\hlstd{(uk_forecasts,} \hlkwc{prediction} \hlstd{= value)}
-\hlstd{covid_uk_data} \hlkwb{<-} \hlkwd{rename}\hlstd{(covid_uk_data,} \hlkwc{true_value} \hlstd{= value)}
-
-\hlstd{combined} \hlkwb{<-} \hlkwd{merge_pred_and_obs}\hlstd{(uk_forecasts, covid_uk_data)}
+\hlstd{> }\hlstd{requirements} \hlkwb{<-} \hlkwd{data.table}\hlstd{(}\hlkwc{Format} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"quantile-based"}\hlstd{,} \hlstr{"sample-based"}\hlstd{,}
+\hlstd{+ }    \hlstr{"binary"}\hlstd{,} \hlstr{"pairwise-comparisons"}\hlstd{),} \hlkwc{`Required columns`} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"'true_value', 'prediction', 'quantile'"}\hlstd{,}
+\hlstd{+ }    \hlstr{"'true_value', 'prediction', 'sample'"}\hlstd{,} \hlstr{"'true_value', 'prediction'"}\hlstd{,}
+\hlstd{+ }    \hlstr{"additionally a column 'model'"}\hlstd{))}
+\hlstd{> }
+\hlstd{> }\hlstd{requirements |>}
+\hlstd{+ }    \hlkwd{kable}\hlstd{(}\hlkwc{format} \hlstd{=} \hlstr{"latex"}\hlstd{)}
 \end{alltt}
 \end{kframe}
+\begin{tabular}{l|l}
+\hline
+Format & Required columns\\
+\hline
+quantile-based & 'true\_value', 'prediction', 'quantile'\\
+\hline
+sample-based & 'true\_value', 'prediction', 'sample'\\
+\hline
+binary & 'true\_value', 'prediction'\\
+\hline
+pairwise-comparisons & additionally a column 'model'\\
+\hline
+\end{tabular}
+
 \end{knitrout}
-% 
-It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance.
+
+We can check whether the data conforms to the requirements by running \fct{check\_forecasts}.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlkwd{check_forecasts}\hlstd{(example_quantile)}
+\end{alltt}
+\begin{verbatim}
+## $target_type
+## [1] "integer"
+## 
+## $prediction_type
+## [1] "quantile"
+## 
+## $forecast_unit
+## [1] "location"        "target_end_date" "target_type"     "location_name"  
+## [5] "forecast_date"   "model"           "horizon"        
+## 
+## $unique_values
+##                    model location target_end_date target_type true_value
+## 1: EuroCOVIDhub-ensemble        4              12           2         96
+## 2: EuroCOVIDhub-baseline        4              12           2         96
+## 3:  epiforecasts-EpiNow2        4              12           2         95
+## 4:       UMass-MechBayes        4              12           1         48
+##    location_name forecast_date quantile prediction horizon
+## 1:             4            11       23       3969       3
+## 2:             4            11       23       3733       3
+## 3:             4            11       23       3903       3
+## 4:             4            11       23       1058       3
+## 
+## $warnings
+## [1] "Some values for `prediction` are NA in the data provided"
+## 
+## 
+## Based on your input, scoringutils thinks:
+## Forecasts are for a `integer` target using a `quantile` prediction format.
+## The unit of a single forecast is defined by `location`, `target_end_date`, `target_type`, `location_name`, `forecast_date`, `model`, `horizon`.
+If this is not as intended, please DELETE UNNECESSARY columns or add new ones.
+## $unique_values shows how many unique values there are per column per model (across the entire data).
+## 
+## You should be aware of the following warnings:
+## Some values for `prediction` are NA in the data provided
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+This returns a list with different entries giving information about what \pkg{scoringutils} infers from the data.
+\code{target_type} \code{prediction_type} refer to the forecast format.
+\code{forecast_unit} contains a vector of the columns which \pkg{scoringutils} thinks denote the unit of a single forecast.
+This means that in this instance a single forecast (with a set of 23 quantiles) can uniquely be identified by the values in the columns "location", "target\_end\_date", "target\_type", "location\_name", "forecast\_date", "model", "horizon".
+In this example, having "location" as well as "location\_name" included does not make a difference, as they contain duplicated information.
+In general, however, it is strongly advised to remove all unnecessary columns that do not help identify a single forecast.
+\code{unique_values} gives an overview of the number of unique values per column across the entire data set, providing a first hint as to whether the forecast set is complete.
+\code{warnings} shows potential warnings about the data.
+In this example, \pkg{scoringutils} warns that there are observed values present for which there is no corresponding forecast.
+These warnings can often be ignored, but may provide important information.
+If there are errors that cannot be ignored, a list entry \code{errors} will appear.
+
+It is helpful to start the evaluation process by visualising data availability, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance.
 The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts.
@@
 \begin{alltt}
-\hlkwd{show_avail_forecasts}\hlstd{(}\hlkwc{data} \hlstd{= combined,} \hlkwc{x} \hlstd{=} \hlstr{"creation_date"}\hlstd{,} \hlkwc{show_numbers} \hlstd{=} \hlnum{FALSE}\hlstd{,}
-    \hlkwc{legend_position} \hlstd{=} \hlstr{"bottom"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc)}
+\hlstd{> }\hlkwd{show_avail_forecasts}\hlstd{(}\hlkwc{data} \hlstd{= example_quantile,} \hlkwc{x} \hlstd{=} \hlstr{"target_end_date"}\hlstd{,}
+\hlstd{+ }    \hlkwc{show_numbers} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{legend_position} \hlstd{=} \hlstr{"bottom"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type)}
 \end{alltt}
 \end{kframe}
-\includegraphics[width=0.98\textwidth]{plots/plot-show-availability-1} 
+\includegraphics[width=\maxwidth]{plots/plot-show-availability-1} 
 \end{knitrout}
@@
 Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data.
-To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: %
+To display, for example, short-term forecasts for COVID-19 cases and deaths made by the EuroCOVIDhub-ensemble model on June 28 2021 as well as 5 weeks of prior data, we can call: %
 \begin{figure}[h!]
@@
 \begin{alltt}
-\hlstd{locations} \hlkwb{<-} \hlstr{"c(\textbackslash{}"Scotland\textbackslash{}", \textbackslash{}"Wales\textbackslash{}", \textbackslash{}"Northern Ireland\textbackslash{}", \textbackslash{}"England\textbackslash{}")"}
-\hlkwd{plot_predictions}\hlstd{(}\hlkwc{truth} \hlstd{= covid_uk_data,} \hlkwc{forecasts} \hlstd{= uk_forecasts,} \hlkwc{filter_both} \hlstd{=} \hlkwd{list}\hlstd{(}\hlkwd{paste}\hlstd{(}\hlstr{"geography %in%"}\hlstd{,}
-    \hlstd{locations,} \hlkwc{collapse} \hlstd{=} \hlstr{""}\hlstd{),} \hlstr{"value_type == \textbackslash{}"death_inc_line\textbackslash{}""}\hlstd{),} \hlkwc{filter_truth} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"value_date <= \textbackslash{}"2020-06-22\textbackslash{}""}\hlstd{,}
-    \hlstr{"value_date > \textbackslash{}"2020-06-01\textbackslash{}""}\hlstd{),} \hlkwc{filter_forecasts} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"model == \textbackslash{}"SIRCOVID\textbackslash{}""}\hlstd{,}
-    \hlstr{"creation_date == \textbackslash{}"2020-06-22\textbackslash{}""}\hlstd{),} \hlkwc{x} \hlstd{=} \hlstr{"value_date"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{geography)} \hlopt{+}
-    \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
+\hlstd{> }\hlkwd{plot_predictions}\hlstd{(}\hlkwc{data} \hlstd{= example_quantile,} \hlkwc{x} \hlstd{=} \hlstr{"target_end_date"}\hlstd{,}
+\hlstd{+ }    \hlkwc{filter_truth} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"target_end_date <= \textbackslash{}"2021-07-15\textbackslash{}""}\hlstd{,}
+\hlstd{+ }        \hlstr{"target_end_date > \textbackslash{}"2021-05-22\textbackslash{}""}\hlstd{),} \hlkwc{filter_forecasts} \hlstd{=} \hlkwd{list}\hlstd{(}\hlstr{"model == 'EuroCOVIDhub-ensemble'"}\hlstd{,}
+\hlstd{+ }        \hlstr{"forecast_date == \textbackslash{}"2021-06-28\textbackslash{}""}\hlstd{),} \hlkwc{facet_formula} \hlstd{= target_type} \hlopt{~}
+\hlstd{+ }        \hlstd{location,} \hlkwc{ncol} \hlstd{=} \hlnum{4}\hlstd{)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
 \end{alltt}
 \end{kframe}
+\includegraphics[width=\maxwidth]{plots/plot-show-forecasts-1} 
 \end{knitrout}
-\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
+\caption{\label{fig:forecast-visualisation} Short-term forecasts for COVID-19 cases and deaths made by the EuroCOVIDhub-ensemble model on June 28 2021.}
 \end{figure}
@@
 The actual scoring of forecasts based on observed values can be performed using the function \fct{eval\_forecasts}.
-The function automatically detects the forecast type and format, applies the appropriate scoring metrics and aggregates results as desired by the user.
+The function automatically detects the forecast type and format (as shown in the output of \fct{check\_forecasts}), applies the appropriate scoring metrics and can in addition also aggregate results as well as perform pairwise comparisons between models.
 Internally, operations are handled using \pkg{data.table} to allow for fast and efficient computation.
-The function requires a \code{data.frame} or similar with some pre-specified columns (e.g.
-a column called "prediction", one called "true\_value" and more depending on the forecast format).
-Additional columns may be present to indicate a grouping of forecasts, for example forecasts made in different locations or over different forecast horizons.
-To illustrate the requirements, \pkg{scoringutils} provides example data for all possible forecast formats.
-It also offers functions to to transform between various formats, e.g.
-from a sample based format to a quantile format.
 
-When conducting an evaluation, we need to decide on the level of grouping over which we want to aggregate scores.
-For the UK short-term forecasts it makes sense to look at the scores achieved by every model separate for all prediction targets.
-This can be achieved by calling
-% ion G
 \begin{knitrout}
@@
 \begin{alltt}
-\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"value_desc"}\hlstd{))}
-
-\hlkwd{glimpse}\hlstd{(scores)}
+\hlstd{> }\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(example_quantile)}
+\hlstd{> }\hlkwd{glimpse}\hlstd{(scores)}
 \end{alltt}
 \begin{verbatim}
-## Rows: 29
-## Columns: 9
-## $ model              <fct> EpiSoon, StructuredODE, Microsimulation, ~
-## $ value_desc         <fct> Hospital admissions, Hospital admissions,~
-## $ interval_score     <dbl> 38.56510, 61.75914, 100.71914, 35.44876, ~
-## $ dispersion         <dbl> 22.886153, 6.133974, 33.985553, 10.365082~
-## $ underprediction    <dbl> 3.66782766, 24.82255866, 33.12350598, 8.5~
-## $ overprediction     <dbl> 12.0111222, 30.8026045, 33.6100814, 16.55~
-## $ coverage_deviation <dbl> 0.034531405, -0.268749094, -0.095652174, ~
-## $ bias               <dbl> 0.26194170, -0.27368253, -0.39217045, -0.~
-## $ aem                <dbl> 45.96333, 73.38096, 138.14395, 53.22563, ~
+## Rows: 887
+## Columns: 14
+## $ location           <chr> "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE", "DE~
+## $ target_end_date    <date> 2021-05-08, 2021-05-08, 2021-05-08, 2021-05-08, 20~
+## $ target_type        <chr> "Cases", "Cases", "Cases", "Deaths", "Deaths", "Dea~
+## $ location_name      <chr> "Germany", "Germany", "Germany", "Germany", "German~
+## $ forecast_date      <chr> "2021-05-03", "2021-05-03", "2021-05-03", "2021-05-~
+## $ model              <chr> "EuroCOVIDhub-baseline", "EuroCOVIDhub-ensemble", "~
+## $ horizon            <dbl> 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 1, 1, 1, 2, 2, 2, 2, ~
+## $ interval_score     <dbl> 16925.04696, 7990.85478, 25395.96087, 46.79304, 53.~
+## $ dispersion         <dbl> 1649.22087, 5440.98522, 8173.70000, 44.66261, 53.27~
+## $ underprediction    <dbl> 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.60869~
+## $ overprediction     <dbl> 15275.826087, 2549.869565, 17222.260870, 2.130435, ~
+## $ coverage_deviation <dbl> -0.38521739, 0.04956522, -0.29826087, 0.22347826, 0~
+## $ bias               <dbl> 0.95, 0.50, 0.90, 0.30, -0.10, -0.50, 0.10, 1.00, 0~
+## $ aem                <dbl> 25620, 12271, 44192, 15, 14, 208, 24, 67622, 45731,~
 \end{verbatim}
@@
 \end{knitrout}
+
+The above produces one score for every forecast. However, we usually like to summarise scores to learn about average performance.
+
+We can either run a separate function to summarise the scores, 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlstd{summarised_scores} \hlkwb{<-} \hlkwd{summarise_scores}\hlstd{(}\hlkwc{scores} \hlstd{= scores,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
+\hlstd{+ }    \hlstr{"target_type"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+or we score and summarise in one function call:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlstd{summarised_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(example_quantile,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
+\hlstd{+ }    \hlstr{"target_type"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlstd{summarised_scores}
+\end{alltt}
+\begin{verbatim}
+##                    model target_type interval_score dispersion underprediction
+## 1: EuroCOVIDhub-baseline       Cases    28483.57465 4102.50094    10284.972826
+## 2: EuroCOVIDhub-ensemble       Cases    17943.82383 3663.52458     4237.177310
+## 3:  epiforecasts-EpiNow2       Cases    20831.55662 5664.37795     3260.355639
+## 4: EuroCOVIDhub-baseline      Deaths      159.40387   91.40625        2.098505
+## 5: EuroCOVIDhub-ensemble      Deaths       41.42249   30.18099        4.103261
+## 6:       UMass-MechBayes      Deaths       52.65195   26.87239       16.800951
+## 7:  epiforecasts-EpiNow2      Deaths       66.64282   31.85692       15.893314
+##    overprediction coverage_deviation        bias         aem
+## 1:   14096.100883        -0.11211957  0.09796875 38473.60156
+## 2:   10043.121943        -0.09785326 -0.05640625 24101.07031
+## 3:   11906.823030        -0.06660326 -0.07890625 27923.81250
+## 4:      65.899117         0.11614130  0.33906250   233.25781
+## 5:       7.138247         0.19528533  0.07265625    53.13281
+## 6:       8.978601        -0.02312500 -0.02234375    78.47656
+## 7:      18.892583        -0.04287176 -0.00512605   104.74790
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
 % 
-\fct{eval\_forecasts} aggregates scores by taking the mean over according to the grouping specified in \code{summarise_by}.
-In the above example, if \code{summarise_by = c("model", "value_type")}, then scores would be averaged over all creation dates, forecast horizons (as represented by the the value dates), locations and quantiles to yield one score per model and forecast target.
-For a more detailed analysis we could for example specifiy \code{summarise_by = c("mode", "value_type", "geography")} to additionally stratify by location.
+\fct{eval\_forecasts} and \fct{summarise\_scores} aggregate scores by taking the mean over according to the grouping specified in \code{summarise_by}.
+In the above example, if \code{summarise_by = c("model", "target_type")}, then scores would be averaged over all other categories to obtain one score per model and forecast target type.
+For a more detailed analysis we could for example specify \code{summarise_by = c("mode", "target_type", "location")} to additionally stratify by country.
+Summarised scores can be visualised using the function \fct{scores\_table} as is shown in Figure \ref{fig:score-table}.
+
 If we wanted to have one score per quantile or one per prediction interval range, we could use something like \code{summarise_by = c("model", "quantile", "range")}.
@@
 In the given example, looking at one score per model (i.e.
-specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by hospital admissions, while errors on death forecasts would have little influence.
+specifying \code{summarise_by = c("model")}) is problematic, as overall aggregate scores would be dominated by case forecasts, while errors on death forecasts would have little influence.
 Similarly, aggregating over different forecast horizons is often not a good idea as the mean will be dominated by further ahead forecast horizons.
 
-The second important argument of the \fct{eval\_forecasts} function is the \code{by} argument that denotes the unit of a single forecast.
-In the above example, the unit of a single forecast would be \code{by = c("model", "geography", "value_type", "creation_date", "value_date", "value_desc")} as there is one distinct forecast for each combination of these categories.
-Quantiles do not need to be included, as several quantiles make up one forecast (and similarly for samples).
-By default, if \code{by = NULL}, \fct{eval\_forecasts} will automatically use all present columns to determine the unit of a single forecast.
-In the above example, we did not have to explicitly specify the \code{by} argument, as all available columns were relevant to the unit of a single forecast.
-If there are additional columns that do not correspond to the unit of a single forecast, like for example the observed temperature in a given location, \code{by} must be specified and these columns must be excluded.
+\begin{figure}[h]
+\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlkwd{score_table}\hlstd{(summarised_scores,} \hlkwc{y} \hlstd{=} \hlstr{"model"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\maxwidth]{plots/plot-score-table-1} 
+\end{knitrout}
+\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
+\end{figure}
 
@@
 the WIS or the CRPS.
+
+Again there are two possible ways to do pairwise comparisons.
+First, the function \fct{pairiwse\_comparison} can be called to compare models against each other.
+Here \code{summarise_by} denotes what relative skill scores shall be computed for.
+\code{summarise_by = "model"} means there will be one relative skill score per model, while \code{summarise_by = c("target_type", "target_type")} would mean there is one relative skill score per model, computed completely separately for different target types.
+
+When using \fct{pairwise\_comparison}, it is important that the scores used as input have not been summarised in any way.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlstd{pairwise} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(scores,} \hlkwc{baseline} \hlstd{=} \hlstr{"EuroCOVIDhub-baseline"}\hlstd{)}
+\hlstd{> }
+\hlstd{> }\hlstd{pairwise2} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(scores,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
+\hlstd{+ }    \hlstr{"target_type"}\hlstd{),} \hlkwc{baseline} \hlstd{=} \hlstr{"EuroCOVIDhub-baseline"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
+The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
+
+\begin{figure}[h!]
+\centering
 \begin{knitrout}
@@
 \begin{alltt}
-\hlcom{# unsummarised scores}
-\hlstd{unsummarised_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined)}
-\hlstd{pairwise} \hlkwb{<-} \hlkwd{pairwise_comparison}\hlstd{(unsummarised_scores,} \hlkwc{summarise_by} \hlstd{=} \hlstr{"value_desc"}\hlstd{)}
+\hlstd{> }\hlkwd{plot_pairwise_comparison}\hlstd{(pairwise2)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{target_type,}
+\hlstd{+ }    \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
 \end{alltt}
 \end{kframe}
+\includegraphics[width=\maxwidth]{plots/plot-pairwise-plot-1} 
 \end{knitrout}
-
-The result is a \code{data.table} with different scores and metrics in a tidy format that can easily be used for further manipulation and plotting.
+\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
+If a tile is blue, then the model on the y-axis performed better.
+If it is red, the model on the x-axis performed better in direct comparison.
+}
+\end{figure}
 
@@
 A good starting point for an evaluation is the following score table that visusalises the scores we produced above. We can facet the table to account for the different forecast targets:
-
-\begin{figure}[h]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{score_table}\hlstd{(scores,} \hlkwc{y} \hlstd{=} \hlstr{"model"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=0.98\textwidth]{plots/plot-score-table-1} 
-\end{knitrout}
-\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
-\end{figure}
 
@@
 
-\begin{figure}[h!]
-\centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{plot_pairwise_comparison}\hlstd{(pairwise)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{value_desc,} \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=0.98\textwidth]{plots/plot-pairwise-plot-1} 
-\end{knitrout}
-\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets.
-If a tile is blue, then the model on the y-axis performed better.
-If it is red, the model on the x-axis performed better in direct comparison.
-}
-\end{figure}
-
 In terms of actually understanding \textit{why} one model performs well or badly, the other metrics shown in Figure \ref{fig:score-table} provide additional insight. We turn to them in the following.
@@
 \begin{alltt}
-\hlkwd{wis_components}\hlstd{(scores,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc,} \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{,}
-    \hlkwc{relative_contributions} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{x_text_angle} \hlstd{=} \hlnum{0}\hlstd{)} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{coord_flip}\hlstd{()} \hlopt{+}
-    \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
+\hlstd{> }\hlkwd{wis_components}\hlstd{(summarised_scores,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type,}
+\hlstd{+ }    \hlkwc{scales} \hlstd{=} \hlstr{"free_x"}\hlstd{,} \hlkwc{relative_contributions} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{x_text_angle} \hlstd{=} \hlnum{0}\hlstd{)} \hlopt{+}
+\hlstd{+ }    \hlstd{ggplot2}\hlopt{::}\hlkwd{coord_flip}\hlstd{()} \hlopt{+} \hlstd{ggplot2}\hlopt{::}\hlkwd{theme}\hlstd{(}\hlkwc{legend.position} \hlstd{=} \hlstr{"bottom"}\hlstd{)}
 \end{alltt}
 \end{kframe}
-\includegraphics[width=0.98\textwidth]{plots/plot-WIS-components-1} 
+\includegraphics[width=\maxwidth]{plots/plot-WIS-components-1} 
 \end{knitrout}
@@
 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlstd{pit} \hlkwb{<-} \hlkwd{pit_df}\hlstd{(example_quantile,} \hlkwc{summarise_by} \hlstd{=} \hlstr{"model"}\hlstd{,} \hlstr{"target"}\hlstd{)}
+\hlstd{> }
+\hlstd{> }\hlkwd{hist_PIT}\hlstd{(pit)}
+\end{alltt}
+\begin{verbatim}
+## $`EuroCOVIDhub-baseline`
+## 
+## $`EuroCOVIDhub-ensemble`
+## 
+## $`epiforecasts-EpiNow2`
+## 
+## $`UMass-MechBayes`
+\end{verbatim}
+\end{kframe}
+
+\includegraphics[width=0.49\linewidth]{plots/plot-pit-plots-1} 
+\includegraphics[width=0.49\linewidth]{plots/plot-pit-plots-2} 
+\includegraphics[width=0.49\linewidth]{plots/plot-pit-plots-3} 
+\includegraphics[width=0.49\linewidth]{plots/plot-pit-plots-4} 
+\end{knitrout}
+
 \begin{figure}[h!]
@@
 \begin{alltt}
-\hlstd{cov_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"range"}\hlstd{,}
-    \hlstr{"quantile"}\hlstd{))}
-\hlstd{scoringutils}\hlopt{::}\hlkwd{interval_coverage}\hlstd{(cov_scores)}
-\hlstd{scoringutils}\hlopt{::}\hlkwd{quantile_coverage}\hlstd{(cov_scores)}
+\hlstd{> }\hlstd{cov_scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(example_quantile,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
+\hlstd{+ }    \hlstr{"target_type"}\hlstd{,} \hlstr{"range"}\hlstd{,} \hlstr{"quantile"}\hlstd{))}
+\hlstd{> }
+\hlstd{> }\hlstd{scoringutils}\hlopt{::}\hlkwd{interval_coverage}\hlstd{(cov_scores,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type)}
 \end{alltt}
 \end{kframe}
-\includegraphics[width=0.4\linewidth]{plots/plot-coverage-1} 
-\includegraphics[width=0.4\linewidth]{plots/plot-coverage-2} 
+\includegraphics[width=\maxwidth]{plots/plot-coverage-1} 
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{> }\hlstd{scoringutils}\hlopt{::}\hlkwd{quantile_coverage}\hlstd{(cov_scores,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\maxwidth]{plots/plot-quantile-coverage-1} 
 \end{knitrout}
@@
 \begin{alltt}
-\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(combined,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,} \hlstr{"value_desc"}\hlstd{,}
-    \hlstr{"geography"}\hlstd{),} \hlkwc{compute_relative_skill} \hlstd{=} \hlnum{FALSE}\hlstd{)}
-\hlstd{scoringutils}\hlopt{::}\hlkwd{score_heatmap}\hlstd{(scores,} \hlkwc{metric} \hlstd{=} \hlstr{"bias"}\hlstd{,} \hlkwc{x} \hlstd{=} \hlstr{"geography"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{value_desc,}
-    \hlkwc{scale} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
+\hlstd{> }\hlstd{scores} \hlkwb{<-} \hlkwd{eval_forecasts}\hlstd{(example_quantile,} \hlkwc{summarise_by} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"model"}\hlstd{,}
+\hlstd{+ }    \hlstr{"target_type"}\hlstd{,} \hlstr{"location_name"}\hlstd{),} \hlkwc{compute_relative_skill} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+\hlstd{> }\hlstd{scoringutils}\hlopt{::}\hlkwd{score_heatmap}\hlstd{(scores,} \hlkwc{metric} \hlstd{=} \hlstr{"interval_score"}\hlstd{,}
+\hlstd{+ }    \hlkwc{x} \hlstd{=} \hlstr{"location_name"}\hlstd{,} \hlkwc{facet_formula} \hlstd{=} \hlopt{~}\hlstd{target_type,} \hlkwc{scale} \hlstd{=} \hlstr{"free_x"}\hlstd{)}
 \end{alltt}
 \end{kframe}
-\includegraphics[width=0.98\textwidth]{plots/plot-calibration-1} 
+\includegraphics[width=\maxwidth]{plots/plot-calibration-1} 
 \end{knitrout}
